### 2018-01-18 Manuscript.Rmd
Added a notebook to collect all the figures and stats we will use for the manuscript.

---
title: "Exposure Consequences - Statistics and Figures"
output:
  pdf_document: default
  html_notebook: default
---
Here I collect the 'right' statistics and figures for the exposure consequences project.

All the scripts doing the statistics and figures are in separate R files, but here we will only see / do the ones that we think should go to the manuscript.

First we load those other scripts:

```{r}
source('shared.R')
source('nocursor.R')
source('localization.R')
source('relateLocalizationNoCursors.R')
```

# Reach Aftereffects

Here is a plot:

```{r fig.width=10, fig.height=4}
plotReachAftereffects()
```

Messages of the figure:

Panel A: there are substantial and persisting reach aftereffects
Panel B: that are somewhat lower than those in classic (but not that much? but only for some targets?)

### 2018-01-18 nocursor.R

@@ -15,7 +15,7 @@ installRequire.Packages(required.packages)
 
 
 # plot both groups reach aftereffects in one figure
-plotReachAftereffects <- function(validate=FALSE) {
+OLDplotReachAftereffects <- function(validate=FALSE) {
   
   points <- c(15,25,35,45,55,65,75)
   
@@ -51,21 +51,71 @@ plotReachAftereffects <- function(validate=FALSE) {
   
   plot(-1000,-1000, main='reach aftereffects', xlab='target angle [deg]', ylab='reach endpoint deviation [deg]', xlim=c(10,80), ylim=c(0,15), axes=F)
   
-  polygon(X,expY,border=NA,col=rgb(1,0,0,.2))
-  polygon(X,claY,border=NA,col=rgb(0,0,1,.2))
+  polygon(X,claY,border=NA,col=rgb(1,0,0,.2))
+  polygon(X,expY,border=NA,col=rgb(0,0,1,.2))
+
+  lines(points,classicAVG$endpoint_angle,col=rgb(1,0,0))
+  lines(points,exposureAVG$endpoint_angle,col=rgb(0,0,1))
+  lines(points,classicAVGrem$endpoint_angle,col=rgb(1,0,0),lty=2)
+  lines(points,exposureAVGrem$endpoint_angle,col=rgb(0,0,1),lty=2)
+
+  axis(1,at=points)
+  axis(2,at=c(0,5,10,15))
   
-  lines(points,exposureAVG$endpoint_angle,col=rgb(1,0,0))
-  lines(points,classicAVG$endpoint_angle,col=rgb(0,0,1))
-  lines(points,exposureAVGrem$endpoint_angle,col=rgb(1,0,0),lty=2)
-  lines(points,classicAVGrem$endpoint_angle,col=rgb(0,0,1),lty=2)
+  legend(10,15,c('exposure','classic'),col=c(rgb(0,0,1),rgb(1,0,0)),lty=c(1,1),bty='n')
+  
+}
+
+plotReachAftereffects <- function() {
+  
+  par(mfrow=c(1,2))
+  
+  points <- c(15,25,35,45,55,65,75)
+  
+  exposure <- getReachAftereffects('exposure',part='initial')
+  classic  <- getReachAftereffects('classic',part='initial')
+  exposure_rem <- getReachAftereffects('exposure',part='remainder')
+  
[…]

### 2018-01-18 relateLocalizationNoCursors.R

@@ -10,93 +10,140 @@ correlateNoCursorsLocalization <- function() {
   
   # LOCALIZATION SHIFT for exposure
   local.exp <- getANOVAlocalization(group='exposure')
-  loc.exp.ro <- aggregate(taperror_deg ~ participant + passive_b + handangle_deg, data=local.exp[local.exp$rotated_b == 1 & local.exp$passive_b == 0,], FUN=mean, drop=FALSE)
-  loc.exp.al <- aggregate(taperror_deg ~ participant + passive_b + handangle_deg, data=local.exp[local.exp$rotated_b == 0 & local.exp$passive_b == 0,], FUN=mean, drop=FALSE)
-  local.exp <- loc.exp.ro
-  local.exp$taperror_deg <- local.exp$taperror_deg - loc.exp.al$taperror_deg
+  # active
+  loc.exp.act.ro <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.exp[local.exp$rotated_b == 1 & local.exp$passive_b == 0,], FUN=mean, drop=FALSE)
+  loc.exp.act.al <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.exp[local.exp$rotated_b == 0 & local.exp$passive_b == 0,], FUN=mean, drop=FALSE)
+  local.exp.act <- loc.exp.act.ro
+  local.exp.act$taperror_deg <- local.exp.act$taperror_deg - loc.exp.act.al$taperror_deg
+  # passive
+  loc.exp.pas.ro <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.exp[local.exp$rotated_b == 1 & local.exp$passive_b == 1,], FUN=mean, drop=FALSE)
+  loc.exp.pas.al <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.exp[local.exp$rotated_b == 0 & local.exp$passive_b == 1,], FUN=mean, drop=FALSE)
+  local.exp.pas <- loc.exp.pas.ro
+  local.exp.pas$taperror_deg <- local.exp.pas$taperror_deg - loc.exp.pas.al$taperror_deg
   
   # REACH AFTEREFFECTS for exposure
   nocur.exp <- getReachAftereffects(group='exposure', part='all', clean=TRUE) 
-  names(nocur.exp)[which(names(nocur.exp) == 'target')] <- 'handangle_deg'
-  
+  names(nocur.exp)[names(nocur.exp) == 'endpoint_angle'] <- 'RAE'
   # combine data frames for exposure
[…]

### 2018-01-18 ManuscriptMaterials.Rmd
Work on Notebook / renamed notebook

---
title: "Exposure Consequences - Statistics and Figures"
output:
  pdf_document: default
  html_notebook: default
---

# Overview

Here I collect the 'right' statistics and figures for the exposure consequences project.

As I think of it right now, we will have these

Figures:

1. Setup & Task
2. Reach Aftereffects
3. Localization
4. Correlation

Maybe not the 4th one.

There will also be a *Table* with tasks, and their order and number of trials in both the classic and exposure version of the task.

Tables:

1. Task order

## Source scripts

All the scripts doing the statistics and figures are in separate R files, but here we will only see / do the ones that we think should go to the manuscript.

First we load those other scripts:

```{r}
source('shared.R') # functions used everywhere
source('nocursor.R') # functions for no-cursor reach data
source('localization.R') # functions for localization data
source('relateLocalizationNoCursors.R') # functions that correlate the two kinds of data
```

## Topics

In the manuscript we'll first show that there are motor changes after exposure training (with the no-cursor data) and then compare that with the classic. Second is the localization: a) localization shifts after exposure b) the shifts are the same/different from those after classic c) the effects are different across the workspace d) this pattern is indistinguishable / different from that after classic. A third topic might be if localization shifts can predict no-cursor changes. Here we'll also follow this order:

1. Reach aftereffects
2. Localization
3. Correlations

# Reach Aftereffects

Here is a plot, probably Figure 2:

```{r fig.width=10, fig.height=4}
plotReachAftereffects()
```

Messages of the figure:

- Panel A: there are substantial and persisting reach aftereffects
- Panel B: that are somewhat lower than those in classic (but not that much? but only for some targets?)

These claims require analyses.

First, we show that no-cursor reaches change direction after exposure training:

### 2018-01-18 nocursor.R

@@ -104,16 +104,16 @@ plotReachAftereffects <- function() {
   
   plot(-1000,-1000, main='reach aftereffects', xlab='target angle [deg]', ylab='reach endpoint deviation [deg]', xlim=c(10,80), ylim=c(0,15), axes=F)
   
-  polygon(X,claY, border=NA,col=rgb(0.75, 0,    0,   .2))
+  polygon(X,claY, border=NA,col=rgb(1,    0,    0,   .2))
   polygon(X,expY, border=NA,col=rgb(0,    0,    .75, .2))
   
-  lines(points,classicAVG$endpoint_angle,col=rgb(.75,0,0))
+  lines(points,classicAVG$endpoint_angle,col=rgb(1.0,0,0))
   lines(points,exposureAVG$endpoint_angle,col=rgb(0,0,.75))
 
   axis(1,at=points)
   axis(2,at=c(0,5,10,15))
   
-  legend(10,15,c('exposure','classic'),col=c(rgb(0,0,.75),rgb(.75,0,0)),lty=c(1,1),bty='n')
+  legend(10,15,c('exposure','classic'),col=c(rgb(0,0,.75),rgb(1,0,0)),lty=c(1,1),bty='n')
   
 }
 

### 2018-01-18 shared.R
changed the function 'getReachAftereffects' to allow it to also return the aligned and rotated session without calculating the difference, and commented a lot in the function

@@ -279,14 +279,17 @@ getLocalizationPoints <- function(subdf, points=c(15,25,35,45,55,65,75), removeO
 
 # for every group, this loads the no-cursor reach directions in all relevant tasks,
 # and calcuates the reach aftereffects from them
-getReachAftereffects <- function(group, part='all', clean=TRUE) {
+getReachAftereffects <- function(group, part='all', clean=TRUE, difference=TRUE) {
   
+  # load pre-processed data for the required group (no default)
   raw.df <- load.DownloadDataframe(url=nocursorURLs[group],filename=sprintf('nocursor_%s.csv',group))
   
+  # remove outliers if requested (default: yes)
   if (clean) {
     clean.df <- removeOutliers(raw.df) 
   }
   
+  # select part of the data if requested (default: use all of it)
   if (part == 'initial') {
     raw.df <- rbind(raw.df[which(raw.df$rotated == 0),], raw.df[which(raw.df$rotated == 1 & raw.df$repetition == 0),])
   }
@@ -294,11 +297,16 @@ getReachAftereffects <- function(group, part='all', clean=TRUE) {
     raw.df <- rbind(raw.df[which(raw.df$rotated == 0),], raw.df[which(raw.df$rotated == 1 & raw.df$repetition > 0),])
   }
   
+  # average across the task iterations and target repetitions
   avg.df <- aggregate(endpoint_angle ~ participant + rotated + target, data=raw.df, FUN=mean)
   
-  RAE <- aggregate(endpoint_angle ~ participant + target, data=avg.df, FUN=diff)
+  # get the difference if requested (default: yes)
+  if (difference) {
+    avg.df <- aggregate(endpoint_angle ~ participant + target, data=avg.df, FUN=diff)
+  }
   
-  return(RAE)
+  # return te result:
+  return(avg.df)
   
 }
 

### 2018-01-18 nocursor.R
Added LME analysis to show effect of training on no-cursor reach directions to the notebook.

@@ -226,3 +226,22 @@ plotReachAftereffectDistributions <- function() {
   
   
 }
+
+assessNoCursorChange <- function() {
+  
+  nocursor_exposure <- getReachAftereffects('exposure',part='initial', difference=FALSE)
+  
+  default.contrasts <- options('contrasts')
+  options(contrasts=c('contr.sum','contr.poly'))
+  
+  cat('\nLME with session and target as fixed effects, and participant as random effect:\n\n')
+  exp_nocursor_omnibus_model <- lme(endpoint_angle ~ rotated * target, data=nocursor_exposure, random = ~1|participant, na.action=na.exclude)
+  print(Anova(exp_nocursor_omnibus_model, type=3))
+  
+  cat('\nLME with session as fixed effect, and participant and target as random effects:\n\n')
+  exp_nocursor_omnibus_model <- lme(endpoint_angle ~ rotated, data=nocursor_exposure, random = ~1|participant/target, na.action=na.exclude)
+  print(Anova(exp_nocursor_omnibus_model, type=3))
+  
+  options('contrasts' <- default.contrasts)
+  
+}

### 2018-01-19 ManuscriptMaterials.Rmd
Added LME analysis to show effect of training on no-cursor reach directions to the notebook.
New plot and three LME analyses for the no-cursor reach deviations / reach aftereffects.

@@
 output:
+  word_document: default
   pdf_document: default
+  html_document:
+    df_print: paged
   html_notebook: default
@@
 
-As I think of it right now, we will have these
+Here are the planned figures and tables:
 
@@
 
-Maybe not the 4th one.
-
-There will also be a *Table* with tasks, and their order and number of trials in both the classic and exposure version of the task.
+(Maybe not the 4th one.)
 
@@
 First, we show that no-cursor reaches change direction after exposure training:
+
+```{r include=TRUE}
+exposureNoCursorChange()
+```
+
+So this shows that whether or not the feedback was rotated mattered for the reach deviations in the exposure training group, i.e.: there are reach aftereffects.
+
+The second analysis is to show that the reach aftereffects aren't all that different for first and later iterations of the rotated no-cursor tasks.
+(The baseline is always all aligned iterations of the task.)
+
+```{r}
+exposureAftereffectsPersistent()
+```
+
+So there is no effect of iteration here, which means the reach aftereffects are the same for the first iteration and the average across the other four iterations.
+We now also have enough power to see an effect of target angle (the generalization curve is not flat), but more importantly the effect of iteration is also not dependent on target.
+
+Finally, we want to see if there is a difference between teh reach aftereffects observed after exposure training and those after classic training.
+
+```{r}
+exposureClassicReachAftereffects()
+```
+
+I'd have expected there to be some difference between the two training types, perhaps only in interaction with target, but that's not the case apparently.

### 2018-01-19 localization.R

@@ -2,8 +2,10 @@
 source('shared.R')
 
 # first make sure we have all necessary packages installed and loaded:
-# not yet sure if we'll use Chi-square, or Satterthwaite estimates of F/p-values:
-installRequire.Packages(c('nlme', 'car', 'lme4', 'lmerTest'))
+# for now we'll use Chi-square (not Satterthwaite estimates of F/p-values):
+installRequire.Packages(c('nlme', 'car'))
+
+# , 'lme4', 'lmerTest'
 
 # localization data can be downloaded from OSF:
 # groupURLs <- c('exposure'='https://osf.io/9qfhp/download', 'classic'='https://osf.io/upw49/download', 'online'='https://osf.io/wjcgk/download')

### 2018-01-19 nocursor.R

@@ -227,20 +227,79 @@ plotReachAftereffectDistributions <- function() {
   
 }
 
-assessNoCursorChange <- function() {
-  
-  nocursor_exposure <- getReachAftereffects('exposure',part='initial', difference=FALSE)
+exposureNoCursorChange <- function() {
   
   default.contrasts <- options('contrasts')
   options(contrasts=c('contr.sum','contr.poly'))
   
-  cat('\nLME with session and target as fixed effects, and participant as random effect:\n\n')
-  exp_nocursor_omnibus_model <- lme(endpoint_angle ~ rotated * target, data=nocursor_exposure, random = ~1|participant, na.action=na.exclude)
-  print(Anova(exp_nocursor_omnibus_model, type=3))
+  nc.exp <- getReachAftereffects('exposure', part='initial', difference=FALSE)
   
-  cat('\nLME with session as fixed effect, and participant and target as random effects:\n\n')
-  exp_nocursor_omnibus_model <- lme(endpoint_angle ~ rotated, data=nocursor_exposure, random = ~1|participant/target, na.action=na.exclude)
-  print(Anova(exp_nocursor_omnibus_model, type=3))
+  nc.exp$participant <- factor(nc.exp$participant)
+  nc.exp$rotated <- factor(nc.exp$rotated)
+  nc.exp$target <- factor(nc.exp$target)
+  
+  attach(nc.exp)
+  
+  cat('\nLME with session and target as fixed effects, and participant as random effect:\n\n')
+  print(Anova(lme(endpoint_angle ~ rotated * target, random = ~1|participant, na.action=na.exclude), type=3))
+  
+  detach(nc.exp)
+
+  options('contrasts' <- default.contrasts)
+  
+}
+
+exposureAftereffectsPersistent <- function() {
+  
+  default.contrasts <- options('contrasts')
+  options(contrasts=c('contr.sum','contr.poly'))
+  
+  exp1 <- getReachAftereffects('exposure', part='initial', difference=TRUE)
+  exp2 <- getReachAftereffects('exposure', part='remainder', difference=TRUE)
+  
+  exp1$iteration <- 1
+  exp2$iteration <- 2
+  
+  exp <- rbind(exp1, exp2)
+  
+  exp$participant <- factor(exp$participant)
+  exp$iteration <- factor(exp$iteration)
+  exp$target <- factor(exp$target)
+  
+  attach(exp)
+  
[…]

### 2018-01-21 ManuscriptMaterials.Rmd
Standardized plot colors, added correlation section to notebook.

@@
 output:
-  word_document: default
-  pdf_document: default
-  html_document:
-    df_print: paged
-  html_notebook: default
+  pdf_document:
+    number_sections: yes
+  html_notebook:
+    number_sections: yes
 ---
+
+```{r setup, cache=FALSE, include=FALSE}
+library(knitr)
+opts_chunk$set(comment='')
+```
 
@@
 
-Figures:
+**Figures**:
 
-1. Setup & Task
-2. Reach Aftereffects
+1. Setup & Task (goal: visualize the experiment)
+    + A: Setup (side view)
+    + B: Rotated training (top view)
+    + C: No-cursor reaches (top view)
+    + D: Localization (top view)
+2. Reach Aftereffects (goal: show motor changes are there, are robust and compare to classic)
+    + A: Exposure reach aftereffects (initial & later)
+    + B: Classic & Exposure reach aftereffects (initial only? or all?)
 3. Localization
+    + A: Exposure, active and passive (goal: small/no difference between active and passive)
+    + B: Classic, active and passive (goal: larger difference between active and passive, different generalization?)
+
+Maybe:
+
 4. Correlation
+    + A: Exposure, active and passive localization predicting reach aftereffects
+    + B: Classic, active and passive localization predicting reach aftereffects
 
-(Maybe not the 4th one.)
+**Tables**:
 
-Tables:
-
-1. Task order
+1. Task order & trial numbers (goal: clarify task order and show similarity between classic and exposure experiment - so probably after Fig 1)
 
@@
 
+## Is there an effect of exposure training on reach aftereffects?
+
 First, we show that no-cursor reaches change direction after exposure training:
@@
 So this shows that whether or not the feedback was rotated mattered for the reach deviations in the exposure training group, i.e.: there are reach aftereffects.
+
+## Do the motor changes persist during localization?
 
@@
 
-Finally, we want to see if there is a difference between teh reach aftereffects observed after exposure training and those after classic training.
+## Are reach aftereffects comparable between classic and exposure training?
+
+Finally, we want to see if there is a difference between the reach aftereffects observed after exposure training and those after classic training.
 
@@
 I'd have expected there to be some difference between the two training types, perhaps only in interaction with target, but that's not the case apparently.
+
+Let's minimize the potential explanatory power of the target by removing it from the model:
+
+```{r}
+exposureClassicReachAftereffects(noTarget=TRUE)
+```
+
+This doesn't seem to change anything, so that whether or not we account for any effects of target direction, there is no overall effect of training with volitional movements (classic) or passive movements (exposure) on -- the amplitude of -- reach aftereffects.
+
+To sum up:
+
+1. Rotated exposure training evokes substantial shifts in no-cursor reach direction,
+2. these changes persist over time somewhat, and
+3. these changes are comparable to the ones we found after classic training.
+
+# Localization
+
+Here is a plot, probably figure 3:
+
+```{r fig.width=10, fig.height=4}
+plotLocalization()
+```
+
+That looks OK to me.
+It seems that just like after classic training, there is a change in localization responses following exposure training.
+There might still be a difference between active and passive localization after exposure training - contrary to our expectations - but it could be smaller than the difference after classic training.
+This will need to be tested.
+
+## Is there a localization shift after exposure training?
+
+We first want to see if there is any overall effect of aligned versus rotated training in the exposure group, with movement type and hand angle as fixed effects, and particiant as random effect.
+
+```{r}
+exposureLocalizationShift()
+```
+
+Wether or not the feedback was rotated makes a difference according to two terms in the model.
+In other words: exposure training caused localization responses to shift (systematically), as we could already guess by looking at the figure.
+
+This means we can look at the difference between repsonses after rotated and after aligned feedback: the training induced shift in localization.
+
+## Effects of movement type and hand angle in exposure localization
+
+Let's first see if there is a difference between active and passive localization after exposure training.
+
+```{r}
+exposureMovementType()
+```
+
+It says there isn't, but let's make sure by not having hand angle in there as possibly stealing lots of variance.
+
+```{r}
+exposureMovementType(noHandAngle = TRUE)
+```
+
+OK, so that is pretty clear: there is no difference between passive and active localization after exposure training, even with double the number of trials as in classic delayed.
+
+## Comparing localization after exposure and classic training
+
+Now, let's see how this absence of a difference compares to that the (previously found difference) after classic training.
+Let's do this with an LME without all possible combinations of factors.
+We're primarily interested in the possibility that the effect of movement type is there after classic training, but not after exposure (that's our hypothesis).
+We're also somewhat interested in the effect of hand angle being different between classic and exposure training.
+
+```{r}
+exposureClassicLocalization(model='restricted')
+```
+
+This works: the interaction between group and movement type indicates that the effect of movement is different for the exposure and for the classic groups.
+Since it wasn't there at all for the exposure group, this means that there should be an effect of movement type in the classic group, which is an easily testable hypothesis.
+We also see an interaction between group and hand angle, which would mean that the shape of the generalization curve is different for the two groups.
+This seems to also be visible in the data, but would be worth probing a little further.
+
+However, we did this with a restricted model that only contained the terms of interest. Let's see if it is still there with a full model:
+
+```{r}
+exposureClassicLocalization(model='full')
+```
+
+Now it is only a trend, so let's follow this up with some deeper probing.
+In the restricted model there was also an interaction between group and hand angle which has all but vanished in the full model.
+
+## Effect of movement type after classic training
+
+We already saw that in the exposure group there was an effect of hand angle, but not of movement type. How is this in the classic group?
+
+```{r}
+classicMovementType()
+classicMovementType(factors='movementtype')
+```
+
+Alright, so there is an effect of movement type here as well as of hand angle.
+Since the exposure group saw no effect of movement type in either analysis, whereas there is an effect of movement type in the classic group in both analyses, we can be fairly confident that if there is any effect of movement type after exposure training it will be smaller than after classic training.
+This is what we would expect from exposure training.
+
+## Hand angle in classic training
+
+We're still not too sure about the effect of hand angle though.
+In the figures the localization shift for the classic groups _appears_ more homogeneous across the workspace as compared to the exposure group, but is that a real effect?
+
+I'm not sure how to focus an analysis on that question, better than already done, except by using a model with only hand angle:
+
+```{r}
+classicMovementType(factors='handangle')
+```
+
+So yes, there is some generalization effect after classic training. We still want to know if it is different from the effect of hand angle in exposure.
+
+```{r}
+exposureClassicLocalization(model='handangle')
+```
+
+According to all these analyses, the generalization is not appreciably different between classic and exposure training.
+
+# Correlation
+
+In this section we look at how recalibrated proprioception (hand localization) correlates with reach aftereffects.
+
+This plot shows regression lines with localization shifts as predictor of reach aftereffects:
+
+```{r fig.width=7.5, fig.height=4}
+correlations <- correlateNoCursorsLocalization()
+```
+
+It looks like for exposure, active localization is the better predictor of reach aftereffects, whereas in classic it is passive localization. But are any of them real?
+
+## Correlations
+
+As one possibility, we can look at the correlations.
+Not just the p-values, but there are confidence intervals for Pearson's $\rho$, which allow to directly see if one correlation is stronger/weaker than another:
+
+```{r}
+for (name in names(correlations)) {
+  cat(name)
+  print(correlations[[name]][['cortest']])
+}
+```
+
+And those linear regression models?
+
+```{r}
+for (name in names(correlations)) {
+  cat(name)
+  print(correlations[[name]][['linearm']])
+}
+```
+
+They get the same p-value, so that seems fine.

### 2018-01-21 localization.R

@@ -119,3 +119,197 @@ localizationLMEsSatterthwaite <- function() {
   
 }
 
+
+
+
+# PLOTS / FIGURES ------
+
+plotLocalization <- function() {
+  
+  # get the data to plot:
+  exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE)
+  cla <- getPointLocalization('classic', difference=TRUE, verbose=FALSE)
+  
+  # get the averages for the line plots:
+  exp.avg <- aggregate(taperror_deg ~ passive_b + handangle_deg, data=exp, FUN=mean)
+  exp.avg.act <- exp.avg[which(exp.avg$passive_b == 0),]
+  exp.avg.pas <- exp.avg[which(exp.avg$passive_b == 1),]
+  cla.avg <- aggregate(taperror_deg ~ passive_b + handangle_deg, data=cla, FUN=mean)
+  cla.avg.act <- cla.avg[which(cla.avg$passive_b == 0),]
+  cla.avg.pas <- cla.avg[which(cla.avg$passive_b == 1),]
+  
+  # get the confidence intervals for polygon areas:
+  exp.act <- exp[which(exp$passive_b == 0),]
+  exp.CI.act <- matrix(unlist(by(exp.act$taperror_deg, INDICES=c(exp.act$handangle_deg), FUN=t.interval)),nrow=2)
+  exp.pas <- exp[which(exp$passive_b == 1),]
+  exp.CI.pas <- matrix(unlist(by(exp.pas$taperror_deg, INDICES=c(exp.pas$handangle_deg), FUN=t.interval)),nrow=2)
+  cla.act <- cla[which(cla$passive_b == 0),]
+  cla.CI.act <- matrix(unlist(by(cla.act$taperror_deg, INDICES=c(cla.act$handangle_deg), FUN=t.interval)),nrow=2)
+  cla.pas <- cla[which(cla$passive_b == 1),]
+  cla.CI.pas <- matrix(unlist(by(cla.pas$taperror_deg, INDICES=c(cla.pas$handangle_deg), FUN=t.interval)),nrow=2)
+
+  
+  par(mfrow=c(1,2))
+  points <- c(15,25,35,45,55,65,75)
+  
+  # panel A: exposure localization (active vs. passive)
+  
[…]

### 2018-01-21 nocursor.R

@@ -91,29 +91,29 @@ plotReachAftereffects <- function() {
   
   plot(-1000,-1000, main='decay of reach aftereffects', xlab='target angle [deg]', ylab='reach endpoint deviation [deg]', xlim=c(10,80), ylim=c(0,15), axes=F)
   
-  polygon(X,expYrem,border=NA,col=rgb(0.25, 0.25, 1,   .2))
-  polygon(X,expY,   border=NA,col=rgb(0,    0,    .75, .2))
+  polygon(X,expYrem,border=NA,col=colorset[['extra1T']])
+  polygon(X,expY,   border=NA,col=colorset[['expActT']])
   
-  lines(points,exposureAVGrem$endpoint_angle,col=rgb(0.25,0.25,1),lty=2)
-  lines(points,exposureAVG$endpoint_angle,col=rgb(0,0,.75))
+  lines(points,exposureAVGrem$endpoint_angle,col=colorset[['extra1S']],lty=2,lwd=2)
+  lines(points,exposureAVG$endpoint_angle,col=colorset[['expActS']],lwd=2)
   
   axis(1,at=points)
   axis(2,at=c(0,5,10,15))
   
-  legend(10,15,c('immediate (iteration 1)','delayed (iterations 2-5)'),col=c(rgb(0,0,0.75),rgb(0.25,0.25,1)),lty=c(1,2),bty='n')
+  legend(10,15,c('immediate (iteration 1)','delayed (iterations 2-5)'),col=c(colorset[['expActS']],colorset[['extra1S']]),lty=c(1,2),lwd=c(2,2),bty='n')
   
   plot(-1000,-1000, main='reach aftereffects', xlab='target angle [deg]', ylab='reach endpoint deviation [deg]', xlim=c(10,80), ylim=c(0,15), axes=F)
   
-  polygon(X,claY, border=NA,col=rgb(1,    0,    0,   .2))
-  polygon(X,expY, border=NA,col=rgb(0,    0,    .75, .2))
+  polygon(X,claY, border=NA,col=colorset[['claActT']])
+  polygon(X,expY, border=NA,col=colorset[['expActT']])
   
-  lines(points,classicAVG$endpoint_angle,col=rgb(1.0,0,0))
-  lines(points,exposureAVG$endpoint_angle,col=rgb(0,0,.75))
+  lines(points,classicAVG$endpoint_angle,col=colorset[['claActS']],lwd=2)
[…]

### 2018-01-21 relateLocalizationNoCursors.R

@@ -9,7 +9,7 @@ correlateNoCursorsLocalization <- function() {
   # get all data, for classic and exposure, we need localization and reach aftereffects
   
   # LOCALIZATION SHIFT for exposure
-  local.exp <- getANOVAlocalization(group='exposure')
+  local.exp <- getPointLocalization(group='exposure', difference=FALSE, verbose=FALSE)
   # active
   loc.exp.act.ro <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.exp[local.exp$rotated_b == 1 & local.exp$passive_b == 0,], FUN=mean, drop=FALSE)
   loc.exp.act.al <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.exp[local.exp$rotated_b == 0 & local.exp$passive_b == 0,], FUN=mean, drop=FALSE)
@@ -30,7 +30,7 @@ correlateNoCursorsLocalization <- function() {
   exposure$passive_localization_shift <- local.exp.pas$taperror_deg
   
   # LOCALIZATION SHIFT for classic
-  local.cla <- getANOVAlocalization(group='classic')
+  local.cla <- getPointLocalization(group='classic', difference=FALSE, verbose=FALSE)
   # active
   loc.cla.act.ro <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.cla[local.cla$rotated_b == 1 & local.cla$passive_b == 0,], FUN=mean, drop=FALSE)
   loc.cla.act.al <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.cla[local.cla$rotated_b == 0 & local.cla$passive_b == 0,], FUN=mean, drop=FALSE)
@@ -45,8 +45,7 @@ correlateNoCursorsLocalization <- function() {
   # REACH AFTEREFFECTS for classic
   nocur.cla <- getReachAftereffects(group='classic', part='all', clean=TRUE) 
   names(nocur.cla)[names(nocur.cla) == 'endpoint_angle'] <- 'RAE'
-  
-  # cbind the dataframes fro classic:
+  # combine the dataframes for classic:
   classic <- nocur.cla[which(nocur.cla$participant %in% local.cla$participant),]
   classic$active_localization_shift <- local.cla.act$taperror_deg
[…]

### 2018-01-21 shared.R

@@ -6,6 +6,21 @@ nocursorURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'=
 # no-cursor data can be downloaded from OSF:
 localizationURLs <- c('exposure'='https://osf.io/pqsdc/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
 
+# we'll control the color in figures centrally from here:
+colorset <- list()
+
+colorset[['expActS']] <- '#005de4ff'
+colorset[['expActT']] <- '#005de42f'
+colorset[['expPasS']] <- '#2ab2f2ff'
+colorset[['expPasT']] <- '#2ab2f22f'
+colorset[['claActS']] <- '#ff0000ff'
+colorset[['claActT']] <- '#ff00002f'
+colorset[['claPasS']] <- '#ff8200ff'
+colorset[['claPasT']] <- '#ff82002f'
+
+colorset[['extra1S']] <- '#c400c4ff'
+colorset[['extra1T']] <- '#c400c42f'
+
 
 installRequire.Packages <- function(packages) {
   
@@ -143,7 +158,49 @@ parGaussian <- function(par,x) {
 
 # handling localization data -----
 
-getANOVAlocalization <- function(group) {
+# getANOVAlocalization <- function(group) {
+#   
+#   df <- load.DownloadDataframe(url=localizationURLs[group],filename=sprintf('localization_%s.csv',group))
+#   
+#   df <- aspligned(df)
+#   
+#   groupdf <- NA
+#   
+#   participants <- unique(df$participant)
+#   
+#   for (pp.no in c(1:length(participants))) {
+#     
+#     pp.id <- participants[pp.no]
+#     
+#     for (rotated in c(0,1)) {
+#       
+#       for (passive in c(0,1)) {
+#         
+#         subdf <- df[which(df$participant == pp.id & df$rotated_b == rotated & df$passive_b == passive),]
+#         
+#         locdf <- getLocalizationPoints(subdf, points=c(15,25,35,45,55,65,75), removeOutliers=TRUE)
+#         
+#         if (any(is.na(locdf$taperror_deg))) {
+#           cat(sprintf('WARNING: NAs in %s, pp:%s, %s/%s\n',group,pp.id,c('aligned','rotated')[rotated+1],c('active','passive')[passive+1]))
+#         }
+#         
+#         if (is.data.frame(groupdf)) {
+#           groupdf <- rbind(groupdf, locdf)
+#         } else {
[…]

### 2018-01-21 ManuscriptMaterials.Rmd
use York Red as color and allow the correlations on task iteration 1 or 2-5 or both

@@
 
-Here I collect the 'right' statistics and figures for the exposure consequences project.
+This is a collection of statistics and figures for the manuscript on the exposure consequences project.
 
@@
 
-```{r fig.width=10, fig.height=4}
+```{r fig.width=7.5, fig.height=4}
 plotReachAftereffects()
 ```
+
+_For Panel B: We're not really interested in reach aftereffects per se.
+One focus of this plot/analyses could be the motor changes evoked by the two different types of training.
+Then it would make more sense to use the initial rotated no-cursor block (I think this is what it does now), as they are directly evoked by the training, and could reflect the motor changes themselves better.
+Or do we want to focus on the motor changes that are present throughout localization?
+In that case it makes more sense to use the four later tasks, as they decay stops fairly quickly, so these might be said to better reflect the motor changes that are present throughout the localization tasks.
+We could also use some average...
+I can see it many different ways, just want to have a clear strategy here._
 
@@
 
-```{r fig.width=10, fig.height=4}
+```{r fig.width=7.5, fig.height=4}
 plotLocalization()
@@
 
+_I'm not sure if we should use the initial reach aftereffects here, or the four later blocks. Unfortunately, it does matter for the results, but I don't want to choose with hindsight._
+
 This plot shows regression lines with localization shifts as predictor of reach aftereffects:
@@
 
-## Correlations
+## Pearons Correlations
 
@@
 
+## Linear regression models
+
 And those linear regression models?

### 2018-01-21 relateLocalizationNoCursors.R

@@ -3,7 +3,7 @@ source('shared.R')
 
 
 
-correlateNoCursorsLocalization <- function() {
+correlateNoCursorsLocalization <- function(NCpart='all') {
   
   # first we do all responses, plan to split it by target later
   # get all data, for classic and exposure, we need localization and reach aftereffects
@@ -22,7 +22,7 @@ correlateNoCursorsLocalization <- function() {
   local.exp.pas$taperror_deg <- local.exp.pas$taperror_deg - loc.exp.pas.al$taperror_deg
   
   # REACH AFTEREFFECTS for exposure
-  nocur.exp <- getReachAftereffects(group='exposure', part='all', clean=TRUE) 
+  nocur.exp <- getReachAftereffects(group='exposure', part=NCpart, clean=TRUE) 
   names(nocur.exp)[names(nocur.exp) == 'endpoint_angle'] <- 'RAE'
   # combine data frames for exposure
   exposure <- nocur.exp[which(nocur.exp$participant %in% local.exp.act$participant),]
@@ -43,7 +43,7 @@ correlateNoCursorsLocalization <- function() {
   local.cla.pas$taperror_deg <- local.cla.pas$taperror_deg - loc.cla.pas.al$taperror_deg
   
   # REACH AFTEREFFECTS for classic
-  nocur.cla <- getReachAftereffects(group='classic', part='all', clean=TRUE) 
+  nocur.cla <- getReachAftereffects(group='classic', part=NCpart, clean=TRUE) 
   names(nocur.cla)[names(nocur.cla) == 'endpoint_angle'] <- 'RAE'
   # combine the dataframes for classic:
   classic <- nocur.cla[which(nocur.cla$participant %in% local.cla$participant),]

### 2018-01-21 shared.R

@@ -9,16 +9,16 @@ localizationURLs <- c('exposure'='https://osf.io/pqsdc/?action=download', 'class
 # we'll control the color in figures centrally from here:
 colorset <- list()
 
-colorset[['expActS']] <- '#005de4ff'
+colorset[['expActS']] <- '#005de4ff' # blue
 colorset[['expActT']] <- '#005de42f'
-colorset[['expPasS']] <- '#2ab2f2ff'
+colorset[['expPasS']] <- '#2ab2f2ff' # lighter blue
 colorset[['expPasT']] <- '#2ab2f22f'
-colorset[['claActS']] <- '#ff0000ff'
-colorset[['claActT']] <- '#ff00002f'
-colorset[['claPasS']] <- '#ff8200ff'
+colorset[['claActS']] <- '#e51636ff' # "York red"
+colorset[['claActT']] <- '#e516362f'
+colorset[['claPasS']] <- '#ff8200ff' # orange
 colorset[['claPasT']] <- '#ff82002f'
 
-colorset[['extra1S']] <- '#c400c4ff'
+colorset[['extra1S']] <- '#c400c4ff' # purple
 colorset[['extra1T']] <- '#c400c42f'
 
 

### 2018-01-22 ManuscriptMaterials.Rmd
updated reach aftereffect plot to use all data for panel B (and for the comparison between classic and exposure too)

--- 2018-01-21 ManuscriptMaterials.Rmd
@@
 
-```{r fig.width=10, fig.height=4}
+```{r fig.width=7.5, fig.height=4}
 plotReachAftereffects()
 ```
+
+_For Panel B: Earlier I used the initial block for this plot, but after discussion, we decided to use all data: there's no significant difference between the initial and other blocks, but using all data allows more accurate estimates._
 
@@
 
-First, we show that no-cursor reaches change direction after exposure training:
+First, we show that no-cursor reaches change direction immediately after exposure training (iteration 1):
 
@@
 
-Let's minimize the potential explanatory power of the target by removing it from the model:
+The absence of an interaction between training and target could be interpreted to mean that the generalization of reach aftereffects is the same after both kinds of training.
+Here are the confidence intervals for both:
+
+```{r}
+cat('\nCLASSIC\n\n')
+getPeakConfidenceInterval('classic',part='all',CIs=c(.95))
+cat('\nEXPOSURE\n\n')
+getPeakConfidenceInterval('exposure',part='all',CIs=c(.95))
+```
+
+Although the two peaks are not different from each other, the 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training), whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training).
+So something might be going on -- but the evidence is far from concrete.
+
+Let's go back to seeing if there is any effect from training type on reach aftereffects at all, and minimize the potential explanatory power of target by removing it from the model:
 
@@
 
-This doesn't seem to change anything, so that whether or not we account for any effects of target direction, there is no overall effect of training with volitional movements (classic) or passive movements (exposure) on -- the amplitude of -- reach aftereffects.
+This doesn't seem to change one iota, so that whether or not we account for any effects of target direction, there is no overall effect of training with volitional movements (classic) or passive movements (exposure) on -- the amplitude of -- reach aftereffects.
 

### 2018-01-22 nocursor.R

@@ -72,35 +72,39 @@ plotReachAftereffects <- function() {
   
   points <- c(15,25,35,45,55,65,75)
   
-  exposure <- getReachAftereffects('exposure',part='initial')
-  classic  <- getReachAftereffects('classic',part='initial')
+  exposure_ini <- getReachAftereffects('exposure',part='initial')
   exposure_rem <- getReachAftereffects('exposure',part='remainder')
+  classic  <- getReachAftereffects('classic',part='all')
+  exposure <- getReachAftereffects('exposure',part='all')
   
-  exposureAVG <- aggregate(endpoint_angle ~ target, data=exposure, FUN=mean)
-  classicAVG <- aggregate(endpoint_angle ~ target, data=classic, FUN=mean)
+  exposureAVGini <- aggregate(endpoint_angle ~ target, data=exposure_ini, FUN=mean)
   exposureAVGrem <- aggregate(endpoint_angle ~ target, data=exposure_rem, FUN=mean)
+  classicAVG <- aggregate(endpoint_angle ~ target, data=classic, FUN=mean)
+  exposureAVG <- aggregate(endpoint_angle ~ target, data=exposure, FUN=mean)
   
-  exposureCI <- matrix(unlist(by(exposure$endpoint_angle, INDICES=c(exposure$target), FUN=t.interval)),nrow=2)
-  classicCI <- matrix(unlist(by(classic$endpoint_angle, INDICES=c(classic$target), FUN=t.interval)),nrow=2)
+  exposureCIini <- matrix(unlist(by(exposure_ini$endpoint_angle, INDICES=c(exposure_ini$target), FUN=t.interval)),nrow=2)
   exposureCIrem <- matrix(unlist(by(exposure_rem$endpoint_angle, INDICES=c(exposure_rem$target), FUN=t.interval)),nrow=2)
+  classicCI <- matrix(unlist(by(classic$endpoint_angle, INDICES=c(classic$target), FUN=t.interval)),nrow=2)
+  exposureCI <- matrix(unlist(by(exposure$endpoint_angle, INDICES=c(exposure$target), FUN=t.interval)),nrow=2)
   
   X <- c(points, rev(points))
-  expY <- c(exposureCI[1,],rev(exposureCI[2,]))
+  expYini <- c(exposureCIini[1,],rev(exposureCIini[2,]))
[…]

### 2018-01-22 shared.R

@@ -81,7 +81,7 @@ BS.interval = function(data, conf.level = .95, bootstraps = 1000) {
   
 }
 
-bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10,offset=4) {
+bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10,offset=4,CIs=c(.95)) {
   
   # parallel for-loop?
   installRequire.Packages(c('foreach','doParallel'))
@@ -104,7 +104,12 @@ bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10
   
   stopCluster(cl)
   
-  return(quantile(mus,probs=c(.025,.05,.10,.50,.90,.95,.975)))
+  probs <- c(.50)
+  for (CI in CIs) {
+    probs <- c(((1-CI)/2), probs, 1-((1-CI)/2))
+  }
+  # c(.025,.05,.10,.50,.90,.95,.975)
+  return(quantile(mus,probs=probs))
   
 }
 

### 2018-01-22 ManuscriptMaterials.Rmd
make parallelization optional, depending on installed packages

@@
 output:
+  html_notebook:
+    number_sections: yes
   pdf_document:
-    number_sections: yes
-  html_notebook:
     number_sections: yes
@@
 ```{r}
-cat('\nCLASSIC\n\n')
 getPeakConfidenceInterval('classic',part='all',CIs=c(.95))
-cat('\nEXPOSURE\n\n')
 getPeakConfidenceInterval('exposure',part='all',CIs=c(.95))
@@
 
-Although the two peaks are not different from each other, the 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training), whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training).
-So something might be going on -- but the evidence is far from concrete.
+Although the two peaks are not different from each other, the 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training) but not the 75 degree target, whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training) but not the 45 degree target.
+Something might be going on -- but the evidence is far from concrete.
 

### 2018-01-22 nocursor.R

@@ -123,6 +123,8 @@ plotReachAftereffects <- function() {
 
 getPeakConfidenceInterval <- function(group,validate=FALSE,part='initial',CIs=c(.95)) {
   
+  cat(sprintf('\n%s\n\n',toupper(group)))
+  
   RAE <- getReachAftereffects(group, part=part)
   
   if (validate) {

### 2018-01-22 shared.R

@@ -25,15 +25,25 @@ colorset[['extra1T']] <- '#c400c42f'
 installRequire.Packages <- function(packages) {
   
   installed.list <- rownames(installed.packages())
+  missingpackages <- c()
   
   for (pkg in packages) {
     
     if (!pkg %in% installed.list) {
-      install.packages(pkg,dep=TRUE)
+      # we can't force people to install packages, instead, we'll warn them
+      #install.packages(pkg,dep=TRUE)
+      missingpackages <- c(missingpackages, pkg)
     }
     
     require(pkg, character.only=TRUE)
     
+  }
+  
+  if (length(missingpackages) > 0) {
+    cat('\nWARNING, some pacakages are missing:\n')
+    cat(missingpackages)
+    cat('\ncar and nlme are required to reproduce the analyses\n')
+    cat('\nforeach and doParallel speed up a few functions\n')
   }
   
 }
@@ -84,25 +94,57 @@ BS.interval = function(data, conf.level = .95, bootstraps = 1000) {
 bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10,offset=4,CIs=c(.95)) {
   
   # parallel for-loop?
-  installRequire.Packages(c('foreach','doParallel'))
+  # installRequire.Packages(c('foreach','doParallel'))
   
-  cores=detectCores()
-  # very friendly for the rest of the system
-  usecores <- max(1, ceiling((cores[1] / 2) - 1))
-  cl <- makeCluster(usecores)
-  registerDoParallel(cl)
-  
-  x <- as.numeric(names(colMeans(data)))
-  
-  mus <- foreach (iteration=1:bootstraps, .combine=rbind, .export=c('getGaussianFit','GaussianErrors','parGaussian')) %dopar% {
-    
-    y <- colMeans(data[sample(row.names(data),size=nrow(data),replace=TRUE),])
-    
-    as.numeric(getGaussianFit(x,y,mu=47.5,sigma=30,scale=10,offset=4)$par['mu'])
-    
+  installed.list <- rownames(installed.packages())
+  packages <- c('foreach','doParallel')
+  allInstalled <- TRUE
+  for (pkg in packages) {
+    if (!pkg %in% installed.list) {
+      allInstalled <- FALSE
+    } else {
+      require(pkg, character.only=TRUE)
+    }
   }
   
-  stopCluster(cl)
[…]

### 2018-01-26 ManuscriptMaterials.Rmd
added multiple regression of RAE from active/passive localization

@@
 output:
+  pdf_document:
+    number_sections: yes
+    toc: true
+    toc_depth: 2
   html_notebook:
     number_sections: yes
-  pdf_document:
-    number_sections: yes
+    toc: true
+    toc_depth: 2
 ---
@@
 
-This is a collection of statistics and figures for the manuscript on the exposure consequences project.
+This document collects statistics and figures to be used in the manuscript on the exposure consequences project.
 
-Here are the planned figures and tables:
+## Figures
 
-**Figures**:
+These are the planned figures. Figure one will not be included here, and can in fact simply be re-used from the previous manuscript.
 
@@
 
-**Tables**:
+## Tables
+
+We expand the table used in the previous manuscript to include this paradigm as well.
 
@@
 
-All the scripts doing the statistics and figures are in separate R files, but here we will only see / do the ones that we think should go to the manuscript.
+The scripts doing the statistics and figures are in separate R files. They allow many options, but here we will only see / do the ones that we think should go to the manuscript.
 
@@
 
-```{r}
+```{r loadsources}
 source('shared.R') # functions used everywhere
@@
 Second is the localization: a) localization shifts after exposure b) the shifts are the same/different from those after classic c) the effects are different across the workspace d) this pattern is indistinguishable / different from that after classic.
-A third topic might be if localization shifts can predict no-cursor changes.
+A third topic might be if localization shifts can predict no-cursor changes, which we'll call the "correlation"  topic.
 Here we'll also follow this order:
@@
 
-```{r fig.width=7.5, fig.height=4}
+```{r fig2_reachaftereffect, fig.width=7.5, fig.height=4}
 plotReachAftereffects()
@@
 
-```{r include=TRUE}
+```{r RAE_exposure,include=TRUE}
 exposureNoCursorChange()
@@
 
-```{r}
+```{r RAE_exp_iterations}
 exposureAftereffectsPersistent()
@@
 
-```{r}
+```{r RAE_exp_cla}
 exposureClassicReachAftereffects()
@@
 
-```{r}
+```{r RAE_peaks}
 getPeakConfidenceInterval('classic',part='all',CIs=c(.95))
@@
 
-Let's go back to seeing if there is any effect from training type on reach aftereffects at all, and minimize the potential explanatory power of target by removing it from the model:
-
-```{r}
-exposureClassicReachAftereffects(noTarget=TRUE)
-```
-
-This doesn't seem to change one iota, so that whether or not we account for any effects of target direction, there is no overall effect of training with volitional movements (classic) or passive movements (exposure) on -- the amplitude of -- reach aftereffects.
-
 To sum up:
 
-1. Rotated exposure training evokes substantial shifts in no-cursor reach direction,
-2. these changes persist over time somewhat, and
+1. Rotated exposure training evokes substantial shifts in no-cursor reach direction across the workspace,
+2. these changes persist over time, and
 3. these changes are comparable to the ones we found after classic training.
@@
 
-```{r fig.width=7.5, fig.height=4}
+```{r fig3_localization, fig.width=7.5, fig.height=4}
 plotLocalization()
@@
 
-```{r}
-exposureLocalizationShift()
+```{r loc_exp}
+exposureLocalization()
 ```
@@
 
-```{r}
-exposureMovementType()
+```{r loc_exp_movtype}
+exposureLocalizationShift()
 ```
 
-It says there isn't, but let's make sure by not having hand angle in there as possibly stealing lots of variance.
-
-```{r}
-exposureMovementType(noHandAngle = TRUE)
-```
-
-OK, so that is pretty clear: there is no difference between passive and active localization after exposure training, even with double the number of trials as in classic delayed.
+There is no difference between passive and active localization after exposure training, even with double the number of trials as in classic delayed.
 
@@
 
-Now, let's see how this absence of a difference compares to that the (previously found difference) after classic training.
-Let's do this with an LME without all possible combinations of factors.
-We're primarily interested in the possibility that the effect of movement type is there after classic training, but not after exposure (that's our hypothesis).
-We're also somewhat interested in the effect of hand angle being different between classic and exposure training.
+Let's see what we get in a model with all data (localization differences between rotated and aligned though) from both groups:
 
-```{r}
-exposureClassicLocalization(model='restricted')
+```{r loc_ombnibus}
+groupLocalization(model='full')
 ```
 
-This works: the interaction between group and movement type indicates that the effect of movement is different for the exposure and for the classic groups.
-Since it wasn't there at all for the exposure group, this means that there should be an effect of movement type in the classic group, which is an easily testable hypothesis.
-We also see an interaction between group and hand angle, which would mean that the shape of the generalization curve is different for the two groups.
-This seems to also be visible in the data, but would be worth probing a little further.
+In the exposure group there is no effect of movement type, but we see one here.
+This could be because the effect of movement type is much larger in the classic group.
+This would predict an interaction between group and movement type, but this is a trend at best (p=.056).
+Hence we tested for an effect of movement type in the classic group:
 
-However, we did this with a restricted model that only contained the terms of interest. Let's see if it is still there with a full model:
-
-```{r}
-exposureClassicLocalization(model='full')
+```{r loc_cla}
+classicLocalizationShift(remove15=TRUE)
+classicLocalizationShift(remove15=FALSE)
 ```
 
-Now it is only a trend, so let's follow this up with some deeper probing.
-In the restricted model there was also an interaction between group and hand angle which has all but vanished in the full model.
+So there is a fairly strong effect of movement type in the classic group.
+Together with the absence of an effect of movement type in the exposure group and a trend of an interaction between group and movement type, this might indicate a sub-threshold effect of movement type in the exposure group in the same direction as the effect of movement type in the classic group.
 
-## Effect of movement type after classic training
+Also note the absence of an effect of hand angle in the classic group here -- potentially because we excluded the 15 degree target.
 
-We already saw that in the exposure group there was an effect of hand angle, but not of movement type. How is this in the classic group?
+To sum up our findings on localization:
 
-```{r}
-classicMovementType()
-classicMovementType(factors='movementtype')
-```
+1. Rotated exposure training leads to changes in localization responses.
 
-Alright, so there is an effect of movement type here as well as of hand angle.
-Since the exposure group saw no effect of movement type in either analysis, whereas there is an effect of movement type in the classic group in both analyses, we can be fairly confident that if there is any effect of movement type after exposure training it will be smaller than after classic training.
-This is what we would expect from exposure training.
+2. In the exposure group, there is no effect of movement type: localization responses with and without efference copies are indistinguishable.
+3. In the classic group, there is an effect of movement type: localization responses informed by efference copies are shifted more than those with only proprioception (as found before).
+4.
+This predicts an interaction between movement type and group, but that is only a trend, suggesting that either the difference in the classic group is not so large, or that there is a small, sub-threshold difference in the exposure group.
 
-## Hand angle in classic training
-
-We're still not too sure about the effect of hand angle though.
-In the figures the localization shift for the classic groups _appears_ more homogeneous across the workspace as compared to the exposure group, but is that a real effect?
-
-I'm not sure how to focus an analysis on that question, better than already done, except by using a model with only hand angle:
-
-```{r}
-classicMovementType(factors='handangle')
-```
-
-So yes, there is some generalization effect after classic training. We still want to know if it is different from the effect of hand angle in exposure.
-
-```{r}
-exposureClassicLocalization(model='handangle')
-```
-
-According to all these analyses, the generalization is not appreciably different between classic and exposure training.
+5. The shift in localization is different across the workspace in the exposure group. I.e., there is a generalization _curve_.
+6.
+This effect is not present in the classic group, so that generalization can be said to be flat; there is no _curve_.
+**NOTE:** When we do include the 15 degree target, there is an effect of hand angle though.
 
@@
 
-It looks like for exposure, active localization is the better predictor of reach aftereffects, whereas in classic it is passive localization. But are any of them real?
+It looks like for exposure, active localization is the better predictor of reach aftereffects, whereas in classic it is passive localization.
+But are any of them real?
+The function also returns the correlation tests and linear regressions for all four datasets.
 
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
+# this code won't run or show up in output, just included here if you want to play with it:
 for (name in names(correlations)) {
@@
 
-They get the same p-value, so that seems fine.
+[code/results hidden] They get the same p-value, so that seems fine.
+
+## Multiple regression models
+
+We've decided to use all the no-cursor blocks for these regression and correlations, so perhaps we shouldn't include the one below where I only use the initial block.
+The idea behind using only the initial block is that that initial block matches the time the localization is done -- relative to the last training.
+
+Here I do step-wise multiple regression, allowing active and passive localization shift as predictor of reach aftereffects.
+If I do this on all the no-cursor block data, the models converge on using the active localization for the exposure group, and passive localization for the classic group.
+
+If we look at only the initial no-cursor block, the story changes.
+It uses both active and passive to predict reach aftereffects for the exposure group, but converges on using only an intercept term for the classic group.
+This might mean that in te classic group there is no relationship between reach aftereffects and localization, or this might mean that in the exposure group we have sufficient amount of measurements to detect the relationship, as we have twice the number of localization trials.
+One way to rule this out is to use only the first half of the localization data from the exposure group.
+
+Here is what the step-wise multiple regressions look like with all available data:
+
+```{r}
+multipleRegressionLocalization(group='exposure',NCpart='initial')
+```
+
+No terms are dropped an the adjusted R squared of the full model is low, but reasonable. _Note that the coefficient for the passive localization is positive._
+
+For classic training, it looks like this:
+
+```{r}
+multipleRegressionLocalization(group='classic',NCpart='initial')
+```
+
+The step-wise multiple regression algorithm thinks an intercept-only model is best.
+Now is this because localization after classic training has little to no information on reach aftereffects, or is this because we have double the data in the exposure group?
+We can test this partially by throwing away the second half of the data for the exposure group:
+
+```{r}
+multipleRegressionLocalization(group='exposure',NCpart='initial', LRpart='first')
+```
+
+Now the passive localization shifts are thrown out by the algorithm and the active localization shifts don't contribute much either.
+This means that perhaps previous papers didn't find any relationships between recalibrated proprioception and reach aftereffects because not enough data was collected to overcome the noisiness of the measurements?
+
+It also makes me wonder if reach aftereffects in classic can be better predicted by the online localization responses:
+
+```{r}
+multipleRegressionLocalization(group='online',NCpart='initial')
+```
+
+So this doesn't work either.
+Does this mean that in previous experiments we just didn't have sufficiently good data to find any link between classic reach aftereffects and (2-AFC) proprioceptive recalibration?
+
+Do we still want to see online localization matches proprioception more, as claimed in Izawa et al., 2012?
+This would imply that it matches the exposure localization and/or reach aftereffects better...

### 2018-01-26 localization.R

@@ -163,8 +163,10 @@ plotLocalization <- function() {
   polygon(X,exp.act.Y,border=NA,col=colorset[['expActT']])
   polygon(X,exp.pas.Y,border=NA,col=colorset[['expPasT']])
   
-  lines(points,exp.avg.act$taperror_deg,col=colorset[['expActS']],lty=1,lwd=2)
-  lines(points,exp.avg.pas$taperror_deg,col=colorset[['expPasS']],lty=1,lwd=2)
+  lines(points[1:2],exp.avg.act$taperror_deg[1:2],col=colorset[['expActS']],lty=2,lwd=2)
+  lines(points[2:7],exp.avg.act$taperror_deg[2:7],col=colorset[['expActS']],lty=1,lwd=2)
+  lines(points[1:2],exp.avg.pas$taperror_deg[1:2],col=colorset[['expPasS']],lty=2,lwd=2)
+  lines(points[2:7],exp.avg.pas$taperror_deg[2:7],col=colorset[['expPasS']],lty=1,lwd=2)
   
   axis(1,at=points)
   axis(2,at=c(0,-5,-10,-15))
@@ -182,8 +184,10 @@ plotLocalization <- function() {
   polygon(X,cla.act.Y,border=NA,col=colorset[['claActT']])
   polygon(X,cla.pas.Y,border=NA,col=colorset[['claPasT']])
   
-  lines(points,cla.avg.act$taperror_deg,col=colorset[['claActS']],lty=1,lwd=2)
-  lines(points,cla.avg.pas$taperror_deg,col=colorset[['claPasS']],lty=1,lwd=2)
+  lines(points[1:2],cla.avg.act$taperror_deg[1:2],col=colorset[['claActS']],lty=2,lwd=2)
+  lines(points[2:7],cla.avg.act$taperror_deg[2:7],col=colorset[['claActS']],lty=1,lwd=2)
+  lines(points[1:2],cla.avg.pas$taperror_deg[1:2],col=colorset[['claPasS']],lty=2,lwd=2)
+  lines(points[2:7],cla.avg.pas$taperror_deg[2:7],col=colorset[['claPasS']],lty=1,lwd=2)
   
   axis(1,at=points)
   axis(2,at=c(0,-5,-10,-15))
@@ -193,12 +197,16 @@ plotLocalization <- function() {
   
 }
 
-exposureLocalizationShift <- function() {
+exposureLocalization <- function(remove15=TRUE) {
   
   default.contrasts <- options('contrasts')
[…]

### 2018-01-26 nocursor.R

@@ -14,57 +14,57 @@ installRequire.Packages(required.packages)
 # groupURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'='https://osf.io/8hm7f/?action=download')
 
 
-# plot both groups reach aftereffects in one figure
-OLDplotReachAftereffects <- function(validate=FALSE) {
-  
-  points <- c(15,25,35,45,55,65,75)
-  
-  exposure <- getReachAftereffects('exposure',part='initial')
-  classic  <- getReachAftereffects('classic',part='initial')
-  exposure_rem <- getReachAftereffects('exposure',part='remainder')
-  classic_rem  <- getReachAftereffects('classic',part='remainder')
-  
-  if (validate) {
-    
-    validExp <- validateReachAftereffects('exposure')
-    validExpParticipants <- validExp$participant[which(validExp$RAE == 1)]
-    exposure <- exposure[which(exposure$participant %in% validExpParticipants),]
-
-    validCla <- validateReachAftereffects('classic')
-    validClaParticipants <- validCla$participant[which(validCla$RAE == 1)]
-    classic <- classic[which(classic$participant %in% validClaParticipants),]
-    
-  }
-  
-  exposureAVG <- aggregate(endpoint_angle ~ target, data=exposure, FUN=mean)
-  classicAVG <- aggregate(endpoint_angle ~ target, data=classic, FUN=mean)
-  
-  exposureAVGrem <- aggregate(endpoint_angle ~ target, data=exposure_rem, FUN=mean)
-  classicAVGrem <- aggregate(endpoint_angle ~ target, data=classic_rem, FUN=mean)
-  
-  exposureCI <- matrix(unlist(by(exposure$endpoint_angle, INDICES=c(exposure$target), FUN=t.interval)),nrow=2)
-  classicCI <- matrix(unlist(by(classic$endpoint_angle, INDICES=c(classic$target), FUN=t.interval)),nrow=2)
-  
-  X <- c(points, rev(points))
-  expY <- c(exposureCI[1,],rev(exposureCI[2,]))
-  claY <- c(classicCI[1,],rev(classicCI[2,]))
-  
[…]

### 2018-01-26 relateLocalizationNoCursors.R

@@ -134,6 +134,51 @@ correlateNoCursorsLocalization <- function(NCpart='all') {
   
 }
 
+
+multipleRegressionLocalization <- function(group,NCpart='all',expart='all',LRpart='all') {
+  
+  # first we do all responses, plan to split it by target later
+  # get all data, for classic and exposure, we need localization and reach aftereffects
+  
+  # LOCALIZATION SHIFT
+  local <- getPointLocalization(group=group, difference=FALSE, verbose=FALSE, LRpart=LRpart)
+  # active
+  loc.act.ro <- aggregate(taperror_deg ~ participant + handangle_deg, data=local[local$rotated_b == 1 & local$passive_b == 0,], FUN=mean, drop=FALSE)
+  loc.act.al <- aggregate(taperror_deg ~ participant + handangle_deg, data=local[local$rotated_b == 0 & local$passive_b == 0,], FUN=mean, drop=FALSE)
+  local.act <- loc.act.ro
+  local.act$taperror_deg <- local.act$taperror_deg - loc.act.al$taperror_deg
+  # passive
+  loc.pas.ro <- aggregate(taperror_deg ~ participant + handangle_deg, data=local[local$rotated_b == 1 & local$passive_b == 1,], FUN=mean, drop=FALSE)
+  loc.pas.al <- aggregate(taperror_deg ~ participant + handangle_deg, data=local[local$rotated_b == 0 & local$passive_b == 1,], FUN=mean, drop=FALSE)
+  local.pas <- loc.pas.ro
+  local.pas$taperror_deg <- local.pas$taperror_deg - loc.pas.al$taperror_deg
+  
+  # REACH AFTEREFFECTS
+  nocur <- getReachAftereffects(group=group, part=NCpart, clean=TRUE) 
+  names(nocur)[names(nocur) == 'endpoint_angle'] <- 'RAE'
+  # combine data frames
+  df <- nocur[which(nocur$participant %in% local.act$participant),]
+  df$active <- local.act$taperror_deg
+  df$passive <- local.pas$taperror_deg
+  
+
+  # the correlations are polluted by the generalization curve... let's remove that!
+  
+  # and for exposure too
+  df.tmp <- df[which(df$target %in% c(35,45,55,65)),]
+  df <- aggregate(RAE ~ participant, data=df.tmp, FUN=mean, na.rm=TRUE, drop=FALSE)
[…]

### 2018-01-26 shared.R

@@ -4,7 +4,7 @@
 nocursorURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'='https://osf.io/8hm7f/?action=download')
 
 # no-cursor data can be downloaded from OSF:
-localizationURLs <- c('exposure'='https://osf.io/pqsdc/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
+localizationURLs <- c('exposure'='https://osf.io/yqkex/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
 
 # we'll control the color in figures centrally from here:
 colorset <- list()
@@ -205,53 +205,26 @@ parGaussian <- function(par,x) {
 
 # handling localization data -----
 
-# getANOVAlocalization <- function(group) {
-#   
-#   df <- load.DownloadDataframe(url=localizationURLs[group],filename=sprintf('localization_%s.csv',group))
-#   
-#   df <- aspligned(df)
-#   
-#   groupdf <- NA
-#   
-#   participants <- unique(df$participant)
-#   
-#   for (pp.no in c(1:length(participants))) {
-#     
-#     pp.id <- participants[pp.no]
-#     
-#     for (rotated in c(0,1)) {
-#       
-#       for (passive in c(0,1)) {
-#         
-#         subdf <- df[which(df$participant == pp.id & df$rotated_b == rotated & df$passive_b == passive),]
-#         
-#         locdf <- getLocalizationPoints(subdf, points=c(15,25,35,45,55,65,75), removeOutliers=TRUE)
-#         
-#         if (any(is.na(locdf$taperror_deg))) {
-#           cat(sprintf('WARNING: NAs in %s, pp:%s, %s/%s\n',group,pp.id,c('aligned','rotated')[rotated+1],c('active','passive')[passive+1]))
-#         }
-#         
-#         if (is.data.frame(groupdf)) {
-#           groupdf <- rbind(groupdf, locdf)
-#         } else {
-#           groupdf <- locdf
-#         }
-#         
-#       }
-#       
-#     }
-#     
-#   }
-#   
-#   return(groupdf)
-#   
-# }
-
-getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55,65,75), movementtype='both', verbose=TRUE) {
[…]

### 2018-01-29 shared.R
made comment

@@ -118,6 +118,7 @@ bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10
     # somewhat friendly for the rest of the system
     usecores <- max(1, (cores[1] - 1))
     
+    # unwarreted comment
     
     cl <- makeCluster(usecores)
     registerDoParallel(cl)

### 2018-02-02 ManuscriptMaterials.Rmd
put some participants back in, as I could not find why they were removed - one had a few missing files, so this resulted in no data at all when running some functions on partial datasets, this has been fixed

@@
 The absence of an interaction between training and target could be interpreted to mean that the generalization of reach aftereffects is the same after both kinds of training.
-Here are the confidence intervals for both:
+Here are the confidence intervals for the peaks of both:
 
@@
 ```{r loc_exp}
-exposureLocalization()
+exposureLocalization(remove15=TRUE)
 ```
@@
 ```{r loc_exp_movtype}
-exposureLocalizationShift()
+exposureLocalizationShift(remove15=TRUE)
 ```
@@
 ```{r loc_ombnibus}
-groupLocalization(model='full')
+groupLocalization(model='full',remove15=TRUE)
 ```
@@
 This could be because the effect of movement type is much larger in the classic group.
-This would predict an interaction between group and movement type, but this is a trend at best (p=.056).
+This would predict an interaction between group and movement type, but this is not even a trend (p=.133).
 Hence we tested for an effect of movement type in the classic group:
@@
 classicLocalizationShift(remove15=TRUE)
-classicLocalizationShift(remove15=FALSE)
 ```
@@
 ```{r}
+# THIS FAILS NOW... maybe because some participants don't have data in the first half?
 multipleRegressionLocalization(group='exposure',NCpart='initial', LRpart='first')
@@
 So this doesn't work either.
-Does this mean that in previous experiments we just didn't have sufficiently good data to find any link between classic reach aftereffects and (2-AFC) proprioceptive recalibration?
-
-Do we still want to see online localization matches proprioception more, as claimed in Izawa et al., 2012?
-This would imply that it matches the exposure localization and/or reach aftereffects better...

### 2018-02-02 localization.R

@@ -336,3 +336,47 @@ classicLocalizationShift <- function(factors='both',remove15=TRUE) {
   options('contrasts' <- default.contrasts)
   
 }
+
+boxPlotLocalization <- function() {
+  
+  # par(mfrow=c(4,1),mar=c(2,3,1,1))
+  
+  for (group in c('exposure','classic','online')) {
+    
+    df <- getPointLocalization(group, difference=FALSE, verbose=FALSE)
+    
+    # print(str(df))
+    
+    for (rotated in c(0,1)) {
+      
+      for (passive in c(0,1)) {
+        
+        # subdf <- df[which(df$rotated_b == rotated & df$passive_b == passive),]
+        
+        # boxplot(taperror_deg ~ handangle_deg, data=df, axes=F, bty='n')
+        
+        for (target in c(15,25,35,45,55,65,75)) {
+          
+          subdf <- df[which(df$rotated_b == rotated & df$passive_b == passive & df$handangle_deg == target),]
+          
+          ppno <- subdf$participant
+          taperror <- subdf$taperror_deg
+          
+          idx <- which(abs(taperror - mean(taperror)) > (3 * sd(taperror)))
+          
+          if (length(idx) > 0) {
+            
+            cat(sprintf('%s %s %s %d deg\n', group, c('aligned','rotated')[rotated+1], c('active','passive')[passive+1], target))
+            print(ppno[idx])
+            
+          }
+          
+        }
+        
+      }
+      
+    }
+    
+  }
+  
+}

### 2018-02-02 shared.R

@@ -4,7 +4,7 @@
 nocursorURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'='https://osf.io/8hm7f/?action=download')
 
 # no-cursor data can be downloaded from OSF:
-localizationURLs <- c('exposure'='https://osf.io/yqkex/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
+localizationURLs <- c('exposure'='https://osf.io/9f6gu/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
 
 # we'll control the color in figures centrally from here:
 colorset <- list()
@@ -212,7 +212,6 @@ getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55
   
   df <- aspligned(df)
   
-  
   # we can select half the localization data... for the exposure group only
   if (group == 'exposure') {
     # print(str(groupdf))
@@ -224,12 +223,44 @@ getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55
     }
     # print(dim(groupdf))
   } else {
-    cat('\nWARNING: Partial localization can not be returned for classic/online data.\n\n')
+    if (LRpart != 'both') {
+      cat('\nWARNING: Partial localization can not be returned for classic/online data.\n\n')
+      # the full dataset is returned anyway?
+    }
   }
   
   groupdf <- NA
   
+  # only get the movement types we need:
+  movementnumbers <- c(0,1)
+  if (movementtype == 'active') {
+    movementnumbers <- c(0)
+  }
+  if (movementtype == 'passive') {
+    movementnumbers <- c(1)
+  }
+  
   participants <- unique(df$participant)
+  
+  # remove participants with no data
+  removepps <- c()
+  for (pp.no in c(1:length(participants))) {
+    pp.id <- participants[pp.no]
+    for (rotated in c(0,1)) {
+      for (passive in movementnumbers) {
+        subrows <- which(df$participant == pp.id & df$rotated_b == rotated & df$passive_b == passive)
+        # could set this to e.g.: '< 10' to remove participants with very little data?
+        if (length(subrows) == 0) {
+          removepps <- c(removepps, pp.id)
+        }
+      }
+    }
+  }
+  if (length(removepps) > 0) {
[…]

### 2018-02-02 ManuscriptMaterials.Rmd
expanded conclusions, added one LME for classic localization that focusses on movement type (ignores hand angle)

@@
 This would predict an interaction between group and movement type, but this is not even a trend (p=.133).
-Hence we tested for an effect of movement type in the classic group:
+
+What if we don't care about hand angle:
+
+```{r loc_groups_movementtype}
+groupLocalization(model='movementtype',remove15=TRUE)
+```
+
+Now there is a trend towards significance for the interaction between training group and movement type. That is closer to what we would expect.
+
+Let's see if the classic group shows an effect of movement type (with the current preprocessing procedures):
 
@@
 
-```{r fig.width=7.5, fig.height=4}
+```{r scattergrams, fig.width=7.5, fig.height=4}
 correlations <- correlateNoCursorsLocalization()
@@
 
-So this doesn't work either.
+But no.
+
+# Conclusions
+
+What can we conclude? Here is an initial list of what we might conclude.
+
+1.
+Exposure training leads to robust reach aftereffects that are similar to those after classic training, although they seem to have a more pronouced generalization curve.
+This means that evoking propceptive recalibration with visual-proprioceptive discrepancies leads to motor changes.
+In other words: proprioception informs movement plans.
+
+2.
+Exposure training leads to shifts in localization, that are not very different between active and passive localization.
+Since both active and passive movements access recalibrated proprioception, while active localization also accesses updated predictions, this indicates that hand localization is makes very good use of proprioception.
+
+    1.
+Additionally, since predictions should not be updated in exposure training, but active localization might be evoking a slightly larger shift, this indicates a non-Bayesian process of combining predictions and perceptions in this case.
+
+    2.
+After classic training (2016 paper) there was an effect of movement type, which indicates that with a training regime that allows updates of predicted sensory consequences, these updates do indeed also affect hand localization.
+
+    3.
+Within the workspace however, there is no interaction between movement type and training group.
+Perhaps there is a small update of predicted proprioceptive consequences, that we can't parcel out.
+A more computational approach would be needed to see if this can explain the pattern of results here.
+
+3.
+Recalibrated proprioception can reliably predict reach aftereffects, but only after **exposure** training and for _active localization_.
+This means that recalibrated proprioception indeed leads to motor changes.
+
+    1.
+There is also a trend towards significance for _passive localization_ to predict reach aftereffects after **classic training**.
+Perhaps active localization is more similar to no-cursor reaches after exposure training, as both of these can only use recalibrated proprioception as well as updated predictions of proprioceptive consequences.
+This does not explain the trend towards significance for predicting classic reach aftereffects with passive localization.
+However, a step-wise multiple regression favours an intercept model for classic reach aftereffects, and a model that mostly relies on active localization for exposure reach aftereffects.
+So perhaps not too much should be read into the trend towards significance in the correlation.
+On the other hand, the exposure group had twice as much data, so that the estimated recalibration would be more reliable within each participant.
+Perhaps with higher quality estimates of proprioceptive recalibration those would be able to predict part of the reach aftereffects even after classic training.

### 2018-02-07 ManuscriptMaterials.Rmd
adressed some of Ahmed's comments

@@
     1.
-Additionally, since predictions should not be updated in exposure training, but active localization might be evoking a slightly larger shift, this indicates a non-Bayesian process of combining predictions and perceptions in this case.
+Additionally, since predictions should not be updated in exposure training, active localization could have resulted in a _lower_ shift in localization if hand location were estimated by combining recalibrated proprioception and updated prediction in a maximum likelihood way.
+If anything however, active localization is slightly larger (non-significant) than passive localization.
+This indicates a non-Bayesian process of combining predictions and perceptions -- in this case.
 
@@
     3.
-Within the workspace however, there is no interaction between movement type and training group.
+Within the workspace however, there is no interaction between movement type and training group (but there is a trend if hand angle is used as a random effect, in all analyses the 15 degree data is omitted as Denise said).
 Perhaps there is a small update of predicted proprioceptive consequences, that we can't parcel out.
@@
 3.
-Recalibrated proprioception can reliably predict reach aftereffects, but only after **exposure** training and for _active localization_.
-This means that recalibrated proprioception indeed leads to motor changes.
+Recalibrated proprioception can somewhat predict reach aftereffects, but only after **exposure** training and for _active localization_.
+This indicates that in some circumstances, recalibrated proprioception might drive motor changes.
+However, it is possible that the active component of the localization is crucial here.
+Perhaps exposure training leads to some updates in predicted sensory consequences after all -- be they visual or proprioceptive.
+These would also factor into the reach aftereffects.
 
@@
 There is also a trend towards significance for _passive localization_ to predict reach aftereffects after **classic training**.
-Perhaps active localization is more similar to no-cursor reaches after exposure training, as both of these can only use recalibrated proprioception as well as updated predictions of proprioceptive consequences.
+Perhaps active localization is more similar to no-cursor reaches after exposure training, as they both use recalibrated proprioception as well as updated predictions of sensory consequences -- which could mostly be proprioceptive consequences after exposure training.
 This does not explain the trend towards significance for predicting classic reach aftereffects with passive localization.
-However, a step-wise multiple regression favours an intercept model for classic reach aftereffects, and a model that mostly relies on active localization for exposure reach aftereffects.
-So perhaps not too much should be read into the trend towards significance in the correlation.
+A step-wise multiple regression favours an intercept model for classic reach aftereffects, hence none of the localization data was deemed informative.
+The same algorithm arrived at a model using active localization to predict exposure reach aftereffects.
+The intercept model for classic reach aftereffects could signify that not too much should be read into the trend towards significance in the correlation.
 On the other hand, the exposure group had twice as much data, so that the estimated recalibration would be more reliable within each participant.

### 2018-02-07 ManuscriptMaterials.Rmd
added peak estimation for localization data

@@
 
+## Localization peaks
+
+Decided to add analyses to get the confidence intervals for the peaks of the localization generalization curves:
+
+```{r localization_peaks}
+getPeakLocConfidenceInterval(group='classic',CIs=c(.95), movementtype='both', LRpart='all')
+getPeakLocConfidenceInterval(group='exposure',CIs=c(.95), movementtype='both', LRpart='all')
+```
+
+These results are strange: the median peak for the exposure localization is around 42 degrees.
+If we look at the plot, it is much higher though, somewhere between 55 and 65 probably.
+This doesn't change much if I only look at the passive or at the active localization data.
+The peak for classic makes sense though.
+
 # Correlation
@@
     3.
-Within the workspace however, there is no interaction between movement type and training group (but there is a trend if hand angle is used as a random effect, in all analyses the 15 degree data is omitted as Denise said).
+Within the workspace however, there is no interaction between movement type and training group (but there is a 'trend', p=.09, if hand angle is used as a random effect, in all analyses the 15 degree data is omitted as Denise said).
 Perhaps there is a small update of predicted proprioceptive consequences, that we can't parcel out.

### 2018-02-07 localization.R

@@ -380,3 +380,15 @@ boxPlotLocalization <- function() {
   }
   
 }
+
+getPeakLocConfidenceInterval <- function(group,part='initial',CIs=c(.95), movementtype='both', LRpart='all') {
+  
+  cat(sprintf('\n%s\n\n',toupper(group)))
+  
+  loc <- getPointLocalization(group, difference=TRUE, points=c(15,25,35,45,55,65,75), movementtype='both', LRpart='all', verbose=FALSE)
+  
+  loc2 <- xtabs(taperror_deg ~ participant + handangle_deg,loc)
+  
+  bootstrapGaussianPeak(data=loc2,bootstraps=1000,mu=47.5,sigma=30,scale=10,offset=4,CIs=CIs)
+  
+}

### 2018-02-07 ManuscriptMaterials.Rmd
fixed peaks for localization

@@
 ```{r localization_peaks}
-getPeakLocConfidenceInterval(group='classic',CIs=c(.95), movementtype='both', LRpart='all')
-getPeakLocConfidenceInterval(group='exposure',CIs=c(.95), movementtype='both', LRpart='all')
+getPeakLocConfidenceInterval(group='classic',CIs=c(.95), movementtype='active', LRpart='all')
+getPeakLocConfidenceInterval(group='exposure',CIs=c(.95), movementtype='active', LRpart='all')
 ```
 
-These results are strange: the median peak for the exposure localization is around 42 degrees.
-If we look at the plot, it is much higher though, somewhere between 55 and 65 probably.
-This doesn't change much if I only look at the passive or at the active localization data.
-The peak for classic makes sense though.
+These results mimic those of the reach aftereffects: the median values are different, but they each fall in the other confidence interval.
+If we want to look at this better, we need to get better estimates of these curves in every participants, by having a wider range of points and more data per point, but probably also larger rotations.
 
@@
 
-```{r}
+```{r pearson_rhos}
 for (name in names(correlations)) {
@@
 
-```{r eval=FALSE, include=FALSE}
+```{r linreg, eval=FALSE, include=FALSE}
 # this code won't run or show up in output, just included here if you want to play with it:
@@
 
-```{r}
+```{r step_exposure}
 multipleRegressionLocalization(group='exposure',NCpart='initial')
@@
 
-```{r}
+```{r step_classic}
 multipleRegressionLocalization(group='classic',NCpart='initial')
@@
 
-```{r}
-# THIS FAILS NOW... maybe because some participants don't have data in the first half?
+```{r step_exposure_firsthalf}
 multipleRegressionLocalization(group='exposure',NCpart='initial', LRpart='first')
@@
 
-```{r}
+```{r step_online}
 multipleRegressionLocalization(group='online',NCpart='initial')

### 2018-02-07 localization.R

@@ -385,9 +385,9 @@ getPeakLocConfidenceInterval <- function(group,part='initial',CIs=c(.95), moveme
   
   cat(sprintf('\n%s\n\n',toupper(group)))
   
-  loc <- getPointLocalization(group, difference=TRUE, points=c(15,25,35,45,55,65,75), movementtype='both', LRpart='all', verbose=FALSE)
+  loc <- getPointLocalization(group, difference=TRUE, points=c(15,25,35,45,55,65,75), movementtype=movementtype, LRpart=LRpart, verbose=FALSE)
   
-  loc2 <- xtabs(taperror_deg ~ participant + handangle_deg,loc)
+  loc2 <- -1 * xtabs(taperror_deg ~ participant + handangle_deg,loc)
   
   bootstrapGaussianPeak(data=loc2,bootstraps=1000,mu=47.5,sigma=30,scale=10,offset=4,CIs=CIs)
   

### 2018-02-07 shared.R

@@ -115,10 +115,8 @@ bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10
     
     # very friendly for the rest of the system
     usecores <- max(1, ceiling((cores[1] / 2) - 1))
-    # somewhat friendly for the rest of the system
+    # somewhat unfriendly for the rest of the system
     usecores <- max(1, (cores[1] - 1))
-    
-    # unwarreted comment
     
     cl <- makeCluster(usecores)
     registerDoParallel(cl)

### 2018-02-25 ManuscriptMaterials.Rmd
Added ideas on getting function peaks.

@@
 
+## Cumulative Function
+
+Potentially, the wide confidence intervals can be explained by two things: 1) large variation across participants, and 2) incomplete sampling over the function: large swaths of the normal function (not distribution!) are sometimes not sampled at all.
+Here are some ideas for solutions.
+
+The first problem -- large differences between participants -- could be dealt with by estimating the function with each participant.
+This way only conditions within groups can be compared, but that still leaves comparisons between active and passive localization, although comparing reach aftereffects between groups would still not be possible.
+We can do a Fisher exact test to see if the proportion people with larger shifts in active than in passive localization is higher than 50% in each group.
+Prediction: should not be different from chance in exposure training, but should be different from chance in classic training.
+
+The second problem -- incomplete sampling of the curve -- could be dealt with by fitting a cumulative function instead of a normal function.
+We could assume the training-induced change is 0 at -90 and at 180 degrees, then at every actually estimated point, we take the sum of everything before it, and then scale so that values are between 0 and 1 (divide by the value at 180).
+(We could also put the extremes at not quite 0, so that at -90 we're at -1 degrees changes and at 180, we add 1 degree to the last measured point / bin, this should leave some wiggle room for a better fit.) At that point a cumulative function can be fit, and we take the 50% point as the centre of the normal function.
+I'm not totally sure if this works, but it's worth a try.
+
 # Correlation

### 2018-05-07 ManuscriptMaterials.Rmd
added online localization to figure

@@
 output:
+  html_document:
+    df_print: paged
+    toc: yes
+    toc_depth: '2'
   pdf_document:
     number_sections: yes
-    toc: true
-    toc_depth: 2
-  html_notebook:
-    number_sections: yes
-    toc: true
+    toc: yes
     toc_depth: 2
@@
 
-```{r fig3_localization, fig.width=7.5, fig.height=4}
-plotLocalization()
+```{r fig3_localization, fig.width=7.5, fig.height=2.5}
+plotLocalization(classicOnline=TRUE)
 ```

### 2018-05-07 localization.R

@@ -124,7 +124,7 @@ localizationLMEsSatterthwaite <- function() {
 
 # PLOTS / FIGURES ------
 
-plotLocalization <- function() {
+plotLocalization <- function(classicOnline=FALSE) {
   
   # get the data to plot:
   exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE)
@@ -148,8 +148,24 @@ plotLocalization <- function() {
   cla.pas <- cla[which(cla$passive_b == 1),]
   cla.CI.pas <- matrix(unlist(by(cla.pas$taperror_deg, INDICES=c(cla.pas$handangle_deg), FUN=t.interval)),nrow=2)
 
+  if (classicOnline) {
+    onl <- getPointLocalization('online', difference=TRUE, verbose=FALSE)
+    
+    onl.avg <- aggregate(taperror_deg ~ passive_b + handangle_deg, data=onl, FUN=mean)
+    onl.avg.act <- onl.avg[which(onl.avg$passive_b == 0),]
+    onl.avg.pas <- onl.avg[which(onl.avg$passive_b == 1),]
+    
+    onl.act <- onl[which(onl$passive_b == 0),]
+    onl.CI.act <- matrix(unlist(by(onl.act$taperror_deg, INDICES=c(onl.act$handangle_deg), FUN=t.interval)),nrow=2)
+    onl.pas <- onl[which(onl$passive_b == 1),]
+    onl.CI.pas <- matrix(unlist(by(onl.pas$taperror_deg, INDICES=c(onl.pas$handangle_deg), FUN=t.interval)),nrow=2)
+  }
   
-  par(mfrow=c(1,2))
+  if (classicOnline) {
+    par(mfrow=c(1,3))
+  } else {
+    par(mfrow=c(1,2))
+  }
   points <- c(15,25,35,45,55,65,75)
   
   # panel A: exposure localization (active vs. passive)
@@ -194,6 +210,28 @@ plotLocalization <- function() {
   
   legend(10,-15,c('passive','active'),col=c(colorset[['claPasS']],colorset[['claActS']]),lty=c(1,1),lwd=c(2,2),bty='n')
   
+  if (classicOnline) {
+    
[…]

### 2018-05-07 nocursor.R

@@ -233,7 +233,7 @@ plotReachAftereffectDistributions <- function() {
   
 }
 
-exposureNoCursorChange <- function() {
+exposureNoCursorChange <- function(printChange=FALSE) {
   
   default.contrasts <- options('contrasts')
   options(contrasts=c('contr.sum','contr.poly'))
@@ -252,6 +252,26 @@ exposureNoCursorChange <- function() {
   detach(nc.exp)
 
   options('contrasts' <- default.contrasts)
+  
+  if (printChange) {
+    
+    nc.exp <- getReachAftereffects('exposure', part='initial', difference=TRUE)
+    
+    nc.exp <- aggregate(endpoint_angle ~ target, data=nc.exp, FUN=mean)
+    
+    cat(sprintf('\nEXPOSURE\nmax reach aftereffects: %0.1f (at: %d deg)\n',max(nc.exp$endpoint_angle), nc.exp$target[which(nc.exp$endpoint_angle == max(nc.exp$endpoint_angle))]))
+    cat(sprintf('avg reach aftereffects: %0.1f (all targets)\n',mean(nc.exp$endpoint_angle, na.rm=TRUE)))
+    cat(sprintf('avg reach aftereffects: %0.1f (targets above 15 degrees)\n',mean(nc.exp$endpoint_angle[which(nc.exp$target > 15)], na.rm=TRUE)))
+    
+    nc.cla <- getReachAftereffects('classic', part='initial', difference=TRUE)
+    
+    nc.cla <- aggregate(endpoint_angle ~ target, data=nc.cla, FUN=mean)
+    
+    cat(sprintf('\nCLASSIC\nmax reach aftereffects: %0.1f (at: %d deg)\n',max(nc.cla$endpoint_angle), nc.cla$target[which(nc.cla$endpoint_angle == max(nc.cla$endpoint_angle))]))
+    cat(sprintf('avg reach aftereffects: %0.1f (all targets)\n',mean(nc.cla$endpoint_angle, na.rm=TRUE)))
+    cat(sprintf('avg reach aftereffects: %0.1f (targets above 15 degrees)\n',mean(nc.cla$endpoint_angle[which(nc.cla$target > 15)], na.rm=TRUE)))
+    
+  }
   
 }
 
@@ -283,13 +303,17 @@ exposureAftereffectsPersistent <- function() {
   
 }
 
-exposureClassicReachAftereffects <- function(noTarget=FALSE) {
[…]

### 2018-05-07 shared.R

@@ -21,6 +21,14 @@ colorset[['claPasT']] <- '#ff82002f'
 colorset[['extra1S']] <- '#c400c4ff' # purple
 colorset[['extra1T']] <- '#c400c42f'
 
+colorset[['onlActS']] <- '#c400c4ff' # purple
+colorset[['onlActT']] <- '#c400c42f'
+# colorset[['onlPasS']] <- '#8266f4ff' # violet
+# colorset[['onlPasT']] <- '#8266ff2f'
+colorset[['onlPasS']] <- '#ff6ec7ff' # pink
+colorset[['onlPasT']] <- '#ff6ec72f'
+
+
 library('ez')
 
 installRequire.Packages <- function(packages) {
@@ -492,10 +500,12 @@ removeOutliers <- function(df, stds=2) {
 
 psychometricGeneralization <- function(df, makefigure=FALSE, response='RAE') {
   
+  # well... psychometric... Ijust use a cumulative normal distribution function
+  
   # does everything use the same dependent variables?
   
   if (makefigure) {
-    par(mfrow=c(4,6))
+    par(mfrow=c(4,6),mar=c(4,4,3,0.5))
   }
   
   participant <- unique(df$participant)
@@ -504,29 +514,23 @@ psychometricGeneralization <- function(df, makefigure=FALSE, response='RAE') {
   
   for (pp.id in participant) {
     
-    # get data fro participant:
-    Pdf <- df[which(df$participant == pp.id),]
-    
-    # get dependent variable (endpoint/tap angle):
+    # let's get the group thing first
+    Gdf <- df[-which(df$participant == pp.id),]
     if (response == 'RAE') {
-      DV <- Pdf$endpoint_angle
+      Gcurve <- aggregate(endpoint_angle ~ target, data=Gdf, FUN=median)
+      Gcurve <- Gcurve$endpoint_angle
     }
     if (response == 'loc') {
-      # NA.idx <- which(is.na(Pdf$taperror_deg))
-      # print(length(NA.idx))
-      Pdf <- Pdf[which(!is.na(Pdf$taperror_deg)),]
-      # print(str(Pdf))
-      DV <- -1 * Pdf$taperror_deg
+      Gcurve <- aggregate(taperror_deg ~ handangle_deg, data=Gdf, FUN=median)
+      Gcurve <- -1 * Gcurve$taperror_deg
     }
     
-    DV <- DV - min(DV)
[…]

### 2018-05-10 ManuscriptMaterials.Rmd
Switched to Satterthwaite estimates for all LME analyses.

@@
 output:
+  pdf_document:
+    number_sections: yes
+    toc: yes
+    toc_depth: 2
   html_document:
@@
     toc_depth: '2'
-  pdf_document:
-    number_sections: yes
-    toc: yes
-    toc_depth: 2
 ---
@@
 
+## Analyses
+
+Since in the localization tasks there is missing data at some points (usually at the extremes) we can't use ANOVAs directly on the data.
+Imputation is going to be very hard because the missing data is systematically missing.
+Hence, we are using Linear Mixed Effects models as an alternative, but this presents a new problem.
+Getting the usual F-values and p-values is not so straightforward.
+Initially we decided to use the $\chi^2$ approach that we get from the `Anova` function from the `car` package.
+However, a paper from Steven G.
+Luke (2017, _Evaluating significance in linear mixed-effects models in R_) suggests using Sattertwaite approximation.
+So all functions running an LME now take an argument that tells them to use the Satterthwaite method, but the default is chi-squared.
+It turns out that it doesn't matter much, so it's a good sanity check.
+For consistency across all the analyses, we set the value of that argument is set here:
+
+```{r}
+LMEmethod <- 'Satterthwaite'
+# options: Satterthwaite, chi-squared
+```
+
 # Reach Aftereffects
@@
 ```{r RAE_exposure,include=TRUE}
-exposureNoCursorChange()
+exposureNoCursorChange(LMEmethod=LMEmethod)
 ```
@@
 ```{r RAE_exp_iterations}
-exposureAftereffectsPersistent()
+exposureAftereffectsPersistent(LMEmethod=LMEmethod)
 ```
@@
 ```{r RAE_exp_cla}
-exposureClassicReachAftereffects()
+exposureClassicReachAftereffects(LMEmethod=LMEmethod)
 ```
@@
 ```{r loc_exp}
-exposureLocalization(remove15=TRUE)
+exposureLocalization(remove15=TRUE, LMEmethod=LMEmethod)
 ```
@@
 ```{r loc_exp_movtype}
-exposureLocalizationShift(remove15=TRUE)
+exposureLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod)
 ```
@@
 ```{r loc_ombnibus}
-groupLocalization(model='full',remove15=TRUE)
+groupLocalization(model='full', remove15=TRUE, LMEmethod=LMEmethod)
 ```
@@
 ```{r loc_groups_movementtype}
-groupLocalization(model='movementtype',remove15=TRUE)
+groupLocalization(model='movementtype', remove15=TRUE, LMEmethod=LMEmethod)
 ```
@@
 ```{r loc_cla}
-classicLocalizationShift(remove15=TRUE)
+classicLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod)
 ```

### 2018-05-10 localization.R

@@ -235,7 +235,7 @@ plotLocalization <- function(classicOnline=FALSE) {
   
 }
 
-exposureLocalization <- function(remove15=TRUE) {
+exposureLocalization <- function(remove15=TRUE, LMEmethod='chi-squared') {
   
   default.contrasts <- options('contrasts')
   options(contrasts=c('contr.sum','contr.poly'))
@@ -254,7 +254,14 @@ exposureLocalization <- function(remove15=TRUE) {
   attach(exp)
   
   cat('\nLME with session, target and movement type as fixed effects, and participant as random effect:\n\n')
-  print(Anova(lme(taperror_deg ~ rotated_b * passive_b * handangle_deg, random = ~1|participant, na.action=na.exclude), type=3))
+  
+  if (LMEmethod=='chi-squared') {
+    print(Anova(lme(taperror_deg ~ rotated_b * passive_b * handangle_deg, random = ~1|participant, na.action=na.exclude), type=3))
+  }
+  if (LMEmethod=='Satterthwaite') {
+    exp_model_lmer <- lmer(taperror_deg ~ rotated_b * passive_b * handangle_deg - (1|participant), na.action=na.exclude)
+    print(anova(exp_model_lmer,ddf='Satterthwaite',type=3))
+  }
   
   detach(exp)
   
@@ -262,7 +269,7 @@ exposureLocalization <- function(remove15=TRUE) {
   
 }
 
-exposureLocalizationShift <- function(noHandAngle=FALSE, remove15=TRUE) {
+exposureLocalizationShift <- function(noHandAngle=FALSE, remove15=TRUE, LMEmethod='chi-squared') {
   
   default.contrasts <- options('contrasts')
   options(contrasts=c('contr.sum','contr.poly'))
@@ -280,12 +287,29 @@ exposureLocalizationShift <- function(noHandAngle=FALSE, remove15=TRUE) {
   attach(exp)
   
   if (noHandAngle) {
+    
     cat('\nLME with movement type as fixed effects - ignoring hand angle, and participant as random effect:\n\n')
-    print(Anova(lme(taperror_deg ~ passive_b, random = ~1|participant, na.action=na.exclude), type=3))
+    
+    if (LMEmethod=='chi-squared') {
+      print(Anova(lme(taperror_deg ~ passive_b, random = ~1|participant, na.action=na.exclude), type=3))
+    }
+    if (LMEmethod=='Satterthwaite') {
[…]

### 2018-05-10 nocursor.R

@@ -7,7 +7,7 @@ source('shared.R')
 # not yet sure if we'll use Chi-square, or Satterthwaite estimates of F/p-values:
 # installRequire.Packages(c('nlme', 'car', 'lme4', 'lmerTest'))
 
-required.packages = c('nlme','car')
+required.packages = c('nlme', 'car', 'lme4', 'lmerTest')
 installRequire.Packages(required.packages)
 
 # nocursor data can be downloaded from OSF:
@@ -233,7 +233,7 @@ plotReachAftereffectDistributions <- function() {
   
 }
 
-exposureNoCursorChange <- function(printChange=FALSE) {
+exposureNoCursorChange <- function(printChange=FALSE,LMEmethod='chi-squared') {
   
   default.contrasts <- options('contrasts')
   options(contrasts=c('contr.sum','contr.poly'))
@@ -247,7 +247,13 @@ exposureNoCursorChange <- function(printChange=FALSE) {
   attach(nc.exp)
   
   cat('\nLME with session and target as fixed effects, and participant as random effect:\n\n')
-  print(Anova(lme(endpoint_angle ~ rotated * target, random = ~1|participant, na.action=na.exclude), type=3))
+  if (LMEmethod=='chi-squared') {
+    print(Anova(lme(endpoint_angle ~ rotated * target, random = ~1|participant, na.action=na.exclude), type=3))
+  }
+  if (LMEmethod=='Satterthwaite') {
+    exp_model_lmer <- lmer(endpoint_angle ~ rotated * target - (1|participant), na.action=na.exclude)
+    print(anova(exp_model_lmer,ddf='Satterthwaite',type=3))
+  }
   
   detach(nc.exp)
 
@@ -275,7 +281,7 @@ exposureNoCursorChange <- function(printChange=FALSE) {
   
 }
 
-exposureAftereffectsPersistent <- function() {
+exposureAftereffectsPersistent <- function(LMEmethod='chi-squared') {
   
   default.contrasts <- options('contrasts')
   options(contrasts=c('contr.sum','contr.poly'))
@@ -295,7 +301,16 @@ exposureAftereffectsPersistent <- function() {
   attach(exp)
   
   cat('\nLME with session and target as fixed effects, and participant as random effect:\n\n')
-  print(Anova(lme(endpoint_angle ~ iteration * target, random = ~1|participant, na.action=na.exclude), type=3))
+  
+  
+  if (LMEmethod=='chi-squared') {
[…]

### 2018-05-10 relateLocalizationNoCursors.R

@@ -172,8 +172,12 @@ multipleRegressionLocalization <- function(group,NCpart='all',expart='all',LRpar
   
   cat(sprintf('\n%s\n\n',toupper(group)))
   
-  fm <- step(lm(RAE ~ active + passive, data=df))
-    
+  str(df)
+  
+  # THIS CAUSES AN ERROR NOW: data argument is of the wrong type...
+  # because STEP is imported from lmerTest, but we want the basic one from 'stats'
+  fm <- stats::step(lm(RAE ~ active + passive, data=df))
+  
   print(summary(fm))
   
 }

### 2018-05-10 shared.R

@@ -11,8 +11,8 @@ colorset <- list()
 
 colorset[['expActS']] <- '#005de4ff' # blue
 colorset[['expActT']] <- '#005de42f'
-colorset[['expPasS']] <- '#2ab2f2ff' # lighter blue
-colorset[['expPasT']] <- '#2ab2f22f'
+colorset[['expPasS']] <- '#0fd2f2ff' # lighter blue
+colorset[['expPasT']] <- '#0fd2f22f'
 colorset[['claActS']] <- '#e51636ff' # "York red"
 colorset[['claActT']] <- '#e516362f'
 colorset[['claPasS']] <- '#ff8200ff' # orange
@@ -21,8 +21,8 @@ colorset[['claPasT']] <- '#ff82002f'
 colorset[['extra1S']] <- '#c400c4ff' # purple
 colorset[['extra1T']] <- '#c400c42f'
 
-colorset[['onlActS']] <- '#c400c4ff' # purple
-colorset[['onlActT']] <- '#c400c42f'
+colorset[['onlActS']] <- '#b400e4ff' # purple
+colorset[['onlActT']] <- '#b400e42f'
 # colorset[['onlPasS']] <- '#8266f4ff' # violet
 # colorset[['onlPasT']] <- '#8266ff2f'
 colorset[['onlPasS']] <- '#ff6ec7ff' # pink

### 2018-05-10 ManuscriptMaterials.Rmd
Added option for plot functions to create SVG files instead of creatng output in notebook / R.

@@
 
-```{r fig2_reachaftereffect, fig.width=7.5, fig.height=4}
+```{r fig2_reachaftereffect, fig.width=7.5, fig.height=2.5}
 plotReachAftereffects()
@@
 
-```{r scattergrams, fig.width=7.5, fig.height=4}
+```{r scattergrams, fig.width=7.5, fig.height=2.5}
 correlations <- correlateNoCursorsLocalization()
@@
 Perhaps with higher quality estimates of proprioceptive recalibration those would be able to predict part of the reach aftereffects even after classic training.
+
+# Figures
+
+Here we simply generate the same figures as above, but now they are saved as SVG files.
+
+```{r create_svgs}
+plotReachAftereffects(generateSVG=TRUE)
+plotLocalization(classicOnline=TRUE, generateSVG=TRUE)
+correlations <- correlateNoCursorsLocalization(generateSVG=TRUE)
+```

### 2018-05-10 localization.R

@@ -124,7 +124,7 @@ localizationLMEsSatterthwaite <- function() {
 
 # PLOTS / FIGURES ------
 
-plotLocalization <- function(classicOnline=FALSE) {
+plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
   
   # get the data to plot:
   exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE)
@@ -161,11 +161,17 @@ plotLocalization <- function(classicOnline=FALSE) {
     onl.CI.pas <- matrix(unlist(by(onl.pas$taperror_deg, INDICES=c(onl.pas$handangle_deg), FUN=t.interval)),nrow=2)
   }
   
-  if (classicOnline) {
-    par(mfrow=c(1,3))
-  } else {
-    par(mfrow=c(1,2))
+  if (generateSVG) {
+    installed.list <- rownames(installed.packages())
+    if ('svglite' %in% installed.list) {
+      svglite(file='Fig2.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
+    } else {
+      generateSVG=FALSE
+    }
   }
+  
+  par(mfrow=c(1,3))
+  
   points <- c(15,25,35,45,55,65,75)
   
   # panel A: exposure localization (active vs. passive)
@@ -231,6 +237,10 @@ plotLocalization <- function(classicOnline=FALSE) {
     
     legend(10,-15,c('passive','active'),col=c(colorset[['onlPasS']],colorset[['onlActS']]),lty=c(1,1),lwd=c(2,2),bty='n')
     
+  }
+  
+  if (generateSVG) {
+    dev.off()
   }
   
 }

### 2018-05-10 nocursor.R

@@ -7,7 +7,7 @@ source('shared.R')
 # not yet sure if we'll use Chi-square, or Satterthwaite estimates of F/p-values:
 # installRequire.Packages(c('nlme', 'car', 'lme4', 'lmerTest'))
 
-required.packages = c('nlme', 'car', 'lme4', 'lmerTest')
+required.packages = c('nlme', 'car', 'lme4', 'lmerTest', 'svglite')
 installRequire.Packages(required.packages)
 
 # nocursor data can be downloaded from OSF:
@@ -66,9 +66,18 @@ installRequire.Packages(required.packages)
 #   
 # }
 
-plotReachAftereffects <- function() {
+plotReachAftereffects <- function(generateSVG=FALSE) {
   
-  par(mfrow=c(1,2))
+  if (generateSVG) {
+    installed.list <- rownames(installed.packages())
+    if ('svglite' %in% installed.list) {
+      svglite(file='Fig1.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
+    } else {
+      generateSVG=FALSE
+    }
+  }
+  
+  par(mfrow=c(1,3))
   
   points <- c(15,25,35,45,55,65,75)
   
@@ -96,15 +105,15 @@ plotReachAftereffects <- function() {
   plot(-1000,-1000, main='decay of reach aftereffects', xlab='target angle [deg]', ylab='reach endpoint deviation [deg]', xlim=c(10,80), ylim=c(0,15), axes=F)
   
   polygon(X,expYrem,border=NA,col=colorset[['expActT']])
-  polygon(X,expYini,border=NA,col=colorset[['extra1T']])
+  polygon(X,expYini,border=NA,col=colorset[['expPasT']])
   
   lines(points,exposureAVGrem$endpoint_angle,col=colorset[['expActS']],lty=2,lwd=2)
-  lines(points,exposureAVGini$endpoint_angle,col=colorset[['extra1S']],lwd=2)
+  lines(points,exposureAVGini$endpoint_angle,col=colorset[['expPasS']],lwd=2)
   
   axis(1,at=points)
   axis(2,at=c(0,5,10,15))
   
-  legend(10,15,c('immediate (iteration 1)','delayed (iterations 2-5)'),col=c(colorset[['extra1S']],colorset[['expActS']]),lty=c(1,2),lwd=c(2,2),bty='n')
[…]

### 2018-05-10 relateLocalizationNoCursors.R

@@ -3,7 +3,7 @@ source('shared.R')
 
 
 
-correlateNoCursorsLocalization <- function(NCpart='all') {
+correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE) {
   
   # first we do all responses, plan to split it by target later
   # get all data, for classic and exposure, we need localization and reach aftereffects
@@ -75,8 +75,16 @@ correlateNoCursorsLocalization <- function(NCpart='all') {
   ylab <- 'reach aftereffect [deg]'
   xlab <- 'localization shift [deg]'
   
+  if (generateSVG) {
+    installed.list <- rownames(installed.packages())
+    if ('svglite' %in% installed.list) {
+      svglite(file='Fig3.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
+    } else {
+      generateSVG=FALSE
+    }
+  }
   
-  par(mfrow=c(1,2))
+  par(mfrow=c(1,3))
   
   dataframes = list()
   dataframes[['exposure']] <- exposure
@@ -128,6 +136,10 @@ correlateNoCursorsLocalization <- function(NCpart='all') {
       axis(2,at=yticks)
     }
     
+  }
+  
+  if (generateSVG) {
+    dev.off()
   }
   
   return(output)

### 2018-05-10 shared.R

@@ -11,8 +11,8 @@ colorset <- list()
 
 colorset[['expActS']] <- '#005de4ff' # blue
 colorset[['expActT']] <- '#005de42f'
-colorset[['expPasS']] <- '#0fd2f2ff' # lighter blue
-colorset[['expPasT']] <- '#0fd2f22f'
+colorset[['expPasS']] <- '#0fd2e2ff' # lighter blue
+colorset[['expPasT']] <- '#0fd2e22f'
 colorset[['claActS']] <- '#e51636ff' # "York red"
 colorset[['claActT']] <- '#e516362f'
 colorset[['claPasS']] <- '#ff8200ff' # orange
@@ -51,8 +51,10 @@ installRequire.Packages <- function(packages) {
   if (length(missingpackages) > 0) {
     cat('\nWARNING, some pacakages are missing:\n')
     cat(missingpackages)
-    cat('\ncar and nlme are required to reproduce the analyses\n')
+    cat('\nlme4 and lmerTest are required to reproduce the Satterthwaite LME analyses\n')
+    cat('\ncar and nlme are required to reproduce the analyses using Chi-squared apprcimation\n')
     cat('\nforeach and doParallel speed up a few functions\n')
+    cat('\nsvglite is used to produce SVG files with figures\n')
   }
   
 }

### 2018-05-10 localization.R
changed figure numbers

@@ -164,7 +164,7 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
     if ('svglite' %in% installed.list) {
-      svglite(file='Fig2.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
+      svglite(file='Fig3.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
     } else {
       generateSVG=FALSE
     }

### 2018-05-10 nocursor.R

@@ -71,7 +71,7 @@ plotReachAftereffects <- function(generateSVG=FALSE) {
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
     if ('svglite' %in% installed.list) {
-      svglite(file='Fig1.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
+      svglite(file='Fig2.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
     } else {
       generateSVG=FALSE
     }

### 2018-05-10 relateLocalizationNoCursors.R

@@ -78,7 +78,7 @@ correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE) {
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
     if ('svglite' %in% installed.list) {
-      svglite(file='Fig3.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
+      svglite(file='Fig4.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
     } else {
       generateSVG=FALSE
     }

### 2018-05-10 ManuscriptMaterials.Rmd
small change

@@
 
-```{r create_svgs}
+```{r create_svgs, cache=FALSE, include=FALSE}
 plotReachAftereffects(generateSVG=TRUE)

### 2018-05-18 localization.R
improved figures

@@ -176,7 +176,9 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
   
   # panel A: exposure localization (active vs. passive)
   
-  plot(-1000,-1000, main='exposure', xlab='hand angle [deg]', ylab='localization shift [deg]', xlim=c(10,80), ylim=c(0,-15), axes=F)
+  plot(-1000,-1000, main='exposure', xlab='hand angle [°]', ylab='localization shift [°]', xlim=c(10,80), ylim=c(0,-15), axes=F)
+  
+  mtext('A', side=3, outer=TRUE, at=c(0,1), line=-1, adj=0, padj=1, family='Arial')
   
   X <- c(points, rev(points))
   exp.act.Y <- c(exp.CI.act[1,],rev(exp.CI.act[2,]))
@@ -185,19 +187,21 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
   polygon(X,exp.act.Y,border=NA,col=colorset[['expActT']])
   polygon(X,exp.pas.Y,border=NA,col=colorset[['expPasT']])
   
-  lines(points[1:2],exp.avg.act$taperror_deg[1:2],col=colorset[['expActS']],lty=2,lwd=2)
-  lines(points[2:7],exp.avg.act$taperror_deg[2:7],col=colorset[['expActS']],lty=1,lwd=2)
-  lines(points[1:2],exp.avg.pas$taperror_deg[1:2],col=colorset[['expPasS']],lty=2,lwd=2)
-  lines(points[2:7],exp.avg.pas$taperror_deg[2:7],col=colorset[['expPasS']],lty=1,lwd=2)
+  lines(points[1:2],exp.avg.act$taperror_deg[1:2],col=colorset[['expActS']],lty=2,lwd=1.5)
+  lines(points[2:7],exp.avg.act$taperror_deg[2:7],col=colorset[['expActS']],lty=1,lwd=1.5)
+  lines(points[1:2],exp.avg.pas$taperror_deg[1:2],col=colorset[['expPasS']],lty=2,lwd=1.5)
+  lines(points[2:7],exp.avg.pas$taperror_deg[2:7],col=colorset[['expPasS']],lty=1,lwd=1.5)
   
   axis(1,at=points)
   axis(2,at=c(0,-5,-10,-15))
   
[…]

### 2018-05-18 nocursor.R

@@ -90,7 +90,7 @@ plotReachAftereffects <- function(generateSVG=FALSE) {
   exposureAVGrem <- aggregate(endpoint_angle ~ target, data=exposure_rem, FUN=mean)
   classicAVG <- aggregate(endpoint_angle ~ target, data=classic, FUN=mean)
   exposureAVG <- aggregate(endpoint_angle ~ target, data=exposure, FUN=mean)
-  
+
   exposureCIini <- matrix(unlist(by(exposure_ini$endpoint_angle, INDICES=c(exposure_ini$target), FUN=t.interval)),nrow=2)
   exposureCIrem <- matrix(unlist(by(exposure_rem$endpoint_angle, INDICES=c(exposure_rem$target), FUN=t.interval)),nrow=2)
   classicCI <- matrix(unlist(by(classic$endpoint_angle, INDICES=c(classic$target), FUN=t.interval)),nrow=2)
@@ -102,31 +102,35 @@ plotReachAftereffects <- function(generateSVG=FALSE) {
   claY <- c(classicCI[1,],rev(classicCI[2,]))
   expY <- c(exposureCI[1,],rev(exposureCI[2,]))
   
-  plot(-1000,-1000, main='decay of reach aftereffects', xlab='target angle [deg]', ylab='reach endpoint deviation [deg]', xlim=c(10,80), ylim=c(0,15), axes=F)
+  plot(-1000,-1000, main='decay of reach aftereffects', xlab='target angle [°]', ylab='reach endpoint deviation [°]', xlim=c(10,80), ylim=c(0,15), axes=F)
+  
+  mtext('A', side=3, outer=TRUE, at=c(0,1), line=-1, adj=0, padj=1, family='Arial')
   
   polygon(X,expYrem,border=NA,col=colorset[['expActT']])
   polygon(X,expYini,border=NA,col=colorset[['expPasT']])
   
-  lines(points,exposureAVGrem$endpoint_angle,col=colorset[['expActS']],lty=2,lwd=2)
-  lines(points,exposureAVGini$endpoint_angle,col=colorset[['expPasS']],lwd=2)
+  lines(points,exposureAVGrem$endpoint_angle,col=colorset[['expActS']],lty=2,lwd=1.5)
+  lines(points,exposureAVGini$endpoint_angle,col=colorset[['expPasS']],lwd=1.5)
   
   axis(1,at=points)
   axis(2,at=c(0,5,10,15))
   
[…]

### 2018-05-18 relateLocalizationNoCursors.R

@@ -72,8 +72,8 @@ correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE) {
   xticks <- seq(from=-20,to=10,by=10)
   yticks <- rev(seq(from=-10,to=20,by=10))
   
-  ylab <- 'reach aftereffect [deg]'
-  xlab <- 'localization shift [deg]'
+  ylab <- 'reach aftereffect [°]'
+  xlab <- 'localization shift [°]'
   
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
@@ -92,9 +92,15 @@ correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE) {
   
   output <- list()
   
-  for (group in c('exposure','classic')) {
+  groups <- c('exposure','classic')
+  
+  for (groupno in c(1:length(groups))) {
+    
+    group <- groups[groupno]
     
     plot(-1000, -1000, main=group, xlab=xlab, ylab=ylab, asp=1, axes=FALSE, xlim=xlimits, ylim=ylimits)
+    
+    mtext(toupper(letters[groupno]), side=3, outer=TRUE, at=c((groupno-1)/3,1), line=-1, adj=0, padj=1, family='Arial')
     
     lines(c(0,0),range(yticks),col=rgb(.5,.5,.5,.5))
     lines(range(xticks),c(0,0),col=rgb(.5,.5,.5,.5))
@@ -127,7 +133,7 @@ correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE) {
       
     }
     
-    legend(-3,-3,c('active','passive'),col=legendlinecolors,lwd=c(2,2),bty='n')
+    legend(-3,-3,c('active','passive'),col=legendlinecolors,lwd=c(1.5,1.5),bty='n')
     
     if (length(xticks)) {
       axis(1,at=xticks)
@@ -225,6 +231,6 @@ plotCorrelationWithCI <- function(X,Y,colors=c('black','red'),main='main title',
   polygon(c(pointlocs,rev(pointlocs)),c(y1,rev(y2)), col=colors[1], border=NA)
   
   # and use that to show a line:
-  lines(range(X), predict(this.lm, newdata=data.frame(X=range(X))), col=colors[2], lwd=2)
+  lines(range(X), predict(this.lm, newdata=data.frame(X=range(X))), col=colors[2], lwd=1.5)
   
 }

### 2018-05-18 shared.R

@@ -28,7 +28,6 @@ colorset[['onlActT']] <- '#b400e42f'
 colorset[['onlPasS']] <- '#ff6ec7ff' # pink
 colorset[['onlPasT']] <- '#ff6ec72f'
 
-
 library('ez')
 
 installRequire.Packages <- function(packages) {

### 2018-05-18 ManuscriptMaterials.Rmd
improved figures

--- 2018-05-10 ManuscriptMaterials.Rmd
@@
 
-Now there is a trend towards significance for the interaction between training group and movement type. That is closer to what we would expect.
+Still no effect...
 
@@
 
-# Figures
-
-Here we simply generate the same figures as above, but now they are saved as SVG files.
-
-```{r create_svgs}
+```{r create_svgs, cache=FALSE, include=FALSE}
+# This chunk creates SVG files but shouldn't show up in the PDF / HTML 
 plotReachAftereffects(generateSVG=TRUE)

### 2018-05-18 localization.R

@@ -178,7 +178,7 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
   
   plot(-1000,-1000, main='exposure', xlab='hand angle [°]', ylab='localization shift [°]', xlim=c(10,80), ylim=c(0,-15), axes=F)
   
-  mtext('A', side=3, outer=TRUE, at=c(0,1), line=-1, adj=0, padj=1, family='Arial')
+  mtext('A', side=3, outer=TRUE, at=c(0,1), line=-1, adj=0, padj=1)
   
   X <- c(points, rev(points))
   exp.act.Y <- c(exp.CI.act[1,],rev(exp.CI.act[2,]))
@@ -201,7 +201,7 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
   
   plot(-1000,-1000, main='classic', xlab='hand angle [°]', ylab='localization shift [°]', xlim=c(10,80), ylim=c(0,-15), axes=F)
   
-  mtext('B', side=3, outer=TRUE, at=c(1/3,1), line=-1, adj=0, padj=1, family='Arial')
+  mtext('B', side=3, outer=TRUE, at=c(1/3,1), line=-1, adj=0, padj=1)
   
   X <- c(points, rev(points))
   cla.act.Y <- c(cla.CI.act[1,],rev(cla.CI.act[2,]))
@@ -224,7 +224,7 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
     
     plot(-1000,-1000, main='online', xlab='hand angle [°]', ylab='localization shift [°]', xlim=c(10,80), ylim=c(0,-15), axes=F)
     
-    mtext('C', side=3, outer=TRUE, at=c(2/3,1), line=-1, adj=0, padj=1, family='Arial')
+    mtext('C', side=3, outer=TRUE, at=c(2/3,1), line=-1, adj=0, padj=1)
     
     X <- c(points, rev(points))
     onl.act.Y <- c(onl.CI.act[1,],rev(onl.CI.act[2,]))

### 2018-05-18 nocursor.R

@@ -104,7 +104,7 @@ plotReachAftereffects <- function(generateSVG=FALSE) {
   
   plot(-1000,-1000, main='decay of reach aftereffects', xlab='target angle [°]', ylab='reach endpoint deviation [°]', xlim=c(10,80), ylim=c(0,15), axes=F)
   
-  mtext('A', side=3, outer=TRUE, at=c(0,1), line=-1, adj=0, padj=1, family='Arial')
+  mtext('A', side=3, outer=TRUE, at=c(0,1), line=-1, adj=0, padj=1)
   
   polygon(X,expYrem,border=NA,col=colorset[['expActT']])
   polygon(X,expYini,border=NA,col=colorset[['expPasT']])
@@ -119,7 +119,7 @@ plotReachAftereffects <- function(generateSVG=FALSE) {
   
   plot(-1000,-1000, main='reach aftereffects', xlab='target angle [°]', ylab='reach endpoint deviation [°]', xlim=c(10,80), ylim=c(0,15), axes=F)
   
-  mtext('B', side=3, outer=TRUE, at=c(1/3,1), line=-1, adj=0, padj=1, family='Arial')
+  mtext('B', side=3, outer=TRUE, at=c(1/3,1), line=-1, adj=0, padj=1)
   
   polygon(X,claY, border=NA,col=colorset[['claActT']])
   polygon(X,expY, border=NA,col=colorset[['expActT']])

### 2018-05-18 relateLocalizationNoCursors.R

@@ -100,7 +100,7 @@ correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE) {
     
     plot(-1000, -1000, main=group, xlab=xlab, ylab=ylab, asp=1, axes=FALSE, xlim=xlimits, ylim=ylimits)
     
-    mtext(toupper(letters[groupno]), side=3, outer=TRUE, at=c((groupno-1)/3,1), line=-1, adj=0, padj=1, family='Arial')
+    mtext(toupper(letters[groupno]), side=3, outer=TRUE, at=c((groupno-1)/3,1), line=-1, adj=0, padj=1)
     
     lines(c(0,0),range(yticks),col=rgb(.5,.5,.5,.5))
     lines(range(xticks),c(0,0),col=rgb(.5,.5,.5,.5))

### 2018-06-23 ManuscriptMaterials.Rmd
small changes

@@
 output:
+  html_document:
+    df_print: paged
+    toc: yes
+    toc_depth: '2'
   pdf_document:
@@
     toc_depth: 2
-  html_document:
-    df_print: paged
-    toc: yes
-    toc_depth: '2'
 ---
@@
     + B: Classic, active and passive (goal: larger difference between active and passive, different generalization?)
+    + C: Online localization?
 
@@
 
-The scripts doing the statistics and figures are in separate R files. They allow many options, but here we will only see / do the ones that we think should go to the manuscript.
+The scripts doing the statistics and creating figures are in separate R files. They allow many options, but here we will only see / do the ones that we think should go to the manuscript.
 
@@
 
-In the manuscript we'll first show that there are motor changes after exposure training (with the no-cursor data) and then compare that with the classic.
-Second is the localization: a) localization shifts after exposure b) the shifts are the same/different from those after classic c) the effects are different across the workspace d) this pattern is indistinguishable / different from that after classic.
+In the manuscript we'll first show that there indeed are motor changes after exposure training (with the no-cursor data) and then compare that with the classic.
+Second is the localization: a) localization shifts after exposure b) the shifts are (not) different from those after classic c) the effects are (not) different across the workspace d) this pattern is (not) different from that after classic.
 A third topic might be if localization shifts can predict no-cursor changes, which we'll call the "correlation"  topic.
@@
 ## Analyses
+
+### ANOVA or LME
 
@@
 
+### Peak location of generalization
+
+Since participants trained with a single target, we could look at generalization curves.
+These appear to be shifted more counter clockwise in most cases, so we want to quantify this.
+But, for most of the curves, and especially the ones that seems to be shifted, we don't have data on the full curve on one side of the peak.
+Or maybe it doesn't even include the peak itself.
+That makes it hard to fit a normal function to the data.
+Instead, we assume that the minimal change in the data, occurs far beyond the boundaries of the measured workspace, make the data cumulative and scale back to a range from 0 to 1.
+This allows fitting a _cumulative_ normal function where the 50% point corresponds to the peak of the generalization curve.
+This method is more stable, and actually works on all the datasets used (but we didn't test online localization from the classic experiment).
+However, the results should still be considered carefully.
+
 # Reach Aftereffects
@@
 ```
-
-_For Panel B: Earlier I used the initial block for this plot, but after discussion, we decided to use all data: there's no significant difference between the initial and other blocks, but using all data allows more accurate estimates._
 
@@
 The second analysis is to show that the reach aftereffects aren't all that different for first and later iterations of the rotated no-cursor tasks.
-(The baseline is always all aligned iterations of the task.)
+(The rotated data is always compared with _all_ aligned iterations of the task.)
 
@@
 
-So there is no effect of iteration here, which means the reach aftereffects are the same for the first iteration and the average across the other four iterations.
-We now also have enough power to see an effect of target angle (the generalization curve is not flat), but more importantly the effect of iteration is also not dependent on target.
+So there is no effect of iteration here, nor does it interact with target, which means the reach aftereffects are the same for the first iteration and the average across the other four iterations.
+We now also have enough power to see an effect of target angle: the generalization curve is not flat.
 
@@
 ```{r RAE_peaks}
-getPeakConfidenceInterval('classic',part='all',CIs=c(.95))
-getPeakConfidenceInterval('exposure',part='all',CIs=c(.95))
+getPeakConfidenceInterval('classic', part='all', CIs=c(.95))
+getPeakConfidenceInterval('exposure', part='all', CIs=c(.95))
 ```
 
-Although the two peaks are not different from each other, the 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training) but not the 75 degree target, whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training) but not the 45 degree target.
-Something might be going on -- but the evidence is far from concrete.
+The 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training) but not the 75 degree target, whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training) but not the 45 degree target.
+Something might be going on -- but the two confidence intervals are not different from each other, which is in line with the LME results.
+We can't conclude much, except that we have no evidence for different generalization curves.
 
@@
 2. these changes persist over time, and
-3. these changes are comparable to the ones we found after classic training.
+3. these changes are comparable to the ones we found after classic training, and
+4. there is some generalization curve across the workspace, but
+5. we have no evidence that this is different from reach aftereffect generalization after classic training.
 
@@
 
-That looks OK to me.
 It seems that just like after classic training, there is a change in localization responses following exposure training.
@@
 
-We first want to see if there is any overall effect of aligned versus rotated training in the exposure group, with movement type and hand angle as fixed effects, and particiant as random effect.
+We first want to see if there is any overall effect of aligned versus rotated training in the exposure group (i.e.: are there any localization shifts at all?), with movement type and hand angle as fixed effects, and particiant as random effect.
 
@@
 In other words: exposure training caused localization responses to shift (systematically), as we could already guess by looking at the figure.
-
 This means we can look at the difference between repsonses after rotated and after aligned feedback: the training induced shift in localization.
+The shifts are also different, given different hand angles, which means there is some form of generalization pattern.
 
@@
 This could be because the effect of movement type is much larger in the classic group.
-This would predict an interaction between group and movement type, but this is not even a trend (p=.133).
+This would predict an interaction between group and movement type, but that's not the case (p=.133).
 
@@
 
-Let's see if the classic group shows an effect of movement type (with the current preprocessing procedures):
+Let's see if the classic group still shows an effect of movement type (with the current preprocessing procedures):
 
@@
 So there is a fairly strong effect of movement type in the classic group.
-Together with the absence of an effect of movement type in the exposure group and a trend of an interaction between group and movement type, this might indicate a sub-threshold effect of movement type in the exposure group in the same direction as the effect of movement type in the classic group.
+Together with the absence of an effect of movement type in the exposure group, but no interaction between group and movement type if we combine both datasets.
 
-Also note the absence of an effect of hand angle in the classic group here -- potentially because we excluded the 15 degree target.
+Also note the absence of an effect of hand angle in the classic group here -- probably because we excluded the 15 degree target, but it seems that after classic training the generalization curve for delayed localization is much flatter.
 
@@
 
-2. In the exposure group, there is no effect of movement type: localization responses with and without efference copies are indistinguishable.
-3. In the classic group, there is an effect of movement type: localization responses informed by efference copies are shifted more than those with only proprioception (as found before).
-4.
-This predicts an interaction between movement type and group, but that is only a trend, suggesting that either the difference in the classic group is not so large, or that there is a small, sub-threshold difference in the exposure group.
+2. In the exposure group, we find no difference between localization responses with and without efference copies.
+3. In the classic group, the localization responses informed by efference copies, are shifted more than those with only proprioception (as found before).
+4. This predicts an interaction between movement type and group in the combined analysis, but we don't see that.
 
@@
 
-Decided to add analyses to get the confidence intervals for the peaks of the localization generalization curves:
+Here are the confidence intervals for the peaks of the generalization curves for _active_ localization:
 
@@
 
-These results mimic those of the reach aftereffects: the median values are different, but they each fall in the other confidence interval.
-If we want to look at this better, we need to get better estimates of these curves in every participants, by having a wider range of points and more data per point, but probably also larger rotations.
-
-## Cumulative Function
-
-Potentially, the wide confidence intervals can be explained by two things: 1) large variation across participants, and 2) incomplete sampling over the function: large swaths of the normal function (not distribution!) are sometimes not sampled at all.
-Here are some ideas for solutions.
-
-The first problem -- large differences between participants -- could be dealt with by estimating the function with each participant.
-This way only conditions within groups can be compared, but that still leaves comparisons between active and passive localization, although comparing reach aftereffects between groups would still not be possible.
-We can do a Fisher exact test to see if the proportion people with larger shifts in active than in passive localization is higher than 50% in each group.
-Prediction: should not be different from chance in exposure training, but should be different from chance in classic training.
-
-The second problem -- incomplete sampling of the curve -- could be dealt with by fitting a cumulative function instead of a normal function.
-We could assume the training-induced change is 0 at -90 and at 180 degrees, then at every actually estimated point, we take the sum of everything before it, and then scale so that values are between 0 and 1 (divide by the value at 180).
-(We could also put the extremes at not quite 0, so that at -90 we're at -1 degrees changes and at 180, we add 1 degree to the last measured point / bin, this should leave some wiggle room for a better fit.) At that point a cumulative function can be fit, and we take the 50% point as the centre of the normal function.
-I'm not totally sure if this works, but it's worth a try.
+These results mimic those of the reach aftereffects: the median values are different, but they each fall in the others confidence interval.
+However, here the LME showed that there was an interaction between training type and hand angle, which suggests that the generalization curves are different.
+If we want to look at this better, we need to get better estimates of these curves in every participant, by having a wider range of points and more data per point, but probably also larger rotations.
 
@@
 
-_I'm not sure if we should use the initial reach aftereffects here, or the four later blocks. Unfortunately, it does matter for the results, but I don't want to choose with hindsight._
-
-This plot shows regression lines with localization shifts as predictor of reach aftereffects:
+This plot shows regression lines with localization shifts as predictor of reach aftereffects, potentially Figure 4:
 
@@
 
-As one possibility, we can look at the correlations.
-Not just the p-values, but there are confidence intervals for Pearson's $\rho$, which allow to directly see if one correlation is stronger/weaker than another:
+As one possibility, we can look at the correlations:
 
@@
 ```
+
+Only one is significant: active localization predicting reach aftereffects in the exposure group.
 
@@
 This means that evoking propceptive recalibration with visual-proprioceptive discrepancies leads to motor changes.
-In other words: proprioception informs movement plans.
+In other words: proprioception informs motor control.
 

### 2018-06-23 shared.R

@@ -4,7 +4,7 @@
 nocursorURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'='https://osf.io/8hm7f/?action=download')
 
 # no-cursor data can be downloaded from OSF:
-localizationURLs <- c('exposure'='https://osf.io/9f6gu/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
+localizationURLs <- c('exposure'='https://osf.io/bk4s6/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
 
 # we'll control the color in figures centrally from here:
 colorset <- list()
@@ -231,7 +231,7 @@ getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55
     }
     # print(dim(groupdf))
   } else {
-    if (LRpart != 'both') {
+    if (LRpart != 'all') {
       cat('\nWARNING: Partial localization can not be returned for classic/online data.\n\n')
       # the full dataset is returned anyway?
     }

### 2018-07-17 ManuscriptMaterials.Rmd
add counts of interpolated localization data for every angle, to see if we want to remove the 15 degree location

--- 2018-06-23 ManuscriptMaterials.Rmd
@@
 This will need to be tested.
+
+## Missing data
+
+Here are the counts of participants with valid localization estimates in all 4 tasks for every hand angle. Missing data occurs here, because we don't extrapolate beyond the range of the data.
+
+```{r localization_count_table}
+kable(getLocCountTable())
+```
+
+Since the majority of missing data appears to be at the 15 degree location, we exclude that angle from analysis, with a switch in analysis functions `remove15=TRUE`, which can be set to false if we'd want.
 

### 2018-07-17 localization.R

@@ -263,7 +263,7 @@ exposureLocalization <- function(remove15=TRUE, LMEmethod='chi-squared') {
   }
   
   exp$participant   <- factor(exp$participant)
-  exp$rotated_b    <- factor(exp$rotated_b)
+  exp$rotated_b     <- factor(exp$rotated_b)
   exp$passive_b     <- factor(exp$passive_b)
   exp$handangle_deg <- factor(exp$handangle_deg)
   
@@ -510,5 +510,66 @@ getPeakLocConfidenceInterval <- function(group,part='initial',CIs=c(.95), moveme
   loc2 <- -1 * xtabs(taperror_deg ~ participant + handangle_deg,loc)
   
   bootstrapGaussianPeak(data=loc2,bootstraps=1000,mu=47.5,sigma=30,scale=10,offset=4,CIs=CIs)
+  
+}
+
+countLocNAs <- function(group='exposure', output='count') {
+  
+  loc <- getPointLocalization(group, difference=FALSE, verbose=FALSE)
+  loc <- loc[is.finite(loc$taperror_deg),]
+  
+  df <- expand.grid(unique(loc$participant), unique(loc$handangle_deg))
+  names(df) <- c('participant', 'handangle_deg')
+  df$count <- 0
+  
+  for (rown in c(1:nrow(df))) {
+    
+    pp    <- df[rown, 'participant']
+    angle <- df[rown, 'handangle_deg']
+    
+    subexp <- loc[which(loc$participant == pp & loc$handangle_deg == angle),]
+    df$count[rown] <- nrow(subexp)
+    
+  }
+  
+  if (output == 'count') {
+    # this is a count of participants with estimates in ALL 4 tasks
+    df$count[which(df$count < 4)] <- 0
+    df$count[which(df$count > 0)] <- 1
+    df <- aggregate(count ~ handangle_deg, data=df, FUN=sum)
+  }
+  
+  if (output == 'percentage') {
+    # this returns a percentage of existing estimates across the 4 tasks
+    df$count <- df$count/4
+    df <- aggregate(count ~ handangle_deg, data=df, FUN=mean)
+  }
+  
+  return(df)
+  
+}
+
+getLocCountTable <- function(output='count') {
+  
+  groups <- c('exposure','classic','online')
+  
+  for (group in groups) {
+    
+    counts <- countLocNAs(group=group, output=output)
+    
+    if (group == groups[1]) {
+      
+      df <- counts
[…]

### 2018-07-25 ManuscriptMaterials.Rmd
updates to reflect extra participants and low performance on blink detection

--- 2018-06-26 ManuscriptMaterials.Rmd
@@
     + D: Localization (top view)
-2. Reach Aftereffects (goal: show motor changes are there, are robust and compare to classic)
-    + A: Exposure reach aftereffects (initial & later)
-    + B: Classic & Exposure reach aftereffects (initial only? or all?)
-3. Localization
+2. Localization
     + A: Exposure, active and passive (goal: small/no difference between active and passive)
@@
     + C: Online localization?
+3. Reach Aftereffects (goal: show motor changes are there, are robust and compare to classic)
+    + A: Exposure reach aftereffects (initial & later)
+    + B: Classic & Exposure reach aftereffects (initial only? or all?)
 
@@
 source('shared.R') # functions used everywhere
+source('localization.R') # functions for localization data
 source('nocursor.R') # functions for no-cursor reach data
-source('localization.R') # functions for localization data
 source('relateLocalizationNoCursors.R') # functions that correlate the two kinds of data
@@
 
-In the manuscript we'll first show that there indeed are motor changes after exposure training (with the no-cursor data) and then compare that with the classic.
-Second is the localization: a) localization shifts after exposure b) the shifts are (not) different from those after classic c) the effects are (not) different across the workspace d) this pattern is (not) different from that after classic.
+In the manuscript we'll first show that there indeed are localization shifts after exposure training (with the no-cursor data) and then compare that with the classic.
+Second are the no-cursor reaches (and the reach aftereffects).
 A third topic might be if localization shifts can predict no-cursor changes, which we'll call the "correlation" topic.
@@
 
-1. Reach aftereffects
-2. Localization
-3. Correlations
+1. Analyses (some general remarks about analyses)
+2. Participants (participant demographics)
+3. Localization
+4. Reach aftereffects
+5. Correlations
 
@@
 However, a paper from Steven G.
-Luke (2017, _Evaluating significance in linear mixed-effects models in R_) suggests using Sattertwaite approximation.
+Luke (2017, _Evaluating significance in linear mixed-effects models in R_) suggests using Satterthwaite approximation.
 So all functions running an LME now take an argument that tells them to use the Satterthwaite method, but the default is chi-squared.
 It turns out that it doesn't matter much, so it's a good sanity check.
-For consistency across all the analyses, we set the value of that argument is set here:
+For consistency across all the analyses, we set the value of that argument here:
 
@@
 LMEmethod <- 'Satterthwaite'
-# options: Satterthwaite, chi-squared
+# options: 'Satterthwaite', 'chi-squared'
 ```
@@
 
-# Reach Aftereffects
+# Participants
 
-Here is a plot, probably Figure 2:
+There are no scripts for this part, as the code is simple and can just be embedded.
 
-```{r fig2_reachaftereffect, fig.width=7.5, fig.height=2.5}
-plotReachAftereffects()
+## Performance
+
+Let's look at the blink detect performance.
+
+```{r}
+plotBlinkDetection()
 ```
 
-Messages of the figure:
+Turns out the blink detection is hard and needs practice. But three participants are still performing around chance in the rotated training task, so we want to remove them from the data:
 
-- Panel A: there are substantial and persisting reach aftereffects
-- Panel B: that are somewhat lower than those in classic (but not that much? but only for some targets?)
-
-These claims require analyses.
-
-## Is there an effect of exposure training on reach aftereffects?
-
-First, we show that no-cursor reaches change direction immediately after exposure training (iteration 1):
-
-```{r RAE_exposure,include=TRUE}
-exposureNoCursorChange(LMEmethod=LMEmethod)
+```{r}
+blinks <- load.DownloadDataframe(informationURLs['blinkdetect'],'blinkdetect_exposure.csv')
+OKparticipants <- blinks$participant[which(blinks$rotated_b == 1 & blinks$performance > 0.65)]
 ```
 
-So this shows that whether or not the feedback was rotated mattered for the reach deviations in the exposure training group, i.e.: there are reach aftereffects.
+We'll also set a variable that is to be passed to all functions to exclude the underperformance:
 
-## Do the motor changes persist during localization?
-
-The second analysis is to show that the reach aftereffects aren't all that different for first and later iterations of the rotated no-cursor tasks.
-(The rotated data is always compared with _all_ aligned iterations of the task.)
-
-```{r RAE_exp_iterations}
-exposureAftereffectsPersistent(LMEmethod=LMEmethod)
+```{r}
+selectPerformance <- TRUE
 ```
 
-So there is no effect of iteration here, nor does it interact with target, which means the reach aftereffects are the same for the first iteration and the average across the other four iterations.
-We now also have enough power to see an effect of target angle: the generalization curve is not flat.
+Which can also be set to FALSE to include those three participants in all analyses and plots.
 
-## Are reach aftereffects comparable between classic and exposure training?
+## Demographics
 
-Finally, we want to see if there is a difference between the reach aftereffects observed after exposure training and those after classic training.
+First load the data, and select only the well performing ones:
 
-```{r RAE_exp_cla}
-exposureClassicReachAftereffects(LMEmethod=LMEmethod)
+```{r eval=F}
+participants <- load.DownloadDataframe(informationURLs['demographics'],'participants_exposure.csv')
+participants <- participants[which(participants$participant %in% OKparticipants),]
 ```
 
-I'd have expected there to be some difference between the two training types, perhaps only in interaction with target, but that's not the case apparently.
+We want to know some demographics, first mean age, and standard deviation:
 
-The absence of an interaction between training and target could be interpreted to mean that the generalization of reach aftereffects is the same after both kinds of training.
-Here are the confidence intervals for the peaks of both:
-
-```{r RAE_peaks}
-getPeakConfidenceInterval('classic', part='all', CIs=c(.95))
-getPeakConfidenceInterval('exposure', part='all', CIs=c(.95))
+```{r eval=F}
+c('mean'=mean(participants$age), 'sd'=sd(participants$age))
 ```
 
-The 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training) but not the 75 degree target, whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training) but not the 45 degree target.
-Something might be going on -- but the two confidence intervals are not different from each other, which is in line with the LME results.
-We can't conclude much, except that we have no evidence for different generalization curves.
+And the distribution of sex:
 
-To sum up:
-
-1. Rotated exposure training evokes substantial shifts in no-cursor reach direction across the workspace,
-2. these changes persist over time, and
-3. these changes are comparable to the ones we found after classic training, and
-4. there is some generalization curve across the workspace, but
-5. we have no evidence that this is different from reach aftereffect generalization after classic training.
+```{r eval=F}
+table(participants$sex)
+```
 
@@
 
-Here is a plot, probably figure 3:
+Here is a plot, probably figure 2:
 
 ```{r fig3_localization, fig.width=7.5, fig.height=2.5}
-plotLocalization(classicOnline=TRUE)
+plotLocalization(classicOnline=TRUE,selectPerformance=selectPerformance)
 ```
@@
 
+## Missing data
+
+Here are the counts of participants with valid localization estimates in all 4 tasks for every hand angle. Missing data occurs here, because we don't extrapolate beyond the range of the data.
+
+```{r localization_count_table}
+kable(getLocCountTable(selectPerformance=selectPerformance))
+```
+
+Since there is a good amount of missing data at the 15 degree location, we exclude that angle from analysis, with a switch in analysis functions `remove15=TRUE`, which can be set to false if we'd want.
+
 ## Is there a localization shift after exposure training?
 
-We first want to see if there is any overall effect of aligned versus rotated training in the exposure group (i.e.: are there any localization shifts at all?), with movement type and hand angle as fixed effects, and particiant as random effect.
+We first want to see if there is any overall effect of aligned versus rotated training in the exposure group (i.e.: are there any localization shifts at all?), with movement type and hand angle added as fixed effects as well, and particiant as random effect.
 
 ```{r loc_exp}
-exposureLocalization(remove15=TRUE, LMEmethod=LMEmethod)
+exposureLocalization(remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
 ```
@@
 ```{r loc_exp_movtype}
-exposureLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod)
+exposureLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
 ```
@@
 ```{r loc_ombnibus}
-groupLocalization(model='full', remove15=TRUE, LMEmethod=LMEmethod)
+groupLocalization(model='full', remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
 ```
@@
 This could be because the effect of movement type is much larger in the classic group.
-This would predict an interaction between group and movement type, but that's not the case (p=.133).
+This would predict an interaction between group and movement type, and we do see that (p=.035).
+(Including the three)
 
-What if we don't care about hand angle:
+Perhaps hand angle factored into this somehow, so let's see what happens if we remove it:
 
 ```{r loc_groups_movementtype}
-groupLocalization(model='movementtype', remove15=TRUE, LMEmethod=LMEmethod)
+groupLocalization(model='movementtype', remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
 ```
 
-Still no effect...
+Still an effect...
 
@@
 
-So there is a fairly strong effect of movement type in the classic group.
-Together with the absence of an effect of movement type in the exposure group, but no interaction between group and movement type if we combine both datasets.
+So there is a reliable effect of movement type in the classic group, which is in line with the interaction between group and movement type.
 
-Also note the absence of an effect of hand angle in the classic group here -- probably because we excluded the 15 degree target, but it seems that after classic training the generalization curve for delayed localization is much flatter.
+Also note the absence of an effect of hand angle in the classic group here -- probably because we excluded the 15 degree target, but it does appear that after classic training the generalization curve for delayed localization is much flatter.
 
@@
 3. In the classic group, the localization responses informed by efference copies, are shifted more than those with only proprioception (as found before).
-4. This predicts an interaction between movement type and group in the combined analysis, but we don't see that.
+4. This predicts an interaction between movement type and group in the combined analysis, and we see that.
 
@@
 
+There was an interaction between group and hand angle on localization shifts above, which suggests different generalization curves in the two groups.
+This could reflect a difference in the shape (more flat for the classic training group as compared to exposure), but also a difference in the locations of the peaks of the generalization curves.
+
 Here are the confidence intervals for the peaks of the generalization curves for _active_ localization:
@@
 ```{r localization_peaks}
-getPeakLocConfidenceInterval(group='classic',CIs=c(.95), movementtype='active', LRpart='all')
-getPeakLocConfidenceInterval(group='exposure',CIs=c(.95), movementtype='active', LRpart='all')
+getPeakLocConfidenceInterval(group='classic',
+                             CIs=c(.95), 
+                             movementtype='active', 
+                             LRpart='all',
+                             selectPerformance=FALSE)
+getPeakLocConfidenceInterval(group='exposure',
+                             CIs=c(.95), 
+                             movementtype='active', 
+                             LRpart='all', 
+                             selectPerformance=selectPerformance)
 ```
 
-These results mimic those of the reach aftereffects: the median values are different, but they each fall in the others confidence interval.
-However, here the LME showed that there was an interaction between training type and hand angle, which suggests that the generalization curves are different.
-If we want to look at this better, we need to get better estimates of these curves in every participant, by having a wider range of points and more data per point, but probably also larger rotations.
+The median values are a little over 13 degrees apart, and they _just_ fall outside each others confidence interval.
+This can explain the interaction between group and hand angle on localization shifts.
+
+# Reach Aftereffects
+
+Here is a plot, probably Figure 3:
+
+```{r fig2_reachaftereffect, fig.width=7.5, fig.height=2.5}
+plotReachAftereffects(selectPerformance=FALSE)
+```
+
+Messages of the figure:
+
+- Panel A: there are substantial and persisting reach aftereffects
+- Panel B: that are somewhat lower than those in classic (but not that much? but only for some targets?)
+
+These claims require analyses.
+
+## Is there an effect of exposure training on reach aftereffects?
+
+First, we look at the no-cursor reaches after exposure training only (see Fig 3a). We show that no-cursor reaches change direction immediately after exposure training (iteration 1):
+
+```{r RAE_exposure,include=TRUE}
+exposureNoCursorChange(LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+```
+
+So this shows that whether or not the feedback was rotated mattered for the reach deviations in the exposure training group, i.e.: there are reach aftereffects.
+
+## Do the motor changes persist during localization?
+
+The second analysis is to show that the reach aftereffects aren't all that different for first and later iterations of the rotated no-cursor tasks.
+(The rotated data is always compared with _all_ aligned iterations of the task.)
+
+```{r RAE_exp_iterations}
+exposureAftereffectsPersistent(LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+```
+
+So there is no effect of iteration here, nor does it interact with target, which means the reach aftereffects are the same for the first iteration and the average across the other four iterations.
+We now also see an effect of target angle: the generalization curve is not flat.
+This seems mostly due to the extra training.
+
+## Are reach aftereffects comparable between classic and exposure training?
+
+Finally, we want to see if there is a difference between the reach aftereffects observed after exposure training and those after classic training.
+Here we combine the data across all iterations (see Fig 3b).
+
+```{r RAE_exp_cla}
+exposureClassicReachAftereffects(LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+```
+
+The figure looked like there might be some difference between the two training types, mainly in where the generalization curves peak.
+There is no main effect of training type, so the overall magnitude of reach aftereffects is not different.
+But there is an interaction between training and target (p=.038), so that perhaps the curves peak at different locations.
+
+So we now investigate the peaks the confidence intervals for the peaks of both:
+
+```{r RAE_peaks}
+getPeakConfidenceInterval('classic', part='all', CIs=c(.95), selectPerformance=selectPerformance)
+getPeakConfidenceInterval('exposure', part='all', CIs=c(.95), selectPerformance=selectPerformance)
+```
+
+The 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training) but not the 75 degree target, whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training) but not the 45 degree target.
+Something might be going on -- but the two confidence intervals are not different from each other, which is not in line with the LME results.
+We can't conclude much, except that we don't have strong evidence for different generalization curves.
+
+To sum up:
+
+1. Rotated exposure training evokes substantial shifts in no-cursor reach direction across the workspace,
+2. these changes persist over time, and
+3. these changes are comparable to the ones we found after classic training, and
+4. there is some generalization curve across the workspace, but
+5. we have no evidence that this is different from reach aftereffect generalization after classic training.
 
@@
 ```{r scattergrams, fig.width=7.5, fig.height=2.5}
-correlations <- correlateNoCursorsLocalization()
+correlations <- correlateNoCursorsLocalization(selectPerformance=selectPerformance)
 ```
 
-It looks like for exposure, active localization is the better predictor of reach aftereffects, whereas in classic it is passive localization.
-But are any of them real?
+It looks like for exposure, both types of localization might be able to predict of reach aftereffects, whereas in classic it is only passive localization.
+But are any of these reliable?
 The function also returns the correlation tests and linear regressions for all four datasets.
@@
 
-Only one is significant: active localization predicting reach aftereffects in the exposure group.
+None are significant... some are "trending".
 
@@
 Here I do step-wise multiple regression, allowing active and passive localization shift as predictor of reach aftereffects.
-If I do this on all the no-cursor block data, the models converge on using the active localization for the exposure group, and passive localization for the classic group.
 
-If we look at only the initial no-cursor block, the story changes.
-It uses both active and passive to predict reach aftereffects for the exposure group, but converges on using only an intercept term for the classic group.
+```{r step_exposure}
+multipleRegressionLocalization(group='exposure', 
+                               NCpart='all', 
+                               selectPerformance=selectPerformance)
+```
+
+The algorithm decides to retain active localization as predictor, but the regression model isn't very good.
+The problem could be that we need no-cursor data and localization data that are both recorded immediately after training.
+This also makes it more comparable to the classic group that had fewer blocks of data.
+
+But if we look at only the initial no-cursor block, the model doesn't change much.
+It still uses only active localization and the p-value goes down a bit and the R-squared goes up a bit.
+For the classic group...
+
+converges on using only an intercept term for the classic group.
 This might mean that in te classic group there is no relationship between reach aftereffects and localization, or this might mean that in the exposure group we have sufficient amount of measurements to detect the relationship, as we have twice the number of localization trials.
@@
 
-```{r step_exposure}
-multipleRegressionLocalization(group='exposure',NCpart='initial')
+```{r step_exposure_initial}
+multipleRegressionLocalization(group='exposure',
+                               NCpart='initial',
+                               LRpart='both', 
+                               selectPerformance=selectPerformance)
 ```
-
-No terms are dropped an the adjusted R squared of the full model is low, but reasonable. _Note that the coefficient for the passive localization is positive._
 
@@
 
-```{r step_classic}
-multipleRegressionLocalization(group='classic',NCpart='initial')
+```{r step_classic_initial}
+multipleRegressionLocalization(group='classic',
+                               NCpart='initial',
+                               LRpart='both')
 ```
@@
 ```{r step_exposure_firsthalf}
-multipleRegressionLocalization(group='exposure',NCpart='initial', LRpart='first')
+multipleRegressionLocalization(group='exposure',
+                               NCpart='initial', 
+                               LRpart='first', 
+                               selectPerformance=selectPerformance)
 ```
 
-Now the passive localization shifts are thrown out by the algorithm and the active localization shifts don't contribute much either.
-This means that perhaps previous papers didn't find any relationships between recalibrated proprioception and reach aftereffects because not enough data was collected to overcome the noisiness of the measurements?
+Now we have an intercept model for the exposure group too.
+Perhaps this means that previous papers didn't find any relationships between recalibrated proprioception and reach aftereffects because not enough data was collected to overcome the noisiness of the measurements?
+
+Maybe including more data for the model for classic might work better:
+
+```{r step_classic_all}
+multipleRegressionLocalization(group='classic',
+                               NCpart='all',
+                               LRpart='both')
+```
+
+And now we get the passive localization as in the initial regressions.
 
@@
 ```{r step_online}
-multipleRegressionLocalization(group='online',NCpart='initial')
+multipleRegressionLocalization(group='online',
+                               NCpart='all')
 ```
 
-But no.
+And that might be the case as one term is preserved by the model, but on the other hand, the predictions are pretty bad.
 
@@
 correlations <- correlateNoCursorsLocalization(generateSVG=TRUE)
+#plotBlinkDetection(generateSVG=TRUE)
 ```

### 2018-07-25 localization.R

@@ -124,11 +124,11 @@ localizationLMEsSatterthwaite <- function() {
 
 # PLOTS / FIGURES ------
 
-plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
+plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerformance=TRUE) {
   
   # get the data to plot:
-  exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE)
-  cla <- getPointLocalization('classic', difference=TRUE, verbose=FALSE)
+  exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE, selectPerformance=selectPerformance)
+  cla <- getPointLocalization('classic', difference=TRUE, verbose=FALSE, selectPerformance=FALSE)
   
   # get the averages for the line plots:
   exp.avg <- aggregate(taperror_deg ~ passive_b + handangle_deg, data=exp, FUN=mean)
@@ -149,7 +149,7 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
   cla.CI.pas <- matrix(unlist(by(cla.pas$taperror_deg, INDICES=c(cla.pas$handangle_deg), FUN=t.interval)),nrow=2)
 
   if (classicOnline) {
-    onl <- getPointLocalization('online', difference=TRUE, verbose=FALSE)
+    onl <- getPointLocalization('online', difference=TRUE, verbose=FALSE, selectPerformance=FALSE)
     
     onl.avg <- aggregate(taperror_deg ~ passive_b + handangle_deg, data=onl, FUN=mean)
     onl.avg.act <- onl.avg[which(onl.avg$passive_b == 0),]
@@ -164,7 +164,7 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
     if ('svglite' %in% installed.list) {
-      svglite(file='Fig3.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
+      svglite(file='Fig2.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
     } else {
       generateSVG=FALSE
     }
@@ -251,12 +251,12 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE) {
   
 }
 
-exposureLocalization <- function(remove15=TRUE, LMEmethod='chi-squared') {
[…]

### 2018-07-25 nocursor.R

@@ -66,12 +66,12 @@ installRequire.Packages(required.packages)
 #   
 # }
 
-plotReachAftereffects <- function(generateSVG=FALSE) {
+plotReachAftereffects <- function(generateSVG=FALSE, selectPerformance=TRUE) {
   
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
     if ('svglite' %in% installed.list) {
-      svglite(file='Fig2.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
+      svglite(file='Fig3.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
     } else {
       generateSVG=FALSE
     }
@@ -81,12 +81,12 @@ plotReachAftereffects <- function(generateSVG=FALSE) {
   
   points <- c(15,25,35,45,55,65,75)
   
-  exposure_ini <- getReachAftereffects('exposure',part='initial')
-  exposure_rem <- getReachAftereffects('exposure',part='remainder')
-  classic  <- getReachAftereffects('classic',part='all')
-  exposure <- getReachAftereffects('exposure',part='all')
+  exposure_ini <- getReachAftereffects('exposure',part='initial', selectPerformance=selectPerformance)
+  exposure_rem <- getReachAftereffects('exposure',part='remainder', selectPerformance=selectPerformance)
+  classic  <- getReachAftereffects('classic',part='all', selectPerformance=selectPerformance)
+  exposure <- getReachAftereffects('exposure',part='all', selectPerformance=selectPerformance)
   
-  exposureAVGini <- aggregate(endpoint_angle ~ target, data=exposure_ini, FUN=mean)
+  exposureAVGini <- aggregate(endpoint_angle ~ target, data=exposure_ini, FUN=mean) # error?
   exposureAVGrem <- aggregate(endpoint_angle ~ target, data=exposure_rem, FUN=mean)
   classicAVG <- aggregate(endpoint_angle ~ target, data=classic, FUN=mean)
   exposureAVG <- aggregate(endpoint_angle ~ target, data=exposure, FUN=mean)
@@ -138,11 +138,11 @@ plotReachAftereffects <- function(generateSVG=FALSE) {
   
 }
 
-getPeakConfidenceInterval <- function(group,validate=FALSE,part='initial',CIs=c(.95)) {
[…]

### 2018-07-25 relateLocalizationNoCursors.R

@@ -3,13 +3,13 @@ source('shared.R')
 
 
 
-correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE) {
+correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE, selectPerformance=TRUE) {
   
   # first we do all responses, plan to split it by target later
   # get all data, for classic and exposure, we need localization and reach aftereffects
   
   # LOCALIZATION SHIFT for exposure
-  local.exp <- getPointLocalization(group='exposure', difference=FALSE, verbose=FALSE)
+  local.exp <- getPointLocalization(group='exposure', difference=FALSE, verbose=FALSE, selectPerformance=selectPerformance)
   # active
   loc.exp.act.ro <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.exp[local.exp$rotated_b == 1 & local.exp$passive_b == 0,], FUN=mean, drop=FALSE)
   loc.exp.act.al <- aggregate(taperror_deg ~ participant + handangle_deg, data=local.exp[local.exp$rotated_b == 0 & local.exp$passive_b == 0,], FUN=mean, drop=FALSE)
@@ -22,7 +22,7 @@ correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE) {
   local.exp.pas$taperror_deg <- local.exp.pas$taperror_deg - loc.exp.pas.al$taperror_deg
   
   # REACH AFTEREFFECTS for exposure
-  nocur.exp <- getReachAftereffects(group='exposure', part=NCpart, clean=TRUE) 
+  nocur.exp <- getReachAftereffects(group='exposure', part=NCpart, clean=TRUE, selectPerformance=selectPerformance) 
   names(nocur.exp)[names(nocur.exp) == 'endpoint_angle'] <- 'RAE'
   # combine data frames for exposure
   exposure <- nocur.exp[which(nocur.exp$participant %in% local.exp.act$participant),]
@@ -30,7 +30,7 @@ correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE) {
   exposure$passive_localization_shift <- local.exp.pas$taperror_deg
   
   # LOCALIZATION SHIFT for classic
-  local.cla <- getPointLocalization(group='classic', difference=FALSE, verbose=FALSE)
[…]

### 2018-07-25 shared.R

@@ -6,6 +6,9 @@ nocursorURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'=
 # no-cursor data can be downloaded from OSF:
 localizationURLs <- c('exposure'='https://osf.io/bk4s6/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
 
+# information on exposure participants
+informationURLs <- c('blinkdetect'='https://osf.io/yrpkm/?action=download', 'demographics'='https://osf.io/7sn4b/?action=download')
+
 # we'll control the color in figures centrally from here:
 colorset <- list()
 
@@ -18,8 +21,8 @@ colorset[['claActT']] <- '#e516362f'
 colorset[['claPasS']] <- '#ff8200ff' # orange
 colorset[['claPasT']] <- '#ff82002f'
 
-colorset[['extra1S']] <- '#c400c4ff' # purple
-colorset[['extra1T']] <- '#c400c42f'
+# colorset[['extra1S']] <- '#c400c4ff' # purple
+# colorset[['extra1T']] <- '#c400c42f'
 
 colorset[['onlActS']] <- '#b400e4ff' # purple
 colorset[['onlActT']] <- '#b400e42f'
@@ -213,10 +216,17 @@ parGaussian <- function(par,x) {
 
 
 # handling localization data -----
+#   exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE, selectPerformance=selectPerformance)
 
-getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55,65,75), movementtype='both', verbose=TRUE, LRpart='all') {
+getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55,65,75), movementtype='both', verbose=TRUE, LRpart='all', selectPerformance=selectPerformance) {
   
   df <- load.DownloadDataframe(url=localizationURLs[group],filename=sprintf('localization_%s.csv',group))
+  
+  if (selectPerformance & group=='exposure') {
+    blinks <- load.DownloadDataframe(informationURLs['blinkdetect'],'blinkdetect_exposure.csv')
+    OKparticipants <- blinks$participant[which(blinks$rotated_b == 1 & blinks$performance > 0.65)]
+    df <- df[which(df$participant %in% OKparticipants),]
+  }
   
   df <- aspligned(df)
   
[…]

### 2018-07-25 shared.R
bug corrected

@@ -218,7 +218,7 @@ parGaussian <- function(par,x) {
 # handling localization data -----
 #   exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE, selectPerformance=selectPerformance)
 
-getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55,65,75), movementtype='both', verbose=TRUE, LRpart='all', selectPerformance=selectPerformance) {
+getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55,65,75), movementtype='both', verbose=TRUE, LRpart='all', selectPerformance=TRUE) {
   
   df <- load.DownloadDataframe(url=localizationURLs[group],filename=sprintf('localization_%s.csv',group))
   

### 2018-07-25 ManuscriptMaterials.Rmd
update URLs of data on OSF

@@
 
-```{r eval=F}
+```{r}
 participants <- load.DownloadDataframe(informationURLs['demographics'],'participants_exposure.csv')
+if (selectPerformance) {
   participants <- participants[which(participants$participant %in% OKparticipants),]
+}
 ```
@@
 
-```{r eval=F}
+```{r}
 c('mean'=mean(participants$age), 'sd'=sd(participants$age))
@@
 
-```{r eval=F}
+```{r}
 table(participants$sex)

### 2018-07-25 shared.R

@@ -4,7 +4,7 @@
 nocursorURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'='https://osf.io/8hm7f/?action=download')
 
 # no-cursor data can be downloaded from OSF:
-localizationURLs <- c('exposure'='https://osf.io/bk4s6/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
+localizationURLs <- c('exposure'='https://osf.io/ysduw/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
 
 # information on exposure participants
 informationURLs <- c('blinkdetect'='https://osf.io/yrpkm/?action=download', 'demographics'='https://osf.io/7sn4b/?action=download')

### 2018-07-25 ManuscriptMaterials.Rmd
bug fix
re-ordered the notebook a bit

@@
 ---
-title: "Exposure Consequences - Statistics and Figures"
+title: "Motor Learning Without Moving"
+subtitle: "Figures and Statistics"
+author:
+- Ahmed A. Mostafa
+- Bernard Marius 't Hart
+- Denise Henriques
 output:
@@
 ```{r}
-c('mean'=mean(participants$age), 'sd'=sd(participants$age))
+print(c('mean'=mean(participants$age), 'sd'=sd(participants$age)))
 ```
@@
 ```{r}
-table(participants$sex)
+print(table(participants$sex))
 ```
@@
 ```{r fig3_localization, fig.width=7.5, fig.height=2.5}
-plotLocalization(classicOnline=TRUE,selectPerformance=selectPerformance)
+plotLocalization(classicOnline=TRUE, selectPerformance=selectPerformance)
 ```
@@
 
+What do these correlations mean?
+I'm not sure.
+There seems to be some relationship but it's not very strong.
+So either there are other factors playing a role, the measurements are too noisy, or we don't have enough power to detect any relationships.
+All of these are very likely, and then there is the other alternative that there is no relationship.
+So it's probably better not to include these (unplanned) analyses in the manuscript and leave this for a better dataset.
+
 # Conclusions
@@
 1.
-Exposure training leads to robust reach aftereffects that are similar to those after classic training, although they seem to have a more pronouced generalization curve.
-This means that evoking propceptive recalibration with visual-proprioceptive discrepancies leads to motor changes.
-In other words: proprioception informs motor control.
-
-2.
 Exposure training leads to shifts in localization, that are not very different between active and passive localization.
@@
     3.
-Within the workspace however, there is no interaction between movement type and training group (but there is a 'trend', p=.09, if hand angle is used as a random effect, in all analyses the 15 degree data is omitted as Denise said).
-Perhaps there is a small update of predicted proprioceptive consequences, that we can't parcel out.
-A more computational approach would be needed to see if this can explain the pattern of results here.
+Within the workspace, there is an interaction between movement type and training group, which is predicted by there being an effect of updated predictions on active localization after classic training, but not after exposure training.
 
-3.
-Recalibrated proprioception can somewhat predict reach aftereffects, but only after **exposure** training and for _active localization_.
-This indicates that in some circumstances, recalibrated proprioception might drive motor changes.
-However, it is possible that the active component of the localization is crucial here.
-Perhaps exposure training leads to some updates in predicted sensory consequences after all -- be they visual or proprioceptive.
-These would also factor into the reach aftereffects.
+    4. The generalization curves for localization are different between the two groups.
+
+2.
+Exposure training leads to robust reach aftereffects that are similar to those after classic training, although they seem to have a different generalization curve.
+This means that evoking propceptive recalibration with visual-proprioceptive discrepancies leads to motor changes.
+In other words: recalibrated proprioception informs motor control.
 
     1.
-There is also a trend towards significance for _passive localization_ to predict reach aftereffects after **classic training**.
-Perhaps active localization is more similar to no-cursor reaches after exposure training, as they both use recalibrated proprioception as well as updated predictions of sensory consequences -- which could mostly be proprioceptive consequences after exposure training.
-This does not explain the trend towards significance for predicting classic reach aftereffects with passive localization.
-A step-wise multiple regression favours an intercept model for classic reach aftereffects, hence none of the localization data was deemed informative.
-The same algorithm arrived at a model using active localization to predict exposure reach aftereffects.
-The intercept model for classic reach aftereffects could signify that not too much should be read into the trend towards significance in the correlation.
-On the other hand, the exposure group had twice as much data, so that the estimated recalibration would be more reliable within each participant.
-Perhaps with higher quality estimates of proprioceptive recalibration those would be able to predict part of the reach aftereffects even after classic training.
+The generalization curves of reach aftereffects look somewhat different for both groups, which can be explained by the effect of updated predictions that factors into the curve for the classic training group, but not the exposure training group.
+However, it's not a large difference, and the analyses are ambiguous.
+Perhaps this is another cue that recalibrated proprioception dominates even reach aftereffects?
+Either way, we find no (statistically reliable) evidence that the updated predictions matter much for the generalization curves, which can have several causes.
 
@@
 # This chunk creates SVG files but shouldn't show up in the PDF / HTML 
-plotReachAftereffects(generateSVG=TRUE)
-plotLocalization(classicOnline=TRUE, generateSVG=TRUE)
-correlations <- correlateNoCursorsLocalization(generateSVG=TRUE)
+plotReachAftereffects(generateSVG=TRUE,
+                      selectPerformance=selectPerformance)
+plotLocalization(classicOnline=TRUE,
+                 generateSVG=TRUE, 
+                 selectPerformance=selectPerformance)
+correlations <- correlateNoCursorsLocalization(generateSVG=TRUE, 
+                                               selectPerformance=selectPerformance)
 #plotBlinkDetection(generateSVG=TRUE)

### 2018-07-30 localization.R
clean up code for loading packages, labeled some chunks

@@ -2,122 +2,118 @@
 source('shared.R')
 
 # first make sure we have all necessary packages installed and loaded:
-# for now we'll use Chi-square (not Satterthwaite estimates of F/p-values):
-installRequire.Packages(c('nlme', 'car'))
+# installRequire.Packages(c('nlme', 'car'))
 
 # , 'lme4', 'lmerTest'
-
-# localization data can be downloaded from OSF:
-# groupURLs <- c('exposure'='https://osf.io/9qfhp/download', 'classic'='https://osf.io/upw49/download', 'online'='https://osf.io/wjcgk/download')
-
-# no-cursor reach data is not uploaded yet (not sure about the format)
-
-localizationLMEsChiSq <- function() {
-  
-  default.contrasts <- options('contrasts')
-  options(contrasts=c('contr.sum','contr.poly'))
-
-  exposure <- getANOVAlocalization('exposure')
-  classic <- getANOVAlocalization('classic')
-  
-  exposure$participant <- exposure$participant + 100
-
-  exposure <- exposure[-which(exposure$handangle_deg == 15),]
-  classic <- classic[-which(classic$handangle_deg == 15),]
-  
-  # they all have this header:
-  # group,online_b,participant,rotated_b,passive_b,handangle_deg,tapangle_deg
-  # (where group is uninformative)
-
-  localization <- rbind(exposure,classic)
-  
-  localization$group         <- factor(localization$group)
-  localization$participant   <- factor(localization$participant)
-  localization$online_b      <- factor(localization$online_b)
-  localization$rotated_b     <- factor(localization$rotated_b)
-  localization$passive_b     <- factor(localization$passive_b)
-  localization$handangle_deg <- factor(localization$handangle_deg)
-  
-  # str(localization)
-  
-  # prepare a data frame with the training-induced changes in hand localization:
-  locshift <- localization[which(localization$rotated_b == 1),]
-  locshift$taperror_deg <- locshift$taperror_deg - localization$taperror_deg[which(localization$rotated_b == 0)]
-  
-  # first a model including the aligned and rotated data
-  # here we look for any effect of session (rotated_b)
-  # and if it is there, we continue with other analyses that are easier to grasp:
-  exp_full_model <- lme(taperror_deg ~ rotated_b * passive_b * group, data=localization, random = ~1|participant/handangle_deg, na.action=na.exclude)
[…]

### 2018-07-30 relateLocalizationNoCursors.R

@@ -1,8 +1,6 @@
 
 source('shared.R')
 
-
-
 correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE, selectPerformance=TRUE) {
   
   # first we do all responses, plan to split it by target later

### 2018-07-30 shared.R

@@ -1,5 +1,10 @@
 # some functions shared by the analyses of the two types of data
 
+# these packages are required:
+required.packages = c('nlme', 'car', 'lme4', 'lmerTest', 'svglite')
+#  , 'ez'
+# they will be loaded later on
+
 # localization data can be downloaded from OSF:
 nocursorURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'='https://osf.io/8hm7f/?action=download')
 
@@ -31,9 +36,8 @@ colorset[['onlActT']] <- '#b400e42f'
 colorset[['onlPasS']] <- '#ff6ec7ff' # pink
 colorset[['onlPasT']] <- '#ff6ec72f'
 
-library('ez')
 
-installRequire.Packages <- function(packages) {
+installRequire.Packages <- function(packages, installmissing=FALSE) {
   
   installed.list <- rownames(installed.packages())
   missingpackages <- c()
@@ -54,12 +58,20 @@ installRequire.Packages <- function(packages) {
     cat('\nWARNING, some pacakages are missing:\n')
     cat(missingpackages)
     cat('\nlme4 and lmerTest are required to reproduce the Satterthwaite LME analyses\n')
-    cat('\ncar and nlme are required to reproduce the analyses using Chi-squared apprcimation\n')
+    cat('\ncar and nlme are required to reproduce the analyses using Chi-squared approximation\n')
+    cat('\nnot required, but helpful:\n')
     cat('\nforeach and doParallel speed up a few functions\n')
     cat('\nsvglite is used to produce SVG files with figures\n')
+    if (installmissing) {
+      install.packages(missingpackages)
+    }
   }
   
 }
+
+
+installRequire.Packages(required.packages)
+
 
 load.DownloadDataframe <- function(url,filename) {
   

### 2018-07-30 ManuscriptMaterials.Rmd
clean up code for loading packages, labeled some chunks

@@
 
+And we try to import some packages we use here:
+
+```{r require_lme_packages}
+required.packages = c('lme4', 'lmerTest', installmissing=FALSE)
+installRequire.Packages(required.packages)
+```
+
+Note: the packages `foreach` and `doParallel` will speed up the bootstrapping of peak fits, but the scripts will also run without.
+
+Note: the package `svglite` is used to generate SVG files with figures, but the scripts will simply skip this if the package is not available.
+
 ## Topics
@@
 
-```{r}
+```{r set_LMEmethod}
 LMEmethod <- 'Satterthwaite'
 # options: 'Satterthwaite', 'chi-squared'
+```
+
+If `LMEmethod` is set as 'chi-squared', we need some other packages to be able to run this:
+
+```{r requireChiSquared_packages}
+required.packages = c('nlme', 'car')
+installRequire.Packages(required.packages, installmissing=FALSE)
 ```
@@
 
-```{r}
+```{r plotBlinkDetectionPerformance}
 plotBlinkDetection()
@@
 
-```{r}
+```{r selectHighPerformingParticipants}
 blinks <- load.DownloadDataframe(informationURLs['blinkdetect'],'blinkdetect_exposure.csv')
@@
 
-```{r}
+```{r set_selectPerformance_variable}
 selectPerformance <- TRUE
@@
 
-```{r}
+```{r load_select_participantInfo}
 participants <- load.DownloadDataframe(informationURLs['demographics'],'participants_exposure.csv')
@@
 
-```{r}
+```{r age_stats}
 print(c('mean'=mean(participants$age), 'sd'=sd(participants$age)))
@@
 
-```{r}
+```{r sex_distribution}
 print(table(participants$sex))
@@
 
-## Pearons Correlations
+## Pearson Correlations
 
@@
                                                selectPerformance=selectPerformance)
-#plotBlinkDetection(generateSVG=TRUE)
 ```

### 2018-07-30 shared.R

@@ -1,13 +1,23 @@
 # some functions shared by the analyses of the two types of data
 
-# localization data can be downloaded from OSF:
+# these packages are required:
+required.packages = c('lme4', 'lmerTest')
+# they will be loaded later on
+
+# these packages or no longer used/required, but might help
+# 'nlme', 'car', 'svglite', 'doParallel', 'foreach'
+
+
+
+# URLs of data files on OSF
+# localization data:
 nocursorURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'='https://osf.io/8hm7f/?action=download')
-
-# no-cursor data can be downloaded from OSF:
+# no-cursor data:
 localizationURLs <- c('exposure'='https://osf.io/ysduw/?action=download', 'classic'='https://osf.io/upw49/?action=download', 'online'='https://osf.io/wjcgk/download')
-
-# information on exposure participants
+# information on exposure participants:
 informationURLs <- c('blinkdetect'='https://osf.io/yrpkm/?action=download', 'demographics'='https://osf.io/7sn4b/?action=download')
+
+
 
 # we'll control the color in figures centrally from here:
 colorset <- list()
@@ -31,9 +41,8 @@ colorset[['onlActT']] <- '#b400e42f'
 colorset[['onlPasS']] <- '#ff6ec7ff' # pink
 colorset[['onlPasT']] <- '#ff6ec72f'
 
-library('ez')
 
-installRequire.Packages <- function(packages) {
+installRequire.Packages <- function(packages, installmissing=FALSE) {
   
   installed.list <- rownames(installed.packages())
   missingpackages <- c()
@@ -54,12 +63,20 @@ installRequire.Packages <- function(packages) {
     cat('\nWARNING, some pacakages are missing:\n')
     cat(missingpackages)
     cat('\nlme4 and lmerTest are required to reproduce the Satterthwaite LME analyses\n')
-    cat('\ncar and nlme are required to reproduce the analyses using Chi-squared apprcimation\n')
+    cat('\ncar and nlme are required to reproduce the analyses using Chi-squared approximation\n')
+    cat('\nnot required, but helpful:\n')
     cat('\nforeach and doParallel speed up a few functions\n')
     cat('\nsvglite is used to produce SVG files with figures\n')
+    if (installmissing) {
[…]

### 2018-07-30 shared.R
fixed some commit errors

@@ -1,20 +1,11 @@
 # some functions shared by the analyses of the two types of data
 
 # these packages are required:
-<<<<<<< HEAD
 required.packages = c('lme4', 'lmerTest')
 # they will be loaded later on
 
 # these packages or no longer used/required, but might help
 # 'nlme', 'car', 'svglite', 'doParallel', 'foreach'
-=======
-required.packages = c('nlme', 'car', 'lme4', 'lmerTest', 'svglite')
-#  , 'ez'
-# they will be loaded later on
-
-# localization data can be downloaded from OSF:
-nocursorURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'='https://osf.io/8hm7f/?action=download')
->>>>>>> 01b71f2e7cdf732f910f96310c7953c133e563f7
 
 
 

### 2018-08-07 ManuscriptMaterials.Rmd
More updates to reflect what is in the manuscript better.

@@
 
-First we load those other scripts:
+We load those other scripts:
 
@@
 
-And we try to import some packages we use here:
+## Packages
+
+We try to import some packages used here.
+This has already been attempted by `shared.R` (for people without R Studio), but in the chunk below we can easily take action if the required packages are not available on the current system.
 
@@
 ```
+
+**If this fails: set the `installmissing` argument to TRUE to install the packages.**
 
@@
 ```{r requireChiSquared_packages}
+if (LMEmethod=='chi-squared') {
   required.packages = c('nlme', 'car')
   installRequire.Packages(required.packages, installmissing=FALSE)
+}
 ```
@@
 But, for most of the curves, and especially the ones that seems to be shifted, we don't have data on the full curve on one side of the peak.
-Or maybe it doesn't even include the peak itself.
-That makes it hard to fit a normal function to the data.
-Instead, we assume that the minimal change in the data, occurs far beyond the boundaries of the measured workspace, make the data cumulative and scale back to a range from 0 to 1.
+We can still bootstrap the __peak__ of Guassian functions fit to the data across many samples of participants for the exposure group and for the delayed localization after classic training, but not for the online localization after classic training.
+This data is presented in the paper, where we only look at the peak of the localization -- not the width of the generalization curve.
+
+The range of the data doesn't include the peak in many cases, which makes it hard to fit a normal function to the data.
+In one alternative approach, we assume that the minimal change in the data, occurs far beyond the boundaries of the measured workspace, make the data cumulative and scale back to a range from 0 to 1.
 This allows fitting a _cumulative_ normal function where the 50% point corresponds to the peak of the generalization curve.
-This method is more stable, and actually works on all the datasets used (but we didn't test online localization from the classic experiment).
-However, the results should still be considered carefully.
+This method is more stable, and seems to works fairly well on the datasets where the more common fits also work (but we didn't test online localization from the classic experiment).
+Code for this approach is available (functions `psychometricGeneralization()`, `pnormMSE()` and `getCDFpeaks()` at the end of the file: `shared.R`), but the results are not presented in the paper or notebook.
 
@@
 ```{r load_select_participantInfo}
-participants <- load.DownloadDataframe(informationURLs['demographics'],'participants_exposure.csv')
+participants <- load.DownloadDataframe(informationURLs['demographics'],
+                                       'participants_exposure.csv')
+
 if (selectPerformance) {
@@
 ```{r age_stats}
-print(c('mean'=mean(participants$age), 'sd'=sd(participants$age)))
+kable( c('mean'=mean(participants$age), 'sd'=sd(participants$age)),
+       col.names=c('age'),
+       caption='age of participants')
 ```
@@
 ```{r sex_distribution}
-print(table(participants$sex))
+kable( table(participants$sex),
+       caption='sex of participants',
+       col.names=c('sex','frequency'))
+```
+
+## Localization selection statistics
+
+We've manually selected localization responses that were unlikely to be true responses. Here are the statistics on how many were removed in each dataset. For exposure:
+
+```{r exposure_loc_selection}
+kable( aggregate(trials ~ rotated * passive, 
+                 data=countSelectedLocalizations('exposure'), 
+                 FUN=mean), 
+       caption='percentages selected localization trials after exposure training')
+```
+
+For classic:
+
+```{r classic_loc_selection}
+kable( aggregate(trials ~ rotated * passive,
+                 data=countSelectedLocalizations('classic'), 
+                 FUN=mean), 
+      caption='percentages selected localization trials after classic training')
+```
+
+We decided not to use any more outlier removal, but it is implemented and can be switched on in the code: set the `removeOutliers` argument to the `getPointLocalization` function in `shared.R` to TRUE.
+
+## No-cursor selection statistics
+
+Similarly, we get the percentage of selected trials in no-cursor reach tasks in exposure training:
+
+```{r exposure_nocursor_selection}
+kable( aggregate(trials ~ 1, 
+                 data=countSelectedNoCursors('exposure'), 
+                 FUN=mean), 
+       caption='percentages selected no-cursor trials after exposure training')
+```
+
+And in the classic training group:
+
+```{r classic_nocursor_selection}
+kable( aggregate(trials ~ 1, 
+                 data=countSelectedNoCursors('classic'), 
+                 FUN=mean), 
+       caption='percentages selected no-cursor trials after classic training')
 ```
@@
 
-```{r fig3_localization, fig.width=7.5, fig.height=2.5}
-plotLocalization(classicOnline=TRUE, selectPerformance=selectPerformance)
+```{r fig3_localization, fig.width=7.5, fig.height=3}
+plotLocalization(selectPerformance=selectPerformance,remove15=FALSE)
 ```
@@
 ```{r localization_count_table}
-kable(getLocCountTable(selectPerformance=selectPerformance))
+kable(getLocCountTable(selectPerformance=selectPerformance),
+      caption='count of participants with localization data at each hand angle')
 ```
@@
 ```{r loc_exp}
-exposureLocalization(remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+exposureLocalization(remove15=TRUE, 
+                     LMEmethod=LMEmethod, 
+                     selectPerformance=selectPerformance)
 ```
@@
 ```{r loc_exp_movtype}
-exposureLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+exposureLocalizationShift(remove15=TRUE, 
+                          LMEmethod=LMEmethod, 
+                          selectPerformance=selectPerformance)
 ```
@@
 ```{r loc_ombnibus}
-groupLocalization(model='full', remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+groupLocalization(model='full', 
+                  remove15=TRUE, 
+                  LMEmethod=LMEmethod, 
+                  selectPerformance=selectPerformance)
 ```
@@
 ```{r loc_groups_movementtype}
-groupLocalization(model='movementtype', remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+groupLocalization(model='movementtype', 
+                  remove15=TRUE, 
+                  LMEmethod=LMEmethod, 
+                  selectPerformance=selectPerformance)
 ```
@@
 ```{r loc_cla}
-classicLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod)
+classicLocalizationShift(remove15=TRUE, 
+                         LMEmethod=LMEmethod)
 ```
@@
 
-```{r fig2_reachaftereffect, fig.width=7.5, fig.height=2.5}
-plotReachAftereffects(selectPerformance=FALSE)
+```{r fig2_reachaftereffect, fig.width=7.5, fig.height=3}
+plotReachAftereffects(selectPerformance=selectPerformance)
 ```
@@
 ```{r RAE_exposure,include=TRUE}
-exposureNoCursorChange(LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+exposureNoCursorChange(LMEmethod=LMEmethod, 
+                       selectPerformance=selectPerformance)
 ```
@@
 ```{r RAE_exp_iterations}
-exposureAftereffectsPersistent(LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+exposureAftereffectsPersistent(LMEmethod=LMEmethod, 
+                               selectPerformance=selectPerformance)
 ```
 
-So there is no effect of iteration here, nor does it interact with target, which means the reach aftereffects are the same for the first iteration and the average across the other four iterations.
+There doesn't seem to be an effect of iteration, and no interaction with target.
+Also, looking at the figure, there would not be decay, as the reach aftereffects seem to get larger.
+We continue assuming there is no substantial change in the reach aftereffects between the first iteration and the average across the other four iterations.
 We now also see an effect of target angle: the generalization curve is not flat.
-This seems mostly due to the extra training.
 
@@
 ```{r RAE_exp_cla}
-exposureClassicReachAftereffects(LMEmethod=LMEmethod, selectPerformance=selectPerformance)
+exposureClassicReachAftereffects(LMEmethod=LMEmethod, 
+                                 selectPerformance=selectPerformance)
 ```
@@
 ```{r RAE_peaks}
-getPeakConfidenceInterval('classic', part='all', CIs=c(.95), selectPerformance=selectPerformance)
-getPeakConfidenceInterval('exposure', part='all', CIs=c(.95), selectPerformance=selectPerformance)
+getPeakConfidenceInterval('classic', 
+                          part='all', 
+                          CIs=c(.95), 
+                          selectPerformance=selectPerformance)
+getPeakConfidenceInterval('exposure', 
+                          part='all', 
+                          CIs=c(.95), 
+                          selectPerformance=selectPerformance)
 ```
@@
                       selectPerformance=selectPerformance)
-plotLocalization(classicOnline=TRUE,
+plotLocalization(classicOnline=FALSE,
                  generateSVG=TRUE, 

### 2018-08-07 localization.R

@@ -1,126 +1,11 @@
 
 source('shared.R')
 
-# first make sure we have all necessary packages installed and loaded:
-# installRequire.Packages(c('nlme', 'car'))
-
-# , 'lme4', 'lmerTest'
-# 
-# # no-cursor reach data is not uploaded yet (not sure about the format)
-# 
-# localizationLMEsChiSq <- function() {
-#   
-#   default.contrasts <- options('contrasts')
-#   options(contrasts=c('contr.sum','contr.poly'))
-# 
-#   exposure <- getANOVAlocalization('exposure')
-#   classic <- getANOVAlocalization('classic')
-#   
-#   exposure$participant <- exposure$participant + 100
-# 
-#   exposure <- exposure[-which(exposure$handangle_deg == 15),]
-#   classic <- classic[-which(classic$handangle_deg == 15),]
-#   
-#   # they all have this header:
-#   # group,online_b,participant,rotated_b,passive_b,handangle_deg,tapangle_deg
-#   # (where group is uninformative)
-# 
-#   localization <- rbind(exposure,classic)
-#   
-#   localization$group         <- factor(localization$group)
-#   localization$participant   <- factor(localization$participant)
-#   localization$online_b      <- factor(localization$online_b)
-#   localization$rotated_b     <- factor(localization$rotated_b)
-#   localization$passive_b     <- factor(localization$passive_b)
-#   localization$handangle_deg <- factor(localization$handangle_deg)
-#   
-#   # str(localization)
-#   
-#   # prepare a data frame with the training-induced changes in hand localization:
-#   locshift <- localization[which(localization$rotated_b == 1),]
-#   locshift$taperror_deg <- locshift$taperror_deg - localization$taperror_deg[which(localization$rotated_b == 0)]
-#   
-#   # first a model including the aligned and rotated data
-#   # here we look for any effect of session (rotated_b)
-#   # and if it is there, we continue with other analyses that are easier to grasp:
-#   exp_full_model <- lme(taperror_deg ~ rotated_b * passive_b * group, data=localization, random = ~1|participant/handangle_deg, na.action=na.exclude)
-#   print(Anova(exp_full_model, type=3))
-# 
[…]

### 2018-08-07 nocursor.R

@@ -1,77 +1,15 @@
 
 source('shared.R')
 
-# first make sure we have all necessary packages installed and loaded:
-# installRequire.Packages(c())
+# FIGURE -----
 
-# not yet sure if we'll use Chi-square, or Satterthwaite estimates of F/p-values:
-# installRequire.Packages(c('nlme', 'car', 'lme4', 'lmerTest'))
-
-required.packages = c('nlme', 'car', 'lme4', 'lmerTest', 'svglite')
-installRequire.Packages(required.packages)
-
-# nocursor data can be downloaded from OSF:
-# groupURLs <- c('exposure'='https://osf.io/9s6au/?action=download', 'classic'='https://osf.io/8hm7f/?action=download')
-
-
-# # plot both groups reach aftereffects in one figure
-# OLDplotReachAftereffects <- function(validate=FALSE) {
-#   
-#   points <- c(15,25,35,45,55,65,75)
-#   
-#   exposure <- getReachAftereffects('exposure',part='initial')
-#   classic  <- getReachAftereffects('classic',part='initial')
-#   exposure_rem <- getReachAftereffects('exposure',part='remainder')
-#   classic_rem  <- getReachAftereffects('classic',part='remainder')
-#   
-#   if (validate) {
-#     
-#     validExp <- validateReachAftereffects('exposure')
-#     validExpParticipants <- validExp$participant[which(validExp$RAE == 1)]
-#     exposure <- exposure[which(exposure$participant %in% validExpParticipants),]
-# 
-#     validCla <- validateReachAftereffects('classic')
-#     validClaParticipants <- validCla$participant[which(validCla$RAE == 1)]
-#     classic <- classic[which(classic$participant %in% validClaParticipants),]
-#     
-#   }
-#   
-#   exposureAVG <- aggregate(endpoint_angle ~ target, data=exposure, FUN=mean)
-#   classicAVG <- aggregate(endpoint_angle ~ target, data=classic, FUN=mean)
-#   
-#   exposureAVGrem <- aggregate(endpoint_angle ~ target, data=exposure_rem, FUN=mean)
-#   classicAVGrem <- aggregate(endpoint_angle ~ target, data=classic_rem, FUN=mean)
-#   
[…]

### 2018-08-07 shared.R

@@ -1,5 +1,7 @@
 # some functions shared by the analyses of the two types of data
 
+# BASICS -----
+
 # these packages are required:
 required.packages = c('lme4', 'lmerTest')
 # they will be loaded later on
@@ -41,6 +43,7 @@ colorset[['onlActT']] <- '#b400e42f'
 colorset[['onlPasS']] <- '#ff6ec7ff' # pink
 colorset[['onlPasT']] <- '#ff6ec72f'
 
+# UTILITY FUNCTIONS -----
 
 installRequire.Packages <- function(packages, installmissing=FALSE) {
   
@@ -126,6 +129,8 @@ bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10
   # parallel for-loop?
   # installRequire.Packages(c('foreach','doParallel'))
   
+  cat(sprintf('with %d bootstraps\n',bootstraps))
+  
   installed.list <- rownames(installed.packages())
   packages <- c('foreach','doParallel')
   allInstalled <- TRUE
@@ -141,6 +146,9 @@ bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10
   x <- as.numeric(names(colMeans(data)))
   
   if (allInstalled) {
+    
+    cat('using parallel processing\n\n')
+    
     cores=detectCores()
     
     # very friendly for the rest of the system
@@ -162,6 +170,8 @@ bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10
     stopCluster(cl)
     
   } else {
+    
+    cat('NOT using parallel processing\n\n')
     
     mus <- c()
     
@@ -207,7 +217,7 @@ bootstrapGaussianPeak <- function(data,bootstraps=1000,mu=47.5,sigma=30,scale=10
 # stopCluster(cl)
 
 
-getGaussianFit <- function(x,y,mu=47.5,sigma=30,scale=10,offset=4) {
+getGaussianFit <- function(x,y,mu=47.5,sigma=15,scale=-10,offset=-4) {
   
   data = data.frame(x,y)
   
@@ -234,7 +244,7 @@ parGaussian <- function(par,x) {
 
 # handling localization data -----
 
-getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55,65,75), movementtype='both', verbose=TRUE, LRpart='all', selectPerformance=TRUE) {
[…]

### 2018-11-28 paper/paper.Rmd
added files to compile the paper

---
title: "Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization After Passive Visuoproprioceptive Discrepancy Training"
author:
  - name: Ahmed A. Mostafa
    affiliation: [YORK,MANSOURA]
#    affiliation: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
  - name: Bernard Marius 't Hart
    affiliation: YORK
    email: thartbm@gmail.com
    corresponding: thartbm@gmail.com
  - name: Denise Y.P. Henriques
    email: deniseh@yorku.ca
    affiliation: YORK
address:
  - code: YORK
    address: CVR / Kinesiology and Health Science, York University, Toronto, Ontario, Canada
  - code: MANSOURA
    address: Faculty of Physical Education, Mansoura University, Mansoura, Egypt

abstract: |
  An accurate estimate of limb position is necessary for movement planning. Where we localize our unseen hand after a reach depends on felt hand position, or proprioception, but this is usually neglected in favour of predicted sensory consequences based on efference copies of motor commands. Both sources of information should contribute, so here we set out to further investigate how much of hand localization depends on proprioception and how much on predicted sensory consequences. We use a passive training paradigm with rotated visual feedback that eliminates the possibility to update predicted sensory consequences, but still recalibrates proprioception. After this training we measure participants’ hand location estimates based on both efference-based predictions and afferent proprioceptive signals with self-generated hand movements as well as based on proprioception only with robot-generated movements. We find indistinguishable shifts in hand localization after robot- and self-generated hand movements, and changes in open-loop reaches. Both motor and proprioceptive changes are only slightly smaller as those after training with self-generated movements, confirming that proprioception plays a large role in estimating limb position and in planning movements. (data: https://doi.org/10.17605/osf.io/zfdth, preprint: https://doi.org/10.1101/384941)

header-includes:
   - \usepackage{threeparttable}
   - \usepackage[table]{xcolor}
   - \usepackage{graphicx}
   - \usepackage{tabularx}
   - \usepackage{longtable}
   - \usepackage{siunitx}
bibliography: bibliography.bib
output: rticles::plos_article
csl: plos.csl
---

# Introduction

Sensory information is central to how we control all our movements. Our brain is even thought to use predicted sensory consequences derived from efferent copies of motor commands for motor control @Duhamel1992a. When training with rotated visual feedback of the hand, we update these predictions @Wolpert1995. Additionally, such training leads to a recalibration of our sense of felt hand position - “proprioception” - to be more aligned with the distorted visual feedback @Cressman2009. Both of these changes have been measured by asking people to localize where their unseen hand is – before and after training [@Clayton2013; @Izawa2012b; @Jones2012a; @Synofzik2008; @Yavari2016]. While our lab has previously found that proprioception accounts for a large portion of the change in hand localization @Hart2016, it is far from clear how much each process contributes or how to tease them apart. Here we use passive training, that removes the need to update predicted sensory consequences, in an attempt to isolate the contribution of proprioception to hand localization.

The predicted sensory consequences of movements may play several important roles in motor learning and control. Predicted sensory consequences allow us to correct our movements before sensory error signals are available, they can be used to select movements that best achieve our goals and they may inform us on the location of our limbs. Hence measuring predicted sensory consequences is valuable in movement research. In visuomotor rotation adaptation tasks, the actual sensory outcome is systematically different from the expected outcome, so that participants update their predictions on the outcome of the trained movements. In previous experiments, people were asked to make a movement and then indicate the location of, or “localize,” their unseen, right hand, before and after training with rotated visual feedback [@Izawa2012b; @Synofzik2008]. Since there was no visual information available to the participants, the predicted sensory consequences of the movement should be used in localizing the unseen hand. Both studies found a significant shift in hand localization, providing evidence that predicted sensory consequences are indeed updated as a result of visuomotor rotation adaptation.

However, our lab has shown that our sense of where we feel our hand to be, proprioception, is also reliably recalibrated after visuomotor rotation adaptation [@Cressman2009; @Hart2016; @Mostafa2014; @Mostafa2015; @Maksimovic2018; @Henriques2012; @Cressman2010a; @Barkley2014; @Salomonczyk2011; @Nourouzpour2014; @Henriques2014]. This has also been shown by other labs @Cameron2012 and a comparable proprioceptive change is induced with force-field adaptation @Ostry2016. As proprioception also informs us on the location of our limbs, we have on occasion used a task that is very similar to hand localization to investigate this [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018]. Although proprioceptive recalibration has been largely ignored as an explanation for changes in hand localization, we and others have shown that it accounts for a substantial part of the changes in localization, along with updates in predicted sensory consequences [@Hart2016; @Cameron2012]. Nevertheless, it is far from clear how much proprioception and prediction each contribute to hand localization.

Here we intend to further examine the contribution of proprioception to hand localization. To do this, we use passive training, where a robot arm moves the participant’s arm out, so that the cursor always directly goes to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013]. This means there is no efference copy available and no visuomotor error-signal, both of which are required to update predicted sensory consequences. However, we impose a discrepancy between vision and proprioception that drives proprioceptive recalibration. Thus, changes in hand localization after this type of training should be due to proprioceptive recalibration only. We use the same experimental protocol as before @Hart2016, so that we can compare localization shifts between the two different types of training, and can better assess the contributions of predicted sensory consequences and proprioception to hand localization.

# Methods

We set out to test the relative contributions of proprioception and efference-based prediction to hand localization. We use visual training with robot-generated hand movements to prevent updates of predicted sensory consequences, but still elicit proprioceptive recalibration.

## Participants

Twenty-five right-handed participants were recruited for this study. One participant was excluded for not following task instructions, and three were excluded for low performance on a task that ensures attention during the passive training. All analyses presented here pertain to the remaining twenty-one participants (13 females and 8 males, mean age: 20.1 +/- 2.3 years), but the data of the three low-performing participants is included in the online \href{https://osf.io/zfdth}{dataset} @Mostafa2017. All had normal or corrected-to-normal vision, and provided prior, written informed consent in accordance with the ethical guidelines set by the York Human subjects Review Subcommittee and received credit toward an undergraduate psychology course. Participants were screened verbally, and all reported being right handed and not having any history of visual, neurological, and/or motor dysfunction.

![**Setup and tasks:** **a)** Participants moved their unseen right hand with visual feedback on hand position provided through a mirror (middle surface) half-way between their hand and the monitor (top surface). A touchscreen located just above the hand was used to collect responses for the localization tasks (bottom surface). **b)** Training task. The target, shown as a yellow disc, is located 10 cm away from the home position at 45°. In the rotated training tasks, the cursor (shown here as a green circle) represents the hand position rotated 30° relative to the home position. **c)** No-cursor reach task. Targets are located 10 cm away from the home position at 15°, 25°, 35°, 45°, 55°, 65°, and 75°, shown by the yellow circles here (only one was shown on each trial). While reaching to one of these targets, no visual feedback on hand position is provided. d) Localization task. The participants’ unseen, right hands moved out, and subsequently participants indicated the direction of the hand movement by indicating a location on an arc using a touch screen with their visible left index finger](Fig1_setup.eps)

## Setup

Participants sat in a height-adjustable chair to ensure that they could easily see and reach all targets presented on a reflective screen (see Fig 1). During all tasks, they held the vertical handle on a two-joint robot manipulandum (Interactive Motion Technologies Inc., Cambridge, MA) with their right hand so that their thumb rested on top of the handle. A monitor (Samsung 510 N, refresh rate 60 Hz) was mounted 11 cm above the reflective screen, such that images displayed on the monitor appeared to lie in the horizontal plane where the right hand was moving. The reflective screen was mounted horizontally 18 cm above the robot manipulandum. A touch screen was mounted 13 cm underneath the reflective surface, so that subjects could indicate the location of the unseen right-hand locations (specifically the unseen thumb) with their visible left hand, which was lit up with a small spot light (only in localization tasks). The room lights were dimmed and the participants’ view of their right hand was blocked by the reflective screen, as well as a dark cloth draped between the touch screen and participants’ right shoulders.

## Procedure

The first part of the experiment used training with a cursor aligned with the hand and the second part had training with a cursor rotated around the start position (Fig 1b; white rows in Table 1). During the training with rotated feedback, the cursor was gradually rotated 30° clockwise. This introduced a discrepancy between the actual, or felt, hand position and the visual feedback, that should evoke proprioceptive recalibration. However, the movements were robot generated, so that there were no predicted sensory consequences based on the outgoing motor command. Hence the prediction errors that are thought to lead to motor learning were absent as well. After both types of training, participants did open-loop reaches as well as two kinds of hand localization tasks, to test the effect of training on proprioceptive and predictive hand estimates.

$

\newcommand{\headrow}{\rowcolor{black!20}}
\newcommand{\thead}[1]{\multicolumn{1}{l}{\bfseries #1\rule[-1.2ex]{0pt}{2em}}}

\definecolor{LGray}{gray}{0.9}
\definecolor{DGray}{gray}{0.8}

\begin{table}[bt]
\caption{Task order}
\begin{threeparttable}
\begin{tabular}{cccc}
\headrow
\multicolumn{2}{c}{\# trials} & \multicolumn{2}{c}{training type} \\
\headrow
%\multicolumn{2}{|c|}{\# trials} & \thead{Exposure training} & \thead{Classic training} \\
\thead{aligned} & \thead{rotated} & \multicolumn{1}{c}{\textbf{Exposure training}} & \multicolumn{1}{c}{\textbf{Classic training}} \\
\rowcolor{white}
50 & 90 & training & training \\
\rowcolor{LGray}
- & 21 & no-cursor & no-cursor \\
\rowcolor{white}
- & 60 & training & training \\
\rowcolor{DGray}
25 & 25 & active delayed localization & active delayed localization \\
\rowcolor{LGray}
21 & 21 & no-cursor & no-cursor \\
\rowcolor{white}
10 & 60 & training & training \\
\rowcolor{DGray}
25 & 25 & passive delayed localization & passive delayed localization \\
\rowcolor{LGray}
21 & 21 & no-cursor & no-cursor \\
\rowcolor{white}
10 & 60 & training & training \\
\rowcolor{DGray}
25 & 25 & active delayed localization & \textit{(active online localization)} \\
\rowcolor{LGray}
21 & 21 & no-cursor & no-cursor \\
\rowcolor{white}
10 & 60 & training & training \\
\rowcolor{DGray}
25 & 25 & passive delayed localization & \textit{(passive online localization)} \\
\rowcolor{LGray}
21 & 21 & no-cursor & no-cursor \\
\hiderowcolors
\hline  % Please only put a hline at the end of the table
\end{tabular}

\begin{tablenotes}
\item 
\end{tablenotes}
\end{threeparttable}
\end{table}
$

## Exposure training

In what we call ‘exposure training’ the participants did not move their hand toward the target, but the robot did. In this task (Table 1), the right hand was represented by a cursor (green disk, 1 cm in diameter, Fig 1b) located directly above participant’s thumb. The robot moved the participant’s unseen right hand (and the cursor) along a direct path toward a visual yellow target disk and back to the starting position (1 cm in diameter, Fig 1b). The home position was located approximately 20 cm in front of participants and the visual target located 10 cm from the home position at 45° (Fig 1b). In order to make sure participants were paying attention to the cursor, the cursor was switched off for 2 screen refreshes (~33.3ms) on 50\% of the trials at a random distance between 4 and 9 cm from the home position and participants were asked to report this using a button press with the left hand. Performance on this task was used to screen participants.

During the first half of the experiment, the cursor and hand path were aligned during exposure training. In the second part of the experiment, the “rotated” session, the same visual training target at 45° was used, and the cursor kept moving straight to this target. However, the robot-generated hand path gradually rotated 30° CCW (Fig 1b) with respect to the visible target and the cursor in increments of 0.75°/trial, so that the full rotation was reached after 45 trials. This mimicked error-free responses to a gradual visuomotor rotation of 30° CW. The initial training consisted of 50 trials in the aligned part and 90 in the rotated part. In between open-loop reach tasks and localization tasks (Table 1) extra training tasks were done, each of which consisted of 10 trials in the aligned part of the experiment and 60 trials in the rotated part (Table 1).

## No-cursor reaching

The trials in no-cursor reaching (Table 1, light gray rows) serve as a classical measure of motor adaptation. On each of these trials participants were asked to reach with their unseen right hand to one of 7 visual targets, without any visual feedback of hand position. The targets were 10 cm from the home position, located radially at: 15°, 25°, 35°, 45°, 55°, 65°, and 75° (Fig 1c). A trial started with the robot handle at the home position and, after 500 ms, the home position disappeared and the target appeared, cuing the participants to reach for the target. Once the participants thought they had reached the target they held their position for 250 ms, and the target and the home position disappeared, cuing participants to move back to the home position along a straight, constrained path, to begin the next trial. If participants tried to move outside of the path, a resistance force, with a stiffness of 2 N/(mm/s) and a viscous damping of 5 N/(mm/s), was generated perpendicular to the path. In every iteration of the no-cursor reach task, each target was reached three times, for a total of 21 reaches in pseudo-random order. The no-cursor reaching task was performed four times in the aligned part of the experiment and five times in the rotated part of the experiment.

## Localization

In this task (Fig 1d; Table 1, dark gray rows) participants indicated where they thought their unseen right hand was after a movement. First, an arc appeared, spanning from 0° to 90° and located 10 cm away from the home position and the participants’ unseen, right hand moved out from the home position in a direction towards towards a point on the arc. The hand was stopped by the robot at 10 cm from the home position and then, to prevent the online proprioceptive signals from overriding the predictive signals [@Izawa2012b; @Hart2016], the hand was moved back to the home position using the same kind of constrained path as used for the return movements in the no-cursor task. Participants indicated with the index finger of their visible, left hand on the touch screen mounted directly above the robot handle where they thought their trained hand had crossed the arc.

Crucially, there were two variations of this task. First, in the ‘active’ localization task participants generated the movement themselves, as they could freely move their unseen right hand from the home position to any point on the arc. Second, there was a ‘passive’ localization task where the robot moved the participants’ hand out and back, to the same locations the participants moved to in the preceding ‘active’ localization task in a shuffled order (hence, active localization is done first). In active localization, participants have access to both proprioceptive information as well as an efference-based prediction of sensory consequences, but in passive localization, only proprioception should be available. The active and passive localization task each consisted of 25 trials, and each of the tasks was done a total of four times; twice after aligned and twice after rotated training.

## Classic training

The paradigm described above is an exact replica of a paradigm we used earlier @Hart2016 with two exceptions. First, we used exposure training here, instead of the standard reach training with volitional movements, which we will call ‘classic’ training. Second, all localization is delayed until the right hand has returned to the home position in this study (see Table 1), so that instead of both delayed and online localization we have two repetitions of each delayed localization task. With this paradigm we can compare changes in localization and no-cursor reaches change after exposure training with changes in the same measures after classic visuomotor adaptation training.

## Analyses

Prior to any analyses, both the localization responses and the no-cursor reach data were visually inspected and trials where the participants did not follow task instruction were removed (e.g. several movements back and forth, or a touch-screen response on the home position, instead of on the white arc).

## Localization

Localization responses were taken as the (signed) angular difference between vectors through the home position and the actual hand position as well as the location indicated on the touch screen. Prior to analyses, idiosyncratic differences in performing this task were countered. Before conversion to degrees angle, a circle with a 10 cm radius was fit to the touch screen responses of each participant and the offset of this circle’s centre was subtracted from all response coordinates, so that all responses fell close to the arc. Then, a smoothed spline was fit to every participant’s response errors in each of the four localization tasks (aligned vs. rotated and active vs. passive) and these were used to obtain localization errors at the same locations used for the no-cursor reaches (15°, 25°, 35°, 45°, 55°, 65° and 75°), but only if that location fell within the range of the data (i.e. we only interpolate). This way localization responses could be compared across participants despite the freely chosen reach directions. At the 15° location 7/21 participants didn’t have an estimate in one or more of the four localization tasks (in the “classic” data it was also 7/21). While that data is shown in the figures, we did not use it for analysis.

First we test if localization responses shifted following rotated exposure training compared to aligned. We then test if the shift in localization responses is different for active and passive localization, and we run analyses comparing localization after exposure training with localization after “classic” training. Finally, we explore the generalization of localization responses and if they are different between the groups doing classic and exposure training.

## Reach aftereffects

To assess any reach adaptation that may have occurred after exposure training we analyzed reach endpoint errors in no-cursor trials. Reach endpoint errors were the (signed) angular difference between a vector from the home position to reach endpoint and a vector from the home position to the target. We obtained reach aftereffects by subtracting reach endpoint errors after aligned training from those after rotated training. No-cursor endpoint errors were analyzed to see if participants adapted the direction of their reaching movements after rotated exposure training. We also tested if any such change decayed, i.e. if it was the same immediately after exposure training, or when a localization task was done in between exposure training and no-cursor reaches. Furthermore, we tested if the generalization of reach aftereffects is different between exposure and classic training and if there is any generalization of reach aftereffects.

Pre-processing and analyses were done in R 3.4.4 @RCoreTeam2017, using lme4, lmerTest and various other packages. Most analyses used linear mixed effects models, since there is some missing localization data. These were “converted” to more readable ANOVA-like output, using a Satterthwaite approximation @Luke2017. Highly similar results were obtained with a Chi-square approximation. Data, scripts and a notebook with analyses have been made available on the Open Science Framework @Mostafa2017 ([https://osf.io/zfdth](https://osf.io/zfdth/)).

In short, this experiment allowed us to test how mere exposure to a visual-proprioceptive discrepancy changes both reach aftereffects and hand localization responses, and compare them with those obtained after more regular, “classic” training.

# Results

In this study we intend to further elucidate the relative contributions of (updated) predicted sensory consequences and (recalibrated) proprioception to hand localization. We can parcel out these contributions by measuring hand localization after both robot-generated and self-generated movements. Finally we compare the data from the current experiment with those obtained in an earlier study that used an identical paradigm, but trained with self-generated movements, or “classic” training.

![**Hand localization:** The shifts of the angles of touchscreen responses in all variations of the localization task, using spline-interpolated estimates for hand angles matching the reach targets in the no-cursor reach block. **a)** Localization shifts after exposure training. Dark blue: active localization shifts, Light blue: passive localization shifts. **b)** Localization shifts after classic training. Dark red: active localization shifts, Orange: passive localization shifts. The dashed line segments illustrate that the 15° data is not used for statistical analyses (see Methods). **c)** Generalization curves of active localization shifts after exposure training (blue) and classic training (red). Shaded areas: 95\% confidence intervals for the peak of the generalization curve (red and blue lines through shaded area indicate 50\% points). Downward black arrow: visual trained target. Upward black arrow: hand location during training.](Fig2_plain.pdf)

## Localization

Here we test our hypothesis that exposure training does not lead to changes in predicted sensory consequences. Since the difference between active and passive localization stems only from the presence or absence of predicted sensory consequences, there should be no difference between the two if predicted sensory consequences are not changed by exposure training. At first glance, it seems there might be a difference between active and passive localization shifts in exposure training (see Fig 2a), although it is smaller than in classic training (Fig 2b).

To test if rotated exposure training induces changes in hand localization, we fit an LME to the localization errors throughout the workspace using session (aligned or rotated), movement type (active and passive) and hand angle (25°, 35°, 45°, 55°, 65° and 75°), and all interactions as fixed effects and participant as random effect. There was an effect of session (F(1,450.5)=155.8; p<0.001), showing that exposure training leads to changes in hand localization. There was also an effect of hand angle (F(5,450.6)=6.54; p<.001) and an interaction between hand angle and session (F(5,450.3)=8.25; p<.001), which we’ll explore below, but no other effects (all p>.60). Since localization responses did shift, we use the difference between hand localization after rotated training and after aligned training (as plotted in Fig 2) for further tests.

If this shift in localization after exposure training partly reflects predicted sensory consequences, then shifts in active localization, that rely on both (recalibrated) proprioception and (updated) predictions should be different from shifts in passive localization that only rely on (recalibrated) proprioception. We fit an LME to the change in localization using movement type (active or passive localization) and hand angle, as well as their interaction as fixed effects and participant as random effect. There was no effect of movement type (F(1,211.8)=0.07; p=0.79). There was an effect of hand angle (F(5,212.2)=10.8; p<0.001), but no interaction between hand angle and movement type (F(5,211.8)=1.23, p=.29). The lack of an effect of movement type suggests that predicted sensory consequences did not contribute to localization in this paradigm.

In order to compare hand localization shifts after exposure training with those after classic training @Hart2016, we fit an LME to localization shift using training type (exposure vs. classic), movement type (active vs. passive) and hand angle and all interactions as fixed effects and participant as random effect. There was a main effect of movement type (F(1,422.0)=6.22; p=.013) and of hand angle (F(5,422.7)=8.19; p<.001), as well as an interaction between training type and hand angle (F(5,422.7)=4.54; p<.001) and between training type and movement type (F(1,422.0)=4.48, p=.035), but there was no main effect of training type (F(1,39.1)=0.92, p=.34) and no other effects (all p>.14). These results also suggest that the magnitude of the shifts in localization are comparable between classic and exposure training, but that the pattern of generalization is different.

To address our main question, we will look at the interaction between training type (exposure vs. classic) and movement type (active vs. passive) we found above. Since there is no difference between active and passive localization shifts after exposure training alone, the interaction between training type and movement type should be caused by an effect of movement type on the localization shifts after classic training, as we found previously @Hart2016. This means that shifts in hand localization after exposure training indeed rely on recalibrated proprioception alone, while after classic training, there also is a contribution of predicted sensory consequences to active localization.

For completeness, we explore the potentially different generalization patterns of localization shifts after classic and exposure training (Fig 2c). The LME indicates no difference in overall amplitude of localization shifts between the groups, so the interaction between training type and hand angle might stem from a generalization that does not peak at the trained location after exposure training. Using the active localization shifts only (which are larger, and arguably more similar to reach aftereffects), we bootstrap a 95\% confidence interval for the peak localization shift across participants in each group. Here we include the data at 15° where it is available. After classic training, the peak localization shift was at 48.8° (95\% confidence: 31.6° - 62.8°; red area in Fig 2c), and after exposure training the peak localization shift was at 62.1° (95\% confidence: 50.0° - 75.6°; blue are in Fig 2c). This means that peak localization after classic training is lower than after exposure training, but not vice versa. Also note that the confidence interval for the peak localization shift after classic training includes the trained target (45°), but not after exposure training @Izawa2012b. In short, the LME for localization shifts indicates a different generalization curve after exposure and classic training, which might be partially explained by a different position of the peak localization shift after exposure or classic training.

## Reach aftereffects

Apart from proprioception and prediction, we want to see if rotated exposure training has any effect on open-loop reaches and if these are robust. We measure whether participants adapted their reach directions by assessing their reach errors in no-cursor reach trials after aligned and rotated exposure training. In Fig 3, the changes in no-cursor endpoint errors, or reach aftereffects, appear to be well over 5°. First, to test if exposure training affects open-loop reach direction, we fit a linear mixed effects model (LME) to reach endpoint error using session (aligned; all blocks, or rotated; only the first block immediately after training) and target (15°, 25°, 35°, 45°, 55°, 65° and 75°), as well as their interactions as fixed effects and participant as random effect. There is an effect of session (F(1,260)=93.81, p<.001), that is: exposure training leads to substantial reach aftereffects. There was no effect of target (F(6,260)=1.07, p=.37) and no interaction (F(6,260)=1.00, p=.42). Since there is an effect of session, we now take the differences in reach endpoint errors between the rotated and aligned session for every participant and target as reach aftereffects, and use those for all further analyses.

![**Reach aftereffects:** Changes of the angle of reach endpoints in the no-cursor tasks. **a)** Reach aftereffects across the experiment. Light blue: first no-cursor task in the rotated session (immediately following training), Dark blue, dashed line: the other four repetitions of the task (with localization in between training and no-cursor tasks). **b)** Reach aftereffects after classic and exposure training. Blue: exposure training, Red: classic training. **c)** Generalization curves of reach aftereffects after exposure training (blue) and classic training (red). Shaded areas: 95\% confidence intervals for the peak of the generalization curve (red and blue lines through shaded area indicate 50\% points). Downward black arrow: visual trained target. Upward black arrow: hand location during training.](Fig3_plain.pdf)

To see if reach aftereffects decayed during the localization tasks, we compared reach aftereffects in the initial no-cursor block, that immediately followed training, with those in the later blocks that followed a localization task. We fit an LME to the reach aftereffects with iteration (initial vs. later no-cursor blocks) and target (as above) as well as their interaction as fixed effects and participant as random effect. There is no effect of iteration (F(1,260)=2.72, p=0.10). There was an effect of target (F(6,260)=6.29, p<.001) but no interaction (F(6,260)=0.58, p=0.74). Hence, reach aftereffects were not appreciably different right after training and after the localization tasks. In other words, there was likely no noticeable decay of reach aftereffects during the localization tasks, so that we can collapse the data across iterations.

Next we compare the reach aftereffects after classic training with those after exposure training (Fig 3b). It appears as if there is little overall difference in magnitude, but there might be a shift of the generalization curve. We fit an LME to reach aftereffects with training type (classic vs. exposure), target (as above) as well as their interaction as fixed effects and participant as random effect. There is no main effect of training (F(1,40)=0.11, p=.74), indicating approximately equal magnitude of reach aftereffects after the two training types. There is an effect of target (F(6,240)=8.36, p<.001), indicating that reach aftereffects exhibit some form of a generalization curve. There is also an interaction between training type and target (F(6,240)=2.27, p=.038), indicating these generalization curves are different after the two training types.

We will explore these potentially different generalization curves here. In Fig 3c we can observe that reach aftereffects after exposure training seem not to peak at the trained target direction of 45° but a more forward direction. We test this by taking the 95\% confidence interval of the centre of a normal curve fit to this data, bootstrapped across participants, and find that the median peak of the generalization curve of reach aftereffects after exposure training is at 66.3°, with a 95\% confidence interval ranging from 49.3° to 78.9°. This would indeed suggest that the reach aftereffects after exposure training do not generalize around the trained direction of 45°, but at a more counter clockwise location. For classic training, generalization of reach aftereffects peaks at 53.2°, with a 95\% confidence interval spanning 42.1° to 66.5°. So for classic training the 95\% confidence interval for peak reach aftereffects does include the trained target. These confidence intervals also indicate that generalization of reach aftereffects does not peak at different target position after exposure and classic training. However, we can also observe that the full curve was not sampled after exposure training, so that curve fitting is not optimal. This means that – given our data – the interaction between target and training type found in the LME above can’t be explained by a shifted generalization curve. This may be because our experiment was not set up to test this, and the similar angles where generalization peaks in both localization and reach aftereffects, suggests there may be a difference in where proprioception and prediction generalize strongest.

In summary, our main hypotheses are confirmed; exposure training leads to shifts in hand localization that are not different for active or passive localization, while movement type does have an effect on localization shift after classic training. Exposure training also causes robust reach aftereffects that are of comparable size to those found with classic training. There is some evidence that the generalization of both localization shifts and reach aftereffects are different after the two training types, and it appears this can partially be explained by a different peak of the generalization curves, but our data and analyses are not definitive.

# Discussion

The position of limbs is important for planning and evaluating movements, and can be estimated through predicted sensory consequences, as well as visual and proprioceptive feedback. As in a previous study @Hart2016 here we quantify the contributions of predicted sensory consequences and proprioceptive recalibration to where we localize our hand after training with altered visual feedback of the hand. In classical adaptation paradigms, both predictions are updated and proprioception is recalibrated. Predictions are updated when they don’t match actual sensory consequences, and proprioception is recalibrated when it doesn’t match visual feedback. In this study we use “exposure” training, where the participants do not have volitional control of their movements. By design, this should eliminate efference copies and prevent updating predicted consequences of movements, but since the proprioceptive and visual feedback is the same, exposure training still allows proprioceptive recalibration. Before and after training, participants localize their hand, both after “active,” self-generated movements that allow using predicted sensory consequences, and after “passive,” robot-generated movements that only allow using proprioception. We calculate the training-induced shift in both types of localization given the same actual hand position. After classical training we previously reported larger shifts in active localization as compared to passive @Hart2016. As we expected, after exposure training there are substantial shifts in localization, but no difference between active and passive localization, indicating that predictions are not updated after exposure training. Furthermore, we find that exposure training evokes substantial and robust reach aftereffects, indicating that recalibrated proprioception is used to plan movements.

Our lab previously investigated proprioceptive recalibration and reach aftereffects following visuomotor adaptation with classic training and matched exposure training. There we also found that proprioceptive recalibration is of similar magnitude in both training paradigms, but unlike here, reach aftereffects are usually much larger with classic training [@Henriques2012; @Henriques2014; @Ruttle2018; @Cressman2010; @Salomonczyk2013]. And while proprioceptive recalibration and reach aftereffects do proportionally increase with gradual increases in rotation size for classical training, they do not for exposure training @Salomonczyk2013. The similar magnitude of proprioceptive recalibration and reach aftereffects following exposure training, but not classical training, suggest that this sensory recalibration is partly driving this modest change in movements. The effect of exposure training on movements is also demonstrated by facilitation by exposure training of subsequent classic training @Sakamoto2015 (but no interference) and transfer of exposure training effects from one hand to the other @Bao2017. In the current study, we further demonstrate that exposure training affects movements and proprioception, but also measure its potential effect on predictive estimates.

Results similar to what we find here were reported in a study by Cameron and colleagues @Cameron2012, using gain modulation of visual feedback of single-joint hand movements around the elbow. Their within-subjects experiment included both training with volitional movements as well as with passive movements and also tested perception of movements that were either passive or active. They too found a robust change in passive perception of hand movement (using a different measure), and these changes did not differ between the two types of training. Similarly, they found shifts in what we might call “active localization,” although the task is different, after both training types. Like here, these shifts are larger after classic training as compared to exposure training. They also found that passive exposure leads to reach aftereffects, although these were smaller than those produced following “classical” training with altered visual gain. Both our findings, and those of Cameron et al. @Cameron2012 indicate that updating predicted sensory consequences requires volitionally controlled movements that lead to prediction errors, while proprioception recalibrates equally in both types of training, and that recalibrated proprioception affects open-loop reaches. Our combined results suggest that updates in predicted sensory consequences only provide a partial explanation for motor learning.

Two related concerns about exposure training and passive localization are that the movements are not fully passive, so that efference-based predictions are still generated or that predicted sensory consequences are generated through another route. Cameron et al. @Cameron2012 measured muscle activity (EMG) during passive movements and found no difference with stationary baseline muscle activity. This suggests that any movements generated in a passive condition are subthreshold, minimizing efference-based predictions. The brain areas generating predicted sensory consequences could also rely on afferent signals. However, such afferent signals are present in both active and passive movements, and if they would result in the same predictions, there would be no difference between active and passive localization after classic training, and no difference between the effects of exposure and classic training, and we find both are different. Hence, while we can not fully exclude any predictive signals in passive localization or exposure training, our data shows that any residual predictive signals in the passive movements we used are qualitatively very different from normal efference-based predicted sensory consequences.

In our classical training group, we not only see shifts for passive localization but even larger shifts for active localization which is consistent with a change in both proprioception and an update in predictions. In our exposure training, we did not find a consistent difference between active and passive localization, and none at the trained direction. Assuming that predictions were not updated in exposure training, a maximum likelihood estimate (MLE) or “optimal integration” @Ernst2002 would predict that active localization should shift less than passive localization after exposure training. But of course this is not the case in our findings (although it is the case for Cameron et al.). This suggests that perhaps these two signals are not optimally integrated which is consistent with our comparisons of the variance between passive and active localization. In ’t Hart and Henriques @Hart2016, we tested the prediction derived from MLE that hand localization with two signals – proprioception and prediction in active localization – should be more reliable, i.e. have lower variance, than hand localization with only one signal – proprioception only in passive localization. However, we found no difference in variance between active and passive localization, and recently replicated this in a much larger dataset @tHart2018. Taken together, this suggests these different sources of information about unseen hand location are not optimally integrated. While localizing the unseen hand is less precise than locating (pointing to) a remembered visual target or a seen and felt hand location, we find that these bimodal estimates are rarely integrated optimally [@Fiehler2010; @Byrne2013; @Jones2010], although others have @VanBeers2002. A more recent study @Mikula2018 has also shed doubt on whether “optimal” or “Bayesian” integration is used for locating the hand with two afferent signals. Analogously, here we again can’t find evidence that afferent and efferent information combine as a maximum likelihood estimate.

It seems clear that the cerebellum plays a role in motor learning as it appears to compute predicted sensory consequences, i.e. it implements a forward model [@Bastian2006; @Shadmehr2008; @Wolpert1998]. People with cerebellar damage do worse on motor learning tasks [@Martin1996; @Maschke2004; @Criscimagna-Hemminger2010; @Donchin2012], and show decreased shifts in hand localization tasks following motor learning [@Izawa2012b; @Synofzik2008]. This highlights that the cerebellum, and likely predicted sensory consequences, are important for motor learning, but does not explain the remaining shifts in hand localization. We previously found that proprioceptive recalibration is intact in people with mild cerebellar ataxia and that it is similar following exposure and classical training with a gradually introduced cursor rotation @Henriques2014. The remaining changes in hand localization found in cerebellar patients can be attributed to recalibrated proprioception which should be intact @Henriques2014. Analogously, here we show that in a paradigm that stops updates of predictions of sensory consequences, as supposedly in people with cerebellar damage, we still see substantial shifts in localization. Again, the remaining localization shifts can be explained if, along with predictions, the human brain uses afferent signals: recalibrated proprioceptive estimates, to localize the hand.

## Generalization

We do find some evidence that, after exposure training, the generalization curves for localization shifts are not centred on the visual location of the trained target; they don’t peak at 45° but at ~62°. In contrast, after classic training the peak of the generalization curve does peak close to the training target. It is possible that proprioceptive recalibration is not anchored to the visual goal of the training task as it is not a requirement to feel your hand at any specific point; rather, in classic training the visual cursor has to be brought to a visual target to end a trial. While in exposure training the movements are executed for the participants without error, their task is still to pay attention to the visual cursor while it moves to the visual target; there are no task demands on proprioception. Although we can’t substantiate this here, the generalization curves of the reach aftereffects seem to mimic the generalization curves of localization shifts, suggesting a relationship between changes in state estimates and changes in movements. This needs to be tested further, but if these effects are true, they may provide insight into how state estimates are used to produce movements, and also may lead to a new method to disentangle the influence of recalibrated proprioception and more traditional updated internal models on motor changes, such as reach aftereffects. Either way, even though the experiment was not designed to investigate this, the shifted generalization curves of changes in localization after classic and exposure training suggest they are generated by different mechanisms. This is in line with our earlier findings that proprioception generalizes differently from reach adaptation @Mostafa2015.

## Conclusion

To sum up, after a training paradigm designed to prevent updating of predicted sensory consequences but allow recalibration of proprioception, we find substantial changes in where people localize their hand. This means that recalibrated proprioceptive estimates can explain shifts in hand localization. The distinct change in the direction of open-loop reaches we observe here, suggests recalibrated proprioception can contribute to motor adaptation in other contexts as well. Finally, we have some evidence that after exposure training, the shift in hand localization does not generalize around the trained target location. All of this confirms that different mechanisms underlie proprioceptive recalibration and motor adaptation.

## Acknowledgements

This work was supported by a Canadian Network for Research and Innovation in Machining Technology NSERC Operating grant (DYPH) and the German Research Foundation (DFG) under grant no. HA 6861/2-1 (BMtH). The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript. We thank Shanaathanan Modchalingam for assistance in collecting the data.

# References {#references .unnumbered}

### 2018-11-28 paper/paper.tex
added files to compile the paper

--- 2018-11-28 paper/paper.Rmd
@@
----
-title: "Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization After Passive Visuoproprioceptive Discrepancy Training"
-author:
-  - name: Ahmed A. Mostafa
-    affiliation: [YORK,MANSOURA]
-#    affiliation: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
-  - name: Bernard Marius 't Hart
-    affiliation: YORK
-    email: thartbm@gmail.com
-    corresponding: thartbm@gmail.com
-  - name: Denise Y.P. Henriques
-    email: deniseh@yorku.ca
-    affiliation: YORK
-address:
-  - code: YORK
-    address: CVR / Kinesiology and Health Science, York University, Toronto, Ontario, Canada
-  - code: MANSOURA
-    address: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
+% Template for PLoS
+% Version 3.4 January 2017
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
 
-abstract: |
-  An accurate estimate of limb position is necessary for movement planning.
-Where we localize our unseen hand after a reach depends on felt hand position, or proprioception, but this is usually neglected in favour of predicted sensory consequences based on efference copies of motor commands.
-Both sources of information should contribute, so here we set out to further investigate how much of hand localization depends on proprioception and how much on predicted sensory consequences.
-We use a passive training paradigm with rotated visual feedback that eliminates the possibility to update predicted sensory consequences, but still recalibrates proprioception.
-After this training we measure participants’ hand location estimates based on both efference-based predictions and afferent proprioceptive signals with self-generated hand movements as well as based on proprioception only with robot-generated movements.
-We find indistinguishable shifts in hand localization after robot- and self-generated hand movements, and changes in open-loop reaches.
-Both motor and proprioceptive changes are only slightly smaller as those after training with self-generated movements, confirming that proprioception plays a large role in estimating limb position and in planning movements.
-(data: https://doi.org/10.17605/osf.io/zfdth, preprint: https://doi.org/10.1101/384941)
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
 
-header-includes:
-   - \usepackage{threeparttable}
-   - \usepackage[table]{xcolor}
-   - \usepackage{graphicx}
-   - \usepackage{tabularx}
-   - \usepackage{longtable}
-   - \usepackage{siunitx}
-bibliography: bibliography.bib
-output: rticles::plos_article
-csl: plos.csl
----
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
 
-# Introduction
+% Use Unicode characters when possible
+\usepackage[utf8x]{inputenc}
 
-Sensory information is central to how we control all our movements.
-Our brain is even thought to use predicted sensory consequences derived from efferent copies of motor commands for motor control @Duhamel1992a.
-When training with rotated visual feedback of the hand, we update these predictions @Wolpert1995.
-Additionally, such training leads to a recalibration of our sense of felt hand position - “proprioception” - to be more aligned with the distorted visual feedback @Cressman2009.
-Both of these changes have been measured by asking people to localize where their unseen hand is – before and after training [@Clayton2013; @Izawa2012b; @Jones2012a; @Synofzik2008; @Yavari2016].
-While our lab has previously found that proprioception accounts for a large portion of the change in hand localization @Hart2016, it is far from clear how much each process contributes or how to tease them apart.
-Here we use passive training, that removes the need to update predicted sensory consequences, in an attempt to isolate the contribution of proprioception to hand localization.
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
 
-The predicted sensory consequences of movements may play several important roles in motor learning and control.
-Predicted sensory consequences allow us to correct our movements before sensory error signals are available, they can be used to select movements that best achieve our goals and they may inform us on the location of our limbs.
-Hence measuring predicted sensory consequences is valuable in movement research.
-In visuomotor rotation adaptation tasks, the actual sensory outcome is systematically different from the expected outcome, so that participants update their predictions on the outcome of the trained movements.
-In previous experiments, people were asked to make a movement and then indicate the location of, or “localize,” their unseen, right hand, before and after training with rotated visual feedback [@Izawa2012b; @Synofzik2008].
-Since there was no visual information available to the participants, the predicted sensory consequences of the movement should be used in localizing the unseen hand.
-Both studies found a significant shift in hand localization, providing evidence that predicted sensory consequences are indeed updated as a result of visuomotor rotation adaptation.
+% cite package, to clean up citations in the main text. Do not remove.
+% \usepackage{cite}
 
-However, our lab has shown that our sense of where we feel our hand to be, proprioception, is also reliably recalibrated after visuomotor rotation adaptation [@Cressman2009; @Hart2016; @Mostafa2014; @Mostafa2015; @Maksimovic2018; @Henriques2012; @Cressman2010a; @Barkley2014; @Salomonczyk2011; @Nourouzpour2014; @Henriques2014].
-This has also been shown by other labs @Cameron2012 and a comparable proprioceptive change is induced with force-field adaptation @Ostry2016.
-As proprioception also informs us on the location of our limbs, we have on occasion used a task that is very similar to hand localization to investigate this [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018].
-Although proprioceptive recalibration has been largely ignored as an explanation for changes in hand localization, we and others have shown that it accounts for a substantial part of the changes in localization, along with updates in predicted sensory consequences [@Hart2016; @Cameron2012].
-Nevertheless, it is far from clear how much proprioception and prediction each contribute to hand localization.
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
 
-Here we intend to further examine the contribution of proprioception to hand localization.
-To do this, we use passive training, where a robot arm moves the participant’s arm out, so that the cursor always directly goes to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013].
-This means there is no efference copy available and no visuomotor error-signal, both of which are required to update predicted sensory consequences.
-However, we impose a discrepancy between vision and proprioception that drives proprioceptive recalibration.
-Thus, changes in hand localization after this type of training should be due to proprioceptive recalibration only.
-We use the same experimental protocol as before @Hart2016, so that we can compare localization shifts between the two different types of training, and can better assess the contributions of predicted sensory consequences and proprioception to hand localization.
+% line numbers
+\usepackage[right]{lineno}
 
-# Methods
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
 
-We set out to test the relative contributions of proprioception and efference-based prediction to hand localization.
-We use visual training with robot-generated hand movements to prevent updates of predicted sensory consequences, but still elicit proprioceptive recalibration.
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
 
-## Participants
+% array package and thick rules for tables
+\usepackage{array}
 
-Twenty-five right-handed participants were recruited for this study.
-One participant was excluded for not following task instructions, and three were excluded for low performance on a task that ensures attention during the passive training.
-All analyses presented here pertain to the remaining twenty-one participants (13 females and 8 males, mean age: 20.1 +/- 2.3 years), but the data of the three low-performing participants is included in the online \href{https://osf.io/zfdth}{dataset} @Mostafa2017.
-All had normal or corrected-to-normal vision, and provided prior, written informed consent in accordance with the ethical guidelines set by the York Human subjects Review Subcommittee and received credit toward an undergraduate psychology course.
-Participants were screened verbally, and all reported being right handed and not having any history of visual, neurological, and/or motor dysfunction.
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
 
-![**Setup and tasks:** **a)** Participants moved their unseen right hand with visual feedback on hand position provided through a mirror (middle surface) half-way between their hand and the monitor (top surface).
-A touchscreen located just above the hand was used to collect responses for the localization tasks (bottom surface).
-**b)** Training task.
-The target, shown as a yellow disc, is located 10 cm away from the home position at 45°.
-In the rotated training tasks, the cursor (shown here as a green circle) represents the hand position rotated 30° relative to the home position.
-**c)** No-cursor reach task.
-Targets are located 10 cm away from the home position at 15°, 25°, 35°, 45°, 55°, 65°, and 75°, shown by the yellow circles here (only one was shown on each trial).
-While reaching to one of these targets, no visual feedback on hand position is provided.
-d) Localization task.
-The participants’ unseen, right hands moved out, and subsequently participants indicated the direction of the hand movement by indicating a location on an arc using a touch screen with their visible left index finger](Fig1_setup.eps)
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
 
-## Setup
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
 
-Participants sat in a height-adjustable chair to ensure that they could easily see and reach all targets presented on a reflective screen (see Fig 1).
-During all tasks, they held the vertical handle on a two-joint robot manipulandum (Interactive Motion Technologies Inc., Cambridge, MA) with their right hand so that their thumb rested on top of the handle.
-A monitor (Samsung 510 N, refresh rate 60 Hz) was mounted 11 cm above the reflective screen, such that images displayed on the monitor appeared to lie in the horizontal plane where the right hand was moving.
-The reflective screen was mounted horizontally 18 cm above the robot manipulandum.
-A touch screen was mounted 13 cm underneath the reflective surface, so that subjects could indicate the location of the unseen right-hand locations (specifically the unseen thumb) with their visible left hand, which was lit up with a small spot light (only in localization tasks).
-The room lights were dimmed and the participants’ view of their right hand was blocked by the reflective screen, as well as a dark cloth draped between the touch screen and participants’ right shoulders.
+% Remove comment for double spacing
+%\usepackage{setspace}
+%\doublespacing
 
-## Procedure
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in
+\textheight 8.75in
 
-The first part of the experiment used training with a cursor aligned with the hand and the second part had training with a cursor rotated around the start position (Fig 1b; white rows in Table 1).
-During the training with rotated feedback, the cursor was gradually rotated 30° clockwise.
-This introduced a discrepancy between the actual, or felt, hand position and the visual feedback, that should evoke proprioceptive recalibration.
-However, the movements were robot generated, so that there were no predicted sensory consequences based on the outgoing motor command.
-Hence the prediction errors that are thought to lead to motor learning were absent as well.
-After both types of training, participants did open-loop reaches as well as two kinds of hand localization tasks, to test the effect of training on proprioceptive and predictive hand estimates.
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
 
-$
+% Use the PLoS provided BiBTeX style
+% \bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+\setlength{\headheight}{27.023pt}
+\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
+
+%% Include all macros below
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
+
+\usepackage{threeparttable}
+\usepackage[table]{xcolor}
+\usepackage{graphicx}
+\usepackage{tabularx}
+\usepackage{longtable}
+\usepackage{siunitx}
+
+\usepackage{forarray}
+\usepackage{xstring}
+\newcommand{\getIndex}[2]{
+  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
+}
+
+\setcounter{secnumdepth}{0}
+
+\newcommand{\getAff}[1]{
+  \getIndex{#1}{YORK,MANSOURA}
+}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\begin{document}
+\vspace*{0.2in}
+
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{Motor Learning Without Moving: Proprioceptive and Predictive Hand
+Localization After Passive Visuoproprioceptive Discrepancy Training} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+}
+\newline
+\\
+Ahmed A. Mostafa\textsuperscript{\getAff{YORK}, \getAff{MANSOURA}},
+Bernard Marius 't Hart\textsuperscript{\getAff{YORK}}\textsuperscript{*},
+Denise Y.P. Henriques\textsuperscript{\getAff{YORK}}\\
+\bigskip
+\textbf{\getAff{YORK}}CVR / Kinesiology and Health Science, York University, Toronto, Ontario,
+Canada\\
+\textbf{\getAff{MANSOURA}}Faculty of Physical Education, Mansoura University, Mansoura, Egypt\\
+\bigskip
+* Corresponding author: thartbm@gmail.com\\
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
+An accurate estimate of limb position is necessary for movement
+planning. Where we localize our unseen hand after a reach depends on
+felt hand position, or proprioception, but this is usually neglected in
+favour of predicted sensory consequences based on efference copies of
+motor commands. Both sources of information should contribute, so here
+we set out to further investigate how much of hand localization depends
+on proprioception and how much on predicted sensory consequences. We use
+a passive training paradigm with rotated visual feedback that eliminates
+the possibility to update predicted sensory consequences, but still
+recalibrates proprioception. After this training we measure
+participants' hand location estimates based on both efference-based
+predictions and afferent proprioceptive signals with self-generated hand
+movements as well as based on proprioception only with robot-generated
+movements. We find indistinguishable shifts in hand localization after
+robot- and self-generated hand movements, and changes in open-loop
+reaches. Both motor and proprioceptive changes are only slightly smaller
+as those after training with self-generated movements, confirming that
+proprioception plays a large role in estimating limb position and in
+planning movements. (data: https://doi.org/10.17605/osf.io/zfdth,
+preprint: https://doi.org/10.1101/384941)
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step.
+% Author Summary not valid for PLOS ONE submissions.
+
+\linenumbers
+
+% Use "Eq" instead of "Equation" for equation citations.
+\section{Introduction}\label{introduction}
+
+Sensory information is central to how we control all our movements. Our
+brain is even thought to use predicted sensory consequences derived from
+efferent copies of motor commands for motor control {[}1{]}. When
+training with rotated visual feedback of the hand, we update these
+predictions {[}2{]}. Additionally, such training leads to a
+recalibration of our sense of felt hand position - ``proprioception'' -
+to be more aligned with the distorted visual feedback {[}3{]}. Both of
+these changes have been measured by asking people to localize where
+their unseen hand is -- before and after training {[}4--8{]}. While our
+lab has previously found that proprioception accounts for a large
+portion of the change in hand localization {[}9{]}, it is far from clear
+how much each process contributes or how to tease them apart. Here we
+use passive training, that removes the need to update predicted sensory
+consequences, in an attempt to isolate the contribution of
+proprioception to hand localization.
+
+The predicted sensory consequences of movements may play several
+important roles in motor learning and control. Predicted sensory
+consequences allow us to correct our movements before sensory error
+signals are available, they can be used to select movements that best
+achieve our goals and they may inform us on the location of our limbs.
+Hence measuring predicted sensory consequences is valuable in movement
+research. In visuomotor rotation adaptation tasks, the actual sensory
+outcome is systematically different from the expected outcome, so that
+participants update their predictions on the outcome of the trained
+movements. In previous experiments, people were asked to make a movement
+and then indicate the location of, or ``localize,'' their unseen, right
+hand, before and after training with rotated visual feedback {[}5,7{]}.
+Since there was no visual information available to the participants, the
+predicted sensory consequences of the movement should be used in
+localizing the unseen hand. Both studies found a significant shift in
+hand localization, providing evidence that predicted sensory
+consequences are indeed updated as a result of visuomotor rotation
+adaptation.
+
+However, our lab has shown that our sense of where we feel our hand to
+be, proprioception, is also reliably recalibrated after visuomotor
+rotation adaptation {[}3,9--18{]}. This has also been shown by other
+labs {[}19{]} and a comparable proprioceptive change is induced with
+force-field adaptation {[}20{]}. As proprioception also informs us on
+the location of our limbs, we have on occasion used a task that is very
+similar to hand localization to investigate this {[}3,6,21,22{]}.
+Although proprioceptive recalibration has been largely ignored as an
+explanation for changes in hand localization, we and others have shown
+that it accounts for a substantial part of the changes in localization,
+along with updates in predicted sensory consequences {[}9,19{]}.
+Nevertheless, it is far from clear how much proprioception and
+prediction each contribute to hand localization.
+
+Here we intend to further examine the contribution of proprioception to
+hand localization. To do this, we use passive training, where a robot
+arm moves the participant's arm out, so that the cursor always directly
+goes to the target {[}19,23,24{]}. This means there is no efference copy
+available and no visuomotor error-signal, both of which are required to
+update predicted sensory consequences. However, we impose a discrepancy
+between vision and proprioception that drives proprioceptive
+recalibration. Thus, changes in hand localization after this type of
+training should be due to proprioceptive recalibration only. We use the
+same experimental protocol as before {[}9{]}, so that we can compare
+localization shifts between the two different types of training, and can
+better assess the contributions of predicted sensory consequences and
+proprioception to hand localization.
+
+\section{Methods}\label{methods}
+
+We set out to test the relative contributions of proprioception and
+efference-based prediction to hand localization. We use visual training
+with robot-generated hand movements to prevent updates of predicted
+sensory consequences, but still elicit proprioceptive recalibration.
+
+\subsection{Participants}\label{participants}
+
+Twenty-five right-handed participants were recruited for this study. One
+participant was excluded for not following task instructions, and three
+were excluded for low performance on a task that ensures attention
+during the passive training. All analyses presented here pertain to the
+remaining twenty-one participants (13 females and 8 males, mean age:
+20.1 +/- 2.3 years), but the data of the three low-performing
+participants is included in the online
+\href{https://osf.io/zfdth}{dataset} {[}25{]}. All had normal or
+corrected-to-normal vision, and provided prior, written informed consent
+in accordance with the ethical guidelines set by the York Human subjects
+Review Subcommittee and received credit toward an undergraduate
+psychology course. Participants were screened verbally, and all reported
+being right handed and not having any history of visual, neurological,
+and/or motor dysfunction.
+
+\begin{figure}
+\centering
+\includegraphics{Fig1_setup.eps}
+\caption{\textbf{Setup and tasks:} \textbf{a)} Participants moved their
+unseen right hand with visual feedback on hand position provided through
+a mirror (middle surface) half-way between their hand and the monitor
+(top surface). A touchscreen located just above the hand was used to
+collect responses for the localization tasks (bottom surface).
+\textbf{b)} Training task. The target, shown as a yellow disc, is
+located 10 cm away from the home position at 45°. In the rotated
+training tasks, the cursor (shown here as a green circle) represents the
+hand position rotated 30° relative to the home position. \textbf{c)}
+No-cursor reach task. Targets are located 10 cm away from the home
+position at 15°, 25°, 35°, 45°, 55°, 65°, and 75°, shown by the yellow
+circles here (only one was shown on each trial). While reaching to one
+of these targets, no visual feedback on hand position is provided. d)
+Localization task. The participants' unseen, right hands moved out, and
+subsequently participants indicated the direction of the hand movement
+by indicating a location on an arc using a touch screen with their
+visible left index finger}
+\end{figure}
+
+\subsection{Setup}\label{setup}
+
+Participants sat in a height-adjustable chair to ensure that they could
+easily see and reach all targets presented on a reflective screen (see
+Fig 1). During all tasks, they held the vertical handle on a two-joint
+robot manipulandum (Interactive Motion Technologies Inc., Cambridge, MA)
+with their right hand so that their thumb rested on top of the handle. A
+monitor (Samsung 510 N, refresh rate 60 Hz) was mounted 11 cm above the
+reflective screen, such that images displayed on the monitor appeared to
+lie in the horizontal plane where the right hand was moving. The
+reflective screen was mounted horizontally 18 cm above the robot
+manipulandum. A touch screen was mounted 13 cm underneath the reflective
+surface, so that subjects could indicate the location of the unseen
+right-hand locations (specifically the unseen thumb) with their visible
+left hand, which was lit up with a small spot light (only in
+localization tasks). The room lights were dimmed and the participants'
+view of their right hand was blocked by the reflective screen, as well
+as a dark cloth draped between the touch screen and participants' right
+shoulders.
+
+\subsection{Procedure}\label{procedure}
+
+The first part of the experiment used training with a cursor aligned
+with the hand and the second part had training with a cursor rotated
+around the start position (Fig 1b; white rows in Table 1). During the
+training with rotated feedback, the cursor was gradually rotated 30°
+clockwise. This introduced a discrepancy between the actual, or felt,
+hand position and the visual feedback, that should evoke proprioceptive
+recalibration. However, the movements were robot generated, so that
+there were no predicted sensory consequences based on the outgoing motor
+command. Hence the prediction errors that are thought to lead to motor
+learning were absent as well. After both types of training, participants
+did open-loop reaches as well as two kinds of hand localization tasks,
+to test the effect of training on proprioceptive and predictive hand
+estimates.
+
+\$
 
@@
 
-\definecolor{LGray}{gray}{0.9}
-\definecolor{DGray}{gray}{0.8}
+\definecolor{LGray}{gray}{0.9} \definecolor{DGray}{gray}{0.8}
 
@@
 \end{table}
-$
 
-## Exposure training
+\$
 
-In what we call ‘exposure training’ the participants did not move their hand toward the target, but the robot did.
-In this task (Table 1), the right hand was represented by a cursor (green disk, 1 cm in diameter, Fig 1b) located directly above participant’s thumb.
-The robot moved the participant’s unseen right hand (and the cursor) along a direct path toward a visual yellow target disk and back to the starting position (1 cm in diameter, Fig 1b).
-The home position was located approximately 20 cm in front of participants and the visual target located 10 cm from the home position at 45° (Fig 1b).
-In order to make sure participants were paying attention to the cursor, the cursor was switched off for 2 screen refreshes (~33.3ms) on 50\% of the trials at a random distance between 4 and 9 cm from the home position and participants were asked to report this using a button press with the left hand.
+\subsection{Exposure training}\label{exposure-training}
+
+In what we call `exposure training' the participants did not move their
+hand toward the target, but the robot did. In this task (Table 1), the
+right hand was represented by a cursor (green disk, 1 cm in diameter,
+Fig 1b) located directly above participant's thumb. The robot moved the
+participant's unseen right hand (and the cursor) along a direct path
+toward a visual yellow target disk and back to the starting position (1
+cm in diameter, Fig 1b). The home position was located approximately 20
+cm in front of participants and the visual target located 10 cm from the
+home position at 45° (Fig 1b). In order to make sure participants were
+paying attention to the cursor, the cursor was switched off for 2 screen
+refreshes (\textasciitilde{}33.3ms) on 50\% of the trials at a random
+distance between 4 and 9 cm from the home position and participants were
+asked to report this using a button press with the left hand.
 Performance on this task was used to screen participants.
 
-During the first half of the experiment, the cursor and hand path were aligned during exposure training.
-In the second part of the experiment, the “rotated” session, the same visual training target at 45° was used, and the cursor kept moving straight to this target.
-However, the robot-generated hand path gradually rotated 30° CCW (Fig 1b) with respect to the visible target and the cursor in increments of 0.75°/trial, so that the full rotation was reached after 45 trials.
-This mimicked error-free responses to a gradual visuomotor rotation of 30° CW.
-The initial training consisted of 50 trials in the aligned part and 90 in the rotated part.
-In between open-loop reach tasks and localization tasks (Table 1) extra training tasks were done, each of which consisted of 10 trials in the aligned part of the experiment and 60 trials in the rotated part (Table 1).
+During the first half of the experiment, the cursor and hand path were
+aligned during exposure training. In the second part of the experiment,
+the ``rotated'' session, the same visual training target at 45° was
+used, and the cursor kept moving straight to this target. However, the
+robot-generated hand path gradually rotated 30° CCW (Fig 1b) with
+respect to the visible target and the cursor in increments of
+0.75°/trial, so that the full rotation was reached after 45 trials. This
+mimicked error-free responses to a gradual visuomotor rotation of 30°
+CW. The initial training consisted of 50 trials in the aligned part and
+90 in the rotated part. In between open-loop reach tasks and
+localization tasks (Table 1) extra training tasks were done, each of
+which consisted of 10 trials in the aligned part of the experiment and
+60 trials in the rotated part (Table 1).
 
-## No-cursor reaching
+\subsection{No-cursor reaching}\label{no-cursor-reaching}
 
-The trials in no-cursor reaching (Table 1, light gray rows) serve as a classical measure of motor adaptation.
-On each of these trials participants were asked to reach with their unseen right hand to one of 7 visual targets, without any visual feedback of hand position.
-The targets were 10 cm from the home position, located radially at: 15°, 25°, 35°, 45°, 55°, 65°, and 75° (Fig 1c).
-A trial started with the robot handle at the home position and, after 500 ms, the home position disappeared and the target appeared, cuing the participants to reach for the target.
-Once the participants thought they had reached the target they held their position for 250 ms, and the target and the home position disappeared, cuing participants to move back to the home position along a straight, constrained path, to begin the next trial.
-If participants tried to move outside of the path, a resistance force, with a stiffness of 2 N/(mm/s) and a viscous damping of 5 N/(mm/s), was generated perpendicular to the path.
-In every iteration of the no-cursor reach task, each target was reached three times, for a total of 21 reaches in pseudo-random order.
-The no-cursor reaching task was performed four times in the aligned part of the experiment and five times in the rotated part of the experiment.
+The trials in no-cursor reaching (Table 1, light gray rows) serve as a
+classical measure of motor adaptation. On each of these trials
+participants were asked to reach with their unseen right hand to one of
+7 visual targets, without any visual feedback of hand position. The
+targets were 10 cm from the home position, located radially at: 15°,
+25°, 35°, 45°, 55°, 65°, and 75° (Fig 1c). A trial started with the
+robot handle at the home position and, after 500 ms, the home position
+disappeared and the target appeared, cuing the participants to reach for
+the target. Once the participants thought they had reached the target
+they held their position for 250 ms, and the target and the home
+position disappeared, cuing participants to move back to the home
+position along a straight, constrained path, to begin the next trial. If
+participants tried to move outside of the path, a resistance force, with
+a stiffness of 2 N/(mm/s) and a viscous damping of 5 N/(mm/s), was
+generated perpendicular to the path. In every iteration of the no-cursor
+reach task, each target was reached three times, for a total of 21
+reaches in pseudo-random order. The no-cursor reaching task was
+performed four times in the aligned part of the experiment and five
+times in the rotated part of the experiment.
 
-## Localization
+\subsection{Localization}\label{localization}
 
-In this task (Fig 1d; Table 1, dark gray rows) participants indicated where they thought their unseen right hand was after a movement.
-First, an arc appeared, spanning from 0° to 90° and located 10 cm away from the home position and the participants’ unseen, right hand moved out from the home position in a direction towards towards a point on the arc.
-The hand was stopped by the robot at 10 cm from the home position and then, to prevent the online proprioceptive signals from overriding the predictive signals [@Izawa2012b; @Hart2016], the hand was moved back to the home position using the same kind of constrained path as used for the return movements in the no-cursor task.
-Participants indicated with the index finger of their visible, left hand on the touch screen mounted directly above the robot handle where they thought their trained hand had crossed the arc.
+In this task (Fig 1d; Table 1, dark gray rows) participants indicated
+where they thought their unseen right hand was after a movement. First,
+an arc appeared, spanning from 0° to 90° and located 10 cm away from the
+home position and the participants' unseen, right hand moved out from
+the home position in a direction towards towards a point on the arc. The
+hand was stopped by the robot at 10 cm from the home position and then,
+to prevent the online proprioceptive signals from overriding the
+predictive signals {[}5,9{]}, the hand was moved back to the home
+position using the same kind of constrained path as used for the return
+movements in the no-cursor task. Participants indicated with the index
+finger of their visible, left hand on the touch screen mounted directly
+above the robot handle where they thought their trained hand had crossed
+the arc.
 
-Crucially, there were two variations of this task.
-First, in the ‘active’ localization task participants generated the movement themselves, as they could freely move their unseen right hand from the home position to any point on the arc.
-Second, there was a ‘passive’ localization task where the robot moved the participants’ hand out and back, to the same locations the participants moved to in the preceding ‘active’ localization task in a shuffled order (hence, active localization is done first).
-In active localization, participants have access to both proprioceptive information as well as an efference-based prediction of sensory consequences, but in passive localization, only proprioception should be available.
-The active and passive localization task each consisted of 25 trials, and each of the tasks was done a total of four times; twice after aligned and twice after rotated training.
+Crucially, there were two variations of this task. First, in the
+`active' localization task participants generated the movement
+themselves, as they could freely move their unseen right hand from the
+home position to any point on the arc. Second, there was a `passive'
+localization task where the robot moved the participants' hand out and
+back, to the same locations the participants moved to in the preceding
+`active' localization task in a shuffled order (hence, active
+localization is done first). In active localization, participants have
+access to both proprioceptive information as well as an efference-based
+prediction of sensory consequences, but in passive localization, only
+proprioception should be available. The active and passive localization
+task each consisted of 25 trials, and each of the tasks was done a total
+of four times; twice after aligned and twice after rotated training.
 
-## Classic training
+\subsection{Classic training}\label{classic-training}
 
-The paradigm described above is an exact replica of a paradigm we used earlier @Hart2016 with two exceptions.
-First, we used exposure training here, instead of the standard reach training with volitional movements, which we will call ‘classic’ training.
-Second, all localization is delayed until the right hand has returned to the home position in this study (see Table 1), so that instead of both delayed and online localization we have two repetitions of each delayed localization task.
-With this paradigm we can compare changes in localization and no-cursor reaches change after exposure training with changes in the same measures after classic visuomotor adaptation training.
+The paradigm described above is an exact replica of a paradigm we used
+earlier {[}9{]} with two exceptions. First, we used exposure training
+here, instead of the standard reach training with volitional movements,
+which we will call `classic' training. Second, all localization is
+delayed until the right hand has returned to the home position in this
+study (see Table 1), so that instead of both delayed and online
+localization we have two repetitions of each delayed localization task.
+With this paradigm we can compare changes in localization and no-cursor
+reaches change after exposure training with changes in the same measures
+after classic visuomotor adaptation training.
 
-## Analyses
+\subsection{Analyses}\label{analyses}
 
-Prior to any analyses, both the localization responses and the no-cursor reach data were visually inspected and trials where the participants did not follow task instruction were removed (e.g.
-several movements back and forth, or a touch-screen response on the home position, instead of on the white arc).
+Prior to any analyses, both the localization responses and the no-cursor
+reach data were visually inspected and trials where the participants did
+not follow task instruction were removed (e.g.~several movements back
+and forth, or a touch-screen response on the home position, instead of
+on the white arc).
 
-## Localization
+\subsection{Localization}\label{localization-1}
 
-Localization responses were taken as the (signed) angular difference between vectors through the home position and the actual hand position as well as the location indicated on the touch screen.
-Prior to analyses, idiosyncratic differences in performing this task were countered.
-Before conversion to degrees angle, a circle with a 10 cm radius was fit to the touch screen responses of each participant and the offset of this circle’s centre was subtracted from all response coordinates, so that all responses fell close to the arc.
-Then, a smoothed spline was fit to every participant’s response errors in each of the four localization tasks (aligned vs.
-rotated and active vs.
-passive) and these were used to obtain localization errors at the same locations used for the no-cursor reaches (15°, 25°, 35°, 45°, 55°, 65° and 75°), but only if that location fell within the range of the data (i.e.
-we only interpolate).
-This way localization responses could be compared across participants despite the freely chosen reach directions.
-At the 15° location 7/21 participants didn’t have an estimate in one or more of the four localization tasks (in the “classic” data it was also 7/21).
-While that data is shown in the figures, we did not use it for analysis.
+Localization responses were taken as the (signed) angular difference
+between vectors through the home position and the actual hand position
+as well as the location indicated on the touch screen. Prior to
+analyses, idiosyncratic differences in performing this task were
+countered. Before conversion to degrees angle, a circle with a 10 cm
+radius was fit to the touch screen responses of each participant and the
+offset of this circle's centre was subtracted from all response
+coordinates, so that all responses fell close to the arc. Then, a
+smoothed spline was fit to every participant's response errors in each
+of the four localization tasks (aligned vs.~rotated and active
+vs.~passive) and these were used to obtain localization errors at the
+same locations used for the no-cursor reaches (15°, 25°, 35°, 45°, 55°,
+65° and 75°), but only if that location fell within the range of the
+data (i.e.~we only interpolate). This way localization responses could
+be compared across participants despite the freely chosen reach
+directions. At the 15° location 7/21 participants didn't have an
+estimate in one or more of the four localization tasks (in the
+``classic'' data it was also 7/21). While that data is shown in the
+figures, we did not use it for analysis.
 
-First we test if localization responses shifted following rotated exposure training compared to aligned.
-We then test if the shift in localization responses is different for active and passive localization, and we run analyses comparing localization after exposure training with localization after “classic” training.
-Finally, we explore the generalization of localization responses and if they are different between the groups doing classic and exposure training.
+First we test if localization responses shifted following rotated
+exposure training compared to aligned. We then test if the shift in
+localization responses is different for active and passive localization,
+and we run analyses comparing localization after exposure training with
+localization after ``classic'' training. Finally, we explore the
+generalization of localization responses and if they are different
+between the groups doing classic and exposure training.
 
-## Reach aftereffects
+\subsection{Reach aftereffects}\label{reach-aftereffects}
 
-To assess any reach adaptation that may have occurred after exposure training we analyzed reach endpoint errors in no-cursor trials.
-Reach endpoint errors were the (signed) angular difference between a vector from the home position to reach endpoint and a vector from the home position to the target.
-We obtained reach aftereffects by subtracting reach endpoint errors after aligned training from those after rotated training.
-No-cursor endpoint errors were analyzed to see if participants adapted the direction of their reaching movements after rotated exposure training.
-We also tested if any such change decayed, i.e.
-if it was the same immediately after exposure training, or when a localization task was done in between exposure training and no-cursor reaches.
-Furthermore, we tested if the generalization of reach aftereffects is different between exposure and classic training and if there is any generalization of reach aftereffects.
+To assess any reach adaptation that may have occurred after exposure
+training we analyzed reach endpoint errors in no-cursor trials. Reach
+endpoint errors were the (signed) angular difference between a vector
+from the home position to reach endpoint and a vector from the home
+position to the target. We obtained reach aftereffects by subtracting
+reach endpoint errors after aligned training from those after rotated
+training. No-cursor endpoint errors were analyzed to see if participants
+adapted the direction of their reaching movements after rotated exposure
+training. We also tested if any such change decayed, i.e.~if it was the
+same immediately after exposure training, or when a localization task
+was done in between exposure training and no-cursor reaches.
+Furthermore, we tested if the generalization of reach aftereffects is
+different between exposure and classic training and if there is any
+generalization of reach aftereffects.
 
-Pre-processing and analyses were done in R 3.4.4 @RCoreTeam2017, using lme4, lmerTest and various other packages.
-Most analyses used linear mixed effects models, since there is some missing localization data.
-These were “converted” to more readable ANOVA-like output, using a Satterthwaite approximation @Luke2017.
-Highly similar results were obtained with a Chi-square approximation.
-Data, scripts and a notebook with analyses have been made available on the Open Science Framework @Mostafa2017 ([https://osf.io/zfdth](https://osf.io/zfdth/)).
+Pre-processing and analyses were done in R 3.4.4 {[}26{]}, using lme4,
+lmerTest and various other packages. Most analyses used linear mixed
+effects models, since there is some missing localization data. These
+were ``converted'' to more readable ANOVA-like output, using a
+Satterthwaite approximation {[}27{]}. Highly similar results were
+obtained with a Chi-square approximation. Data, scripts and a notebook
+with analyses have been made available on the Open Science Framework
+{[}25{]} (\href{https://osf.io/zfdth/}{https://osf.io/zfdth}).
 
-In short, this experiment allowed us to test how mere exposure to a visual-proprioceptive discrepancy changes both reach aftereffects and hand localization responses, and compare them with those obtained after more regular, “classic” training.
+In short, this experiment allowed us to test how mere exposure to a
+visual-proprioceptive discrepancy changes both reach aftereffects and
+hand localization responses, and compare them with those obtained after
+more regular, ``classic'' training.
 
-# Results
+\section{Results}\label{results}
 
-In this study we intend to further elucidate the relative contributions of (updated) predicted sensory consequences and (recalibrated) proprioception to hand localization.
-We can parcel out these contributions by measuring hand localization after both robot-generated and self-generated movements.
-Finally we compare the data from the current experiment with those obtained in an earlier study that used an identical paradigm, but trained with self-generated movements, or “classic” training.
+In this study we intend to further elucidate the relative contributions
+of (updated) predicted sensory consequences and (recalibrated)
+proprioception to hand localization. We can parcel out these
+contributions by measuring hand localization after both robot-generated
+and self-generated movements. Finally we compare the data from the
+current experiment with those obtained in an earlier study that used an
+identical paradigm, but trained with self-generated movements, or
+``classic'' training.
 
-![**Hand localization:** The shifts of the angles of touchscreen responses in all variations of the localization task, using spline-interpolated estimates for hand angles matching the reach targets in the no-cursor reach block.
-**a)** Localization shifts after exposure training.
-Dark blue: active localization shifts, Light blue: passive localization shifts.
-**b)** Localization shifts after classic training.
-Dark red: active localization shifts, Orange: passive localization shifts.
-The dashed line segments illustrate that the 15° data is not used for statistical analyses (see Methods).
-**c)** Generalization curves of active localization shifts after exposure training (blue) and classic training (red).
-Shaded areas: 95\% confidence intervals for the peak of the generalization curve (red and blue lines through shaded area indicate 50\% points).
-Downward black arrow: visual trained target.
-Upward black arrow: hand location during training.](Fig2_plain.pdf)
+\begin{figure}
+\centering
+\includegraphics{Fig2_plain.pdf}
+\caption{\textbf{Hand localization:} The shifts of the angles of
+touchscreen responses in all variations of the localization task, using
+spline-interpolated estimates for hand angles matching the reach targets
+in the no-cursor reach block. \textbf{a)} Localization shifts after
+exposure training. Dark blue: active localization shifts, Light blue:
+passive localization shifts. \textbf{b)} Localization shifts after
+classic training. Dark red: active localization shifts, Orange: passive
+localization shifts. The dashed line segments illustrate that the 15°
+data is not used for statistical analyses (see Methods). \textbf{c)}
+Generalization curves of active localization shifts after exposure
+training (blue) and classic training (red). Shaded areas: 95\%
+confidence intervals for the peak of the generalization curve (red and
+blue lines through shaded area indicate 50\% points). Downward black
+arrow: visual trained target. Upward black arrow: hand location during
+training.}
+\end{figure}
 
-## Localization
+\subsection{Localization}\label{localization-2}
 
-Here we test our hypothesis that exposure training does not lead to changes in predicted sensory consequences.
-Since the difference between active and passive localization stems only from the presence or absence of predicted sensory consequences, there should be no difference between the two if predicted sensory consequences are not changed by exposure training.
-At first glance, it seems there might be a difference between active and passive localization shifts in exposure training (see Fig 2a), although it is smaller than in classic training (Fig 2b).
+Here we test our hypothesis that exposure training does not lead to
+changes in predicted sensory consequences. Since the difference between
+active and passive localization stems only from the presence or absence
+of predicted sensory consequences, there should be no difference between
+the two if predicted sensory consequences are not changed by exposure
+training. At first glance, it seems there might be a difference between
+active and passive localization shifts in exposure training (see Fig
+2a), although it is smaller than in classic training (Fig 2b).
 
-To test if rotated exposure training induces changes in hand localization, we fit an LME to the localization errors throughout the workspace using session (aligned or rotated), movement type (active and passive) and hand angle (25°, 35°, 45°, 55°, 65° and 75°), and all interactions as fixed effects and participant as random effect.
-There was an effect of session (F(1,450.5)=155.8; p<0.001), showing that exposure training leads to changes in hand localization.
-There was also an effect of hand angle (F(5,450.6)=6.54; p<.001) and an interaction between hand angle and session (F(5,450.3)=8.25; p<.001), which we’ll explore below, but no other effects (all p>.60).
-Since localization responses did shift, we use the difference between hand localization after rotated training and after aligned training (as plotted in Fig 2) for further tests.
+To test if rotated exposure training induces changes in hand
+localization, we fit an LME to the localization errors throughout the
+workspace using session (aligned or rotated), movement type (active and
+passive) and hand angle (25°, 35°, 45°, 55°, 65° and 75°), and all
+interactions as fixed effects and participant as random effect. There
+was an effect of session (F(1,450.5)=155.8; p\textless{}0.001), showing
+that exposure training leads to changes in hand localization. There was
+also an effect of hand angle (F(5,450.6)=6.54; p\textless{}.001) and an
+interaction between hand angle and session (F(5,450.3)=8.25;
+p\textless{}.001), which we'll explore below, but no other effects (all
+p\textgreater{}.60). Since localization responses did shift, we use the
+difference between hand localization after rotated training and after
+aligned training (as plotted in Fig 2) for further tests.
 
-If this shift in localization after exposure training partly reflects predicted sensory consequences, then shifts in active localization, that rely on both (recalibrated) proprioception and (updated) predictions should be different from shifts in passive localization that only rely on (recalibrated) proprioception.
-We fit an LME to the change in localization using movement type (active or passive localization) and hand angle, as well as their interaction as fixed effects and participant as random effect.
-There was no effect of movement type (F(1,211.8)=0.07; p=0.79).
-There was an effect of hand angle (F(5,212.2)=10.8; p<0.001), but no interaction between hand angle and movement type (F(5,211.8)=1.23, p=.29).
-The lack of an effect of movement type suggests that predicted sensory consequences did not contribute to localization in this paradigm.
+If this shift in localization after exposure training partly reflects
+predicted sensory consequences, then shifts in active localization, that
+rely on both (recalibrated) proprioception and (updated) predictions
+should be different from shifts in passive localization that only rely
+on (recalibrated) proprioception. We fit an LME to the change in
+localization using movement type (active or passive localization) and
+hand angle, as well as their interaction as fixed effects and
+participant as random effect. There was no effect of movement type
+(F(1,211.8)=0.07; p=0.79). There was an effect of hand angle
+(F(5,212.2)=10.8; p\textless{}0.001), but no interaction between hand
+angle and movement type (F(5,211.8)=1.23, p=.29). The lack of an effect
+of movement type suggests that predicted sensory consequences did not
+contribute to localization in this paradigm.
 
-In order to compare hand localization shifts after exposure training with those after classic training @Hart2016, we fit an LME to localization shift using training type (exposure vs.
-classic), movement type (active vs.
-passive) and hand angle and all interactions as fixed effects and participant as random effect.
-There was a main effect of movement type (F(1,422.0)=6.22; p=.013) and of hand angle (F(5,422.7)=8.19; p<.001), as well as an interaction between training type and hand angle (F(5,422.7)=4.54; p<.001) and between training type and movement type (F(1,422.0)=4.48, p=.035), but there was no main effect of training type (F(1,39.1)=0.92, p=.34) and no other effects (all p>.14).
-These results also suggest that the magnitude of the shifts in localization are comparable between classic and exposure training, but that the pattern of generalization is different.
+In order to compare hand localization shifts after exposure training
+with those after classic training {[}9{]}, we fit an LME to localization
+shift using training type (exposure vs.~classic), movement type (active
+vs.~passive) and hand angle and all interactions as fixed effects and
+participant as random effect. There was a main effect of movement type
+(F(1,422.0)=6.22; p=.013) and of hand angle (F(5,422.7)=8.19;
+p\textless{}.001), as well as an interaction between training type and
+hand angle (F(5,422.7)=4.54; p\textless{}.001) and between training type
+and movement type (F(1,422.0)=4.48, p=.035), but there was no main
+effect of training type (F(1,39.1)=0.92, p=.34) and no other effects
+(all p\textgreater{}.14). These results also suggest that the magnitude
+of the shifts in localization are comparable between classic and
+exposure training, but that the pattern of generalization is different.
 
-To address our main question, we will look at the interaction between training type (exposure vs.
-classic) and movement type (active vs.
-passive) we found above.
-Since there is no difference between active and passive localization shifts after exposure training alone, the interaction between training type and movement type should be caused by an effect of movement type on the localization shifts after classic training, as we found previously @Hart2016.
-This means that shifts in hand localization after exposure training indeed rely on recalibrated proprioception alone, while after classic training, there also is a contribution of predicted sensory consequences to active localization.
+To address our main question, we will look at the interaction between
+training type (exposure vs.~classic) and movement type (active
+vs.~passive) we found above. Since there is no difference between active
+and passive localization shifts after exposure training alone, the
+interaction between training type and movement type should be caused by
+an effect of movement type on the localization shifts after classic
+training, as we found previously {[}9{]}. This means that shifts in hand
+localization after exposure training indeed rely on recalibrated
+proprioception alone, while after classic training, there also is a
+contribution of predicted sensory consequences to active localization.
 
-For completeness, we explore the potentially different generalization patterns of localization shifts after classic and exposure training (Fig 2c).
-The LME indicates no difference in overall amplitude of localization shifts between the groups, so the interaction between training type and hand angle might stem from a generalization that does not peak at the trained location after exposure training.
-Using the active localization shifts only (which are larger, and arguably more similar to reach aftereffects), we bootstrap a 95\% confidence interval for the peak localization shift across participants in each group.
-Here we include the data at 15° where it is available.
-After classic training, the peak localization shift was at 48.8° (95\% confidence: 31.6° - 62.8°; red area in Fig 2c), and after exposure training the peak localization shift was at 62.1° (95\% confidence: 50.0° - 75.6°; blue are in Fig 2c).
-This means that peak localization after classic training is lower than after exposure training, but not vice versa.
-Also note that the confidence interval for the peak localization shift after classic training includes the trained target (45°), but not after exposure training @Izawa2012b.
-In short, the LME for localization shifts indicates a different generalization curve after exposure and classic training, which might be partially explained by a different position of the peak localization shift after exposure or classic training.
+For completeness, we explore the potentially different generalization
+patterns of localization shifts after classic and exposure training (Fig
+2c). The LME indicates no difference in overall amplitude of
+localization shifts between the groups, so the interaction between
+training type and hand angle might stem from a generalization that does
+not peak at the trained location after exposure training. Using the
+active localization shifts only (which are larger, and arguably more
+similar to reach aftereffects), we bootstrap a 95\% confidence interval
+for the peak localization shift across participants in each group. Here
+we include the data at 15° where it is available. After classic
+training, the peak localization shift was at 48.8° (95\% confidence:
+31.6° - 62.8°; red area in Fig 2c), and after exposure training the peak
+localization shift was at 62.1° (95\% confidence: 50.0° - 75.6°; blue
+are in Fig 2c). This means that peak localization after classic training
+is lower than after exposure training, but not vice versa. Also note
+that the confidence interval for the peak localization shift after
+classic training includes the trained target (45°), but not after
+exposure training {[}5{]}. In short, the LME for localization shifts
+indicates a different generalization curve after exposure and classic
+training, which might be partially explained by a different position of
+the peak localization shift after exposure or classic training.
 
-## Reach aftereffects
+\subsection{Reach aftereffects}\label{reach-aftereffects-1}
 
-Apart from proprioception and prediction, we want to see if rotated exposure training has any effect on open-loop reaches and if these are robust.
-We measure whether participants adapted their reach directions by assessing their reach errors in no-cursor reach trials after aligned and rotated exposure training.
-In Fig 3, the changes in no-cursor endpoint errors, or reach aftereffects, appear to be well over 5°.
-First, to test if exposure training affects open-loop reach direction, we fit a linear mixed effects model (LME) to reach endpoint error using session (aligned; all blocks, or rotated; only the first block immediately after training) and target (15°, 25°, 35°, 45°, 55°, 65° and 75°), as well as their interactions as fixed effects and participant as random effect.
-There is an effect of session (F(1,260)=93.81, p<.001), that is: exposure training leads to substantial reach aftereffects.
-There was no effect of target (F(6,260)=1.07, p=.37) and no interaction (F(6,260)=1.00, p=.42).
-Since there is an effect of session, we now take the differences in reach endpoint errors between the rotated and aligned session for every participant and target as reach aftereffects, and use those for all further analyses.
+Apart from proprioception and prediction, we want to see if rotated
+exposure training has any effect on open-loop reaches and if these are
+robust. We measure whether participants adapted their reach directions
+by assessing their reach errors in no-cursor reach trials after aligned
+and rotated exposure training. In Fig 3, the changes in no-cursor
+endpoint errors, or reach aftereffects, appear to be well over 5°.
+First, to test if exposure training affects open-loop reach direction,
+we fit a linear mixed effects model (LME) to reach endpoint error using
+session (aligned; all blocks, or rotated; only the first block
+immediately after training) and target (15°, 25°, 35°, 45°, 55°, 65° and
+75°), as well as their interactions as fixed effects and participant as
+random effect. There is an effect of session (F(1,260)=93.81,
+p\textless{}.001), that is: exposure training leads to substantial reach
+aftereffects. There was no effect of target (F(6,260)=1.07, p=.37) and
+no interaction (F(6,260)=1.00, p=.42). Since there is an effect of
+session, we now take the differences in reach endpoint errors between
+the rotated and aligned session for every participant and target as
+reach aftereffects, and use those for all further analyses.
 
-![**Reach aftereffects:** Changes of the angle of reach endpoints in the no-cursor tasks.
-**a)** Reach aftereffects across the experiment.
-Light blue: first no-cursor task in the rotated session (immediately following training), Dark blue, dashed line: the other four repetitions of the task (with localization in between training and no-cursor tasks).
-**b)** Reach aftereffects after classic and exposure training.
-Blue: exposure training, Red: classic training.
-**c)** Generalization curves of reach aftereffects after exposure training (blue) and classic training (red).
-Shaded areas: 95\% confidence intervals for the peak of the generalization curve (red and blue lines through shaded area indicate 50\% points).
-Downward black arrow: visual trained target.
-Upward black arrow: hand location during training.](Fig3_plain.pdf)
+\begin{figure}
+\centering
+\includegraphics{Fig3_plain.pdf}
+\caption{\textbf{Reach aftereffects:} Changes of the angle of reach
+endpoints in the no-cursor tasks. \textbf{a)} Reach aftereffects across
+the experiment. Light blue: first no-cursor task in the rotated session
+(immediately following training), Dark blue, dashed line: the other four
+repetitions of the task (with localization in between training and
+no-cursor tasks). \textbf{b)} Reach aftereffects after classic and
+exposure training. Blue: exposure training, Red: classic training.
+\textbf{c)} Generalization curves of reach aftereffects after exposure
+training (blue) and classic training (red). Shaded areas: 95\%
+confidence intervals for the peak of the generalization curve (red and
+blue lines through shaded area indicate 50\% points). Downward black
+arrow: visual trained target. Upward black arrow: hand location during
+training.}
+\end{figure}
 
-To see if reach aftereffects decayed during the localization tasks, we compared reach aftereffects in the initial no-cursor block, that immediately followed training, with those in the later blocks that followed a localization task.
-We fit an LME to the reach aftereffects with iteration (initial vs.
-later no-cursor blocks) and target (as above) as well as their interaction as fixed effects and participant as random effect.
-There is no effect of iteration (F(1,260)=2.72, p=0.10).
-There was an effect of target (F(6,260)=6.29, p<.001) but no interaction (F(6,260)=0.58, p=0.74).
-Hence, reach aftereffects were not appreciably different right after training and after the localization tasks.
-In other words, there was likely no noticeable decay of reach aftereffects during the localization tasks, so that we can collapse the data across iterations.
+To see if reach aftereffects decayed during the localization tasks, we
+compared reach aftereffects in the initial no-cursor block, that
+immediately followed training, with those in the later blocks that
+followed a localization task. We fit an LME to the reach aftereffects
+with iteration (initial vs.~later no-cursor blocks) and target (as
+above) as well as their interaction as fixed effects and participant as
+random effect. There is no effect of iteration (F(1,260)=2.72, p=0.10).
+There was an effect of target (F(6,260)=6.29, p\textless{}.001) but no
+interaction (F(6,260)=0.58, p=0.74). Hence, reach aftereffects were not
+appreciably different right after training and after the localization
+tasks. In other words, there was likely no noticeable decay of reach
+aftereffects during the localization tasks, so that we can collapse the
+data across iterations.
 
-Next we compare the reach aftereffects after classic training with those after exposure training (Fig 3b).
-It appears as if there is little overall difference in magnitude, but there might be a shift of the generalization curve.
-We fit an LME to reach aftereffects with training type (classic vs.
-exposure), target (as above) as well as their interaction as fixed effects and participant as random effect.
-There is no main effect of training (F(1,40)=0.11, p=.74), indicating approximately equal magnitude of reach aftereffects after the two training types.
-There is an effect of target (F(6,240)=8.36, p<.001), indicating that reach aftereffects exhibit some form of a generalization curve.
-There is also an interaction between training type and target (F(6,240)=2.27, p=.038), indicating these generalization curves are different after the two training types.
+Next we compare the reach aftereffects after classic training with those
+after exposure training (Fig 3b). It appears as if there is little
+overall difference in magnitude, but there might be a shift of the
+generalization curve. We fit an LME to reach aftereffects with training
+type (classic vs.~exposure), target (as above) as well as their
+interaction as fixed effects and participant as random effect. There is
+no main effect of training (F(1,40)=0.11, p=.74), indicating
+approximately equal magnitude of reach aftereffects after the two
+training types. There is an effect of target (F(6,240)=8.36,
+p\textless{}.001), indicating that reach aftereffects exhibit some form
+of a generalization curve. There is also an interaction between training
+type and target (F(6,240)=2.27, p=.038), indicating these generalization
+curves are different after the two training types.
 
 We will explore these potentially different generalization curves here.
-In Fig 3c we can observe that reach aftereffects after exposure training seem not to peak at the trained target direction of 45° but a more forward direction.
-We test this by taking the 95\% confidence interval of the centre of a normal curve fit to this data, bootstrapped across participants, and find that the median peak of the generalization curve of reach aftereffects after exposure training is at 66.3°, with a 95\% confidence interval ranging from 49.3° to 78.9°.
-This would indeed suggest that the reach aftereffects after exposure training do not generalize around the trained direction of 45°, but at a more counter clockwise location.
-For classic training, generalization of reach aftereffects peaks at 53.2°, with a 95\% confidence interval spanning 42.1° to 66.5°.
-So for classic training the 95\% confidence interval for peak reach aftereffects does include the trained target.
-These confidence intervals also indicate that generalization of reach aftereffects does not peak at different target position after exposure and classic training.
-However, we can also observe that the full curve was not sampled after exposure training, so that curve fitting is not optimal.
-This means that – given our data – the interaction between target and training type found in the LME above can’t be explained by a shifted generalization curve.
-This may be because our experiment was not set up to test this, and the similar angles where generalization peaks in both localization and reach aftereffects, suggests there may be a difference in where proprioception and prediction generalize strongest.
+In Fig 3c we can observe that reach aftereffects after exposure training
+seem not to peak at the trained target direction of 45° but a more
+forward direction. We test this by taking the 95\% confidence interval
+of the centre of a normal curve fit to this data, bootstrapped across
+participants, and find that the median peak of the generalization curve
+of reach aftereffects after exposure training is at 66.3°, with a 95\%
+confidence interval ranging from 49.3° to 78.9°. This would indeed
+suggest that the reach aftereffects after exposure training do not
+generalize around the trained direction of 45°, but at a more counter
+clockwise location. For classic training, generalization of reach
+aftereffects peaks at 53.2°, with a 95\% confidence interval spanning
+42.1° to 66.5°. So for classic training the 95\% confidence interval for
+peak reach aftereffects does include the trained target. These
+confidence intervals also indicate that generalization of reach
+aftereffects does not peak at different target position after exposure
+and classic training. However, we can also observe that the full curve
+was not sampled after exposure training, so that curve fitting is not
+optimal. This means that -- given our data -- the interaction between
+target and training type found in the LME above can't be explained by a
+shifted generalization curve. This may be because our experiment was not
+set up to test this, and the similar angles where generalization peaks
+in both localization and reach aftereffects, suggests there may be a
+difference in where proprioception and prediction generalize strongest.
 
-In summary, our main hypotheses are confirmed; exposure training leads to shifts in hand localization that are not different for active or passive localization, while movement type does have an effect on localization shift after classic training.
-Exposure training also causes robust reach aftereffects that are of comparable size to those found with classic training.
-There is some evidence that the generalization of both localization shifts and reach aftereffects are different after the two training types, and it appears this can partially be explained by a different peak of the generalization curves, but our data and analyses are not definitive.
+In summary, our main hypotheses are confirmed; exposure training leads
+to shifts in hand localization that are not different for active or
+passive localization, while movement type does have an effect on
+localization shift after classic training. Exposure training also causes
+robust reach aftereffects that are of comparable size to those found
+with classic training. There is some evidence that the generalization of
+both localization shifts and reach aftereffects are different after the
+two training types, and it appears this can partially be explained by a
+different peak of the generalization curves, but our data and analyses
+are not definitive.
 
-# Discussion
+\section{Discussion}\label{discussion}
 
-The position of limbs is important for planning and evaluating movements, and can be estimated through predicted sensory consequences, as well as visual and proprioceptive feedback.
-As in a previous study @Hart2016 here we quantify the contributions of predicted sensory consequences and proprioceptive recalibration to where we localize our hand after training with altered visual feedback of the hand.
-In classical adaptation paradigms, both predictions are updated and proprioception is recalibrated.
-Predictions are updated when they don’t match actual sensory consequences, and proprioception is recalibrated when it doesn’t match visual feedback.
-In this study we use “exposure” training, where the participants do not have volitional control of their movements.
-By design, this should eliminate efference copies and prevent updating predicted consequences of movements, but since the proprioceptive and visual feedback is the same, exposure training still allows proprioceptive recalibration.
-Before and after training, participants localize their hand, both after “active,” self-generated movements that allow using predicted sensory consequences, and after “passive,” robot-generated movements that only allow using proprioception.
-We calculate the training-induced shift in both types of localization given the same actual hand position.
-After classical training we previously reported larger shifts in active localization as compared to passive @Hart2016.
-As we expected, after exposure training there are substantial shifts in localization, but no difference between active and passive localization, indicating that predictions are not updated after exposure training.
-Furthermore, we find that exposure training evokes substantial and robust reach aftereffects, indicating that recalibrated proprioception is used to plan movements.
+The position of limbs is important for planning and evaluating
+movements, and can be estimated through predicted sensory consequences,
+as well as visual and proprioceptive feedback. As in a previous study
+{[}9{]} here we quantify the contributions of predicted sensory
+consequences and proprioceptive recalibration to where we localize our
+hand after training with altered visual feedback of the hand. In
+classical adaptation paradigms, both predictions are updated and
+proprioception is recalibrated. Predictions are updated when they don't
+match actual sensory consequences, and proprioception is recalibrated
+when it doesn't match visual feedback. In this study we use ``exposure''
+training, where the participants do not have volitional control of their
+movements. By design, this should eliminate efference copies and prevent
+updating predicted consequences of movements, but since the
+proprioceptive and visual feedback is the same, exposure training still
+allows proprioceptive recalibration. Before and after training,
+participants localize their hand, both after ``active,'' self-generated
+movements that allow using predicted sensory consequences, and after
+``passive,'' robot-generated movements that only allow using
+proprioception. We calculate the training-induced shift in both types of
+localization given the same actual hand position. After classical
+training we previously reported larger shifts in active localization as
+compared to passive {[}9{]}. As we expected, after exposure training
+there are substantial shifts in localization, but no difference between
+active and passive localization, indicating that predictions are not
+updated after exposure training. Furthermore, we find that exposure
+training evokes substantial and robust reach aftereffects, indicating
+that recalibrated proprioception is used to plan movements.
 
-Our lab previously investigated proprioceptive recalibration and reach aftereffects following visuomotor adaptation with classic training and matched exposure training.
-There we also found that proprioceptive recalibration is of similar magnitude in both training paradigms, but unlike here, reach aftereffects are usually much larger with classic training [@Henriques2012; @Henriques2014; @Ruttle2018; @Cressman2010; @Salomonczyk2013].
-And while proprioceptive recalibration and reach aftereffects do proportionally increase with gradual increases in rotation size for classical training, they do not for exposure training @Salomonczyk2013.
-The similar magnitude of proprioceptive recalibration and reach aftereffects following exposure training, but not classical training, suggest that this sensory recalibration is partly driving this modest change in movements.
-The effect of exposure training on movements is also demonstrated by facilitation by exposure training of subsequent classic training @Sakamoto2015 (but no interference) and transfer of exposure training effects from one hand to the other @Bao2017.
-In the current study, we further demonstrate that exposure training affects movements and proprioception, but also measure its potential effect on predictive estimates.
+Our lab previously investigated proprioceptive recalibration and reach
+aftereffects following visuomotor adaptation with classic training and
+matched exposure training. There we also found that proprioceptive
+recalibration is of similar magnitude in both training paradigms, but
+unlike here, reach aftereffects are usually much larger with classic
+training {[}13,18,22--24{]}. And while proprioceptive recalibration and
+reach aftereffects do proportionally increase with gradual increases in
+rotation size for classical training, they do not for exposure training
+{[}24{]}. The similar magnitude of proprioceptive recalibration and
+reach aftereffects following exposure training, but not classical
+training, suggest that this sensory recalibration is partly driving this
+modest change in movements. The effect of exposure training on movements
+is also demonstrated by facilitation by exposure training of subsequent
+classic training {[}28{]} (but no interference) and transfer of exposure
+training effects from one hand to the other {[}29{]}. In the current
+study, we further demonstrate that exposure training affects movements
+and proprioception, but also measure its potential effect on predictive
+estimates.
 
-Results similar to what we find here were reported in a study by Cameron and colleagues @Cameron2012, using gain modulation of visual feedback of single-joint hand movements around the elbow.
-Their within-subjects experiment included both training with volitional movements as well as with passive movements and also tested perception of movements that were either passive or active.
-They too found a robust change in passive perception of hand movement (using a different measure), and these changes did not differ between the two types of training.
-Similarly, they found shifts in what we might call “active localization,” although the task is different, after both training types.
-Like here, these shifts are larger after classic training as compared to exposure training.
-They also found that passive exposure leads to reach aftereffects, although these were smaller than those produced following “classical” training with altered visual gain.
-Both our findings, and those of Cameron et al.
-@Cameron2012 indicate that updating predicted sensory consequences requires volitionally controlled movements that lead to prediction errors, while proprioception recalibrates equally in both types of training, and that recalibrated proprioception affects open-loop reaches.
-Our combined results suggest that updates in predicted sensory consequences only provide a partial explanation for motor learning.
+Results similar to what we find here were reported in a study by Cameron
+and colleagues {[}19{]}, using gain modulation of visual feedback of
+single-joint hand movements around the elbow. Their within-subjects
+experiment included both training with volitional movements as well as
+with passive movements and also tested perception of movements that were
+either passive or active. They too found a robust change in passive
+perception of hand movement (using a different measure), and these
+changes did not differ between the two types of training. Similarly,
+they found shifts in what we might call ``active localization,''
+although the task is different, after both training types. Like here,
+these shifts are larger after classic training as compared to exposure
+training. They also found that passive exposure leads to reach
+aftereffects, although these were smaller than those produced following
+``classical'' training with altered visual gain. Both our findings, and
+those of Cameron et al. {[}19{]} indicate that updating predicted
+sensory consequences requires volitionally controlled movements that
+lead to prediction errors, while proprioception recalibrates equally in
+both types of training, and that recalibrated proprioception affects
+open-loop reaches. Our combined results suggest that updates in
+predicted sensory consequences only provide a partial explanation for
+motor learning.
 
-Two related concerns about exposure training and passive localization are that the movements are not fully passive, so that efference-based predictions are still generated or that predicted sensory consequences are generated through another route.
-Cameron et al.
-@Cameron2012 measured muscle activity (EMG) during passive movements and found no difference with stationary baseline muscle activity.
-This suggests that any movements generated in a passive condition are subthreshold, minimizing efference-based predictions.
-The brain areas generating predicted sensory consequences could also rely on afferent signals.
-However, such afferent signals are present in both active and passive movements, and if they would result in the same predictions, there would be no difference between active and passive localization after classic training, and no difference between the effects of exposure and classic training, and we find both are different.
-Hence, while we can not fully exclude any predictive signals in passive localization or exposure training, our data shows that any residual predictive signals in the passive movements we used are qualitatively very different from normal efference-based predicted sensory consequences.
+Two related concerns about exposure training and passive localization
+are that the movements are not fully passive, so that efference-based
+predictions are still generated or that predicted sensory consequences
+are generated through another route. Cameron et al. {[}19{]} measured
+muscle activity (EMG) during passive movements and found no difference
+with stationary baseline muscle activity. This suggests that any
+movements generated in a passive condition are subthreshold, minimizing
+efference-based predictions. The brain areas generating predicted
+sensory consequences could also rely on afferent signals. However, such
+afferent signals are present in both active and passive movements, and
+if they would result in the same predictions, there would be no
+difference between active and passive localization after classic
+training, and no difference between the effects of exposure and classic
+training, and we find both are different. Hence, while we can not fully
+exclude any predictive signals in passive localization or exposure
+training, our data shows that any residual predictive signals in the
+passive movements we used are qualitatively very different from normal
+efference-based predicted sensory consequences.
 
-In our classical training group, we not only see shifts for passive localization but even larger shifts for active localization which is consistent with a change in both proprioception and an update in predictions.
-In our exposure training, we did not find a consistent difference between active and passive localization, and none at the trained direction.
-Assuming that predictions were not updated in exposure training, a maximum likelihood estimate (MLE) or “optimal integration” @Ernst2002 would predict that active localization should shift less than passive localization after exposure training.
-But of course this is not the case in our findings (although it is the case for Cameron et al.).
-This suggests that perhaps these two signals are not optimally integrated which is consistent with our comparisons of the variance between passive and active localization.
-In ’t Hart and Henriques @Hart2016, we tested the prediction derived from MLE that hand localization with two signals – proprioception and prediction in active localization – should be more reliable, i.e.
-have lower variance, than hand localization with only one signal – proprioception only in passive localization.
-However, we found no difference in variance between active and passive localization, and recently replicated this in a much larger dataset @tHart2018.
-Taken together, this suggests these different sources of information about unseen hand location are not optimally integrated.
-While localizing the unseen hand is less precise than locating (pointing to) a remembered visual target or a seen and felt hand location, we find that these bimodal estimates are rarely integrated optimally [@Fiehler2010; @Byrne2013; @Jones2010], although others have @VanBeers2002.
-A more recent study @Mikula2018 has also shed doubt on whether “optimal” or “Bayesian” integration is used for locating the hand with two afferent signals.
-Analogously, here we again can’t find evidence that afferent and efferent information combine as a maximum likelihood estimate.
+In our classical training group, we not only see shifts for passive
+localization but even larger shifts for active localization which is
+consistent with a change in both proprioception and an update in
+predictions. In our exposure training, we did not find a consistent
+difference between active and passive localization, and none at the
+trained direction. Assuming that predictions were not updated in
+exposure training, a maximum likelihood estimate (MLE) or ``optimal
+integration'' {[}30{]} would predict that active localization should
+shift less than passive localization after exposure training. But of
+course this is not the case in our findings (although it is the case for
+Cameron et al.). This suggests that perhaps these two signals are not
+optimally integrated which is consistent with our comparisons of the
+variance between passive and active localization. In 't Hart and
+Henriques {[}9{]}, we tested the prediction derived from MLE that hand
+localization with two signals -- proprioception and prediction in active
+localization -- should be more reliable, i.e.~have lower variance, than
+hand localization with only one signal -- proprioception only in passive
+localization. However, we found no difference in variance between active
+and passive localization, and recently replicated this in a much larger
+dataset {[}31{]}. Taken together, this suggests these different sources
+of information about unseen hand location are not optimally integrated.
+While localizing the unseen hand is less precise than locating (pointing
+to) a remembered visual target or a seen and felt hand location, we find
+that these bimodal estimates are rarely integrated optimally
+{[}32--34{]}, although others have {[}35{]}. A more recent study
+{[}36{]} has also shed doubt on whether ``optimal'' or ``Bayesian''
+integration is used for locating the hand with two afferent signals.
+Analogously, here we again can't find evidence that afferent and
+efferent information combine as a maximum likelihood estimate.
 
-It seems clear that the cerebellum plays a role in motor learning as it appears to compute predicted sensory consequences, i.e.
-it implements a forward model [@Bastian2006; @Shadmehr2008; @Wolpert1998].
-People with cerebellar damage do worse on motor learning tasks [@Martin1996; @Maschke2004; @Criscimagna-Hemminger2010; @Donchin2012], and show decreased shifts in hand localization tasks following motor learning [@Izawa2012b; @Synofzik2008].
-This highlights that the cerebellum, and likely predicted sensory consequences, are important for motor learning, but does not explain the remaining shifts in hand localization.
-We previously found that proprioceptive recalibration is intact in people with mild cerebellar ataxia and that it is similar following exposure and classical training with a gradually introduced cursor rotation @Henriques2014.
-The remaining changes in hand localization found in cerebellar patients can be attributed to recalibrated proprioception which should be intact @Henriques2014.
-Analogously, here we show that in a paradigm that stops updates of predictions of sensory consequences, as supposedly in people with cerebellar damage, we still see substantial shifts in localization.
-Again, the remaining localization shifts can be explained if, along with predictions, the human brain uses afferent signals: recalibrated proprioceptive estimates, to localize the hand.
+It seems clear that the cerebellum plays a role in motor learning as it
+appears to compute predicted sensory consequences, i.e.~it implements a
+forward model {[}37--39{]}. People with cerebellar damage do worse on
+motor learning tasks {[}40--43{]}, and show decreased shifts in hand
+localization tasks following motor learning {[}5,7{]}. This highlights
+that the cerebellum, and likely predicted sensory consequences, are
+important for motor learning, but does not explain the remaining shifts
+in hand localization. We previously found that proprioceptive
+recalibration is intact in people with mild cerebellar ataxia and that
+it is similar following exposure and classical training with a gradually
+introduced cursor rotation {[}18{]}. The remaining changes in hand
+localization found in cerebellar patients can be attributed to
+recalibrated proprioception which should be intact {[}18{]}.
+Analogously, here we show that in a paradigm that stops updates of
+predictions of sensory consequences, as supposedly in people with
+cerebellar damage, we still see substantial shifts in localization.
+Again, the remaining localization shifts can be explained if, along with
+predictions, the human brain uses afferent signals: recalibrated
+proprioceptive estimates, to localize the hand.
 
-## Generalization
+\subsection{Generalization}\label{generalization}
 
-We do find some evidence that, after exposure training, the generalization curves for localization shifts are not centred on the visual location of the trained target; they don’t peak at 45° but at ~62°.
-In contrast, after classic training the peak of the generalization curve does peak close to the training target.
-It is possible that proprioceptive recalibration is not anchored to the visual goal of the training task as it is not a requirement to feel your hand at any specific point; rather, in classic training the visual cursor has to be brought to a visual target to end a trial.
-While in exposure training the movements are executed for the participants without error, their task is still to pay attention to the visual cursor while it moves to the visual target; there are no task demands on proprioception.
-Although we can’t substantiate this here, the generalization curves of the reach aftereffects seem to mimic the generalization curves of localization shifts, suggesting a relationship between changes in state estimates and changes in movements.
-This needs to be tested further, but if these effects are true, they may provide insight into how state estimates are used to produce movements, and also may lead to a new method to disentangle the influence of recalibrated proprioception and more traditional updated internal models on motor changes, such as reach aftereffects.
-Either way, even though the experiment was not designed to investigate this, the shifted generalization curves of changes in localization after classic and exposure training suggest they are generated by different mechanisms.
-This is in line with our earlier findings that proprioception generalizes differently from reach adaptation @Mostafa2015.
+We do find some evidence that, after exposure training, the
+generalization curves for localization shifts are not centred on the
+visual location of the trained target; they don't peak at 45° but at
+\textasciitilde{}62°. In contrast, after classic training the peak of
+the generalization curve does peak close to the training target. It is
+possible that proprioceptive recalibration is not anchored to the visual
+goal of the training task as it is not a requirement to feel your hand
+at any specific point; rather, in classic training the visual cursor has
+to be brought to a visual target to end a trial. While in exposure
+training the movements are executed for the participants without error,
+their task is still to pay attention to the visual cursor while it moves
+to the visual target; there are no task demands on proprioception.
+Although we can't substantiate this here, the generalization curves of
+the reach aftereffects seem to mimic the generalization curves of
+localization shifts, suggesting a relationship between changes in state
+estimates and changes in movements. This needs to be tested further, but
+if these effects are true, they may provide insight into how state
+estimates are used to produce movements, and also may lead to a new
+method to disentangle the influence of recalibrated proprioception and
+more traditional updated internal models on motor changes, such as reach
+aftereffects. Either way, even though the experiment was not designed to
+investigate this, the shifted generalization curves of changes in
+localization after classic and exposure training suggest they are
+generated by different mechanisms. This is in line with our earlier
+findings that proprioception generalizes differently from reach
+adaptation {[}11{]}.
 
-## Conclusion
+\subsection{Conclusion}\label{conclusion}
 
-To sum up, after a training paradigm designed to prevent updating of predicted sensory consequences but allow recalibration of proprioception, we find substantial changes in where people localize their hand.
-This means that recalibrated proprioceptive estimates can explain shifts in hand localization.
-The distinct change in the direction of open-loop reaches we observe here, suggests recalibrated proprioception can contribute to motor adaptation in other contexts as well.
-Finally, we have some evidence that after exposure training, the shift in hand localization does not generalize around the trained target location.
-All of this confirms that different mechanisms underlie proprioceptive recalibration and motor adaptation.
+To sum up, after a training paradigm designed to prevent updating of
+predicted sensory consequences but allow recalibration of
+proprioception, we find substantial changes in where people localize
+their hand. This means that recalibrated proprioceptive estimates can
+explain shifts in hand localization. The distinct change in the
+direction of open-loop reaches we observe here, suggests recalibrated
+proprioception can contribute to motor adaptation in other contexts as
+well. Finally, we have some evidence that after exposure training, the
+shift in hand localization does not generalize around the trained target
+location. All of this confirms that different mechanisms underlie
+proprioceptive recalibration and motor adaptation.
 
-## Acknowledgements
+\subsection{Acknowledgements}\label{acknowledgements}
 
-This work was supported by a Canadian Network for Research and Innovation in Machining Technology NSERC Operating grant (DYPH) and the German Research Foundation (DFG) under grant no.
-HA 6861/2-1 (BMtH).
-The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
-We thank Shanaathanan Modchalingam for assistance in collecting the data.
+This work was supported by a Canadian Network for Research and
+Innovation in Machining Technology NSERC Operating grant (DYPH) and the
+German Research Foundation (DFG) under grant no. HA 6861/2-1 (BMtH). The
+funders had no role in study design, data collection and analysis,
+decision to publish, or preparation of the manuscript. We thank
+Shanaathanan Modchalingam for assistance in collecting the data.
 
-# References {#references .unnumbered}
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-Duhamel1992a}{}
+1. Duhamel J-R, Colby CL, Goldberg ME. The updating of the
+representation of visual space in parietal cortex by intended eye
+movements. Science. 1992;255: 90--92.
+doi:\href{https://doi.org/10.1126/science.1553535}{10.1126/science.1553535}
+
+\hypertarget{ref-Wolpert1995}{}
+2. Wolpert DM, Ghahramani Z, Jordan MI. An internal model for
+sensorimotor integration. Science. 1995;269: 1880--2.
+doi:\href{https://doi.org/10.1126/science.7569931}{10.1126/science.7569931}
+
+\hypertarget{ref-Cressman2009}{}
+3. Cressman EK, Henriques DYP. Sensory recalibration of hand position
+following visuomotor adaptation. Journal of Neurophysiology. 2009;102:
+3505--3518.
+doi:\href{https://doi.org/10.1152/jn.00514.2009}{10.1152/jn.00514.2009}
+
+\hypertarget{ref-Clayton2013}{}
+4. Clayton HA, Cressman EK, Henriques DY. Proprioceptive sensitivity in
+ehlers-danlos syndrome patients. Experimental Brain Research. 2013;230:
+311--321.
+doi:\href{https://doi.org/10.1007/s00221-013-3656-4}{10.1007/s00221-013-3656-4}
+
+\hypertarget{ref-Izawa2012b}{}
+5. Izawa J, Criscimagna-Hemminger SE, Shadmehr R. Cerebellar
+contributions to reach adaptation and learning sensory consequences of
+action. Journal of Neuroscience. 2012;32: 4230--4239.
+doi:\href{https://doi.org/10.1523/JNEUROSCI.6353-11.2012}{10.1523/JNEUROSCI.6353-11.2012}
+
+\hypertarget{ref-Jones2012a}{}
+6. Jones SAH, Fiehler K, Henriques DYP. A task-dependent effect of
+memory and hand-target on proprioceptive localization. Neuropsychologia.
+2012;50: 1462--1470.
+doi:\href{https://doi.org/10.1016/j.neuropsychologia.2012.02.031}{10.1016/j.neuropsychologia.2012.02.031}
+
+\hypertarget{ref-Synofzik2008}{}
+7. Synofzik M, Lindner A, Thier P. The cerebellum updates predictions
+about the visual consequences of one's behavior. Current Biology.
+2008;18: 814--818.
+doi:\href{https://doi.org/10.1016/j.cub.2008.04.071}{10.1016/j.cub.2008.04.071}
+
+\hypertarget{ref-Yavari2016}{}
+8. Yavari F, Mahdavi S, Towhidkhah F, Ahmadi-Pajouh M-A, Ekhtiari H,
+Darainy M. Cerebellum as a forward but not inverse model in visuomotor
+adaptation task: A tDCS-based and modeling study. Experimental Brain
+Research. 2016;234: 997--1012.
+doi:\href{https://doi.org/10.1007/s00221-015-4523-2}{10.1007/s00221-015-4523-2}
+
+\hypertarget{ref-Hart2016}{}
+9. 't Hart BM, Henriques DY. Separating predicted and perceived sensory
+consequences of motor learning. PLoS ONE. 2016;11.
+doi:\href{https://doi.org/10.1371/journal.pone.0163556}{10.1371/journal.pone.0163556}
+
+\hypertarget{ref-Mostafa2014}{}
+10. Mostafa AA, Salomonczyk D, Cressman EK, Henriques DY. Intermanual
+transfer and proprioceptive recalibration following training with
+translated visual feedback of the hand. Experimental Brain Research.
+2014;232: 1639--1651.
+doi:\href{https://doi.org/10.1007/s00221-014-3833-0}{10.1007/s00221-014-3833-0}
+
+\hypertarget{ref-Mostafa2015}{}
+11. Mostafa AA, Kamran-Disfani R, Bahari-Kashani G, Cressman EK,
+Henriques DYP. Generalization of reach adaptation and proprioceptive
+recalibration at different distances in the workspace. Experimental
+Brain Research. 2015;233: 817--827.
+doi:\href{https://doi.org/10.1007/s00221-014-4157-9}{10.1007/s00221-014-4157-9}
+
+\hypertarget{ref-Maksimovic2018}{}
+12. Maksimovic S, Cressman EK. Long-term retention of proprioceptive
+recalibration. Neuropsychologia. 2018;114: 65--76.
+doi:\href{https://doi.org/10.1016/j.neuropsychologia.2018.04.009}{10.1016/j.neuropsychologia.2018.04.009}
+
+\hypertarget{ref-Henriques2012}{}
+13. Henriques DYP, Cressman EK. Visuomotor adaptation and proprioceptive
+recalibration. Journal of Motor Behavior. 2012;44: 435--444.
+doi:\href{https://doi.org/10.1080/00222895.2012.659232}{10.1080/00222895.2012.659232}
+
+\hypertarget{ref-Cressman2010a}{}
+14. Cressman EK, Salomonczyk D, Henriques DYP. Visuomotor adaptation and
+proprioceptive recalibration in older adults. Experimental Brain
+Research. 2010;205: 533--544.
+doi:\href{https://doi.org/10.1007/s00221-010-2392-2}{10.1007/s00221-010-2392-2}
+
+\hypertarget{ref-Barkley2014}{}
+15. Barkley V, Salomonczyk D, Cressman EK, Henriques DYP. Reach
+adaptation and proprioceptive recalibration following terminal visual
+feedback of the hand. Frontiers in Human Neuroscience. 2014;8: 1--11.
+doi:\href{https://doi.org/10.3389/fnhum.2014.00705}{10.3389/fnhum.2014.00705}
+
+\hypertarget{ref-Salomonczyk2011}{}
+16. Salomonczyk D, Cressman EK, Henriques DYP. Proprioceptive
+recalibration following prolonged training and increasing distortions in
+visuomotor adaptation. Neuropsychologia. 2011;49: 3053--3062.
+doi:\href{https://doi.org/10.1016/j.neuropsychologia.2011.07.006}{10.1016/j.neuropsychologia.2011.07.006}
+
+\hypertarget{ref-Nourouzpour2014}{}
+17. Nourouzpour N, Salomonczyk D, Cressman EK, Henriques DYP. Retention
+of proprioceptive recalibration following visuomotor adaptation.
+Experimental Brain Research. 2014;233: 1019--1029.
+doi:\href{https://doi.org/10.1007/s00221-014-4176-6}{10.1007/s00221-014-4176-6}
+
+\hypertarget{ref-Henriques2014}{}
+18. Henriques DYP, Filippopulos F, Straube A, Eggert T. The cerebellum
+is not necessary for visually driven recalibration of hand
+proprioception. Neuropsychologia. 2014;64: 195--204.
+doi:\href{https://doi.org/10.1016/j.neuropsychologia.2014.09.029}{10.1016/j.neuropsychologia.2014.09.029}
+
+\hypertarget{ref-Cameron2012}{}
+19. Cameron BD, Franks IM, Inglis JT, Chua R. The adaptability of
+self-action perception and movement control when the limb is passively
+versus actively moved. Consciousness and Cognition. 2012;21: 4--17.
+doi:\href{https://doi.org/10.1016/j.concog.2010.11.006}{10.1016/j.concog.2010.11.006}
+
+\hypertarget{ref-Ostry2016}{}
+20. Ostry DJ, Gribble PL. Sensory plasticity in human motor learning.
+Trends in Neurosciences. 2016;39: 114--123.
+doi:\href{https://doi.org/10.1016/j.tins.2015.12.006}{10.1016/j.tins.2015.12.006}
+
+\hypertarget{ref-Clayton2014}{}
+21. Clayton HA, Cressman EK, Henriques DYP. The effect of visuomotor
+adaptation on proprioceptive localization: The contributions of
+perceptual and motor changes. Experimental Brain Research. 2014;232:
+2073--2086.
+doi:\href{https://doi.org/10.1007/s00221-014-3896-y}{10.1007/s00221-014-3896-y}
+
+\hypertarget{ref-Ruttle2018}{}
+22. Ruttle JE, 't Hart BM, Henriques DYP. The fast contribution of
+visual-proprioceptive discrepancy to reach aftereffects and
+proprioceptive recalibration. PLOS ONE. 2018;13: e0200621.
+doi:\href{https://doi.org/10.1371/JOURNAL.PONE.0200621}{10.1371/JOURNAL.PONE.0200621}
+
+\hypertarget{ref-Cressman2010}{}
+23. Cressman EK, Henriques DYP. Reach adaptation and proprioceptive
+recalibration following exposure to misaligned sensory input. Journal of
+Neurophysiology. 2010;103: 1888--1895.
+doi:\href{https://doi.org/10.1152/jn.01002.2009}{10.1152/jn.01002.2009}
+
+\hypertarget{ref-Salomonczyk2013}{}
+24. Salomonczyk D, Cressman EK, Henriques DYP. The role of the
+cross-sensory error signal in visuomotor adaptation. Experimental Brain
+Research. 2013;228: 313--325.
+doi:\href{https://doi.org/10.1007/s00221-013-3564-7}{10.1007/s00221-013-3564-7}
+
+\hypertarget{ref-Mostafa2017}{}
+25. Mostafa AA, 't Hart BM, Henriques DY. Motor learning without moving
+{[}Internet{]}. Open Science Framework; 2017.
+doi:\href{https://doi.org/10.17605/osf.io/zfdth}{10.17605/osf.io/zfdth}
+
+\hypertarget{ref-RCoreTeam2017}{}
+26. Team RC. R: A language and environment for statistical computing
+{[}Internet{]}. Vienna, Austria: R Foundation for Statistical Computing.
+2017. Available: \url{https://www.r-project.org/}
+
+\hypertarget{ref-Luke2017}{}
+27. Luke SG. Evaluating significance in linear mixed-effects models in
+r. Behavior Research Methods. 2017;49: 1494--1502.
+doi:\href{https://doi.org/10.3758/s13428-016-0809-y}{10.3758/s13428-016-0809-y}
+
+\hypertarget{ref-Sakamoto2015}{}
+28. Sakamoto T, Kondo T. Visuomotor learning by passive motor
+experience. Frontiers in Human Neuroscience. 2015;9: 279.
+doi:\href{https://doi.org/10.3389/fnhum.2015.00279}{10.3389/fnhum.2015.00279}
+
+\hypertarget{ref-Bao2017}{}
+29. Bao S, Lei Y, Wang J. Experiencing a reaching task passively with
+one arm while adapting to a visuomotor rotation with the other can lead
+to substantial transfer of motor learning across the arms. Neuroscience
+letters. 2017;638: 109--113.
+doi:\href{https://doi.org/10.1016/j.neulet.2016.12.028}{10.1016/j.neulet.2016.12.028}
+
+\hypertarget{ref-Ernst2002}{}
+30. Ernst MO, Banks MS. Humans integrate visual and haptic information
+in a statistically optimal fashion. Nature. 2002;
+doi:\href{https://doi.org/10.1038/415429a}{10.1038/415429a}
+
+\hypertarget{ref-tHart2018}{}
+31. 't Hart BM, Ayala M, Henriques DY. Testing maximum likelihood
+estimates of hand position based on proprioception and prediction
+{[}Internet{]}. Open Science Framework; 2018.
+doi:\href{https://doi.org/10.17605/OSF.IO/WV3DB}{10.17605/OSF.IO/WV3DB}
+
+\hypertarget{ref-Fiehler2010}{}
+32. Fiehler K, Rösler F, Henriques DYP. Interaction between gaze and
+visual and proprioceptive position judgements. Experimental Brain
+Research. 2010;203: 485--498.
+doi:\href{https://doi.org/10.1007/s00221-010-2251-1}{10.1007/s00221-010-2251-1}
+
+\hypertarget{ref-Byrne2013}{}
+33. Byrne PA, Henriques DY. When more is less: Increasing allocentric
+visual information can switch visual-proprioceptive combination from an
+optimal to sub-optimal process. Neuropsychologia. 2013;51: 26--37.
+doi:\href{https://doi.org/10.1016/j.neuropsychologia.2012.10.008}{10.1016/j.neuropsychologia.2012.10.008}
+
+\hypertarget{ref-Jones2010}{}
+34. Jones SAH, Henriques DYP. Memory for proprioceptive and multisensory
+targets is partially coded relative to gaze. Neuropsychologia. 2010;
+doi:\href{https://doi.org/10.1016/j.neuropsychologia.2010.10.001}{10.1016/j.neuropsychologia.2010.10.001}
+
+\hypertarget{ref-VanBeers2002}{}
+35. Beers RJ van, Wolpert DM, Haggard P. When feeling is more important
+than seeing in sensorimotor adaptation. Current Biology. 2002;12:
+834--837.
+doi:\href{https://doi.org/10.1016/S0960-9822(02)00836-9}{10.1016/S0960-9822(02)00836-9}
+
+\hypertarget{ref-Mikula2018}{}
+36. Mikula L, Gaveau V, Pisella L, Khan AZ, Blohm G. Learned rather than
+online relative weighting of visual-proprioceptive sensory cues. Journal
+of Neurophysiology. 2018;119: 1981--1992.
+doi:\href{https://doi.org/10.1152/jn.00338.2017}{10.1152/jn.00338.2017}
+
+\hypertarget{ref-Bastian2006}{}
+37. Bastian AJ. Learning to predict the future: The cerebellum adapts
+feedforward movement control. Current Opinion in Neurobiology. 2006;16:
+645--649.
+doi:\href{https://doi.org/10.1016/j.conb.2006.08.016}{10.1016/j.conb.2006.08.016}
+
+\hypertarget{ref-Shadmehr2008}{}
+38. Shadmehr R, Krakauer JW. A computational neuroanatomy for motor
+control. Experimental Brain Research. 2008;185: 359--381.
+doi:\href{https://doi.org/10.1007/s00221-008-1280-5}{10.1007/s00221-008-1280-5}
+
+\hypertarget{ref-Wolpert1998}{}
+39. Wolpert DM, Miall RC, Kawato M. Internal models in the cerebellum.
+Trends in Cognitive Sciences. 1998;2: 338--347.
+doi:\href{https://doi.org/10.1016/S1364-6613(98)01221-2}{10.1016/S1364-6613(98)01221-2}
+
+\hypertarget{ref-Martin1996}{}
+40. Martin TA, Keating JG, Goodkin HP, Bastian AJ, Thach WT. Throwing
+while looking through prisms. ii. specificity and storage of multiple
+gaze-throw calibrations. Brain: a journal of neurology. 1996;119:
+1199--211.
+
+\hypertarget{ref-Maschke2004}{}
+41. Maschke M, Gomez CM, Ebner TJ, Konczak J. Hereditary cerebellar
+ataxia progressively impairs force adaptation during goal-directed arm
+movements. Journal of Neurophysiology. 2004;91: 230--238.
+doi:\href{https://doi.org/10.1152/jn.00557.2003}{10.1152/jn.00557.2003}
+
+\hypertarget{ref-Criscimagna-Hemminger2010}{}
+42. Criscimagna-Hemminger SE, Bastian AJ, Shadmehr R. Size of error
+affects cerebellar contributions to motor learning. Journal of
+Neurophysiology. 2010;103: 2275--2284.
+doi:\href{https://doi.org/10.1152/jn.00822.2009}{10.1152/jn.00822.2009}
+
+\hypertarget{ref-Donchin2012}{}
+43. Donchin O, Rabe K, Diedrichsen J, Lally N, Schoch B, Gizewski ER, et
+al. Cerebellar regions involved in adaptation to force field and
+visuomotor perturbation. Journal of Neurophysiology. 2012;107: 134--147.
+doi:\href{https://doi.org/10.1152/jn.00007.2011}{10.1152/jn.00007.2011}
+
+\nolinenumbers
+
+\end{document}

### 2018-11-28 localization.R

@@ -7,6 +7,8 @@ source('shared.R')
 
 plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerformance=TRUE, remove15=FALSE) {
   
+  classicOnline = FALSE # the third panel is now used for generalization curves
+  
   # get the data to plot:
   exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE, selectPerformance=selectPerformance)
   cla <- getPointLocalization('classic', difference=TRUE, verbose=FALSE, selectPerformance=FALSE)
@@ -213,6 +215,96 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerfo
   
   if (generateSVG) {
     dev.off()
+  }
+  
+}
+
+plotALignedRotatedLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerformance=TRUE, remove15=FALSE) {
+  
+  groups <- c('exposure','classic')
+  
+  if (classicOnline) {
+    
+    groups <- c(groups,'online')
+    
+  }
+  
+  if (generateSVG) {
+    installed.list <- rownames(installed.packages())
+    if ('svglite' %in% installed.list) {
+      # nothing
+    } else {
+      generateSVG=FALSE
+    }
+  }
+  
+  par(mfrow=c(7,3),mai=c(.6,.5,.01,.01))
+  
+  points <- c(15,25,35,45,55,65,75)
+  
+  
+  for (group in groups) {
+    
+    SP <- FALSE
+    if (group == 'exposure') {
+      SP <- selectPerformance
+    }
+    
+    df <- load.DownloadDataframe(url=localizationURLs[group],filename=sprintf('localization_%s.csv',group))
+    
+    if (selectPerformance & group=='exposure') {
+      blinks <- load.DownloadDataframe(informationURLs['blinkdetect'],'blinkdetect_exposure.csv')
+      OKparticipants <- blinks$participant[which(blinks$rotated_b == 1 & blinks$performance > 0.65)]
+      df <- df[which(df$participant %in% OKparticipants),]
+    }
+    
+    participants <- unique(df$participant)
+    
+    for (passive in c(0,1)) {
+      
+      if (generateSVG) {
[…]

### 2018-12-05 paper/paper.Rmd
removed dollars
changed ethics for plos

@@
 
-Twenty-five right-handed participants were recruited for this study.
+Twenty-five right-handed participants were recruited for the passive learning task in this study.
 One participant was excluded for not following task instructions, and three were excluded for low performance on a task that ensures attention during the passive training.
 All analyses presented here pertain to the remaining twenty-one participants (13 females and 8 males, mean age: 20.1 +/- 2.3 years), but the data of the three low-performing participants is included in the online \href{https://osf.io/zfdth}{dataset} @Mostafa2017.
-All had normal or corrected-to-normal vision, and provided prior, written informed consent in accordance with the ethical guidelines set by the York Human subjects Review Subcommittee and received credit toward an undergraduate psychology course.
+Participants provided prior, written informed consent in accordance with the ethical guidelines set by the York Human subjects Review Subcommittee (#2014-240).
+All had normal or corrected-to-normal vision, and received credit toward an undergraduate psychology course.
 Participants were screened verbally, and all reported being right handed and not having any history of visual, neurological, and/or motor dysfunction.
+Published data from 21 further participants is used for comparison @Hart2016.
 
@@
 After both types of training, participants did open-loop reaches as well as two kinds of hand localization tasks, to test the effect of training on proprioceptive and predictive hand estimates.
-
-$
 
@@
 \end{table}
-$
 

### 2019-03-30 ManuscriptMaterials.Rmd
organized scripts to reflect revisions

@@
 
-These are the planned figures. Figure one will not be included here, and can in fact simply be re-used from the previous manuscript.
+These are the planned figures. Figures one and two will not be included here as they are not generated from data but illustrate the setup, tasks and experimental design.
 
@@
     + D: Localization (top view)
-2. Localization
+2. Experimental Design
+    + A: Task order (instead of the table in the previous paper)
+    + B: Main localization measures
+3. Localization
     + A: Exposure, active and passive (goal: small/no difference between active and passive)
     + B: Classic, active and passive (goal: larger difference between active and passive, different generalization?)
-    + C: Online localization?
-3. Reach Aftereffects (goal: show motor changes are there, are robust and compare to classic)
+4. Reach Aftereffects (goal: show motor changes are there, are robust and compare to classic)
     + A: Exposure reach aftereffects (initial & later)
     + B: Classic & Exposure reach aftereffects (initial only? or all?)
+5. Generalization Maxima
+    + A: Maxima of active localization after classic and exposure training
+    + B: Maxima of reach aftereffects after classic and exposure training
 
@@
 
-4. Correlation
+6. Correlation
     + A: Exposure, active and passive localization predicting reach aftereffects
     + B: Classic, active and passive localization predicting reach aftereffects
-
-## Tables
-
-We expand the table used in the previous manuscript to include this paradigm as well.
-
-1. Task order & trial numbers (goal: clarify task order and show similarity between classic and exposure experiment - so probably after Fig 1)
+7. Individual localization responses
 
@@
 
-The scripts doing the statistics and creating figures are in separate R files. They allow many options, but here we will only see / do the ones that we think should go to the manuscript.
+The scripts doing the statistics and creating figures are in a few separate R files. They allow many options, but here we will only see / do the ones that we think should go to the manuscript.
 
@@
 ```{r require_lme_packages}
-required.packages = c('lme4', 'lmerTest', installmissing=FALSE)
-installRequire.Packages(required.packages)
+required.packages = c('lme4', 'lmerTest')
+installRequire.Packages(required.packages, installmissing=FALSE)
+if (!all(required.packages %in% installed.packages())) {
+  warning('packages required for analysis are not installed, later chunks will fail\n')
+}
 ```
@@
 Second are the no-cursor reaches (and the reach aftereffects).
-A third topic might be if localization shifts can predict no-cursor changes, which we'll call the "correlation" topic.
-Here we'll also follow this order:
+Then we'll look at generalization patterns of localization shifts and reach aftereffects after both classic and exposure training.
+
+A further topic we only explore in this notebook is if localization shifts can predict no-cursor changes:
 
@@
 4. Reach aftereffects
-5. Correlations
+5. Generalization Maxima
+6. Correlations
 
@@
 Since in the localization tasks there is missing data at some points (usually at the extremes) we can't use ANOVAs directly on the data.
-Imputation is going to be very hard because the missing data is systematically missing.
-Hence, we are using Linear Mixed Effects models as an alternative, but this presents a new problem.
-Getting the usual F-values and p-values is not so straightforward.
-Initially we decided to use the $\chi^2$ approach that we get from the `Anova` function from the `car` package.
+Imputation is going to be very hard because the missing data is systematically missing, and there is no (good) model for the localization data (for reach aftereffects it might work).
+An alternative to imputation is to either remove participants or removed measurement locations, but that would mean deleting 7/21 participants or 4/7 locations, so at least 1/3 of the data, which would mean a big loss in power.
+As the data is systematically missing, this might also skew results away from what would be observed in the general population.
+
+Hence, we are using Linear Mixed Effects models as an alternative, but the output of those models are not readily interpretable by everybody.
+Initially we decided to use the $\chi^2$ approach (from the `Anova` function from the `car` package).
 However, a paper from Steven G.
 Luke (2017, _Evaluating significance in linear mixed-effects models in R_) suggests using Satterthwaite approximation.
-So all functions running an LME now take an argument that tells them to use the Satterthwaite method, but the default is chi-squared.
+So all functions running an LME now take an argument that tells them to use the Satterthwaite method, but it can be set to chi-squared as well.
 It turns out that it doesn't matter much, so it's a good sanity check.
@@
   installRequire.Packages(required.packages, installmissing=FALSE)
+  # we check if that worked and otherwise
+  if (!all(required.packages %in% installed.packages())) {
+    LMEmethod <- 'Satterthwaite'
+  }
 }
@@
 
-### Peak location of generalization
+If the packages were not installed (we won't force you), we do switch back to the Satterthwaite method so that we don't generate errors instead of statistical results:
+
+LME analyses have fixed effects, usually the effects we are interested in, as well as random effects, which are usually not the effects we are interested in.
+Fixed effects are comparable to 'factors' in ANOVA's, and including random effects is conceptually (but not mathematically) comparable to making it a repeated measures design, or including covariates.
+We'll spell out the fixed and random effects included in each analyses right before showing the results, so that those results can easily be put in the proper context by the reader, and hence not all details are mentioned in the methods section.
+
+### Generalization Maxima
 
 Since participants trained with a single target, we could look at generalization curves.
-These appear to be shifted more counter clockwise in most cases, so we want to quantify this.
-But, for most of the curves, and especially the ones that seems to be shifted, we don't have data on the full curve on one side of the peak.
-We can still bootstrap the __peak__ of Guassian functions fit to the data across many samples of participants for the exposure group and for the delayed localization after classic training, but not for the online localization after classic training.
+These appear to be shifted more counter clockwise after exposure training compared to classic, so we want to quantify this.
+But, for most of the curves, and especially the ones that seems to be shifted, we don't have data on the full curve on one side of the peak, which limits the interpretability of the analysis somewhat.
+We can still bootstrap the __peak__ of Guassian functions fit to the data across many samples of participants for the exposure group and for the delayed localization after classic training, but not for the online localization after classic training (the shifts are even larger there...).
 This data is presented in the paper, where we only look at the peak of the localization -- not the width of the generalization curve.
 
-The range of the data doesn't include the peak in many cases, which makes it hard to fit a normal function to the data.
-In one alternative approach, we assume that the minimal change in the data, occurs far beyond the boundaries of the measured workspace, make the data cumulative and scale back to a range from 0 to 1.
-This allows fitting a _cumulative_ normal function where the 50% point corresponds to the peak of the generalization curve.
+The range of the data doesn't include the peak in some participants, which makes it hard to fit a normal function to every individual participant's data.
+In one alternative approach, we assumed that the minimal change in the data, occurs far beyond the boundaries of the measured workspace, and converted the data to a cumulative distribution, assuming that at some point it would revert back to the base level we see on the other side of the peak.
+We can scale back to a range from 0 to 1, to allow fitting a _cumulative_ normal function where the 50% point corresponds to the peak of the generalization curve.
 This method is more stable, and seems to works fairly well on the datasets where the more common fits also work (but we didn't test online localization from the classic experiment).
-Code for this approach is available (functions `psychometricGeneralization()`, `pnormMSE()` and `getCDFpeaks()` at the end of the file: `shared.R`), but the results are not presented in the paper or notebook.
+Code for this approach is available (functions `psychometricGeneralization()`, `pnormMSE()` and `getCDFpeaks()` at the end of the file: `shared.R`), but *these analysis are not presented* in the paper or notebook, but the code is left here for the interested reader.
 
@@
 
-There are no scripts for this part, as the code is simple and can just be embedded.
+Here are some simple descriptives of participants, as well as the selection of participants based on how they performed in detecting the very brief blink of the cursor (2 frames ~33 ms) during the exposure training.
+This ensures that the participants paid attention to the cursor during training, which is probably required for the cross-sensory error signal to have the desired effect of proprioceptive recalibration.
 
@@
 
-We'll also set a variable that is to be passed to all functions to exclude the underperformance:
+We'll also set a variable that is to be passed to all functions to exclude the underperformers:
 
@@
 
+This is highly similar between conditions and training paradigms, and at an acceptable level.
+
 We decided not to use any more outlier removal, but it is implemented and can be switched on in the code: set the `removeOutliers` argument to the `getPointLocalization` function in `shared.R` to TRUE.
@@
 ```
+
+These high percentages may indicate there are some additional no-cursor reaches that could be considered outliers and romeving them would not constitute a big loss of data.
+However, as it is, the variance in the dataset appears to mostly reflect inidividual differences and not noise, so we kept the data as is.
 
@@
 
-## Localization peaks
-
-There was an interaction between group and hand angle on localization shifts above, which suggests different generalization curves in the two groups.
-This could reflect a difference in the shape (more flat for the classic training group as compared to exposure), but also a difference in the locations of the peaks of the generalization curves.
-
-Here are the confidence intervals for the peaks of the generalization curves for _active_ localization:
-
-```{r localization_peaks}
-getPeakLocConfidenceInterval(group='classic',
-                             CIs=c(.95), 
-                             movementtype='active', 
-                             LRpart='all',
-                             selectPerformance=FALSE)
-getPeakLocConfidenceInterval(group='exposure',
-                             CIs=c(.95), 
-                             movementtype='active', 
-                             LRpart='all', 
-                             selectPerformance=selectPerformance)
-```
-
-The median values are a little over 13 degrees apart, and they _just_ fall outside each others confidence interval.
-This can explain the interaction between group and hand angle on localization shifts.
-
 # Reach Aftereffects
 
-Here is a plot, probably Figure 3:
+Here is a plot, probably Figure 4:
 
-```{r fig2_reachaftereffect, fig.width=7.5, fig.height=3}
+```{r fig4_reachaftereffect, fig.width=7.5, fig.height=3}
 plotReachAftereffects(selectPerformance=selectPerformance)
@@
 
+So there is no effect of training (overall reach aftereffect magnitudes are not very different between classic and exposure training), but there is an effect of target (there is a generalization curve, as we can clearly see in the data) and there is an interaction between training and target (the generalization curves are different between classic and exposure training).
+This last bit is very similar to the pattern observed in localization, but we'll get back to that later.
+
 The figure looked like there might be some difference between the two training types, mainly in where the generalization curves peak.
 There is no main effect of training type, so the overall magnitude of reach aftereffects is not different.
-But there is an interaction between training and target (p=.038), so that perhaps the curves peak at different locations.
 
-So we now investigate the peaks the confidence intervals for the peaks of both:
+However, the main take-away message from this analysis is that the motor changes measured with changes in open-loop reach direction are of similar magnitude in both types of training.
+That is, passive exposure training (proprioceptive recalibration) drives motor changes.
+
+To sum up:
+
+1. Rotated exposure training evokes substantial shifts in no-cursor reach direction across the workspace,
+2. these changes persist over time, and
+3. these changes are comparable to the ones we found after classic training, and
+4. there is some generalization curve across the workspace, but
+
+# Generalization Maxima
+
+Here we'll look at the differences in generalization suggested by both the LME on localization shift and on reach aftereffects.
+In particular, we estimate the peak of each curve by bootstrapping participants and fitting a Gaussian (normal) function a 1000 times.
+This disregards other possible differences in generalization curves: the width of the curve, the peak amplitude or the offset.
+There may be effects there, but we are most interested in the location of the maxima.
+
+First, a figure:
+
+```{r fig5_generalization_maxima, fig.width=7.5, fig.height=3}
+plotGeneralization(generateSVG=FALSE,
+                   selectPerformance=selectPerformance)
+```
+
+## Localization
+
+There was an interaction between group and hand angle on localization shifts in the LME (far) above, which suggests different generalization curves in the two groups.
+Here are the confidence intervals for the peaks of the generalization curves for _active_ localization:
+
+```{r localization_peaks}
+getPeakLocConfidenceInterval(group='classic',
+                             CIs=c(.95), 
+                             movementtype='active', 
+                             LRpart='all',
+                             selectPerformance=FALSE)
+getPeakLocConfidenceInterval(group='exposure',
+                             CIs=c(.95), 
+                             movementtype='active', 
+                             LRpart='all', 
+                             selectPerformance=selectPerformance)
+```
+
+The median values are a little over 13 degrees apart, and median for classic falls just outside the exposure confidence interval.
+This could explain the interaction between group and hand angle on localization shifts, but the other way around it's not so clear.
+
+## Reach aftereffects
+
+Now we look at the confidence intervals for the maxima of the generalization curves of reach aftereffects after both exposure training and after classic training:
 
@@
 Something might be going on -- but the two confidence intervals are not different from each other, which is not in line with the LME results.
-We can't conclude much, except that we don't have strong evidence for different generalization curves.
-
-To sum up:
-
-1. Rotated exposure training evokes substantial shifts in no-cursor reach direction across the workspace,
-2. these changes persist over time, and
-3. these changes are comparable to the ones we found after classic training, and
-4. there is some generalization curve across the workspace, but
-5. we have no evidence that this is different from reach aftereffect generalization after classic training.
+We can't conclude much, except that we don't know if the differences in generalization curves comes from a difference in the location of the maxima, or perhaps the width of the curves, or both.
 
@@
 1.
-Exposure training leads to shifts in localization, that are not very different between active and passive localization.
-Since both active and passive movements access recalibrated proprioception, while active localization also accesses updated predictions, this indicates that hand localization is makes very good use of proprioception.
+**Exposure training leads to shifts in localization**, that are _indistinguishable_ for active and passive localization.
+Since both active and passive movements access recalibrated proprioception, while active localization also accesses updated predictions, this indicates that: _hand localization relies on (recalibrated) proprioception_.
 
@@
 2.
-Exposure training leads to robust reach aftereffects that are similar to those after classic training, although they seem to have a different generalization curve.
+**Exposure training leads to robust reach aftereffects** that are _indistinguishable_ in magnitude to those after classic training (although they seem to have a different generalization curve).
 This means that evoking propceptive recalibration with visual-proprioceptive discrepancies leads to motor changes.
-In other words: recalibrated proprioception informs motor control.
+In other words: _recalibrated proprioception informs motor control_.
 
@@
 However, it's not a large difference, and the analyses are ambiguous.
-Perhaps this is another cue that recalibrated proprioception dominates even reach aftereffects?
-Either way, we find no (statistically reliable) evidence that the updated predictions matter much for the generalization curves, which can have several causes.
+Either way, we find no (statistically reliable) evidence that the updated predictions matter much for the generalization curves, which can have several causes all outside the scope of this paper.
 
@@
 # This chunk creates SVG files but shouldn't show up in the PDF / HTML 
+plotLocalization(classicOnline=FALSE,
+                 generateSVG=TRUE, 
+                 selectPerformance=selectPerformance)
 plotReachAftereffects(generateSVG=TRUE,
                       selectPerformance=selectPerformance)
-plotLocalization(classicOnline=FALSE,
-                 generateSVG=TRUE, 
+plotGeneralization(generateSVG=TRUE,
                    selectPerformance=selectPerformance)

### 2019-03-30 localization.R

@@ -2,12 +2,14 @@
 source('shared.R')
 
 
-
 # PLOT / FIGURE ------
 
-plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerformance=TRUE, remove15=FALSE) {
+plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerformance=TRUE, remove15=FALSE, thirdPanel='2x2') {
   
-  classicOnline = FALSE # the third panel is now used for generalization curves
+  # thirdPanel=
+  # 'peakCIs'
+  # 'classicOnline'
+  # '2x2'
   
   # get the data to plot:
   exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE, selectPerformance=selectPerformance)
@@ -31,7 +33,7 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerfo
   cla.pas <- cla[which(cla$passive_b == 1),]
   cla.CI.pas <- matrix(unlist(by(cla.pas$taperror_deg, INDICES=c(cla.pas$handangle_deg), FUN=t.interval)),nrow=2)
 
-  if (classicOnline) {
+  if (thirdPanel == 'classicOnline') {
     onl <- getPointLocalization('online', difference=TRUE, verbose=FALSE, selectPerformance=FALSE)
     
     onl.avg <- aggregate(taperror_deg ~ passive_b + handangle_deg, data=onl, FUN=mean)
@@ -47,7 +49,8 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerfo
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
     if ('svglite' %in% installed.list) {
-      svglite(file='Fig2.svg', width=7.5, height=3, system_fonts=list(sans='Arial', mono='Times New Roman'))
+      library('svglite')
+      svglite(file='Fig3.svg', width=7.5, height=3, system_fonts=list(sans='Arial', mono='Times New Roman'))
     } else {
       generateSVG=FALSE
     }
@@ -103,7 +106,7 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerfo
   
   legend(10,-15,c('passive','active'),col=c(colorset[['claPasS']],colorset[['claActS']]),lty=c(1,1),lwd=c(1.5,1.5),bty='n')
   
-  if (classicOnline) {
[…]

### 2019-03-30 nocursor.R

@@ -3,13 +3,13 @@ source('shared.R')
 
 # FIGURE -----
 
-plotReachAftereffects <- function(generateSVG=FALSE, selectPerformance=TRUE, addNormalFits=TRUE) {
+plotReachAftereffects <- function(generateSVG=FALSE, selectPerformance=TRUE, addNormalFits=FALSE) {
   
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
     if ('svglite' %in% installed.list) {
       library('svglite')
-      svglite(file='Fig3.svg', width=7.5, height=3, system_fonts=list(sans='Arial', mono='Times New Roman'))
+      svglite(file='Fig4.svg', width=7.5, height=3, system_fonts=list(sans='Arial', mono='Times New Roman'))
     } else {
       generateSVG=FALSE
     }
@@ -380,7 +380,8 @@ plotBlinkDetection <- function(generateSVG=FALSE) {
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
     if ('svglite' %in% installed.list) {
-      svglite(file='Fig1.svg', width=4, height=4, system_fonts=list(sans='Arial', mono='Times New Roman'))
+      library('svglite')
+      svglite(file='Fig2.5.svg', width=4, height=4, system_fonts=list(sans='Arial', mono='Times New Roman'))
     } else {
       generateSVG=FALSE
     }

### 2019-03-30 relateLocalizationNoCursors.R

@@ -76,7 +76,8 @@ correlateNoCursorsLocalization <- function(NCpart='all', generateSVG=FALSE, sele
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
     if ('svglite' %in% installed.list) {
-      svglite(file='Fig4.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
+      library('svglite')
+      svglite(file='Fig6.svg', width=7.5, height=2.5, system_fonts=list(sans='Arial', mono='Times New Roman'))
     } else {
       generateSVG=FALSE
     }
@@ -230,5 +231,204 @@ plotCorrelationWithCI <- function(X,Y,colors=c('black','red'),main='main title',
   
   # and use that to show a line:
   lines(range(X), predict(this.lm, newdata=data.frame(X=range(X))), col=colors[2], lwd=1.5)
+  
+}
+
+# Generalization -----
+
+plotGeneralization <- function(generateSVG=FALSE, selectPerformance=TRUE) {
+  
+  if (generateSVG) {
+    installed.list <- rownames(installed.packages())
+    if ('svglite' %in% installed.list) {
+      library('svglite')
+      svglite(file='Fig5.svg', width=7.5, height=3, system_fonts=list(sans='Arial', mono='Times New Roman'))
+    } else {
+      generateSVG=FALSE
+    }
+  }
+  
+  
+  par(mfrow=c(1,3))
+  
+  # get the data to plot:
+  exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE, selectPerformance=selectPerformance)
+  cla <- getPointLocalization('classic', difference=TRUE, verbose=FALSE, selectPerformance=FALSE)
+  
+  # get the averages for the line plots:
+  exp.avg <- aggregate(taperror_deg ~ passive_b + handangle_deg, data=exp, FUN=mean)
+  exp.avg.act <- exp.avg[which(exp.avg$passive_b == 0),]
+  exp.avg.pas <- exp.avg[which(exp.avg$passive_b == 1),]
+  cla.avg <- aggregate(taperror_deg ~ passive_b + handangle_deg, data=cla, FUN=mean)
+  cla.avg.act <- cla.avg[which(cla.avg$passive_b == 0),]
[…]

### 2019-03-31 ManuscriptMaterials.Rmd
reorganizing bootstrapping maxima

@@
 Here we'll look at the differences in generalization suggested by both the LME on localization shift and on reach aftereffects.
-In particular, we estimate the peak of each curve by bootstrapping participants and fitting a Gaussian (normal) function a 1000 times.
+In particular, we estimate the peak of each curve by bootstrapping participants and fitting a Gaussian (normal) function 100000 times.
 This disregards other possible differences in generalization curves: the width of the curve, the peak amplitude or the offset.
-There may be effects there, but we are most interested in the location of the maxima.
+There may be effects there, but we didn't check, as we are mostly interested in the location of the maxima.
+
+Since this is a very time-consuming process (more than an hour on my laptop), the results of the bootstrapping analyses are saved in files and can be downloaded from the OSF repository.
+
+Alternatively, if you want to check the results, you can run this chunk to delete the pre-calculated results, which will force the code to re-calculate it.
+However, by default, the code here will only use 5000 iterations for bootstrapping, which will ensure that the confidence intervals are less precise than the ones we pre-calculated.
+
+```{r delete_download_genmaxfiles}
+# set delete=TRUE to force redoing the bootstrapping 
+# set download=TRUE to download pre-calculated confidence intervals instead
+generalizationMaximaFiles(delete=FALSE, download=TRUE)
+```
 
@@
 ```{r create_svgs, cache=FALSE, include=FALSE}
+
+# #########
 # This chunk creates SVG files but shouldn't show up in the PDF / HTML 
+# #########
+
 plotLocalization(classicOnline=FALSE,

### 2019-03-31 localization.R

@@ -611,6 +611,8 @@ getPeakLocConfidenceInterval <- function(group, CIs=c(.95), movementtype='active
   
   if (file.exists(filename)) {
     
+    cat(sprintf('\nloading peak LOC generalization from file for: %s\n',toupper(group)))
+    
     df <- read.csv(filename, stringsAsFactors=FALSE)
     
   } else {
@@ -625,7 +627,7 @@ getPeakLocConfidenceInterval <- function(group, CIs=c(.95), movementtype='active
     
     loc2 <- -1 * xtabs(taperror_deg ~ participant + handangle_deg,loc)
     
-    data <- bootstrapGaussianPeak(data=loc2,bootstraps=5000,mu=47.5,sigma=15,scale=10,offset=4,CIs=CIs)
+    data <- bootstrapGaussianPeak(data=loc2,bootstraps=iterations,mu=47.5,sigma=15,scale=10,offset=4,CIs=CIs)
     
     df <- data.frame('level'=names(data),'value'=data)
     

### 2019-03-31 nocursor.R

@@ -419,6 +419,8 @@ getPeakConfidenceInterval <- function(group,validate=FALSE,part='all',CIs=c(.95)
   
   if (file.exists(filename)) {
     
+    cat(sprintf('\nloading peak RAE generalization from file for: %s\n',toupper(group)))
+    
     df <- read.csv(filename, stringsAsFactors=FALSE)
     
   } else {

### 2019-03-31 relateLocalizationNoCursors.R

@@ -236,7 +236,7 @@ plotCorrelationWithCI <- function(X,Y,colors=c('black','red'),main='main title',
 
 # Generalization -----
 
-plotGeneralization <- function(generateSVG=FALSE, selectPerformance=TRUE) {
+plotGeneralization <- function(generateSVG=FALSE, selectPerformance=TRUE, remove15=FALSE) {
   
   if (generateSVG) {
     installed.list <- rownames(installed.packages())

### 2019-03-31 shared.R

@@ -99,6 +99,22 @@ load.DownloadDataframe <- function(url,filename) {
   
 }
 
+generalizationMaximaFiles <- function(delete=FALSE, download=FALSE) {
+  
+  if (delete) {
+    file.remove('LOC_peakCI_classic.csv',
+                'LOC_peakCI_exposure.csv',
+                'RAE_peakCI_classic.csv',
+                'RAE_peakCI_exposure.csv')
+  }
+  
+  if (download) {
+    cat('\nNOT IMPLEMENTED YET\n')
+  }
+  
+}
+
+
 t.interval = function(data, variance = var(data, na.rm=TRUE), conf.level = 0.95) {
   
   data <- data[!is.na(data)]

### 2019-04-26 localization.R
updated code for new figures/analyses

@@ -4,7 +4,7 @@ source('shared.R')
 
 # PLOT / FIGURE ------
 
-plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerformance=TRUE, remove15=FALSE, thirdPanel='2x2') {
+plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerformance=TRUE, remove15=FALSE, thirdPanel='2x2',  points=c(15,25,35,45,55,65,75)) {
   
   # thirdPanel=
   # 'peakCIs'
@@ -12,8 +12,8 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerfo
   # '2x2'
   
   # get the data to plot:
-  exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE, selectPerformance=selectPerformance)
-  cla <- getPointLocalization('classic', difference=TRUE, verbose=FALSE, selectPerformance=FALSE)
+  exp <- getPointLocalization('exposure', difference=TRUE, verbose=FALSE, selectPerformance=selectPerformance, points=points)
+  cla <- getPointLocalization('classic', difference=TRUE, verbose=FALSE, selectPerformance=FALSE, points=points)
   
   # get the averages for the line plots:
   exp.avg <- aggregate(taperror_deg ~ passive_b + handangle_deg, data=exp, FUN=mean)
@@ -58,15 +58,16 @@ plotLocalization <- function(classicOnline=FALSE, generateSVG=FALSE, selectPerfo
   
   par(mfrow=c(1,3))
   
-  points <- c(15,25,35,45,55,65,75)
+  #points <- c(15,25,35,45,55,65,75,85)
   
   # panel A: exposure localization (active vs. passive)
   
-  plot(-1000,-1000, main='exposure', xlab='hand angle [°]', ylab='localization shift [°]', xlim=c(10,80), ylim=c(0,-15), axes=F)
+  plot(-1000,-1000, main='exposure', xlab='hand angle [°]', ylab='localization shift [°]', xlim=c(min(points)-5,max(points)+5), ylim=c(0,-15), axes=F)
   
   mtext('A', side=3, outer=TRUE, at=c(0,1), line=-1, adj=0, padj=1)
   
   X <- c(points, rev(points))
+  
   exp.act.Y <- c(exp.CI.act[1,],rev(exp.CI.act[2,]))
   exp.pas.Y <- c(exp.CI.pas[1,],rev(exp.CI.pas[2,]))
   
[…]

### 2019-04-26 nocursor.R

@@ -40,7 +40,7 @@ plotReachAftereffects <- function(generateSVG=FALSE, selectPerformance=TRUE, add
   claY <- c(classicCI[1,],rev(classicCI[2,]))
   expY <- c(exposureCI[1,],rev(exposureCI[2,]))
   
-  plot(-1000,-1000, main='decay of reach aftereffects', xlab='target angle [°]', ylab='reach endpoint deviation [°]', xlim=c(10,80), ylim=c(0,15), axes=F)
+  plot(-1000,-1000, main='reach aftereffect decay', xlab='target angle [°]', ylab='reach endpoint deviation [°]', xlim=c(10,80), ylim=c(0,15), axes=F)
   
   mtext('A', side=3, outer=TRUE, at=c(0,1), line=-1, adj=0, padj=1)
   
@@ -413,13 +413,13 @@ plotBlinkDetection <- function(generateSVG=FALSE) {
 
 # GENERALIZATION PEAK CI -----
 
-getPeakConfidenceInterval <- function(group,validate=FALSE,part='all',CIs=c(.95), selectPerformance=selectPerformance,iterations=100000) {
+getPeakConfidenceInterval <- function(group,validate=FALSE,part='all',CIs=c(.95), selectPerformance=selectPerformance,iterations=1000) {
   
-  filename <- sprintf('RAE_peakCI_%s.csv', group)
+  filename <- sprintf('maxima_RAE_%s.csv', group)
   
   if (file.exists(filename)) {
     
-    cat(sprintf('\nloading peak RAE generalization from file for: %s\n',toupper(group)))
+    #cat(sprintf('\nloading peak RAE generalization from file for: %s\n',toupper(group)))
     
     df <- read.csv(filename, stringsAsFactors=FALSE)
     

### 2019-04-26 relateLocalizationNoCursors.R

@@ -236,7 +236,7 @@ plotCorrelationWithCI <- function(X,Y,colors=c('black','red'),main='main title',
 
 # Generalization -----
 
-plotGeneralization <- function(generateSVG=FALSE, selectPerformance=TRUE, remove15=FALSE) {
+plotGeneralization <- function(generateSVG=FALSE, selectPerformance=TRUE, remove15=FALSE, iterations=1000) {
   
   if (generateSVG) {
     installed.list <- rownames(installed.packages())
@@ -288,13 +288,15 @@ plotGeneralization <- function(generateSVG=FALSE, selectPerformance=TRUE, remove
                                                movementtype='active', 
                                                LRpart='all',
                                                selectPerformance=FALSE,
-                                               remove15=remove15)
+                                               remove15=remove15,
+                                               iterations=iterations)
   exp.locshift <- getPeakLocConfidenceInterval(group='exposure',
                                                CIs=c(.95), 
                                                movementtype='active', 
                                                LRpart='all', 
                                                selectPerformance=selectPerformance,
-                                               remove15=remove15)
+                                               remove15=remove15,
+                                               iterations=iterations)
   
   plot(-1000,-1000, main='localization shifts', xlab='hand angle [°]', ylab='localization shift [°]', xlim=c(10,80), ylim=c(0,-15), axes=F)
   
@@ -373,8 +375,8 @@ plotGeneralization <- function(generateSVG=FALSE, selectPerformance=TRUE, remove
   cla.fit <- getGaussianFit(x=classic$target,classic$endpoint_angle,mu=50,sigma=30,scale=50,offset=4)
   
   # get confidence intervals for the peak of the generalization curve for localization shifts:
-  cla.RAEshift <- getPeakConfidenceInterval('classic', part='all', CIs=c(.95), selectPerformance=selectPerformance)
-  exp.RAEshift <- getPeakConfidenceInterval('exposure', part='all', CIs=c(.95), selectPerformance=selectPerformance)
+  cla.RAEshift <- getPeakConfidenceInterval('classic', part='all', CIs=c(.95), selectPerformance=selectPerformance,iterations=iterations)
+  exp.RAEshift <- getPeakConfidenceInterval('exposure', part='all', CIs=c(.95), selectPerformance=selectPerformance,iterations=iterations)
   
   plot(-1000,-1000, main='reach aftereffects', xlab='target angle [°]', ylab='reach endpoint deviation [°]', xlim=c(10,80), ylim=c(0,15), axes=F)
   

### 2019-04-26 shared.R

@@ -19,6 +19,10 @@ localizationURLs <- c('exposure'='https://osf.io/ysduw/?action=download', 'class
 # information on exposure participants:
 informationURLs <- c('blinkdetect'='https://osf.io/yrpkm/?action=download', 'demographics'='https://osf.io/7sn4b/?action=download')
 
+maximaURLs <- c('LOC_classic'='https://osf.io/rp5gm/?action=download',
+                'LOC_exposure'='https://osf.io/s987y/?action=download',
+                'RAE_classic'='https://osf.io/bm4r5/?action=download',
+                'RAE_exposure'='https://osf.io/2m8g3/?action=download')
 
 
 # we'll control the color in figures centrally from here:
@@ -102,14 +106,23 @@ load.DownloadDataframe <- function(url,filename) {
 generalizationMaximaFiles <- function(delete=FALSE, download=FALSE) {
   
   if (delete) {
-    file.remove('LOC_peakCI_classic.csv',
-                'LOC_peakCI_exposure.csv',
-                'RAE_peakCI_classic.csv',
-                'RAE_peakCI_exposure.csv')
+    filenames <-  c('maxima_LOC_classic.csv',
+                    'maxima_LOC_exposure.csv',
+                    'maxima_RAE_classic.csv',
+                    'maxima_RAE_exposure.csv')
+    file.remove(filenames[file.exists(filenames)])
   }
   
   if (download) {
-    cat('\nNOT IMPLEMENTED YET\n')
+    
+    for (dv in c('LOC','RAE')) {
+      for (group in c('classic','exposure')) {
+        df <- read.csv(url(maximaURLs[sprintf('%s_%s',dv,group)]), stringsAsFactors=F)
+      
+        write.csv(df, sprintf('maxima_%s_%s.csv',dv,group), row.names=F, quote=F)
+      }
+    }
+    
   }
   
 }
@@ -309,7 +322,7 @@ getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55
     df <- df[which(df$participant %in% OKparticipants),]
   }
   
-  df <- aspligned(df)
+  df <- aspligned(df) # subtract smooth spline fitted on aligned data only from all data
   
   # we can select half the localization data... for the exposure group only
   if (group == 'exposure') {
@@ -598,6 +611,9 @@ removeOutliers <- function(df, stds=2) {
 }
 
 # CDF fits of generalization -----
+
+# these functions are not used but provide a first attempt at getting some reasonable
[…]

### 2019-04-26 ManuscriptMaterials.Rmd
restructured for revisions

@@
 This document collects statistics and figures to be used in the manuscript on the exposure consequences project.
+It downloads files from the accompanying Open Science Framework project: [https://osf.io/zfdth/](https://osf.io/zfdth/)
 
@@
 
+Is the age of this group of participants lower than that of the classic group? We can do a t-test against 24:
+
+```{r age_group_test}
+t.test(participants$age, mu=24)
+
+```
+
 And the distribution of sex:
@@
 
+So there are also more females now. Let's see if that is significantly different with a binomial exact test:
+
+```{r sex_group_test}
+binom.test(x=c(13,8),p=11/21)
+```
+
 ## Localization selection statistics
 
-We've manually selected localization responses that were unlikely to be true responses. Here are the statistics on how many were removed in each dataset. For exposure:
+We've manually selected localization responses that were unlikely to be true responses, and we removed data from reaches where the hand was outside the 0-90 degree span.
+Here are the statistics on how many were removed in each dataset.
+For exposure:
 
@@
                  data=countSelectedLocalizations('exposure'), 
-                 FUN=mean), 
-       caption='percentages selected localization trials after exposure training')
+                 FUN=mean), caption='percentages selected localization trials after exposure training')
 ```
@@
 We decided not to use any more outlier removal, but it is implemented and can be switched on in the code: set the `removeOutliers` argument to the `getPointLocalization` function in `shared.R` to TRUE.
+
+## Distribution of data
+
+Let's see how the data is distributed through the workspace. The plot below bins the data, with bins centred on the 7 angles we're interested in, using bins of 10 degrees wide.
+
+```{r loc_bin_counts, fig.width=8, fig.height=5}
+plotLocalizationBinCounts()
+```
+
+So we have at well over 100 data points close to every hand angle.
+But we want to know if there are participants without estimates of localization at some points in the workspace.
+So here are the counts of participants with valid localization estimates in all 4 tasks for every hand angle.
+Missing data occurs here, because we don't extrapolate beyond the range of the data.
+
+```{r localization_count_table}
+kable(getLocCountTable(selectPerformance=selectPerformance),
+      caption='count of participants with localization data at each hand angle')
+```
 
@@
 
-Here is a plot, probably figure 2:
+Here is a plot, probably figure 3:
 
 ```{r fig3_localization, fig.width=7.5, fig.height=3}
-plotLocalization(selectPerformance=selectPerformance,remove15=FALSE)
+plotLocalization(selectPerformance=selectPerformance,remove15=FALSE,points=c(15,25,35,45,55,65,75))
 ```
@@
 ## Missing data
-
-Here are the counts of participants with valid localization estimates in all 4 tasks for every hand angle. Missing data occurs here, because we don't extrapolate beyond the range of the data.
-
-```{r localization_count_table}
-kable(getLocCountTable(selectPerformance=selectPerformance),
-      caption='count of participants with localization data at each hand angle')
-```
 
@@
 
+## Localization precision
+
+Many people keep being interested in the precision of active versus passive localization.
+If active movements before localization result in higher precision, then the variance of the response errors should be lower for active movements.
+Let's test this in the aligned data for exposure:
+
+```{r exp_loc_var_check}
+exp_a <- getLocalizationVariance(group='exposure')
+t.test(taperror_deg ~ passive_b, data=exp_a, paired=T)
+```
+
+An alternative is to see how many participants were more precise in the active versus the passive localization task, using a binomial exact test.
+We compare the observed data to a 50% chance which would express the hypothesis that it doesn't matter for localization precision.
+
+```{r}
+binom.test( x=sum(aggregate(taperror_deg ~ participant, data=exp_a, FUN=diff)$taperror_deg < 0), n=21, p=.5 )
+```
+
+And for the classic group, we do the same:
+
+```{r cla_loc_var_check}
+loc_a <- getLocalizationVariance(group='classic')
+t.test(taperror_deg ~ passive_b, data=loc_a, paired=T)
+binom.test( x=sum(aggregate(taperror_deg ~ participant, data=loc_a, FUN=diff)$taperror_deg < 0), n=21, p=.5 )
+```
+
+How about the rotated session for the exposure group:
+
+```{r exp_rot_loc_var_check}
+exp_r <- getLocalizationVariance(group='exposure', session='rotated')
+t.test(taperror_deg ~ passive_b, data=exp_r, paired=T)
+binom.test( x=sum(aggregate(taperror_deg ~ participant, data=exp_r, FUN=diff)$taperror_deg < 0), n=21, p=.5 )
+```
+
+Or the rotated session for the classic group:
+
+```{r cla_rot_loc_var_check}
+loc_r <- getLocalizationVariance(group='classic', session='rotated')
+t.test(taperror_deg ~ passive_b, data=loc_r, paired=T)
+binom.test( x=sum(aggregate(taperror_deg ~ participant, data=loc_r, FUN=diff)$taperror_deg < 0), n=21, p=.5 )
+```
+
+So in all four cases, making active movements doesn't change the variance: it is neither increased nor decreased. As expected, this confirms all our other tests in previous/upcoming papers.
+
 # Reach Aftereffects
@@
 
-## Is there an effect of exposure training on reach aftereffects?
+## Is there an effect of exposure training on open-loop reaches?
 
@@
 So this shows that whether or not the feedback was rotated mattered for the reach deviations in the exposure training group, i.e.: there are reach aftereffects.
+This means that we subtract the aligned no-cursor reach end-point deviation for each participant and target from the rotated no-cursor reach end-point deviation from the same participant and target, to get the training induced _change_ in reach deviations.
 
@@
 The second analysis is to show that the reach aftereffects aren't all that different for first and later iterations of the rotated no-cursor tasks.
-(The rotated data is always compared with _all_ aligned iterations of the task.)
 
@@
 
-Alternatively, if you want to check the results, you can run this chunk to delete the pre-calculated results, which will force the code to re-calculate it.
-However, by default, the code here will only use 5000 iterations for bootstrapping, which will ensure that the confidence intervals are less precise than the ones we pre-calculated.
+Alternatively, if you want to thoroughly check the results, you can run this chunk to delete the pre-calculated results (set delete=TRUE and download=FALSE), which will force the code to re-calculate it.
+By default, the code here will use "only" 1000 iterations for bootstrapping, which will run 20 times faster, but also guarantees that the confidence intervals are less precise than the ones we pre-calculated.
 
@@
 plotGeneralization(generateSVG=FALSE,
-                   selectPerformance=selectPerformance)
+                   selectPerformance=selectPerformance,
+                   iterations=200)
 ```
@@
 
-There was an interaction between group and hand angle on localization shifts in the LME (far) above, which suggests different generalization curves in the two groups.
+There was an interaction between group and hand angle on localization shifts in the LME described (far) above, which suggests different generalization curves in the two groups.
 Here are the confidence intervals for the peaks of the generalization curves for _active_ localization:
@@
                              LRpart='all',
-                             selectPerformance=FALSE)
+                             selectPerformance=selectPerformance,
+                             iterations=1000)
 getPeakLocConfidenceInterval(group='exposure',
@@
                              LRpart='all', 
-                             selectPerformance=selectPerformance)
+                             selectPerformance=selectPerformance,
+                             iterations=1000)
 ```
@@
                           CIs=c(.95), 
-                          selectPerformance=selectPerformance)
+                          selectPerformance=selectPerformance,
+                          iterations=1000)
 getPeakConfidenceInterval('exposure', 
@@
                           CIs=c(.95), 
-                          selectPerformance=selectPerformance)
+                          selectPerformance=selectPerformance,
+                          iterations=1000)
 ```
@@
 # #########
-# This chunk creates SVG files but shouldn't show up in the PDF / HTML 
+# This chunk creates SVG files for all the figures but shouldn't show up in the PDF / HTML 
 # #########
@@
                  generateSVG=TRUE, 
-                 selectPerformance=selectPerformance)
+                 selectPerformance=selectPerformance,points=c(15,25,35,45,55,65,75))
 plotReachAftereffects(generateSVG=TRUE,
@@
 plotGeneralization(generateSVG=TRUE,
-                   selectPerformance=selectPerformance)
+                   selectPerformance=selectPerformance,
+                   iterations=1000)
 correlations <- correlateNoCursorsLocalization(generateSVG=TRUE, 

### 2019-04-26 paper/paper.Rmd
revisions of paper

@@
 ---
-title: "Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization After Passive Visuoproprioceptive Discrepancy Training"
+title: 'Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization
+  After Passive Visuoproprioceptive Discrepancy Training'
 author:
-  - name: Ahmed A. Mostafa
-    affiliation: [YORK,MANSOURA]
-#    affiliation: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
-  - name: Bernard Marius 't Hart
-    affiliation: YORK
+- affiliation:
+  - YORK
+  - MANSOURA
+  name: Ahmed A. Mostafa
+- affiliation: YORK
+  corresponding: thartbm@gmail.com
   email: thartbm@gmail.com
-    corresponding: thartbm@gmail.com
-  - name: Denise Y.P. Henriques
+  name: Bernard Marius 't Hart
+- affiliation: YORK
   email: deniseh@yorku.ca
-    affiliation: YORK
+  name: Denise Y.P. Henriques
 address:
-  - code: YORK
-    address: CVR / Kinesiology and Health Science, York University, Toronto, Ontario, Canada
-  - code: MANSOURA
-    address: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
-
-abstract: |
-  An accurate estimate of limb position is necessary for movement planning.
-Where we localize our unseen hand after a reach depends on felt hand position, or proprioception, but this is usually neglected in favour of predicted sensory consequences based on efference copies of motor commands.
-Both sources of information should contribute, so here we set out to further investigate how much of hand localization depends on proprioception and how much on predicted sensory consequences.
-We use a passive training paradigm with rotated visual feedback that eliminates the possibility to update predicted sensory consequences, but still recalibrates proprioception.
-After this training we measure participants’ hand location estimates based on both efference-based predictions and afferent proprioceptive signals with self-generated hand movements as well as based on proprioception only with robot-generated movements.
-We find indistinguishable shifts in hand localization after robot- and self-generated hand movements, and changes in open-loop reaches.
-Both motor and proprioceptive changes are only slightly smaller as those after training with self-generated movements, confirming that proprioception plays a large role in estimating limb position and in planning movements.
-(data: https://doi.org/10.17605/osf.io/zfdth, preprint: https://doi.org/10.1101/384941)
-
+- address: CVR / Kinesiology and Health Science, York University, Toronto, Ontario,
+    Canada
+  code: YORK
+- address: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
+  code: MANSOURA
+output:
+  html_document:
+    df_print: paged
+  pdf_document: default
+  rticles::plos_article: default
+  word_document: default
+csl: plos.csl
+fig_caption: yes
 header-includes:
@@
 bibliography: bibliography.bib
-output: rticles::plos_article
-csl: plos.csl
+abstract: |
+  An accurate estimate of limb position is necessary for movement planning, before and after motor learning.
+Where we localize our unseen hand after a reach depends on felt hand position, or proprioception, but in studies and theories on motor adaptation this is quite often neglected in favour of predicted sensory consequences based on efference copies of motor commands.
+Both sources of information should contribute, so here we set out to further investigate how much of hand localization depends on proprioception and how much on predicted sensory consequences.
+We use a training paradigm combining robot controlled hand movements with rotated visual feedback that eliminates the possibility to update predicted sensory consequences (‘exposure training’), but still recalibrates proprioception, as well as a classic training paradigm with self-generated movements in another set of participants.
+After each kind of training we measure participants’ hand location estimates based on both efference-based predictions and afferent proprioceptive signals with self-generated hand movements (‘active localization’) as well as based on proprioception only with robot-generated movements (‘passive localization’).
+In the exposure training group, we find indistinguishable shifts in passive and active hand localization, but after classic training, active localization shifts more than passive, indicating a contribution from updated predicted sensory consequences.
+Both changes in open-loop reaches and hand localization are only slightly smaller after exposure training as compared to after classic training, confirming that proprioception plays a large role in estimating limb position and in planning movements, even after adaptation.
+(data: https://doi.org/10.17605/osf.io/zfdth, preprint: https://doi.org/10.1101/384941)
 ---
@@
 Sensory information is central to how we control all our movements.
-Our brain is even thought to use predicted sensory consequences derived from efferent copies of motor commands for motor control @Duhamel1992a.
-When training with rotated visual feedback of the hand, we update these predictions @Wolpert1995.
-Additionally, such training leads to a recalibration of our sense of felt hand position - “proprioception” - to be more aligned with the distorted visual feedback @Cressman2009.
+Our brain is even thought to use predicted sensory consequences derived from efference copies of motor commands for motor control @Duhamel1992a.
+When training with rotated visual feedback of the hand, we update these predictions @Wolpert1995, and these updated predictions contribute to motor control.
+Additionally, such training leads to a recalibration of our sense of felt hand position based on afferent signals (that we call “proprioception” here) to be more aligned with the distorted visual feedback @Cressman2009.
 Both of these changes have been measured by asking people to localize where their unseen hand is – before and after training [@Clayton2013; @Izawa2012b; @Jones2012a; @Synofzik2008; @Yavari2016].
-While our lab has previously found that proprioception accounts for a large portion of the change in hand localization @Hart2016, it is far from clear how much each process contributes or how to tease them apart.
-Here we use passive training, that removes the need to update predicted sensory consequences, in an attempt to isolate the contribution of proprioception to hand localization.
+While our lab has previously found that proprioception accounts for a substantial portion of the change in hand localization, and should hence contribute to motor adaptation @Hart2016, it is far from clear how much each process contributes or how to tease them apart.
+Here we use training with robot-generated reaches, that prevents efference-based predictions and eliminates differences between actual and expected visual feedback and hence prevents updating predicted sensory consequences.
+We then compare hand localization data from this experiment to hand localization in a classic training paradigm with self-generated movements, from an – otherwise identical – earlier study @Hart2016, in an attempt to isolate the contribution of proprioception to hand localization prior to and after adaptation.
 
-The predicted sensory consequences of movements may play several important roles in motor learning and control.
+The predicted sensory consequences of movements likely play several important roles in motor learning and control.
 Predicted sensory consequences allow us to correct our movements before sensory error signals are available, they can be used to select movements that best achieve our goals and they may inform us on the location of our limbs.
 Hence measuring predicted sensory consequences is valuable in movement research.
-In visuomotor rotation adaptation tasks, the actual sensory outcome is systematically different from the expected outcome, so that participants update their predictions on the outcome of the trained movements.
-In previous experiments, people were asked to make a movement and then indicate the location of, or “localize,” their unseen, right hand, before and after training with rotated visual feedback [@Izawa2012b; @Synofzik2008].
+In motor adaptation, the actual sensory outcome is systematically different from the expected outcome, so that participants update their predictions on the outcome of the trained movements.
+In two previous visuomotor adaptation experiments, people were asked to make a movement and then indicate the location of, or “localize,” their unseen, right hand, before and after training with rotated visual feedback [@Izawa2012b; @Synofzik2008].
 Since there was no visual information available to the participants, the predicted sensory consequences of the movement should be used in localizing the unseen hand.
@@
 
-However, our lab has shown that our sense of where we feel our hand to be, proprioception, is also reliably recalibrated after visuomotor rotation adaptation [@Cressman2009; @Hart2016; @Mostafa2014; @Mostafa2015; @Maksimovic2018; @Henriques2012; @Cressman2010a; @Barkley2014; @Salomonczyk2011; @Nourouzpour2014; @Henriques2014].
-This has also been shown by other labs @Cameron2012 and a comparable proprioceptive change is induced with force-field adaptation @Ostry2016.
-As proprioception also informs us on the location of our limbs, we have on occasion used a task that is very similar to hand localization to investigate this [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018].
-Although proprioceptive recalibration has been largely ignored as an explanation for changes in hand localization, we and others have shown that it accounts for a substantial part of the changes in localization, along with updates in predicted sensory consequences [@Hart2016; @Cameron2012].
+However, there is another explanation for shifts in hand localization after visuomotor adaptation.
+Our lab, and others, have shown that our sense of where we feel our hand to be, proprioception, is also reliably recalibrated after visuomotor adaptation [@Cressman2009; @Hart2016; @Mostafa2014; @Mostafa2015; @Maksimovic2018; @Henriques2012; @Cressman2010a; @Barkley2014; @Salomonczyk2011; @Nourouzpour2014; @Henriques2014; @Cameron2012], and a comparable proprioceptive change is induced with force-field adaptation [@Ostry2016; @Ohashi2019].
+As proprioception, like predicted sensory consequences, also informs us on the location of our limbs, we have on occasion used a task that is very similar to hand localization to investigate this [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018].
+Although proprioceptive recalibration has been largely ignored as an explanation for changes in hand localization, we and others have shown that it accounts for a substantial part of the changes in localization, along with, but separate from, updates in predicted sensory consequences [@Hart2016; @Cameron2012].
 Nevertheless, it is far from clear how much proprioception and prediction each contribute to hand localization.
 
-Here we intend to further examine the contribution of proprioception to hand localization.
-To do this, we use passive training, where a robot arm moves the participant’s arm out, so that the cursor always directly goes to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013].
+Here we intend to further examine the contribution of proprioception to hand localization, relative to the contribution of predicted sensory consequences.
+To do this, we use a passive training paradigm with robot-generated  movements, which we refer to as ‘exposure training’, where the cursor always goes directly goes to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013].
 This means there is no efference copy available and no visuomotor error-signal, both of which are required to update predicted sensory consequences.
-However, we impose a discrepancy between vision and proprioception that drives proprioceptive recalibration.
-Thus, changes in hand localization after this type of training should be due to proprioceptive recalibration only.
-We use the same experimental protocol as before @Hart2016, so that we can compare localization shifts between the two different types of training, and can better assess the contributions of predicted sensory consequences and proprioception to hand localization.
+That is, exposure training to visuoproprioceptive discrepancy should not lead to updates of predicted sensory consequences.
+However, the discrepancy between vision and proprioception does drive proprioceptive recalibration.
+Thus, changes in any kind of hand localization after this type of training should be due to proprioceptive recalibration only.
+We use the same experimental protocol as before @Hart2016, where classic visuomotor training should have changed predicted sensory consequences along with proprioception.
+This allows comparing shifts in hand localization between the two different types of training, and a better assessment of the contributions of predicted sensory consequences and proprioception to hand localization.
+
+Primarily, we will compare the magnitude of shifts in passive and active hand localization after visuoproprioceptive discrepancy exposure and classic visuomotor training.
+After exposure training there should be no difference, while after classic training, predicted sensory consequences should be updated but only contribute to active localization, so that active and passive localization changes should be different.
+For a secondary analysis we will compare the magnitude of reach aftereffects between exposure and classic training.
+As exposure training should not lead to updates of internal models, exposure training should lead to lower reach aftereffects than classic training.
+However, if there are any reach aftereffects at all after exposure training this means that afferent-based localization shifts drive motor changes too, and here we can see to what degree that is the case.
 
@@
 We set out to test the relative contributions of proprioception and efference-based prediction to hand localization.
-We use visual training with robot-generated hand movements to prevent updates of predicted sensory consequences, but still elicit proprioceptive recalibration.
+We use exposure training, with robot-generated reaches to prevent updates of predicted sensory consequences, but still elicit proprioceptive recalibration, and compare this with a classic training paradigm that should change both predicted sensory consequences and proprioception.
+We then tested this in both active localization, that should use both sources – updated or not – and in passive localization, that should rely only on proprioception, which should be recalibrated equally by both training paradigms.
 
@@
 
-Twenty-five right-handed participants were recruited for the passive learning task in this study.
-One participant was excluded for not following task instructions, and three were excluded for low performance on a task that ensures attention during the passive training.
-All analyses presented here pertain to the remaining twenty-one participants (13 females and 8 males, mean age: 20.1 +/- 2.3 years), but the data of the three low-performing participants is included in the online \href{https://osf.io/zfdth}{dataset} @Mostafa2017.
+Twenty-five right-handed participants were recruited for the exposure training group in this study.
+One participant was excluded for not following task instructions, and three were excluded for low performance on a task that ensures attention during the exposure training.
+All analyses presented here pertain to the remaining twenty-one participants (mean age: 20.1±2.3, age range: 18-25, 13 females), but the data of the three low-performing participants is included in the online dataset @Mostafa2017 ([https://osf.io/zfdth](https://osf.io/zfdth/)).
 Participants provided prior, written informed consent in accordance with the ethical guidelines set by the York Human subjects Review Subcommittee (#2014-240).
@@
 Participants were screened verbally, and all reported being right handed and not having any history of visual, neurological, and/or motor dysfunction.
-Published data from 21 further participants is used for comparison @Hart2016.
+Published data from 21 further participants (mean age: 24, age range: 18-38, 11 females) that completed classic visuomotor training is used for comparison @Hart2016.
 
-![**Setup and tasks:** **a)** Participants moved their unseen right hand with visual feedback on hand position provided through a mirror (middle surface) half-way between their hand and the monitor (top surface).
-A touchscreen located just above the hand was used to collect responses for the localization tasks (bottom surface).
+![**Setup and tasks** **a)** The participants held on to the handle of a robotic manipulandum with their unseen right hand.
+Visual feedback on hand position was provided through a mirror (_middle surface_) half-way between the hand and the monitor (_top surface_).
+A touchscreen located just above the hand was used to collect responses for the localization tasks (_bottom surface_).
 **b)** Training task.
@@
 While reaching to one of these targets, no visual feedback on hand position is provided.
-d) Localization task.
-The participants’ unseen, right hands moved out, and subsequently participants indicated the direction of the hand movement by indicating a location on an arc using a touch screen with their visible left index finger](Fig1_setup.eps)
+**d)** Localization task.
+The participants’ unseen, right hands moved out and back, and subsequently participants indicated the direction of the hand movement by indicating a location on an arc using a touch screen with their visible left index finger.](Fig1_setup.eps)
 
@@
 
-Participants sat in a height-adjustable chair to ensure that they could easily see and reach all targets presented on a reflective screen (see Fig 1).
+Participants sat in a height-adjustable chair to ensure that they could easily see and reach all targets.
+To ensure no visual information on right hand position was available to localize the right hand, the room lights were dimmed and the participants’ view of their right hand was blocked by a horizontal surface, as well as a dark cloth draped between a touch screen and participants’ right shoulders.
 During all tasks, they held the vertical handle on a two-joint robot manipulandum (Interactive Motion Technologies Inc., Cambridge, MA) with their right hand so that their thumb rested on top of the handle.
-A monitor (Samsung 510 N, refresh rate 60 Hz) was mounted 11 cm above the reflective screen, such that images displayed on the monitor appeared to lie in the horizontal plane where the right hand was moving.
-The reflective screen was mounted horizontally 18 cm above the robot manipulandum.
-A touch screen was mounted 13 cm underneath the reflective surface, so that subjects could indicate the location of the unseen right-hand locations (specifically the unseen thumb) with their visible left hand, which was lit up with a small spot light (only in localization tasks).
-The room lights were dimmed and the participants’ view of their right hand was blocked by the reflective screen, as well as a dark cloth draped between the touch screen and participants’ right shoulders.
+A horizontal touch screen, reflective surface and monitor were mounted 5, 18 and 29 cm above the handle, respectively.
+Stimuli were presented using the monitor (Samsung 510 N, refresh rate 60 Hz), which was mounted facing downward, 11 cm above the reflective surface, such that images displayed on the monitor and viewed in the reflective surface appeared to lie in the horizontal plane where the right hand was moving.
+The reflective surface was mounted horizontally 18 cm above the robot manipulandum.
+The touch screen was mounted 13 cm underneath the reflective surface (5 cm above the manipulandum), so that subjects could indicate the location of the unseen right hand (specifically the thumb on top of the handle) with their left hand.
+The touch screen always blocked vision of the right hand, but the left hand was lit up with a small spot light (only in localization tasks) in order to be visible through the reflective surface for accurate touch screen responses.
+For a schematic illustration of the setup, see Fig 1a.
 
@@
 
-The first part of the experiment used training with a cursor aligned with the hand and the second part had training with a cursor rotated around the start position (Fig 1b; white rows in Table 1).
+The first part of the experiment used training with a cursor aligned with the hand and the second part had training with a cursor rotated around the start position (Fig 1b; Fig 2a).
 During the training with rotated feedback, the cursor was gradually rotated 30° clockwise.
 This introduced a discrepancy between the actual, or felt, hand position and the visual feedback, that should evoke proprioceptive recalibration.
-However, the movements were robot generated, so that there were no predicted sensory consequences based on the outgoing motor command.
-Hence the prediction errors that are thought to lead to motor learning were absent as well.
-After both types of training, participants did open-loop reaches as well as two kinds of hand localization tasks, to test the effect of training on proprioceptive and predictive hand estimates.
+However, the movements were robot generated, and always brought the cursor straight to the target, so that there were no motor commands generated, and no predicted sensory consequences based on this outgoing command.
+In the absence of predicted sensory consequences and with perfect, error-free reach performance, there were no sensory prediction errors to drive classic visuomotor adaptation either @Wong2019.
+After both types of training, participants did two kinds of hand localization tasks, to test the effect of training on proprioceptive and predictive hand estimates, as well as open-loop reaches as a classic measure of implicit adaptation.
 
-\newcommand{\headrow}{\rowcolor{black!20}}
-\newcommand{\thead}[1]{\multicolumn{1}{l}{\bfseries #1\rule[-1.2ex]{0pt}{2em}}}
-
-\definecolor{LGray}{gray}{0.9}
-\definecolor{DGray}{gray}{0.8}
-
-\begin{table}[bt]
-\caption{Task order}
-\begin{threeparttable}
-\begin{tabular}{cccc}
-\headrow
-\multicolumn{2}{c}{\# trials} & \multicolumn{2}{c}{training type} \\
-\headrow
-%\multicolumn{2}{|c|}{\# trials} & \thead{Exposure training} & \thead{Classic training} \\
-\thead{aligned} & \thead{rotated} & \multicolumn{1}{c}{\textbf{Exposure training}} & \multicolumn{1}{c}{\textbf{Classic training}} \\
-\rowcolor{white}
-50 & 90 & training & training \\
-\rowcolor{LGray}
-- & 21 & no-cursor & no-cursor \\
-\rowcolor{white}
-- & 60 & training & training \\
-\rowcolor{DGray}
-25 & 25 & active delayed localization & active delayed localization \\
-\rowcolor{LGray}
-21 & 21 & no-cursor & no-cursor \\
-\rowcolor{white}
-10 & 60 & training & training \\
-\rowcolor{DGray}
-25 & 25 & passive delayed localization & passive delayed localization \\
-\rowcolor{LGray}
-21 & 21 & no-cursor & no-cursor \\
-\rowcolor{white}
-10 & 60 & training & training \\
-\rowcolor{DGray}
-25 & 25 & active delayed localization & \textit{(active online localization)} \\
-\rowcolor{LGray}
-21 & 21 & no-cursor & no-cursor \\
-\rowcolor{white}
-10 & 60 & training & training \\
-\rowcolor{DGray}
-25 & 25 & passive delayed localization & \textit{(passive online localization)} \\
-\rowcolor{LGray}
-21 & 21 & no-cursor & no-cursor \\
-\hiderowcolors
-\hline  % Please only put a hline at the end of the table
-\end{tabular}
-
-\begin{tablenotes}
-\item 
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
+![**Experimental design** **a)** Task order.
+All participants performed active and passive localization as well as no-cursor reaching tasks in both a first, aligned session and a second, rotated session.
+There is an extra no-cursor reach block in the rotated session, to assess decay.
+The number of trials (numbers above the rectangles) in the no-cursor reaching tasks and both types of localization were the same in the aligned and rotated sessions, but while aligned training was kept to a minimum, rotated training was increased to ensure sufficient adaptation.
+The Exposure training group did passive and active localization twice, but the Classic training group did different localization tasks in the second half of each session, the data of which is not used here.
+**b)** Our main question was about the signals contributing to hand localization in the four different kinds of localization.
+According to our hypothesis, updated predicted sensory consequences should have only contributed to active hand localization after Classic training.](Fig2.eps)
 
@@
 
-In what we call ‘exposure training’ the participants did not move their hand toward the target, but the robot did.
-In this task (Table 1), the right hand was represented by a cursor (green disk, 1 cm in diameter, Fig 1b) located directly above participant’s thumb.
-The robot moved the participant’s unseen right hand (and the cursor) along a direct path toward a visual yellow target disk and back to the starting position (1 cm in diameter, Fig 1b).
+In what we called ‘exposure training’ the participants did not move their hand toward the target, but the robot did.
+In this task (Fig 2a, gray boxes), the right hand was represented by a cursor (green disk, 1 cm in diameter, Fig 1b) located directly above participant’s thumb.
+The robot moved the participant’s unseen right hand (and the cursor) along a direct path, with a constant velocity of 10 cm/s, toward a visual yellow target disk and back to the starting position (1 cm in diameter, Fig 1b).
 The home position was located approximately 20 cm in front of participants and the visual target located 10 cm from the home position at 45° (Fig 1b).
@@
 
-During the first half of the experiment, the cursor and hand path were aligned during exposure training.
-In the second part of the experiment, the “rotated” session, the same visual training target at 45° was used, and the cursor kept moving straight to this target.
-However, the robot-generated hand path gradually rotated 30° CCW (Fig 1b) with respect to the visible target and the cursor in increments of 0.75°/trial, so that the full rotation was reached after 45 trials.
+During the first half of the experiment, the “aligned” session, the cursor and hand path were aligned during exposure training.
+In the second part of the experiment, the “rotated” session, the visuoproprioceptive discrepancy was introduced.
+The same visual training target at 45° was used, and the cursor kept moving straight to this target.
+However, the robot-generated hand path gradually rotated 30° CCW (Fig 1b) with respect to the cursor and target, in increments of 0.75°/trial, to prevent awareness of the rotation, so that the full rotation was reached after 45 trials.
 This mimicked error-free responses to a gradual visuomotor rotation of 30° CW.
 The initial training consisted of 50 trials in the aligned part and 90 in the rotated part.
-In between open-loop reach tasks and localization tasks (Table 1) extra training tasks were done, each of which consisted of 10 trials in the aligned part of the experiment and 60 trials in the rotated part (Table 1).
+In between open-loop reach tasks and localization tasks (Fig 2, outlined and blue boxes) extra training tasks were done, each of which consisted of 10 trials in the aligned part of the experiment and 60 trials in the rotated part (Fig 2, gray boxes).
 
@@
 
-The trials in no-cursor reaching (Table 1, light gray rows) serve as a classical measure of motor adaptation.
+The trials in no-cursor reaching (Fig 1c; Fig 2, outlined boxes) served as a standard measure of motor adaptation.
 On each of these trials participants were asked to reach with their unseen right hand to one of 7 visual targets, without any visual feedback of hand position.
@@
 A trial started with the robot handle at the home position and, after 500 ms, the home position disappeared and the target appeared, cuing the participants to reach for the target.
-Once the participants thought they had reached the target they held their position for 250 ms, and the target and the home position disappeared, cuing participants to move back to the home position along a straight, constrained path, to begin the next trial.
-If participants tried to move outside of the path, a resistance force, with a stiffness of 2 N/(mm/s) and a viscous damping of 5 N/(mm/s), was generated perpendicular to the path.
-In every iteration of the no-cursor reach task, each target was reached three times, for a total of 21 reaches in pseudo-random order.
+Once the participants thought they had reached the target they held their position for 250 ms, and the target disappeared, while the home position re-appeared, cuing participants to move back to the home position along a straight, constrained path, to begin the next trial.
+The path was constrained so that if participants tried to move outside of the path, a resistance force, with a stiffness of 2 N/(mm/s) and a viscous damping of 5 N/(mm/s), was generated perpendicular to the path.
+In every iteration of the no-cursor reach task, each target was reached for three times, for a total of 21 reaches in pseudo-random order.
 The no-cursor reaching task was performed four times in the aligned part of the experiment and five times in the rotated part of the experiment.
+The end-points of no-cursor reaches, in particular the angles relative to the start position and their deviations from the angles of the targets were analyzed, as these typically change in classic visuomotor adaptation.
 
@@
 
-In this task (Fig 1d; Table 1, dark gray rows) participants indicated where they thought their unseen right hand was after a movement.
-First, an arc appeared, spanning from 0° to 90° and located 10 cm away from the home position and the participants’ unseen, right hand moved out from the home position in a direction towards towards a point on the arc.
-The hand was stopped by the robot at 10 cm from the home position and then, to prevent the online proprioceptive signals from overriding the predictive signals [@Izawa2012b; @Hart2016], the hand was moved back to the home position using the same kind of constrained path as used for the return movements in the no-cursor task.
-Participants indicated with the index finger of their visible, left hand on the touch screen mounted directly above the robot handle where they thought their trained hand had crossed the arc.
+In this task (Fig 1d; Fig 2, blue boxes) participants indicated where they thought their unseen right hand was after a movement.
+First, an arc appeared, spanning from 0° to 90° and located 10 cm away from the home position and the participants’ unseen, right hand moved out from the home position in a direction towards a point on the arc.
+The hand was stopped by the robot at 10 cm from the home position and then, to prevent online proprioceptive signals from overriding the predictive signals [@Izawa2012b; @Hart2016], the hand was moved back to the home position using the same kind of constrained path as used for the return movements in the no-cursor task.
+Participants indicated with the index finger of their visible, left hand on the touch screen mounted directly above the robot handle where they thought their trained hand had ended its movement under the arc.
 
 Crucially, there were two variations of this task.
-First, in the ‘active’ localization task participants generated the movement themselves, as they could freely move their unseen right hand from the home position to any point on the arc.
-Second, there was a ‘passive’ localization task where the robot moved the participants’ hand out and back, to the same locations the participants moved to in the preceding ‘active’ localization task in a shuffled order (hence, active localization is done first).
-In active localization, participants have access to both proprioceptive information as well as an efference-based prediction of sensory consequences, but in passive localization, only proprioception should be available.
+First, in the ‘active’ localization task participants generated the movement themselves, as they could freely move their unseen right hand from the home position to any point on the arc (Fig 2a, dark blue boxes).
+Second, there was a ‘passive’ localization task where the robot moved the participants’ hand out and back (Fig 2a, light blue boxes), to the same locations the participants moved to in the preceding ‘active’ localization task in a shuffled order.
+Hence, active localization was always done first.
+In active localization, participants had access to both afferent, proprioceptive information as well as an efference-based prediction of sensory consequences, but in passive localization, only proprioception should have been available.
 The active and passive localization task each consisted of 25 trials, and each of the tasks was done a total of four times; twice after aligned and twice after rotated training.
+From these tasks, the localization responses were analyzed, i.e.
+where participants indicated the endpoint of their hand movement was, in particular the angles of the indicated points relative to the home position in comparison to the true reach endpoints’ angle.
 
@@
 
-The paradigm described above is an exact replica of a paradigm we used earlier @Hart2016 with two exceptions.
-First, we used exposure training here, instead of the standard reach training with volitional movements, which we will call ‘classic’ training.
-Second, all localization is delayed until the right hand has returned to the home position in this study (see Table 1), so that instead of both delayed and online localization we have two repetitions of each delayed localization task.
-With this paradigm we can compare changes in localization and no-cursor reaches change after exposure training with changes in the same measures after classic visuomotor adaptation training.
+The paradigm described above was an exact replica of a paradigm we used earlier @Hart2016 with three exceptions.
+First, we used exposure training here, instead of the standard reach training with volitional movements, which we called ‘classic’ training.
+Second, all localization was delayed until the right hand has returned to the home position in this study, whereas the previous study also included a second set of “online” localization tasks, where touch screen responses were collected when the unseen, right hand was still at the furthest point of the movement.
+Instead of both delayed and online localization we had two repetitions of each delayed localization task here.
+Third, we added cursor blink detection to the training tasks, to make sure participants attended to the task.
+In the present study we compared changes in localization responses and no-cursor reach endpoints after exposure training with changes in the same measures after classic visuomotor adaptation training, as collected earlier.
 
@@
 
-Localization responses were taken as the (signed) angular difference between vectors through the home position and the actual hand position as well as the location indicated on the touch screen.
-Prior to analyses, idiosyncratic differences in performing this task were countered.
-Before conversion to degrees angle, a circle with a 10 cm radius was fit to the touch screen responses of each participant and the offset of this circle’s centre was subtracted from all response coordinates, so that all responses fell close to the arc.
-Then, a smoothed spline was fit to every participant’s response errors in each of the four localization tasks (aligned vs.
+We primarily wanted to test if hand localization shifts after exposure training, whether or not there were differences between active and passive localization shifts, and finally compare localization shifts after exposure training to those after classic visuomotor training.
+
+There were some idiosyncratic differences in performing localization (e.g.
+systematic under- or overshoot of the arc by some participants).
+To counter this, we fitted a circle with a 10 cm radius to the touch screen responses of each participant (using a least-squared errors algorithm) and the offset of this circle’s centre was subtracted from all response coordinates, so that all responses fell close to the arc.
+Localization responses were taken as the (signed) angular difference between vectors through the home position and both the actual hand position as well as the location indicated on the touch screen.
+For each localization response, we retain the angle of the actual reach endpoint, as well as the signed angular error of the localization response.
+Since participants freely chose their movement directions, we interpolated angular localization errors at the same angles used for the no-cursor reaches (15°, 25°, 35°, 45°, 55°, 65° and 75°), but only if that location fell within the range of the data (i.e.
+we did not extrapolate).
+For interpolation we used a smoothed spline that was fit to every participant’s response errors in each of the four localization tasks (aligned vs.
 rotated and active vs.
-passive) and these were used to obtain localization errors at the same locations used for the no-cursor reaches (15°, 25°, 35°, 45°, 55°, 65° and 75°), but only if that location fell within the range of the data (i.e.
-we only interpolate).
+passive).
 This way localization responses could be compared across participants despite the freely chosen reach directions.
 At the 15° location 7/21 participants didn’t have an estimate in one or more of the four localization tasks (in the “classic” data it was also 7/21).
-While that data is shown in the figures, we did not use it for analysis.
+While that data was shown in the figures, we did not use it for analysis.
 
-First we test if localization responses shifted following rotated exposure training compared to aligned.
-We then test if the shift in localization responses is different for active and passive localization, and we run analyses comparing localization after exposure training with localization after “classic” training.
-Finally, we explore the generalization of localization responses and if they are different between the groups doing classic and exposure training.
+First we compared localization responses shifted following rotated exposure training to those following aligned, and confirmed that training shifted localization.
+We then tested if the shift in localization responses is different for active and passive localization, and we ran analyses comparing localization after exposure training with localization after “classic” training.
+Finally, we test if the variance of active localization responses is lower than the variance of passive localization responses.
 
@@
 
-To assess any reach adaptation that may have occurred after exposure training we analyzed reach endpoint errors in no-cursor trials.
+For our secondary question, we wanted to assess if any reach adaptation had occurred after exposure training, so we analyzed reach endpoint errors in no-cursor trials.
 Reach endpoint errors were the (signed) angular difference between a vector from the home position to reach endpoint and a vector from the home position to the target.
-We obtained reach aftereffects by subtracting reach endpoint errors after aligned training from those after rotated training.
-No-cursor endpoint errors were analyzed to see if participants adapted the direction of their reaching movements after rotated exposure training.
+We obtained reach aftereffects by subtracting median reach endpoint errors for each of the 7 targets after aligned training from those after rotated training.
+No-cursor endpoint errors were analyzed to test if participants adapted the direction of their reaching movements after rotated exposure training, which would indicate that recalibrated proprioception contributes to motor changes.
 We also tested if any such change decayed, i.e.
 if it was the same immediately after exposure training, or when a localization task was done in between exposure training and no-cursor reaches.
-Furthermore, we tested if the generalization of reach aftereffects is different between exposure and classic training and if there is any generalization of reach aftereffects.
+Finally, we compare reach aftereffects between the exposure and classic training groups.
 
-Pre-processing and analyses were done in R 3.4.4 @RCoreTeam2017, using lme4, lmerTest and various other packages.
-Most analyses used linear mixed effects models, since there is some missing localization data.
-These were “converted” to more readable ANOVA-like output, using a Satterthwaite approximation @Luke2017.
+## Generalization maxima
+
+We noticed that the generalization curves after exposure and classic training looked different, and the planned analyses indeed showed interactions between hand angle and training paradigm.
+Specifically, the maxima after classic training are very close to the target, as expected, but for exposure training, the maxima are shifted (see Results).
+Since this is potentially interesting, we added two purely exploratory (i.e.
+unplanned) analyses of the location of the maxima in the workspace.
+The same analysis was done for active localization and reach aftereffects, comparing generalization after exposure and classic training.
+Note that these analyses do not answer our main questions.
+
+Pre-processing and analyses were done in R 3.4.4 @RCoreTeam2017 using lme4, lmerTest and various other packages.
+There is some systematically missing localization data, which means we can’t run ANOVA’s on the full data set.
+We also can’t reliably extrapolate or impute beyond the range of responses, and removing hand locations (4/7) or participants (7/21) would remove too much data, decreasing power.
+Hence we decided not to use regular ANOVAs but linear mixed effects models instead, as these are robust against missing data.
+The output was converted to more readable ANOVA-like output, using a Satterthwaite approximation @Luke2017.
 Highly similar results were obtained with a Chi-square approximation.
@@
 
-In short, this experiment allowed us to test how mere exposure to a visual-proprioceptive discrepancy changes both reach aftereffects and hand localization responses, and compare them with those obtained after more regular, “classic” training.
+In short, this experiment allowed us to test how mere exposure to a visual-proprioceptive discrepancy changed both reach aftereffects and hand localization responses, and compare them with those obtained after more common, “classic” training.
 
@@
 
-In this study we intend to further elucidate the relative contributions of (updated) predicted sensory consequences and (recalibrated) proprioception to hand localization.
-We can parcel out these contributions by measuring hand localization after both robot-generated and self-generated movements.
-Finally we compare the data from the current experiment with those obtained in an earlier study that used an identical paradigm, but trained with self-generated movements, or “classic” training.
+In this study we intend to further elucidate the relative contributions of (updated) predicted sensory consequences and (recalibrated) proprioception to training-induced shifts in hand localization.
+We can parcel out these contributions by measuring hand localization after both robot-generated (afferent signals only) and self-generated movements (both afferent and efferent signals).
+Finally, all analyses are performed with the data from the current experiment and those obtained in an earlier study that used an identical paradigm, but trained with self-generated movements, or “classic” training.
 
-![**Hand localization:** The shifts of the angles of touchscreen responses in all variations of the localization task, using spline-interpolated estimates for hand angles matching the reach targets in the no-cursor reach block.
+```{r fig3, echo=FALSE, message=FALSE, fig.width=5.25, fig.height=2.5, fig.cap="\\label{fig:fig3}**Hand Localization** The shifts of the angles of touchscreen responses in all variations of the localization task, using spline-interpolated estimates for hand angles matching the reach targets in the no-cursor reach tasks.
 **a)** Localization shifts after exposure training.
-Dark blue: active localization shifts, Light blue: passive localization shifts.
+_Dark blue_: active localization shifts, _Light blue_: passive localization shifts.
 **b)** Localization shifts after classic training.
-Dark red: active localization shifts, Orange: passive localization shifts.
+_Dark red_: active localization shifts, _Orange_: passive localization shifts.
 The dashed line segments illustrate that the 15° data is not used for statistical analyses (see Methods).
-**c)** Generalization curves of active localization shifts after exposure training (blue) and classic training (red).
-Shaded areas: 95\% confidence intervals for the peak of the generalization curve (red and blue lines through shaded area indicate 50\% points).
-Downward black arrow: visual trained target.
-Upward black arrow: hand location during training.](Fig2_plain.pdf)
+**c)** The overall localization shifts for active and passive localization after both exposure and classic training.
+_Blue_: localization shifts after exposure training, _Red_: localization shifts after classic training."}
+setwd('..')
+source('localization.R')
+plotLocalization(selectPerformance=TRUE,remove15=FALSE)
+```
 
@@
 
-Here we test our hypothesis that exposure training does not lead to changes in predicted sensory consequences.
+Here we test our hypothesis that exposure training with a visual-proprioceptive discrepancy does not lead to changes in predicted sensory consequences.
+First, we can see that the difference between rotated and aligned localization responses is different from zero, after exposure training (Fig 3), so that it seems there are _shifts_ in localization that we can compare between tasks and training paradigms.
+To verify this we fit an LME to the localization errors throughout the workspace using _session_ (aligned or rotated), _movement type_ (active and passive) and _hand angle_ (25°, 35°, 45°, 55°, 65° and 75°), and all interactions as fixed effects and _participant_ as random effect.
+There is an effect of _session_ (F(1,450.5)=155.8; p<0.001), showing that training with exposure to a visual-proprioceptive discrepancy leads to changes in hand localization.
+There is also an effect of _hand angle_ (F(5,450.6)=6.54; p<.001) and an interaction between _hand angle_ and _session_ (F(5,450.3)=8.25; p<.001), but no other effects (all p>.60).
+Since localization responses shift, we use the difference between hand localization after rotated training and after aligned training as a measure of localization shift.
+
 Since the difference between active and passive localization stems only from the presence or absence of predicted sensory consequences, there should be no difference between the two if predicted sensory consequences are not changed by exposure training.
-At first glance, it seems there might be a difference between active and passive localization shifts in exposure training (see Fig 2a), although it is smaller than in classic training (Fig 2b).
+At first glance, it seems there might be a difference between active and passive localization shifts in exposure training (see Fig 3a), although it is smaller than in classic training (Fig 3b).
+Also noteworthy is that the overall magnitude of the localization shifts seems higher after exposure training as compared to classic (Fig 3c).
+This is unexpected, but we do find a similar pattern in other work @Ruttle2018.
 
-To test if rotated exposure training induces changes in hand localization, we fit an LME to the localization errors throughout the workspace using session (aligned or rotated), movement type (active and passive) and hand angle (25°, 35°, 45°, 55°, 65° and 75°), and all interactions as fixed effects and participant as random effect.
-There was an effect of session (F(1,450.5)=155.8; p<0.001), showing that exposure training leads to changes in hand localization.
-There was also an effect of hand angle (F(5,450.6)=6.54; p<.001) and an interaction between hand angle and session (F(5,450.3)=8.25; p<.001), which we’ll explore below, but no other effects (all p>.60).
-Since localization responses did shift, we use the difference between hand localization after rotated training and after aligned training (as plotted in Fig 2) for further tests.
+If, contrary to our hypothesis, this shift in localization after exposure training partly reflects predicted sensory consequences, then shifts in active localization, that rely on both (recalibrated) proprioception and (updated) predictions should be different from shifts in passive localization that only rely on (recalibrated) proprioception.
+We fit a linear mixed effects model (LME) to the change in localization using _movement type_ (active or passive localization) and _hand angle_, as well as their interaction as fixed effects and _participant_ as random effect.
+As we expected, there was no effect of _movement type_ (F(1,211.8)=0.07; p=0.79).
+There was an effect of _hand angle_ (F(5,212.2)=10.8; p<0.001), but no interaction between _hand angle_ and _movement type_ (F(5,211.8)=1.23, p=.29).
+The lack of an effect of movement type means we don’t have evidence that predicted sensory consequences contribute to localization in this paradigm.
 
-If this shift in localization after exposure training partly reflects predicted sensory consequences, then shifts in active localization, that rely on both (recalibrated) proprioception and (updated) predictions should be different from shifts in passive localization that only rely on (recalibrated) proprioception.
-We fit an LME to the change in localization using movement type (active or passive localization) and hand angle, as well as their interaction as fixed effects and participant as random effect.
-There was no effect of movement type (F(1,211.8)=0.07; p=0.79).
-There was an effect of hand angle (F(5,212.2)=10.8; p<0.001), but no interaction between hand angle and movement type (F(5,211.8)=1.23, p=.29).
-The lack of an effect of movement type suggests that predicted sensory consequences did not contribute to localization in this paradigm.
-
-In order to compare hand localization shifts after exposure training with those after classic training @Hart2016, we fit an LME to localization shift using training type (exposure vs.
-classic), movement type (active vs.
-passive) and hand angle and all interactions as fixed effects and participant as random effect.
-There was a main effect of movement type (F(1,422.0)=6.22; p=.013) and of hand angle (F(5,422.7)=8.19; p<.001), as well as an interaction between training type and hand angle (F(5,422.7)=4.54; p<.001) and between training type and movement type (F(1,422.0)=4.48, p=.035), but there was no main effect of training type (F(1,39.1)=0.92, p=.34) and no other effects (all p>.14).
+In order to compare hand localization shifts after exposure training with those after classic training @Hart2016, we fit an LME to localization shift using _training type_ (exposure vs.
+classic), _movement type_ (active vs.
+passive) and _hand angle_ and all interactions as fixed effects and _participant_ as random effect.
+There is a main effect of _movement type_ (F(1,422.0)=6.22; p=.013) and of _hand angle_ (F(5,422.7)=8.19; p<.001), as well as an interaction between _training type_ and _hand angle_ (F(5,422.7)=4.54; p<.001) and between _training type_ and _movement type_ (F(1,422.0)=4.48, p=.035), but there is no main effect of _training type_ (F(1,39.1)=0.92, p=.34) and no other effects (all p>.14).
 These results also suggest that the magnitude of the shifts in localization are comparable between classic and exposure training, but that the pattern of generalization is different.
 
-To address our main question, we will look at the interaction between training type (exposure vs.
-classic) and movement type (active vs.
-passive) we found above.
-Since there is no difference between active and passive localization shifts after exposure training alone, the interaction between training type and movement type should be caused by an effect of movement type on the localization shifts after classic training, as we found previously @Hart2016.
+To address our main question, we will consider the interaction between _training type_ (exposure vs.
+classic) and _movement type_ (active vs.
+passive) we report above (see also Fig 3c).
+The overall larger shifts after exposure training could be due to inherent differences between the groups, or because either exposure training magnifies proprioceptive recalibration, or because the sensory prediction error based learning present in classic training dampens other mechanisms @Wong2019, such as proprioceptive recalibration.
+However, since there is no difference between active and passive localization shifts after exposure training alone, the interaction between _training type_ and _movement type_ should be caused by an effect of _movement type_ on the localization shifts after classic training, as we found previously @Hart2016.
 This means that shifts in hand localization after exposure training indeed rely on recalibrated proprioception alone, while after classic training, there also is a contribution of predicted sensory consequences to active localization.
 
-For completeness, we explore the potentially different generalization patterns of localization shifts after classic and exposure training (Fig 2c).
-The LME indicates no difference in overall amplitude of localization shifts between the groups, so the interaction between training type and hand angle might stem from a generalization that does not peak at the trained location after exposure training.
-Using the active localization shifts only (which are larger, and arguably more similar to reach aftereffects), we bootstrap a 95\% confidence interval for the peak localization shift across participants in each group.
-Here we include the data at 15° where it is available.
-After classic training, the peak localization shift was at 48.8° (95\% confidence: 31.6° - 62.8°; red area in Fig 2c), and after exposure training the peak localization shift was at 62.1° (95\% confidence: 50.0° - 75.6°; blue are in Fig 2c).
-This means that peak localization after classic training is lower than after exposure training, but not vice versa.
-Also note that the confidence interval for the peak localization shift after classic training includes the trained target (45°), but not after exposure training @Izawa2012b.
-In short, the LME for localization shifts indicates a different generalization curve after exposure and classic training, which might be partially explained by a different position of the peak localization shift after exposure or classic training.
+### Localization variance
+
+In active movements there could have been more information available to localize the hand, which could have lead to more precise responses.
+To test this, we calculated the variance of localization responses relative to a smoothed spline expressing the bias of localization responses in each participant across the workspace.
+A paired-sample t-test showed that the variance of active and passive localization were not different from each other in the aligned session (t(20)=1.178, p=.253) as well as the rotated session (t(20)=-1.066, p=.299) in the exposure training group.
+This confirmed earlier findings where we also did not find that active localization was more precise than passive localization [@Jones2010b; @Jones2012a; @Hart2016].
 
@@
 
-Apart from proprioception and prediction, we want to see if rotated exposure training has any effect on open-loop reaches and if these are robust.
-We measure whether participants adapted their reach directions by assessing their reach errors in no-cursor reach trials after aligned and rotated exposure training.
-In Fig 3, the changes in no-cursor endpoint errors, or reach aftereffects, appear to be well over 5°.
-First, to test if exposure training affects open-loop reach direction, we fit a linear mixed effects model (LME) to reach endpoint error using session (aligned; all blocks, or rotated; only the first block immediately after training) and target (15°, 25°, 35°, 45°, 55°, 65° and 75°), as well as their interactions as fixed effects and participant as random effect.
-There is an effect of session (F(1,260)=93.81, p<.001), that is: exposure training leads to substantial reach aftereffects.
-There was no effect of target (F(6,260)=1.07, p=.37) and no interaction (F(6,260)=1.00, p=.42).
-Since there is an effect of session, we now take the differences in reach endpoint errors between the rotated and aligned session for every participant and target as reach aftereffects, and use those for all further analyses.
+Apart from our main question on the contribution of proprioception and prediction to hand localization, we want to see if rotated exposure training has any effect on open-loop reaches and if these persisted.
+We quantify whether participants adapt their reach directions by assessing their reach errors in no-cursor reach trials after aligned and rotated exposure training.
+In Fig 4, the changes in no-cursor endpoint errors, or reach aftereffects, appear to be well over 5°.
+First, to confirm that exposure training affects open-loop reach direction, we fit a linear mixed effects model (LME) to reach endpoint error using _session_ (aligned; all blocks, or rotated; only the first block immediately after training) and _target_ (15°, 25°, 35°, 45°, 55°, 65° and 75°), as well as their interactions as fixed effects and _participant_ as random effect.
+There is an effect of _session_ (F(1,260)=93.81, p<.001), but there is no effect of _target_ (F(6,260)=1.07, p=.37) and no interaction (F(6,260)=1.00, p=.42).
+Since there is an effect of _session_, we can say there are reach aftereffects, and now take the differences in reach endpoint errors between the rotated and aligned session for every participant and target as “reach aftereffects”, and use those for all further analyses.
+Importantly, the effect of _session_; i.e.
+open-loop reaches after rotated exposure training are different from those after aligned training, also indicates that exposure training evokes motor changes usually associated with other adaptation mechanisms.
 
-![**Reach aftereffects:** Changes of the angle of reach endpoints in the no-cursor tasks.
-**a)** Reach aftereffects across the experiment.
-Light blue: first no-cursor task in the rotated session (immediately following training), Dark blue, dashed line: the other four repetitions of the task (with localization in between training and no-cursor tasks).
+```{r fig4, echo=FALSE, fig.width=7.5, fig.height=3,fig.cap="\\label{fig:fig4}**Reach aftereffects** Changes of the angle of reach endpoints in the no-cursor tasks after rotated training.
+**a)** Reach aftereffects at different time points in the rotated session.
+_Light blue_: first no-cursor task in the rotated session (immediately following training), _Dark blue, dashed_ line: the other four repetitions of the task (with localization in between training and no-cursor tasks).
 **b)** Reach aftereffects after classic and exposure training.
-Blue: exposure training, Red: classic training.
-**c)** Generalization curves of reach aftereffects after exposure training (blue) and classic training (red).
-Shaded areas: 95\% confidence intervals for the peak of the generalization curve (red and blue lines through shaded area indicate 50\% points).
-Downward black arrow: visual trained target.
-Upward black arrow: hand location during training.](Fig3_plain.pdf)
+_Blue_: exposure training, _Red_: classic training.
+"}
+setwd('..')
+source('nocursor.R')
+plotReachAftereffects(selectPerformance=TRUE)
+```
 
-To see if reach aftereffects decayed during the localization tasks, we compared reach aftereffects in the initial no-cursor block, that immediately followed training, with those in the later blocks that followed a localization task.
-We fit an LME to the reach aftereffects with iteration (initial vs.
-later no-cursor blocks) and target (as above) as well as their interaction as fixed effects and participant as random effect.
-There is no effect of iteration (F(1,260)=2.72, p=0.10).
-There was an effect of target (F(6,260)=6.29, p<.001) but no interaction (F(6,260)=0.58, p=0.74).
-Hence, reach aftereffects were not appreciably different right after training and after the localization tasks.
-In other words, there was likely no noticeable decay of reach aftereffects during the localization tasks, so that we can collapse the data across iterations.
+To see if reach aftereffects decay during the localization tasks, we compare reach aftereffects in the initial no-cursor block, that immediately follow training, with those in the later blocks that follow a localization task.
+We fit an LME to the reach aftereffects with _iteration_ (initial vs.
+later no-cursor blocks) and _target_ (as above) as well as their interaction as fixed effects and participant as random effect.
+There is no effect of _iteration_ (F(1,260)=2.72, p=0.10).
+There is an effect of _target_ (F(6,260)=6.29, p<.001) but no interaction (F(6,260)=0.58, p=0.74).
+In sum, reach aftereffects are not appreciably different right after training and after the localization tasks.
+In other words, there is likely no noticeable decay of reach aftereffects during the localization tasks, so that we can collapse the data across iterations.
 
-Next we compare the reach aftereffects after classic training with those after exposure training (Fig 3b).
+Next we compare the reach aftereffects after classic training with those after exposure training (Fig 4b).
 It appears as if there is little overall difference in magnitude, but there might be a shift of the generalization curve.
-We fit an LME to reach aftereffects with training type (classic vs.
-exposure), target (as above) as well as their interaction as fixed effects and participant as random effect.
-There is no main effect of training (F(1,40)=0.11, p=.74), indicating approximately equal magnitude of reach aftereffects after the two training types.
-There is an effect of target (F(6,240)=8.36, p<.001), indicating that reach aftereffects exhibit some form of a generalization curve.
-There is also an interaction between training type and target (F(6,240)=2.27, p=.038), indicating these generalization curves are different after the two training types.
+We fit an LME to reach aftereffects with _training type_ (classic vs.
+exposure), _target_ (as above) as well as their interaction as fixed effects and _participant_ as random effect.
+Contrary to our expectation, and earlier findings @Henriques2012, there is no main effect of _training type_ (F(1,40)=0.11, p=.74), indicating approximately equal overall magnitude of reach aftereffects after the two training types.
+There is an effect of _target_ (F(6,240)=8.36, p<.001), indicating that reach aftereffects exhibit some form of a generalization curve.
+There is also an interaction between _training type_ and _target_ (F(6,240)=2.27, p=.038), indicating these generalization curves are different after the two training types.
+To summarize, we find that exposure training elicits substantial and persistent motor changes and we don’t find a difference in the magnitude of reach aftereffects between those after classic and exposure training.
 
-We will explore these potentially different generalization curves here.
-In Fig 3c we can observe that reach aftereffects after exposure training seem not to peak at the trained target direction of 45° but a more forward direction.
-We test this by taking the 95\% confidence interval of the centre of a normal curve fit to this data, bootstrapped across participants, and find that the median peak of the generalization curve of reach aftereffects after exposure training is at 66.3°, with a 95\% confidence interval ranging from 49.3° to 78.9°.
-This would indeed suggest that the reach aftereffects after exposure training do not generalize around the trained direction of 45°, but at a more counter clockwise location.
-For classic training, generalization of reach aftereffects peaks at 53.2°, with a 95\% confidence interval spanning 42.1° to 66.5°.
-So for classic training the 95\% confidence interval for peak reach aftereffects does include the trained target.
-These confidence intervals also indicate that generalization of reach aftereffects does not peak at different target position after exposure and classic training.
-However, we can also observe that the full curve was not sampled after exposure training, so that curve fitting is not optimal.
-This means that – given our data – the interaction between target and training type found in the LME above can’t be explained by a shifted generalization curve.
-This may be because our experiment was not set up to test this, and the similar angles where generalization peaks in both localization and reach aftereffects, suggests there may be a difference in where proprioception and prediction generalize strongest.
+## Generalization maxima
 
-In summary, our main hypotheses are confirmed; exposure training leads to shifts in hand localization that are not different for active or passive localization, while movement type does have an effect on localization shift after classic training.
-Exposure training also causes robust reach aftereffects that are of comparable size to those found with classic training.
-There is some evidence that the generalization of both localization shifts and reach aftereffects are different after the two training types, and it appears this can partially be explained by a different peak of the generalization curves, but our data and analyses are not definitive.
+While not our main question, we will add exploratory analyses on the potentially different generalization patterns of localization shifts and reach aftereffects after classic and exposure training (Fig 5).
+The LMEs above indicate no difference in overall amplitude of localization shifts or reach aftereffects between the groups, so the interaction between training type and hand angle (or target for reach aftereffects) might stem from a shifted generalization curve after exposure training, in contrast to classic training where the maximum is close to the trained target @Krakauer2000.
+Using the active localization shifts only (which are larger, and should arguably be more similar to reach aftereffects, because they include predicted sensory consequences), we bootstrap a 95\% confidence interval for the maximum localization shift, by resampling participants within each group 100,000 times and taking the centre of a normal curve fit.
+This procedure indicates that after classic training, the maximum localization shift is at 48.3° (95\% confidence: 36.9° - 57.8°; red area in Fig 5a), and after exposure training the maximum localization shift is at 62.5° (95\% confidence: 55.4° - 70.8°; blue area in Fig 5a).
+This means that the maximum effect of classic training on localization, occurs at a different point in the workspace than that of exposure training.
+
+```{r fig5, echo=FALSE, fig.width=7.5, fig.height=3,fig.cap="\\label{fig:fig5}**Generalization maxima** Locations of generalization maxima in the workspace for active localization and reach aftereffects.
+_Blue_: Exposure training, and _red_: classic training, with shaded areas at the bottom marking the 95% confidence intervals for the location of the maximum of each curve (vertical red and blue lines in shaded areas indicate the median; colored arrows pointing downward from the curves indicate the maxima of the curves fitted to all data).
+Downward black arrows: visual trained target at 45°.
+Upward black arrows: hand location during training at 75°.
+**a)** Generalization maxima of active localization shifts.
+**b)** Generalization maxima of reach aftereffects."}
+setwd('..')
+source('relateLocalizationNoCursors.R')
+plotGeneralization(selectPerformance=TRUE)
+```
+
+In Fig 5b we can observe a similar pattern for reach aftereffects.
+Using the same bootstrapping procedure, we find that the maximum of the generalization curve of reach aftereffects after exposure training is at 70.2°, with a 95\% confidence interval ranging from 65.7° to 153.5°.
+This very high upper bound can be explained by the fact that our experiment didn’t sample the full distribution (nor was it meant to), so that a curve with a peak at such a location could be an accurate fit of the data.
+However, we can at least take the lower bound as the minimum location in the workspace where reach aftereffects maximally generalize, as repeated bootstrapping procedures all converge on values around 65°.
+In comparison, the generalization of reach aftereffects after classic training, peaks at 52.0°, with a 95\% confidence interval spanning 47.0° to 56.3°.
+This means that the generalization curves for reach aftereffects after exposure and classic training have maxima at different directions relative to the trained direction.
+
+The LMEs for localization shifts and reach aftereffects indicate a different generalization curve after exposure and classic training, and this can be explained by a different position of the generalization curves after the two types training.
+The unrealistically high upper bound for the confidence interval of the peak of the generalization curve of reach aftereffects after exposure training, indicates that we should sample a wider part of the workspace in future experiments to get more accurate fits.
+However, the different generalization functions after classic and exposure training indicate that the mechanisms of adaptation that are employed by participants in each of the training paradigm are substantially different, as we have found before @Mostafa2015, even when both affect hand localization and open-loop reach directions.
+In classic training, sensory prediction errors mainly drive adaptation @Wong2019, and these are more or less anchored to the visual target at 45° [@Krakauer2000; @Day2016].
+In exposure training, cross-sensory discrepancy drives adaptation, such that the experienced afferent signals from the trained hand position (75°) are newly associated with the rotated visual feedback.
+Most likely, the different error signals underlying each type of training are the source of the differences in generalization.
+An open question is how these two generalization functions interact in behavior.
+
+The differences between the generalization curves for both reach aftereffects and hand localization shifts after both classic and exposure training are not well understood.
+Perhaps the error-signals underlying both may provide a hypothesis as the basis for future work.
+On the one hand, in classic training, what gets updated is the movement that should be used to reach a certain visual target using a visual cursor.
+Hence the reach aftereffects, and perhaps hand localization shifts should be strongest close to the visual target [@Krakauer2000].
+On the other hand, in exposure training, people do not have to make any movements, but the proprioceptive feedback they experience is associated with different visual feedback.
+Hence the hand localization shifts, and perhaps reach aftereffects should be strongest close to the experienced hand position.
+This hypothesis matches the patterns we observe here, but requires experimental verification.
+
+Regardless of the differences in generalization maxima, exposure training leads to substantial shifts in hand localization that are not different for active or passive localization, while movement type does have an effect on localization shift after classic training.
+Exposure training also causes persistent reach aftereffects that are not different in magnitude to those found with classic training.
 
@@
 
-The position of limbs is important for planning and evaluating movements, and can be estimated through predicted sensory consequences, as well as visual and proprioceptive feedback.
+Accurate information on where our limbs are is important for planning and evaluating movements, and can be estimated through predicted sensory consequences, as well as visual and proprioceptive feedback.
 As in a previous study @Hart2016 here we quantify the contributions of predicted sensory consequences and proprioceptive recalibration to where we localize our hand after training with altered visual feedback of the hand.
@@
 After classical training we previously reported larger shifts in active localization as compared to passive @Hart2016.
-As we expected, after exposure training there are substantial shifts in localization, but no difference between active and passive localization, indicating that predictions are not updated after exposure training.
+As we expected, after exposure training there are substantial shifts in localization, but no differences between active and passive localization, indicating that predictions are not updated after exposure training.
+Somewhat unexpectedly, localization shifts after exposure training are slightly larger than after classic training.
 Furthermore, we find that exposure training evokes substantial and robust reach aftereffects, indicating that recalibrated proprioception is used to plan movements.
 
-Our lab previously investigated proprioceptive recalibration and reach aftereffects following visuomotor adaptation with classic training and matched exposure training.
-There we also found that proprioceptive recalibration is of similar magnitude in both training paradigms, but unlike here, reach aftereffects are usually much larger with classic training [@Henriques2012; @Henriques2014; @Ruttle2018; @Cressman2010; @Salomonczyk2013].
+Our lab previously investigated proprioceptive recalibration and reach aftereffects following visuomotor adaptation with classic training and matched exposure training, however without comparing active with passive localization.
+In those studies we usually find that proprioceptive recalibration is of similar magnitude in both training paradigms, but unlike here, reach aftereffects are usually much larger with classic training [@Henriques2012; @Henriques2014; @Ruttle2018; @Cressman2010; @Salomonczyk2013].
+In one study @Ruttle2018, we observed that localization shifts after exposure training are larger than after classic training – like here.
+Perhaps focusing on the sensory information increases proprioceptive recalibration in exposure training, or engaging in sensory prediction error based learning reduces the need to engage other adaptive mechanisms, such as proprioceptive recalibration.
 And while proprioceptive recalibration and reach aftereffects do proportionally increase with gradual increases in rotation size for classical training, they do not for exposure training @Salomonczyk2013.
-The similar magnitude of proprioceptive recalibration and reach aftereffects following exposure training, but not classical training, suggest that this sensory recalibration is partly driving this modest change in movements.
-The effect of exposure training on movements is also demonstrated by facilitation by exposure training of subsequent classic training @Sakamoto2015 (but no interference) and transfer of exposure training effects from one hand to the other @Bao2017.
-In the current study, we further demonstrate that exposure training affects movements and proprioception, but also measure its potential effect on predictive estimates.
+The similar magnitude of proprioceptive recalibration and reach aftereffects following exposure training, but not classical training, suggest that sensory recalibration is driving this modest change in movements, with additional motor changes after classic training through updates of internal models.
+The effect of exposure training on movements is also demonstrated by savings and interference from exposure training carrying over to subsequent classic training @Sakamoto2015 and transfer of exposure training effects from one hand to the other @Bao2017.
+In the current study, we further demonstrate that exposure training affects both movements and proprioception, and verify that it likely has little effect on predictive estimates.
 
 Results similar to what we find here were reported in a study by Cameron and colleagues @Cameron2012, using gain modulation of visual feedback of single-joint hand movements around the elbow.
-Their within-subjects experiment included both training with volitional movements as well as with passive movements and also tested perception of movements that were either passive or active.
+Their within-subjects experiment included both training with volitional movements as well as with machine-generated movements and also tested perception of movements that were either passive or active.
 They too found a robust change in passive perception of hand movement (using a different measure), and these changes did not differ between the two types of training.
@@
 Like here, these shifts are larger after classic training as compared to exposure training.
-They also found that passive exposure leads to reach aftereffects, although these were smaller than those produced following “classical” training with altered visual gain.
+They also found that exposure to a visuo-proprioceptive discrepancy leads to reach aftereffects, although these were smaller than those produced following “classical” training with altered visual gain.
 Both our findings, and those of Cameron et al.
-@Cameron2012 indicate that updating predicted sensory consequences requires volitionally controlled movements that lead to prediction errors, while proprioception recalibrates equally in both types of training, and that recalibrated proprioception affects open-loop reaches.
-Our combined results suggest that updates in predicted sensory consequences only provide a partial explanation for motor learning.
+@Cameron2012 indicate that updating predicted sensory consequences requires volitionally controlled movements that lead to sensory prediction errors, while proprioception recalibrates equally in both types of training, and that recalibrated proprioception affects open-loop reaches.
+Our combined results suggest that updates in predicted sensory consequences only provide a partial explanation for visuomotor adaptation.
 
-Two related concerns about exposure training and passive localization are that the movements are not fully passive, so that efference-based predictions are still generated or that predicted sensory consequences are generated through another route.
+Two related concerns about exposure training and passive localization are that first, the movements might not be not fully passive, so that efference-based predictions are still generated, and second, that predicted sensory consequences could be generated through another route, e.g., based on afferent signals.
 Cameron et al.
-@Cameron2012 measured muscle activity (EMG) during passive movements and found no difference with stationary baseline muscle activity.
-This suggests that any movements generated in a passive condition are subthreshold, minimizing efference-based predictions.
+@Cameron2012 measured muscle activity (EMG) during passive` movements and found no difference with stationary baseline muscle activity.
+This implies that no movements are generated in any passive condition or they are subthreshold, so that efference-based predictions are at worst minimal.
 The brain areas generating predicted sensory consequences could also rely on afferent signals.
-However, such afferent signals are present in both active and passive movements, and if they would result in the same predictions, there would be no difference between active and passive localization after classic training, and no difference between the effects of exposure and classic training, and we find both are different.
+However, such afferent signals are present in both active and passive movements, and if they would result in the same predictions, there would be no difference between active and passive localization after classic training, but there is.
 Hence, while we can not fully exclude any predictive signals in passive localization or exposure training, our data shows that any residual predictive signals in the passive movements we used are qualitatively very different from normal efference-based predicted sensory consequences.
 
-In our classical training group, we not only see shifts for passive localization but even larger shifts for active localization which is consistent with a change in both proprioception and an update in predictions.
-In our exposure training, we did not find a consistent difference between active and passive localization, and none at the trained direction.
+Apart from these concerns, it might be that active localization is more precise than passive localization, as it is based on more information.
+However, here we confirm previous findings [@Jones2010; @Jones2012a, @Hart2016] that show this is not the case, and recently replicated this in a much larger dataset @tHart2018.
+This sets limits for the kinds of mechanism we can hypothesize are used by the brain to combine efferent with afferent signals, as well as perhaps the nature of the afferent signals used in localization.
 Assuming that predictions were not updated in exposure training, a maximum likelihood estimate (MLE) or “optimal integration” @Ernst2002 would predict that active localization should shift less than passive localization after exposure training.
-But of course this is not the case in our findings (although it is the case for Cameron et al.).
-This suggests that perhaps these two signals are not optimally integrated which is consistent with our comparisons of the variance between passive and active localization.
-In ’t Hart and Henriques @Hart2016, we tested the prediction derived from MLE that hand localization with two signals – proprioception and prediction in active localization – should be more reliable, i.e.
-have lower variance, than hand localization with only one signal – proprioception only in passive localization.
-However, we found no difference in variance between active and passive localization, and recently replicated this in a much larger dataset @tHart2018.
-Taken together, this suggests these different sources of information about unseen hand location are not optimally integrated.
+But this is not the case in our findings (although it is the case for Cameron et al., @Cameron2012).
+Taken together, this suggests these different sources of information about unseen hand location are probably not optimally integrated.
+It also suggests that the afferent signals in active localization are not more informative than afferent signals in passive localization.
 While localizing the unseen hand is less precise than locating (pointing to) a remembered visual target or a seen and felt hand location, we find that these bimodal estimates are rarely integrated optimally [@Fiehler2010; @Byrne2013; @Jones2010], although others have @VanBeers2002.
@@
 
-## Generalization
-
-We do find some evidence that, after exposure training, the generalization curves for localization shifts are not centred on the visual location of the trained target; they don’t peak at 45° but at ~62°.
-In contrast, after classic training the peak of the generalization curve does peak close to the training target.
-It is possible that proprioceptive recalibration is not anchored to the visual goal of the training task as it is not a requirement to feel your hand at any specific point; rather, in classic training the visual cursor has to be brought to a visual target to end a trial.
-While in exposure training the movements are executed for the participants without error, their task is still to pay attention to the visual cursor while it moves to the visual target; there are no task demands on proprioception.
-Although we can’t substantiate this here, the generalization curves of the reach aftereffects seem to mimic the generalization curves of localization shifts, suggesting a relationship between changes in state estimates and changes in movements.
-This needs to be tested further, but if these effects are true, they may provide insight into how state estimates are used to produce movements, and also may lead to a new method to disentangle the influence of recalibrated proprioception and more traditional updated internal models on motor changes, such as reach aftereffects.
-Either way, even though the experiment was not designed to investigate this, the shifted generalization curves of changes in localization after classic and exposure training suggest they are generated by different mechanisms.
-This is in line with our earlier findings that proprioception generalizes differently from reach adaptation @Mostafa2015.
-
 ## Conclusion
 
-To sum up, after a training paradigm designed to prevent updating of predicted sensory consequences but allow recalibration of proprioception, we find substantial changes in where people localize their hand.
+To sum up, after a training paradigm designed to prevent updating of predicted sensory consequences but allowing recalibration of proprioception, we find substantial changes in where people localize their hand.
 This means that recalibrated proprioceptive estimates can explain shifts in hand localization.
-The distinct change in the direction of open-loop reaches we observe here, suggests recalibrated proprioception can contribute to motor adaptation in other contexts as well.
-Finally, we have some evidence that after exposure training, the shift in hand localization does not generalize around the trained target location.
-All of this confirms that different mechanisms underlie proprioceptive recalibration and motor adaptation.
+Since our participants also changed the direction of open-loop reaches, recalibrated proprioception seems to guide motor planning.
+This study confirms that sensory prediction error based learning and proprioceptive recalibration are different mechanisms that separately contribute to motor adaptation.
 

### 2019-05-09 paper/paper.Rmd
small revisions

@@
 output:
+  word_document: default
+  rticles::plos_article: default
+  pdf_document: default
   html_document:
     df_print: paged
-  pdf_document: default
-  rticles::plos_article: default
-  word_document: default
 csl: plos.csl
@@
 ---
+
+```{r, include=FALSE}
+options(tinytex.verbose = TRUE)
+```
 
@@
 Our lab, and others, have shown that our sense of where we feel our hand to be, proprioception, is also reliably recalibrated after visuomotor adaptation [@Cressman2009; @Hart2016; @Mostafa2014; @Mostafa2015; @Maksimovic2018; @Henriques2012; @Cressman2010a; @Barkley2014; @Salomonczyk2011; @Nourouzpour2014; @Henriques2014; @Cameron2012], and a comparable proprioceptive change is induced with force-field adaptation [@Ostry2016; @Ohashi2019].
-As proprioception, like predicted sensory consequences, also informs us on the location of our limbs, we have on occasion used a task that is very similar to hand localization to investigate this [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018].
+Just like efference-based predictions, proprioception also informs us on the location of our limbs.
+To investigate recalibrated proprioception we have on occasion used a task that is very similar to hand localization [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018].
 Although proprioceptive recalibration has been largely ignored as an explanation for changes in hand localization, we and others have shown that it accounts for a substantial part of the changes in localization, along with, but separate from, updates in predicted sensory consequences [@Hart2016; @Cameron2012].
@@
 Here we intend to further examine the contribution of proprioception to hand localization, relative to the contribution of predicted sensory consequences.
-To do this, we use a passive training paradigm with robot-generated  movements, which we refer to as ‘exposure training’, where the cursor always goes directly goes to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013].
+To do this, we use a training paradigm with robot-generated  movements, which we refer to as ‘exposure training’, where the cursor always goes directly goes to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013].
 This means there is no efference copy available and no visuomotor error-signal, both of which are required to update predicted sensory consequences.
@@
 
-## Exposure training
+### Exposure training
 
@@
 
-## No-cursor reaching
+### No-cursor reaching
 
@@
 
-## Localization
+### Localization
 
@@
 
-## Reach aftereffects
+### Reach aftereffects
 
@@
 
-## Generalization maxima
+### Generalization maxima
 

### 2019-05-11 paper/paper.Rmd
Plos needed a better ethics statement...

@@
 output:
+  rticles::plos_article: default
   word_document: default
-  rticles::plos_article: default
   pdf_document: default
@@
 All analyses presented here pertain to the remaining twenty-one participants (mean age: 20.1±2.3, age range: 18-25, 13 females), but the data of the three low-performing participants is included in the online dataset @Mostafa2017 ([https://osf.io/zfdth](https://osf.io/zfdth/)).
-Participants provided prior, written informed consent in accordance with the ethical guidelines set by the York Human subjects Review Subcommittee (#2014-240).
+Participants provided prior, written informed consent in accordance with the ethical guidelines set by the York Human Participants Review Committee (#2014-240).
 All had normal or corrected-to-normal vision, and received credit toward an undergraduate psychology course.

### 2019-05-25 ManuscriptMaterials.Rmd
Updated for R 3.6

@@
 
+Apparently it is lower. We don't think it matters much though.
+
 And the distribution of sex:
@@
 ```
+
+There is no difference in sex distribution between the groups.
 
@@
 ```
+
+Not great, but acceptable, except for the 15 degree location. We've decided to leave that location out of the analyses on hand localization (shifts).
 
@@
 This would predict an interaction between group and movement type, and we do see that (p=.035).
-(Including the three)
 
@@
 
-Also note the absence of an effect of hand angle in the classic group here -- probably because we excluded the 15 degree target, but it does appear that after classic training the generalization curve for delayed localization is much flatter.
+Also note the absence of an effect of hand angle in the classic group here -- probably because we excluded the 15 degree target, but it does appear that after classic training the generalization curve for delayed localization is much flatter/wider.
 
@@
 1. Rotated exposure training leads to changes in localization responses.
-
 2. In the exposure group, we find no difference between localization responses with and without efference copies.
@@
 5. The shift in localization is different across the workspace in the exposure group. I.e., there is a generalization _curve_.
-6.
-This effect is not present in the classic group, so that generalization can be said to be flat; there is no _curve_.
-**NOTE:** When we do include the 15 degree target, there is an effect of hand angle though.
+6. This effect is not present in the classic group, so that we don't see a generalization _curve_. **NOTE:** We see a curve when we include the 15 degree target though.
 
@@
 ```
+
+See the variance is not different between the two conditions.
 
@@
 - Panel A: there are substantial and persisting reach aftereffects
-- Panel B: that are somewhat lower than those in classic (but not that much? but only for some targets?)
+- Panel B: that are somewhat lower than those in classic (but not that much? or only for some targets?)
 
@@
 
-There doesn't seem to be an effect of iteration, and no interaction with target.
-Also, looking at the figure, there would not be decay, as the reach aftereffects seem to get larger.
-We continue assuming there is no substantial change in the reach aftereffects between the first iteration and the average across the other four iterations.
+There seems to be an effect of iteration, but no interaction with target.
+Looking at the figure, there would not be decay, as the reach aftereffects seem to get larger.
+We continue under the assumption there is **no decay** in the reach aftereffects between the first iteration and the average across the other four iterations, so we pool the data across iterations.
 We now also see an effect of target angle: the generalization curve is not flat.
@@
 The figure looked like there might be some difference between the two training types, mainly in where the generalization curves peak.
-There is no main effect of training type, so the overall magnitude of reach aftereffects is not different.
-
 However, the main take-away message from this analysis is that the motor changes measured with changes in open-loop reach direction are of similar magnitude in both types of training.
-That is, passive exposure training (proprioceptive recalibration) drives motor changes.
+That is, passive exposure training (proprioceptive recalibration) drives motor changes to a similar degree as classic training.
 
@@
 Alternatively, if you want to thoroughly check the results, you can run this chunk to delete the pre-calculated results (set delete=TRUE and download=FALSE), which will force the code to re-calculate it.
-By default, the code here will use "only" 1000 iterations for bootstrapping, which will run 20 times faster, but also guarantees that the confidence intervals are less precise than the ones we pre-calculated.
+By default, the code here will use "only" 1000 iterations for bootstrapping, which will run much faster, but also guarantees that the confidence intervals are less precise than the ones we pre-calculated.
 
@@
                    selectPerformance=selectPerformance,
-                   iterations=200)
+                   iterations=1000)
 ```
@@
 
-The algorithm decides to retain active localization as predictor, but the regression model isn't very good.
+The algorithm ends up with an intercept model.
 The problem could be that we need no-cursor data and localization data that are both recorded immediately after training.
@@
 
-The step-wise multiple regression algorithm thinks an intercept-only model is best.
+The step-wise multiple regression algorithm thinks an intercept-only model is best here too.
 Now is this because localization after classic training has little to no information on reach aftereffects, or is this because we have double the data in the exposure group?
@@
 
-And now we get the passive localization as in the initial regressions.
+And now we get the passive localization as the only predictor.
 
@@
 I'm not sure.
-There seems to be some relationship but it's not very strong.
+There might be some relationship but it's not very strong.
 So either there are other factors playing a role, the measurements are too noisy, or we don't have enough power to detect any relationships.

### 2019-05-25 shared.R

@@ -368,6 +368,7 @@ getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55
       }
     }
   }
+  
   if (length(removepps) > 0) {
     participants <- participants[-which(participants %in% unique(removepps))]
     cat('\nremoved these participants:\n')
@@ -415,10 +416,17 @@ getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55
   # get rotated - aligned difference, if required:
   if (difference) {
     groupdf <- aggregate(taperror_deg ~ group + participant + passive_b + handangle_deg, data=groupdf, FUN=diff, drop=FALSE)
-    groupdf$taperror_deg <- as.numeric(groupdf$taperror_deg)
+    ted <- groupdf$taperror_deg
+    ted_new <- c()
+    for (el in ted) {
+      if (is.numeric(el) & length(el) > 0) {
+        ted_new <- c(ted_new, el)
+      } else {
+        ted_new <- c(ted_new, NA)
+      }
+    }
+    groupdf$taperror_deg <- ted_new
   }
-  
-  
   
   # select only one movement type, if required:
   if (movementtype == 'active') {
@@ -429,8 +437,6 @@ getPointLocalization <- function(group, difference=TRUE, points=c(15,25,35,45,55
     groupdf <- groupdf[which(groupdf$passive_b == 1),]
     groupdf <- groupdf[,-which(names(groupdf) == c("passive_b"))]
   }
-  
-  
   
   return(groupdf)
   

### 2019-07-12 nocursor.R
calculate # trials removed (changes undone)

@@ -498,6 +498,7 @@ countSelectedNoCursors <- function(group, ignoreRepetitions=FALSE, selectPerform
         rotated     <- c(rotated, session)
         repetition  <- c(repetition, iter)
         trials      <- c(trials, (Ntrials/.21))
+        #trials      <- c(trials, Ntrials)
         
       }
       

### 2019-07-14 paper/paper.Rmd
plos one ethics rules
minor revisions to address review

--- 2019-05-09 paper/paper.Rmd
@@
 output:
+  pdf_document: default
+  rticles::plos_article: default
   word_document: default
-  rticles::plos_article: default
-  pdf_document: default
   html_document:
@@
 Additionally, such training leads to a recalibration of our sense of felt hand position based on afferent signals (that we call “proprioception” here) to be more aligned with the distorted visual feedback @Cressman2009.
-Both of these changes have been measured by asking people to localize where their unseen hand is – before and after training [@Clayton2013; @Izawa2012b; @Jones2012a; @Synofzik2008; @Yavari2016].
-While our lab has previously found that proprioception accounts for a substantial portion of the change in hand localization, and should hence contribute to motor adaptation @Hart2016, it is far from clear how much each process contributes or how to tease them apart.
+Both of these changes are measured by asking people to localize where their unseen hand is – before and after training [@Clayton2013; @Izawa2012b; @Jones2012a; @Synofzik2008; @Yavari2016].
+While our lab finds that proprioception accounts for a substantial portion of the change in hand localization, and should hence contribute to motor adaptation @Hart2016, it is far from clear how much each process contributes or how to tease them apart.
 Here we use training with robot-generated reaches, that prevents efference-based predictions and eliminates differences between actual and expected visual feedback and hence prevents updating predicted sensory consequences.
@@
 In motor adaptation, the actual sensory outcome is systematically different from the expected outcome, so that participants update their predictions on the outcome of the trained movements.
-In two previous visuomotor adaptation experiments, people were asked to make a movement and then indicate the location of, or “localize,” their unseen, right hand, before and after training with rotated visual feedback [@Izawa2012b; @Synofzik2008].
-Since there was no visual information available to the participants, the predicted sensory consequences of the movement should be used in localizing the unseen hand.
-Both studies found a significant shift in hand localization, providing evidence that predicted sensory consequences are indeed updated as a result of visuomotor rotation adaptation.
+In two previous visuomotor adaptation experiments, people are asked to make a movement and then indicate the location of, or “localize,” their unseen, right hand, before and after training with rotated visual feedback [@Izawa2012b; @Synofzik2008].
+Since there is no visual information available to the participants, the predicted sensory consequences of the movement should be used in localizing the unseen hand.
+Both studies find a significant shift in hand localization, providing evidence that predicted sensory consequences are indeed updated as a result of visuomotor rotation adaptation.
 
@@
 Just like efference-based predictions, proprioception also informs us on the location of our limbs.
-To investigate recalibrated proprioception we have on occasion used a task that is very similar to hand localization [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018].
-Although proprioceptive recalibration has been largely ignored as an explanation for changes in hand localization, we and others have shown that it accounts for a substantial part of the changes in localization, along with, but separate from, updates in predicted sensory consequences [@Hart2016; @Cameron2012].
+To investigate recalibrated proprioception we on occasion use a task that is very similar to hand localization [@Cressman2009; @Jones2012a; @Clayton2014; @Ruttle2018].
+Although proprioceptive recalibration is largely ignored as an explanation for changes in hand localization, we and others have shown that it accounts for a substantial part of the changes in localization, along with, but separate from, updates in predicted sensory consequences [@Hart2016; @Cameron2012].
 Nevertheless, it is far from clear how much proprioception and prediction each contribute to hand localization.
@@
 Here we intend to further examine the contribution of proprioception to hand localization, relative to the contribution of predicted sensory consequences.
-To do this, we use a training paradigm with robot-generated  movements, which we refer to as ‘exposure training’, where the cursor always goes directly goes to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013].
+To do this, we use a training paradigm with robot-generated  movements, which we refer to as ‘exposure training’, where the cursor always goes directly to the target [@Cameron2012; @Cressman2010; @Salomonczyk2013].
 This means there is no efference copy available and no visuomotor error-signal, both of which are required to update predicted sensory consequences.
@@
 Primarily, we will compare the magnitude of shifts in passive and active hand localization after visuoproprioceptive discrepancy exposure and classic visuomotor training.
-After exposure training there should be no difference, while after classic training, predicted sensory consequences should be updated but only contribute to active localization, so that active and passive localization changes should be different.
+This is because after exposure training only proprioception should be recalibrated, and since we think this contributes to active and passive hand localization equally, there should be no difference between active and passive localization shifts after exposure training.
+After classic training, on the other hand, proprioception is recalibrated too but in addition, predicted sensory consequences should also be updated, which only contributes to active localization, so that active and passive localization shifts should be different after classic training.
 For a secondary analysis we will compare the magnitude of reach aftereffects between exposure and classic training.
@@
 We set out to test the relative contributions of proprioception and efference-based prediction to hand localization.
-We use exposure training, with robot-generated reaches to prevent updates of predicted sensory consequences, but still elicit proprioceptive recalibration, and compare this with a classic training paradigm that should change both predicted sensory consequences and proprioception.
+We used exposure training, with robot-generated reaches to prevent updates of predicted sensory consequences, but still elicit proprioceptive recalibration, and compared this with a classic training paradigm that should change both predicted sensory consequences and proprioception.
 We then tested this in both active localization, that should use both sources – updated or not – and in passive localization, that should rely only on proprioception, which should be recalibrated equally by both training paradigms.
@@
 One participant was excluded for not following task instructions, and three were excluded for low performance on a task that ensures attention during the exposure training.
-All analyses presented here pertain to the remaining twenty-one participants (mean age: 20.1±2.3, age range: 18-25, 13 females), but the data of the three low-performing participants is included in the online dataset @Mostafa2017 ([https://osf.io/zfdth](https://osf.io/zfdth/)).
-Participants provided prior, written informed consent in accordance with the ethical guidelines set by the York Human subjects Review Subcommittee (#2014-240).
+All analyses presented here pertain to the remaining twenty-one participants (mean age: 20.1±2.3, age range: 18-25, 13 females), but the data of the three low-performing participants was included in the online dataset @Mostafa2017 ([https://osf.io/zfdth](https://osf.io/zfdth/)).
+Participants provided prior, written informed consent as approved by the York Human Participants Review Committee (#2014-240).
 All had normal or corrected-to-normal vision, and received credit toward an undergraduate psychology course.
 Participants were screened verbally, and all reported being right handed and not having any history of visual, neurological, and/or motor dysfunction.
-Published data from 21 further participants (mean age: 24, age range: 18-38, 11 females) that completed classic visuomotor training is used for comparison @Hart2016.
+Published data from 21 further participants (mean age: 24, age range: 18-38, 11 females) that completed classic visuomotor training was used for comparison @Hart2016.
 
@@
 The target, shown as a yellow disc, is located 10 cm away from the home position at 45°.
-In the rotated training tasks, the cursor (shown here as a green circle) represents the hand position rotated 30° relative to the home position.
+In the rotated training tasks, the cursor (shown here as a green circle) represented the hand position rotated 30° relative to the home position.
 **c)** No-cursor reach task.
-Targets are located 10 cm away from the home position at 15°, 25°, 35°, 45°, 55°, 65°, and 75°, shown by the yellow circles here (only one was shown on each trial).
-While reaching to one of these targets, no visual feedback on hand position is provided.
+Targets were located 10 cm away from the home position at 15°, 25°, 35°, 45°, 55°, 65°, and 75°, shown by the yellow circles here (only one was shown on each trial).
+While reaching to one of these targets, no visual feedback on hand position was provided.
 **d)** Localization task.
@@
 All participants performed active and passive localization as well as no-cursor reaching tasks in both a first, aligned session and a second, rotated session.
-There is an extra no-cursor reach block in the rotated session, to assess decay.
+There was an extra no-cursor reach block in the rotated session, to assess decay.
 The number of trials (numbers above the rectangles) in the no-cursor reaching tasks and both types of localization were the same in the aligned and rotated sessions, but while aligned training was kept to a minimum, rotated training was increased to ensure sufficient adaptation.
-The Exposure training group did passive and active localization twice, but the Classic training group did different localization tasks in the second half of each session, the data of which is not used here.
+The exposure training group did passive and active localization twice, but the classic training group did different localization tasks in the second half of each session, the data of which is not used here.
 **b)** Our main question was about the signals contributing to hand localization in the four different kinds of localization.
-According to our hypothesis, updated predicted sensory consequences should have only contributed to active hand localization after Classic training.](Fig2.eps)
+According to our hypothesis, updated predicted sensory consequences should have only contributed to active hand localization after classic training.](Fig2.eps)
 
@@
 In what we called ‘exposure training’ the participants did not move their hand toward the target, but the robot did.
-In this task (Fig 2a, gray boxes), the right hand was represented by a cursor (green disk, 1 cm in diameter, Fig 1b) located directly above participant’s thumb.
-The robot moved the participant’s unseen right hand (and the cursor) along a direct path, with a constant velocity of 10 cm/s, toward a visual yellow target disk and back to the starting position (1 cm in diameter, Fig 1b).
+In this task (Fig 2a, gray boxes), the right hand was represented by a cursor (green disk, 1 cm in diameter, Fig 1b) located directly above the participant’s thumb.
+The robot moved the participant’s unseen right hand (and the cursor) along a direct path, toward a visual yellow target disk (1 cm in diameter, Fig 1b).
+The velocity profile was bell-shaped, with a peak velocity of ~44 cm/s after ~300 ms.
+The hand reached the target after ~600 ms and was then kept in place for an additional ~400 ms.
+The hand was then placed back at the home position using an identical movement.
 The home position was located approximately 20 cm in front of participants and the visual target located 10 cm from the home position at 45° (Fig 1b).
@@
 The same visual training target at 45° was used, and the cursor kept moving straight to this target.
-However, the robot-generated hand path gradually rotated 30° CCW (Fig 1b) with respect to the cursor and target, in increments of 0.75°/trial, to prevent awareness of the rotation, so that the full rotation was reached after 45 trials.
+However, the robot-generated hand path gradually rotated 30° CCW (Fig 1b) with respect to the cursor and target, in increments of 0.75°/trial, to prevent awareness of the rotation, so that the full rotation was reached after 40 trials.
 This mimicked error-free responses to a gradual visuomotor rotation of 30° CW.
@@
 
-The trials in no-cursor reaching (Fig 1c; Fig 2, outlined boxes) served as a standard measure of motor adaptation.
+The trials in no-cursor reaching (Fig 1c; Fig 2, outlined boxes) served to elicit open-loop reaching as a standard measure of motor adaptation.
 On each of these trials participants were asked to reach with their unseen right hand to one of 7 visual targets, without any visual feedback of hand position.
@@
 A trial started with the robot handle at the home position and, after 500 ms, the home position disappeared and the target appeared, cuing the participants to reach for the target.
-Once the participants thought they had reached the target they held their position for 250 ms, and the target disappeared, while the home position re-appeared, cuing participants to move back to the home position along a straight, constrained path, to begin the next trial.
-The path was constrained so that if participants tried to move outside of the path, a resistance force, with a stiffness of 2 N/(mm/s) and a viscous damping of 5 N/(mm/s), was generated perpendicular to the path.
+Once the participants thought they had reached the target they held their position (criterion: 4 cm or further from the home position, velocity under 0.5 cm/s for 250 ms), and subsequently the target disappeared, while the home position re-appeared, cuing participants to move back to the home position along a straight, constrained path, to begin the next trial.
+The path was constrained so that if participants tried to move outside of the path, a resistant force, with a stiffness of 2 N/(mm/s) and a viscous damping of 5 N/(mm/s), was generated perpendicular to the path.
 In every iteration of the no-cursor reach task, each target was reached for three times, for a total of 21 reaches in pseudo-random order.
@@
 several movements back and forth, or a touch-screen response on the home position, instead of on the white arc).
+From the localization responses 1024 / 16800 (~6.1\%) trials were removed from the exposure group's data, and 220 / 8400 (~2.6\%) from the classic group's data.
+From the no-cursor reaches, 2 / 3984 (~0.05\%) were removed from the exposure group's data, and 1 / 3969 (~.025\%) was removed from the classic group's data.
 
@@
 
-We primarily wanted to test if hand localization shifts after exposure training, whether or not there were differences between active and passive localization shifts, and finally compare localization shifts after exposure training to those after classic visuomotor training.
+We primarily wanted to test if hand localization shifts after exposure training; whether or not there were differences between active and passive localization shifts; and finally compare localization shifts after exposure training to those after classic visuomotor training.
 
@@
 We noticed that the generalization curves after exposure and classic training looked different, and the planned analyses indeed showed interactions between hand angle and training paradigm.
-Specifically, the maxima after classic training are very close to the target, as expected, but for exposure training, the maxima are shifted (see Results).
-Since this is potentially interesting, we added two purely exploratory (i.e.
+Specifically, the maxima after classic training are very close to the target, as expected, but for exposure training, the maxima seemed shifted (see Results).
+Since this was potentially interesting, we added two purely exploratory (i.e.
 unplanned) analyses of the location of the maxima in the workspace.
 The same analysis was done for active localization and reach aftereffects, comparing generalization after exposure and classic training.
-Note that these analyses do not answer our main questions.
+Note that these analyses did not answer our main questions.
 
@@
 We can parcel out these contributions by measuring hand localization after both robot-generated (afferent signals only) and self-generated movements (both afferent and efferent signals).
-Finally, all analyses are performed with the data from the current experiment and those obtained in an earlier study that used an identical paradigm, but trained with self-generated movements, or “classic” training.
+Finally, all analyses are performed with the data from the current experiment and those obtained in an earlier study that uses an identical paradigm, but training with self-generated movements, or “classic” training.
 
 ```{r fig3, echo=FALSE, message=FALSE, fig.width=5.25, fig.height=2.5, fig.cap="\\label{fig:fig3}**Hand Localization** The shifts of the angles of touchscreen responses in all variations of the localization task, using spline-interpolated estimates for hand angles matching the reach targets in the no-cursor reach tasks.
+Shaded areas denote 95% confidence intervals.
 **a)** Localization shifts after exposure training.
@@
 
-Here we test our hypothesis that exposure training with a visual-proprioceptive discrepancy does not lead to changes in predicted sensory consequences.
-First, we can see that the difference between rotated and aligned localization responses is different from zero, after exposure training (Fig 3), so that it seems there are _shifts_ in localization that we can compare between tasks and training paradigms.
-To verify this we fit an LME to the localization errors throughout the workspace using _session_ (aligned or rotated), _movement type_ (active and passive) and _hand angle_ (25°, 35°, 45°, 55°, 65° and 75°), and all interactions as fixed effects and _participant_ as random effect.
-There is an effect of _session_ (F(1,450.5)=155.8; p<0.001), showing that training with exposure to a visual-proprioceptive discrepancy leads to changes in hand localization.
+Here we test our hypothesis that exposure training with a visual-proprioceptive discrepancy does not lead to changes in predicted sensory consequences, which predicts equal shifts in active and passive localization.
+First, we can see that the difference between rotated and aligned localization responses is different from zero, after exposure training (Fig 3a), so that it seems there are _shifts_ in localization that we can compare between tasks and training paradigms.
+To verify there is a shift in localization responses, we fit an LME model to the localization errors throughout the workspace using _session_ (aligned or rotated), _movement type_ (active and passive) and _hand angle_ (25°, 35°, 45°, 55°, 65° and 75°), and all interactions as fixed effects and _participant_ as random effect.
+There is an effect of _session_ (F(1,450.5)=155.8; p<0.001), showing that training with exposure to a visual-proprioceptive discrepancy leads to a shift in hand localization.
 There is also an effect of _hand angle_ (F(5,450.6)=6.54; p<.001) and an interaction between _hand angle_ and _session_ (F(5,450.3)=8.25; p<.001), but no other effects (all p>.60).
-Since localization responses shift, we use the difference between hand localization after rotated training and after aligned training as a measure of localization shift.
+Since exposure training led to a shift in localization responses, we use the difference between hand localization after rotated training and after aligned training as a measure of localization shift.
 
-Since the difference between active and passive localization stems only from the presence or absence of predicted sensory consequences, there should be no difference between the two if predicted sensory consequences are not changed by exposure training.
+Since the difference between active and passive localization stems only from the presence or absence of predicted sensory consequences, there should be no difference between the two if predicted sensory consequences are not updated by exposure training.
 At first glance, it seems there might be a difference between active and passive localization shifts in exposure training (see Fig 3a), although it is smaller than in classic training (Fig 3b).
@@
 We fit a linear mixed effects model (LME) to the change in localization using _movement type_ (active or passive localization) and _hand angle_, as well as their interaction as fixed effects and _participant_ as random effect.
-As we expected, there was no effect of _movement type_ (F(1,211.8)=0.07; p=0.79).
-There was an effect of _hand angle_ (F(5,212.2)=10.8; p<0.001), but no interaction between _hand angle_ and _movement type_ (F(5,211.8)=1.23, p=.29).
-The lack of an effect of movement type means we don’t have evidence that predicted sensory consequences contribute to localization in this paradigm.
+As we expected, there is no effect of _movement type_ (F(1,211.8)=0.07; p=0.79).
+There is an effect of _hand angle_ (F(5,212.2)=10.8; p<0.001), but no interaction between _hand angle_ and _movement type_ (F(5,211.8)=1.23, p=.29).
+The lack of an effect of _movement type_ means we don’t have evidence that predicted sensory consequences contribute to localization in this paradigm.
 
-In order to compare hand localization shifts after exposure training with those after classic training @Hart2016, we fit an LME to localization shift using _training type_ (exposure vs.
+Active localization shifts are larger than passive localization shifts after classic training (Fig 3b).
+In order to compare hand localization shifts after exposure training with those after classic training @Hart2016, we fit an LME model to localization shifts using _training type_ (exposure vs.
 classic), _movement type_ (active vs.
-passive) and _hand angle_ and all interactions as fixed effects and _participant_ as random effect.
+passive) and _hand angle_ and all interactions as fixed effects and _participant_ as random effect, and expect an interaction between _training type_ and _movement type_.
 There is a main effect of _movement type_ (F(1,422.0)=6.22; p=.013) and of _hand angle_ (F(5,422.7)=8.19; p<.001), as well as an interaction between _training type_ and _hand angle_ (F(5,422.7)=4.54; p<.001) and between _training type_ and _movement type_ (F(1,422.0)=4.48, p=.035), but there is no main effect of _training type_ (F(1,39.1)=0.92, p=.34) and no other effects (all p>.14).
-These results also suggest that the magnitude of the shifts in localization are comparable between classic and exposure training, but that the pattern of generalization is different.
+These results suggest that the magnitude of the shifts in localization are comparable between classic and exposure training (no main effect of _training type_), but that the pattern of generalization is different (interaction between _training type_ and _hand angle_).
+Importantly, the interaction between _training type_ and _movement type_ is in line with our main prediction that the difference between active and passive localization shifts should be different in the two training types.
 
-To address our main question, we will consider the interaction between _training type_ (exposure vs.
+To further address our main question, we will consider the interaction between _training type_ (exposure vs.
 classic) and _movement type_ (active vs.
@@
 However, since there is no difference between active and passive localization shifts after exposure training alone, the interaction between _training type_ and _movement type_ should be caused by an effect of _movement type_ on the localization shifts after classic training, as we found previously @Hart2016.
-This means that shifts in hand localization after exposure training indeed rely on recalibrated proprioception alone, while after classic training, there also is a contribution of predicted sensory consequences to active localization.
+This strongly suggests that shifts in hand localization after exposure training indeed rely on recalibrated proprioception alone, while after classic training, there also is a contribution of predicted sensory consequences to active localization.
 
@@
 
-In active movements there could have been more information available to localize the hand, which could have lead to more precise responses.
-To test this, we calculated the variance of localization responses relative to a smoothed spline expressing the bias of localization responses in each participant across the workspace.
-A paired-sample t-test showed that the variance of active and passive localization were not different from each other in the aligned session (t(20)=1.178, p=.253) as well as the rotated session (t(20)=-1.066, p=.299) in the exposure training group.
-This confirmed earlier findings where we also did not find that active localization was more precise than passive localization [@Jones2010b; @Jones2012a; @Hart2016].
+In active movements there could be more information available to localize the hand, which could lead to more precise responses.
+To test this, we calculate the variance of localization responses relative to a smoothed spline expressing the bias of localization responses in each participant across the workspace.
+A paired-sample t-test shows that the variance of active and passive localization are not different from each other in the aligned session (t(20)=1.178, p=.253) as well as the rotated session (t(20)=1.066, p=.299) in the exposure training group.
+This confirms earlier studies where we also do not find that active localization is more precise than passive localization [@Jones2010b; @Jones2012a; @Hart2016].
 
@@
 ```{r fig4, echo=FALSE, fig.width=7.5, fig.height=3,fig.cap="\\label{fig:fig4}**Reach aftereffects** Changes of the angle of reach endpoints in the no-cursor tasks after rotated training.
+Shaded areas denote 95% confidence intervals.
 **a)** Reach aftereffects at different time points in the rotated session.
@@
 
-To see if reach aftereffects decay during the localization tasks, we compare reach aftereffects in the initial no-cursor block, that immediately follow training, with those in the later blocks that follow a localization task.
-We fit an LME to the reach aftereffects with _iteration_ (initial vs.
+To see if reach aftereffects decay during the localization tasks, we compare reach aftereffects in the initial no-cursor block, that immediately follow training, with those in the later blocks that follow a localization task (Fig 4a).
+We fit an LME model to the reach aftereffects with _iteration_ (initial vs.
 later no-cursor blocks) and _target_ (as above) as well as their interaction as fixed effects and participant as random effect.
@@
 There is an effect of _target_ (F(6,260)=6.29, p<.001) but no interaction (F(6,260)=0.58, p=0.74).
-In sum, reach aftereffects are not appreciably different right after training and after the localization tasks.
+In sum, reach aftereffects are not appreciably different right after training or after the localization tasks.
 In other words, there is likely no noticeable decay of reach aftereffects during the localization tasks, so that we can collapse the data across iterations.
@@
 It appears as if there is little overall difference in magnitude, but there might be a shift of the generalization curve.
-We fit an LME to reach aftereffects with _training type_ (classic vs.
+We fit an LME model to reach aftereffects with _training type_ (classic vs.
 exposure), _target_ (as above) as well as their interaction as fixed effects and _participant_ as random effect.
@@
 While not our main question, we will add exploratory analyses on the potentially different generalization patterns of localization shifts and reach aftereffects after classic and exposure training (Fig 5).
-The LMEs above indicate no difference in overall amplitude of localization shifts or reach aftereffects between the groups, so the interaction between training type and hand angle (or target for reach aftereffects) might stem from a shifted generalization curve after exposure training, in contrast to classic training where the maximum is close to the trained target @Krakauer2000.
+The LME models above indicate no difference in overall amplitude of localization shifts or reach aftereffects between the groups, so the interaction between training type and hand angle (or target for reach aftereffects) might stem from a shifted generalization curve after exposure training, in contrast to classic training where the maximum is close to the trained target @Krakauer2000.
 Using the active localization shifts only (which are larger, and should arguably be more similar to reach aftereffects, because they include predicted sensory consequences), we bootstrap a 95\% confidence interval for the maximum localization shift, by resampling participants within each group 100,000 times and taking the centre of a normal curve fit.
@@
 ```{r fig5, echo=FALSE, fig.width=7.5, fig.height=3,fig.cap="\\label{fig:fig5}**Generalization maxima** Locations of generalization maxima in the workspace for active localization and reach aftereffects.
-_Blue_: Exposure training, and _red_: classic training, with shaded areas at the bottom marking the 95% confidence intervals for the location of the maximum of each curve (vertical red and blue lines in shaded areas indicate the median; colored arrows pointing downward from the curves indicate the maxima of the curves fitted to all data).
+_Blue_: Exposure training, and _red_: classic training.
+The average data is shown as faint lines, and the curves are Gaussians fitted to all the data.
+Shaded areas at the bottom mark the 95% confidence intervals for the location of the maximum of each curve (vertical red and blue lines in shaded areas indicate the median; colored arrows pointing downward from the curves indicate the maxima of the curves fitted to all data).
 Downward black arrows: visual trained target at 45°.
@@
 
-The LMEs for localization shifts and reach aftereffects indicate a different generalization curve after exposure and classic training, and this can be explained by a different position of the generalization curves after the two types training.
+The LME models for localization shifts and reach aftereffects indicate a different generalization curve after exposure and classic training, and this can be explained by a different position of the generalization curves after the two types of training.
 The unrealistically high upper bound for the confidence interval of the peak of the generalization curve of reach aftereffects after exposure training, indicates that we should sample a wider part of the workspace in future experiments to get more accurate fits.
@@
 Results similar to what we find here were reported in a study by Cameron and colleagues @Cameron2012, using gain modulation of visual feedback of single-joint hand movements around the elbow.
-Their within-subjects experiment included both training with volitional movements as well as with machine-generated movements and also tested perception of movements that were either passive or active.
-They too found a robust change in passive perception of hand movement (using a different measure), and these changes did not differ between the two types of training.
-Similarly, they found shifts in what we might call “active localization,” although the task is different, after both training types.
-Like here, these shifts are larger after classic training as compared to exposure training.
-They also found that exposure to a visuo-proprioceptive discrepancy leads to reach aftereffects, although these were smaller than those produced following “classical” training with altered visual gain.
+Their within-subjects experiment includes both training with volitional movements as well as with machine-generated movements and also tests perception of movements that were either passive or active.
+They too find a robust change in passive perception of hand movement (using a different measure), and these changes do not differ between the two types of training.
+Similarly, they find shifts in what we might call “active localization,” although the task is different, after both training types.
+These shifts are larger after classic training as compared to exposure training.
+They also find that exposure to a visuo-proprioceptive discrepancy leads to reach aftereffects, although these are smaller than those produced following “classical” training with altered visual gain.
 Both our findings, and those of Cameron et al.
